---
title: "Tourism (TR) Goal Data Preparation"
output:
  github_document:
    toc: true
    toc_depth: 3
params:
    datasource: csv
---

<br>

```{r preamble prep, message = FALSE}
loc <- here::here("prep", "TR")

source(here::here("R", "setup.R"))
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = TRUE, results = "hide", fig.width = 9.5, fig.height = 6)

bkgd_path <- here::here("supplement", "goal_summaries", "TR.Rmd")
data_path <- here::here("data", "TR", version_year, "tr_data.rmd")
refs_path <- file.path(loc, "tr_references.Rmd")
```

## 1. Background

```{r tr goal background, child = bkgd_path, results = "asis", echo = FALSE}
```

<br/>

## 2. Data

This prep document is used to generate and explore the following data layers:

- `tr_accommodations_bhi2019.csv` 
- `tr_coastal_tourism_gva_bhi2019.csv` 

These are saved to the `layers` folder. These are derived from or informed by the raw datasets from Eurostat and coastal tourism revenue data from the EU Blue Growth 2020 Report.

<br>

```{r tr goal data, child = data_path, results = "asis", echo = FALSE}
```

<br/>

## 3. Prep: Wrangling & Derivations, Checks/Evaluation, Gapfilling

#### 3.1 Save Coastal Tourism Data Layer

Already wrangled and combined tourism datasets in `tr_data.rmd` so just need to visualize to plan for status calculation and reference points...

```{r wrangle and save coastal tourism data layers}
## already wrangled and combined tourism datasets in tr_data.rmd
## read data saved in data intermediate folder
coastal_tourism_value <- read_csv(file.path(dir_prep, "data", "TR", version_year, "intermediate", "tourism_data_combined.csv"))


## join with BHI regions
bhi_tourism_value <- coastal_tourism_value %>% 
  rename(eez = country) %>% 
  left_join(readr::read_csv(file.path(dir_prep, "supplement", "lookup_tabs", "rgns_complete.csv")), by = "eez") %>% 
  select(region_id, year, cntry_tourism_gva = coastal_tour_country_gva_estimate)

bhi_tourism_accomodations <- coastal_tourism_value %>% 
  rename(eez = country) %>% 
  left_join(readr::read_csv(file.path(dir_prep, "supplement", "lookup_tabs", "rgns_complete.csv")), by = "eez") %>% 
  select(region_id, year, accom_per_area = avg_accomm_per_area)


## save TR data layers
write_csv(bhi_tourism_value, file.path(dir_layers, "tr_coastal_tourism_gva_bhi2019.csv"))
write_csv(bhi_tourism_accomodations, file.path(dir_layers, "tr_accommodations_bhi2019.csv"))
```
<br/>


## 4. Visualizing Data Layers

#### 4.1 Timeseries Plots: check regional tourism accommodations time series

```{r plot and check regional employment time series, results = "show", fig.width = 9.5, fig.height = 4}
gridExtra::grid.arrange(
  ggplot(coastal_tourism_value) +
    geom_col(aes(x = year, y = avg_accomm_per_area), size = 0.8) +
    facet_wrap(~country, ncol = 2) +
    labs(x = NULL, y = "Accommodations per sq. km \n"),
  ggplot(coastal_tourism_value) +
    geom_col(aes(x = year, y = coastal_tour_country_gva_estimate), size = 0.8) +
    facet_wrap(~country, ncol = 2) +
    labs(x = NULL, y = "Coastal Tourism Revenue (GVA, M€) \n"),
  ggplot(coastal_tourism_value) +
    geom_col(aes(x = year, y = tourism_gva_per_accom), size = 0.8) +
    facet_wrap(~country, ncol = 2) +
    labs(x = NULL, y = "Coastal Tourism Revenue relative to \nAccommodations Density (M€/Nights/Sq.km) \n"),
  nrow = 1
)
```

<br>

## 5. Considerations for `BHI3.0`

An interesting [report form 2008 about how tourism and recreation industries in the Baltic Sea area are affected by the state of the marine environment](https://www.naturvardsverket.se/Documents/publikationer/978-91-620-5878-4.pdf). May contain interesting information relevant to out Tourism & recreation goal, particularly if we want to incorporate in the future some kind of sustainability measure.

The [Baltic Sea Tourism Center](https://bstc.eu/) may also be a good source of information, to start upgrading the BHI TR goal.

Also the World Economic Forum, [Travel and Tourism competitiveness reports](https://www.weforum.org/reports?utf8=%E2%9C%93&query=travel+and+tourism+competitiveness) contain many key preformance indicators, including and environmental sustainability indicator (though these are not specific to coastal tourism).

<br>

## 6. References

```{r tr goal references, child = refs_path, results = "asis", echo = FALSE}
```

<br>