---
title: "Invasive Species Pressure Layer Data Preparation"
output:
  github_document:
    toc: true
    toc_depth: 3
params: 
    datasource: csv
---
<br>

```{r invasive spp setup, message = FALSE}
## root location of the raw data
source(here::here("R", "setup.R"))
source(here::here("R", "spatial.R"))
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = TRUE, results = "hide", fig.width = 9.5, fig.height = 6)
source(here::here("R", "setup.R"))
loc <- here::here("prep", "pressures", "invasive_spp")
data_path <- here::here("data", "pressures", "invasive_spp", version_year, "invasive_spp_data.rmd")

dir_A <- file.path(dirname(dir_prep), "bhi-data")
dir_B <- file.path(dirname(dir_prep), "bhi-data", "BHI 2.0")
dir_rawdata <- file.path(dir_B, "Pressure", "invasive_spp")
```
<br>

## 1. Background {-}

Due to increasing shipping, more alien species are finding their way into the Baltic Sea than ever before. 
These invasive species can induce considerable changes in the structure and dynamics of marine ecosystems. Among the goals affected by this pressure, the Iconic Species (ICO) sub-goal and Biodiversity (BD) goal are particularly affected.

## 2. Data {-}

This prep document is used to generate and explore the following data layer:

- `sp_invasives_bhi2019.csv` 

This is saved to the `layers` folder. These are derived from or informed by the raw data obtained from Aquanis database.

<br>

```{r invasive species pressure prep data, child = data_path, results = "asis", echo = FALSE}
```

<br/>

## 3. Prep: Wrangling & Derivations, Checks/Evaluation, Gapfilling {-} 

```{r read in invasive spp data}
country_score <- read_delim(file.path(dir_rawdata, "invasive_spp.csv"), delim = ";") %>% 
  mutate(ref_pt = 10, 
         score = min(1, num_invasive/ref_pt))

bhi_basin_lookup <- read_delim(file.path(dir_prep, "supplement", "lookup_tabs", "bhi_basin_country_lookup.csv"), delim = ";") %>% 
  rename(country = rgn_nam) 
```
<br>

## 4 Calculate Invasive Species Pressure Scores {-}

**Score** = #_of_species_introduction_event / reference_point

**Reference point** = 10 introduction events (with no additional information, this is an arbitrary number set for now, as for BHI 1.0) 

```{r calculate pressure score}
bhi_score <- full_join(country_score, bhi_basin_lookup, by = 'country') %>% 
  dplyr::select(rgn_id = BHI_ID, 
         pressure_score = score)

# Save pressure layer
#write_csv(bhi_score, file.path(dir_layers, "sp_invasives_bhi2019.csv"))
```
<br>
