---
title: "Livelihoods and Economies (LE) - Livelihoods (LIV) Subgoal Data Preparation"
output:
  github_document:
    toc: true
    toc_depth: 3
params:
    datasource: csv
---


```{r liv prep preamble, include = FALSE}
loc <- here::here("prep", "LIV")

source(here::here("R", "setup.R"))
knitr::opts_chunk$set(message = FALSE, warning = FALSE, results = "hide")

dir_rawdata <- file.path(dir_B, "Goals", "LE", "LIV")
bkgd_path <- here::here("supplement", "goal_summaries", "LIV.Rmd")
data_path <- here::here("data", "LIV", version_year, "liv_data.rmd")
refs_path <- file.path(loc, "liv_references.Rmd")
# file.exists(c(loc, bkgd_path, data_path, refs_path))
```

## 1. Background

```{r liv prep background, child = bkgd_path, results = "asis", echo = FALSE}
```

<br/>

## 2. Data

This prep document is used to generate and explore the following data layers:

- `le_liv_coastal_employ_bhi2019.csv`
- `le_liv_coastal_pop_bhi2019.csv` 
- `le_liv_national_pop_bhi2019.csv` 

These are saved to the `layers` folder. Saved to `data/LIV/v2019/intermediate` are intermediate datasets: `rgn_coastalpop_bhi2019.csv` and `DATASETINTEMED2.csv`. All these are derived from or informed by the raw datasets from Eurostat employment rates and population change, and population densities from the [HYDE database](https://themasites.pbl.nl/tridion/en/themasites/hyde/).

<br>

```{r liv data, child = data_path, results = "asis", echo = FALSE}
```

<br/>

## 3. Prep: Wrangling & Derivations, Checks/Evaluation, Gapfilling

#### 3.1 Spatial Data {-}

[Eurostat NUTS regions shapefiles can be found here](https://ec.europa.eu/eurostat/web/gisco/geodata/reference-data/administrative-units-statistical-units/nuts). The NUTS 2016, 2013, 2010, 2006, and 2003 were downloaded at 1:10 million resolution (NUTS_RG_10M_year_4326_LEVL_2or3 files within the zipped downloads are the polygons in espg 4326 projection).

```{r nuts polygons, echo = TRUE, message = FALSE, warning = FALSE}
##  NUTS3 shapefiles because coastal regions are distinguished from non-coastal at this level
nutsshpfiles <- grep(
  "NUTS_RG_10M_[0-9]{4}_4326_LEVL_3/.*\\.shp", 
  list.files(file.path(dirname(dir_B), "Shapefiles", "all_NUTS_shapefiles"), recursive = TRUE),
  value = TRUE
)
## will check 2013 in cases where matching with 2016 version doesn't work
## gives GDAL Errors / CPL_read_ogr warnings, but still reads in the shapefiles...
for(f in grep("2013|2016", nutsshpfiles, value = TRUE)){
  shp <- sf::st_read(
    dsn = file.path(dirname(dir_B), "Shapefiles", "all_NUTS_shapefiles", stringr::str_extract(f, "ref.*LEVL_3(?=/)")),
    stringsAsFactors = FALSE
  )
  shp_baltic <- shp %>% filter(CNTR_CODE %in% c("SE", "FI", "EE", "LV", "LT", "PL", "DE", "DK"))
  assign(sprintf("nuts3_%s_shp", stringr::str_extract(f, "(?<=10M_)[0-9]{4}")), shp_baltic)
}

## 25km inland buffer for BHI regions
buffer_25km_inland <- sf::st_read(
  file.path(dirname(dir_B), "Shapefiles", "BHI_shapefile_25km_buffer"),
  "BHI_shapefile_25km_buffer"
)
buffer_25km_inland <- cbind(
  buffer_25km_inland,
  buffer_area_km2 = as.numeric(sf::st_area(buffer_25km_inland)/(1000^2))
)


## create some spatial objects for making maps later
buffer_25k_simple <- rmapshaper::ms_simplify(input = buffer_25km_inland) %>% sf::st_as_sf()

basemap <- ggplot2::ggplot(rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")) +
  geom_sf(size = 0.1, color = "burlywood", alpha = 0.4) +
  theme(panel.background = element_rect(fill = "#F8FBFC", color = "#E2EEF3")) +
  scale_x_continuous(limit = c(4, 32)) +
  scale_y_continuous(limit = c(53.5, 66)) 
```

<br>

### 3.2 Clean and Merge Employment Rate Data: standardize units, rename fields/variables {-}

```{r read in datasets with liv employment by country, echo = TRUE, message = FALSE, warning = FALSE}
employ_data <- do.call(rbind, lapply(
  grep("emp-rate.csv", list.files(dir_rawdata, recursive = TRUE, full.names = TRUE), value = TRUE),
  function(x){
    joindf <- read.delim(x, sep = ";")
    colnames(joindf) <- c("sector_or_metric", "2009", "2011", "2013", "2015", "2016", "2017", "2018")
    joindf <- joindf %>% 
      filter(str_length(sector_or_metric) > 0) %>% 
      tidyr::pivot_longer(
        cols = starts_with("20"), 
        names_to = "year", 
        values_to = "persons_emp_thousand"
      ) %>% 
      mutate(sector_or_metric = str_to_lower(str_replace_all(sector_or_metric, " ", "_"))) %>% 
      mutate(persons_emp_thousand = str_replace(persons_emp_thousand, ",", ".")) %>% 
      mutate(persons_emp_thousand = str_replace(persons_emp_thousand, " ", "")) %>% 
      tidyr::pivot_wider(names_from = sector_or_metric, values_from = persons_emp_thousand) %>% 
      mutate(country = str_extract(x, "(?<=NUTS3/)[a-z]+(?=_emp-rate\\.csv)"))
    return(joindf)
  }
))
```


### 3.3 Clean Population Data: standardize units, rename fields/variables {-}

```{r BHI coastal population estimation from NUTS3 population data, echo = TRUE, message = FALSE, warning = FALSE}
## regional employment per NUTS3 assessemnt areas, raw data from eurostat
regional_pop_data <- file.path(
    dir_rawdata, "Population_NUTS3", 
    "demo_r_pjangrp3_w_nutscodes", 
    "demo_r_pjangrp3_1_Data.csv"
  )
regional_pop <- read.csv(
  regional_pop_data, 
  stringsAsFactors = FALSE
)
```

```{r clean regional pop data, echo = TRUE, message = FALSE, warning = FALSE}
rgn_pop_clean <- regional_pop %>%
  filter(SEX == "Total") %>% 
  dplyr::select(
    year = TIME, nuts3 = GEO,
    value = Value,
    flag_notes = Flag.and.Footnotes
  ) %>%
  mutate(
    value = as.numeric(stringr::str_remove_all(value, ",")),
    flag_notes = case_when(
      flag_notes == "b" ~ "break in timeseries",
      flag_notes == "u" ~ "low reliability",
      flag_notes == "bu" ~ "break in timeseries and low reliability",
      !flag_notes %in% c("b", "u", "bu") ~ ""
    )
  )

## check dataflags
unique(rgn_pop_clean$flag_notes)
## no rows "low reliability" or "break in timeseries and low reliability"
## breaks in timeseries a concern?
filter(rgn_pop_clean, flag_notes == "break in timeseries")

## remove flags_notes
rgn_pop_clean <- rgn_pop_clean %>% 
  dplyr::select(-flag_notes) %>% 
  filter(substr(nuts3, 1, 2) %in% c("SE", "FI", "EE", "LV", "LT", "PL", "DE", "DK"))
# summary(rgn_pop_clean$value)
# summary(filter(rgn_pop_clean, str_length(nuts3)>4)$value)
```


**Join population info with NUTS3 shapefiles**

```{r join pop data with nuts3 shapefile, results = "show", fig.width = 9.5, fig.height = 8}
rgn_pop_nuts_spatial <- tidyr::pivot_wider(
  rgn_pop_clean,
  names_from = year, names_prefix = "pop_", 
  values_from = value
)
rgn_pop_nuts_spatial <- nuts3_2016_shp %>% 
  sf::st_transform(crs = 4326) %>% 
  rename(nuts3 = NUTS_ID) %>% 
  left_join(rgn_pop_nuts_spatial, by = "nuts3") %>% 
  filter(COAST_TYPE %in% 1:2)

## plotting initial values, raw NUTS2 regions employment rates
## loop through all years to check population data alignment with reporting regions...
basemap + 
  geom_sf(data = rgn_pop_nuts_spatial, aes(fill = pop_2019), size = 0.1) +
  geom_sf(data = buffer_25k_simple, fill = NA, color = "red", size = 0.2) +
  scale_fill_viridis_c() +
  labs(fill = "Population") +
  theme(legend.position = c(0.15, 0.85)) +
  ggtitle("Population 2019")
```



### 3.4 Wrangle and Save LIV data Layers {-}

```{r merge with bhi regions save data layers for the liv subgoal, echo = TRUE, message = FALSE, warning = FALSE}
rgns_complete <- read_csv(file.path(dir_prep, "supplement", "lookup_tabs", "rgns_complete.csv"))

## join bhi regions with employment rates ----
emp_lyr <- left_join(
  employ_data %>% 
    select(year, eez = country, blue_economy_pct_of_natl = `blue_economy_(%_of_national_jobs)`) %>% 
    mutate(blue_economy_pct_of_natl = str_extract(blue_economy_pct_of_natl, "[0-9]\\.[0-9]")) %>% 
    mutate(blue_economy_pct_of_natl = as.numeric(blue_economy_pct_of_natl)) %>% 
    mutate(
      eez = str_to_sentence(eez),
      year = as.numeric(year)
    ),
  rgns_complete %>% 
    select(eez, region_id, region_name) %>% 
    mutate(region_id = as.numeric(region_id)),
  by = "eez"
)
emp_lyr <- emp_lyr %>% filter(!is.na(region_id))

## join bhi regions with national coastal population ----
rgn_coast_pop <- left_join(
  rgn_pop_nuts_spatial %>% 
    st_drop_geometry() %>% 
    tidyr::pivot_longer(
      cols = starts_with("pop_"), 
      names_to = "year", names_prefix = "pop_", 
      values_to = "population"
    ) %>% 
    group_by(year, CNTR_CODE) %>% 
    summarize(natl_coastal_pop = sum(population)) %>% 
    ungroup() %>% 
    mutate(eez = case_when(
      CNTR_CODE == "SE" ~ "Sweden",
      CNTR_CODE == "FI" ~ "Finland",
      CNTR_CODE == "EE" ~ "Estonia",
      CNTR_CODE == "LV" ~ "Latvia",
      CNTR_CODE == "LT" ~ "Lithuania",
      CNTR_CODE == "PL" ~ "Poland",
      CNTR_CODE == "DE" ~ "Germany",
      CNTR_CODE == "DK" ~ "Denmark"
    )),
  rgns_complete %>% 
    select(eez, region_id, region_name),
  by = "eez"
)

## join bhi regions with national total population ----
natl_pop <- left_join(
  rgn_pop_clean %>% 
    rename(natl_population = value) %>% 
    filter(nuts3 %in% c("SE", "FI", "EE", "LV", "LT", "PL", "DE", "DK")) %>% 
    mutate(eez = case_when(
      nuts3 == "SE" ~ "Sweden",
      nuts3 == "FI" ~ "Finland",
      nuts3 == "EE" ~ "Estonia",
      nuts3 == "LV" ~ "Latvia",
      nuts3 == "LT" ~ "Lithuania",
      nuts3 == "PL" ~ "Poland",
      nuts3 == "DE" ~ "Germany",
      nuts3 == "DK" ~ "Denmark"
    )),
  rgns_complete %>% 
    select(eez, region_id, region_name),
  by = "eez"
)


## save layers ----
write_csv(
  select(emp_lyr, region_id, year, blue_economy_pct_of_natl),
  file.path(dir_layers, "le_liv_coastal_employ_bhi2019.csv")
)
write_csv(
  select(rgn_coast_pop, region_id, year, natl_coastal_pop),
  file.path(dir_layers, "le_liv_coastal_pop_bhi2019.csv")
)
write_csv(
  select(natl_pop, region_id, year, natl_population),
  file.path(dir_layers, "le_liv_national_pop_bhi2019.csv")
)
```


<br/>

## 4. Visualizing Data Layers {-}

**If LIV Status was calculated using blue economy Employment rates**

```{r calculate liv status, echo = TRUE, message = FALSE, warning = FALSE}
coastal_pop_percentage <- left_join(
  rgn_coast_pop %>% 
    select(region_id, region_name, eez, year, natl_coastal_pop) %>% 
    mutate(year = as.numeric(year)),
  natl_pop %>% 
    select(region_id, region_name, eez, year, natl_population),
  by = c("region_id", "region_name", "eez", "year")
)
coastal_pop_percentage <- coastal_pop_percentage %>% 
  mutate(coastal_pop_pct = natl_coastal_pop/natl_population)

liv_status <- emp_lyr %>% 
  filter(year %in% unique(coastal_pop_percentage$year)) %>% 
  left_join(coastal_pop_percentage, by = c("region_id", "region_name", "eez", "year")) %>% 
  select(region_id, region_name, eez, year, blue_economy_pct_of_natl, coastal_pop_pct) %>% 
  mutate(coastal_emp_vs_coastal_pop = blue_economy_pct_of_natl/coastal_pop_pct)

# distinct(liv_status, eez, year, blue_economy_pct_of_natl, coastal_pop_pct, coastal_emp_vs_coastal_pop)
```

```{r plot of liv status, fig.width = 9.5, fig.height = 9, results = "show"}
gridExtra::grid.arrange(
  ggplot(liv_status) + 
    geom_col(aes(x = year, y = blue_economy_pct_of_natl, fill = blue_economy_pct_of_natl)) +
    scale_fill_gradient(low = "grey", high = "paleturquoise") +
    facet_wrap(~region_name, labeller = label_wrap_gen(width = 15), ncol = 10) +
    labs(x = NULL, y = NULL, fill = "%", title = "Blue Econ. Empl. % of National Empl.") +
    theme(legend.direction = "horizontal", legend.position = c(0.9, 0.01), axis.text.x = element_text(angle = 90)),
  ggplot(liv_status) + 
    geom_col(aes(x = year, y = coastal_emp_vs_coastal_pop, fill = coastal_emp_vs_coastal_pop)) +
    scale_fill_gradient(low = "grey", high = "paleturquoise") +
    facet_wrap(~region_name, labeller = label_wrap_gen(width = 15), ncol = 10) +
    labs(x = NULL, y = NULL, fill = "Ratio", title = "Blue Econ. Empl. %\n Normalized by Coastal % of National Pop.") +
    theme(legend.direction = "horizontal", legend.position = c(0.9, 0.01), axis.text.x = element_text(angle = 90)),
  nrow = 2
)
```


<br/>

## 5. Considerations for `BHI3.0` {-}

In the EU Blue Growth Reports, the Blue Economy metrics by country don't distinguish between coasts i.e. Baltic vs North Sea.

Consider: should employment in marine-related sectors be normalized just by coastal/total national population, or is potential for coastal jobs more related to geography (e.g. length coastline vs total area)?

Use [Eurostat API for Administrative Units](https://ec.europa.eu/eurostat/web/gisco/geodata/reference-data/administrative-units-statistical-units) and [R package 'eurostat: Tools for Eurostat Open Data'](https://cran.r-project.org/web/packages/eurostat/index.html). See tutorial about the R package here: https://ropengov.github.io/eurostat/articles/eurostat_tutorial.html

[Eurostat regional yearbook 2018 > Background Maps > GEOSTAT population grid 2011](https://ec.europa.eu/eurostat/statistical-atlas/gis/viewer/?mids=BKGCNT,BKGPGR,CNTOVL&o=1,1,0.7&ch=BKG,C02&center=56.62571,21.14265,5&) shows how population density along coast is typically higher than inland; using pop. reported by NUTS2 or even NUTS3 spatial units to cannnot result in accurate representation of the coastal population...


## 6. References

```{r References, child = refs_path, results = "asis", echo = FALSE}
```
