---
title: "Livelihoods (LIV) Subgoal Data Preparation"
output:
  github_document:
    toc: true
    toc_depth: 3
params: 
    datasource: csv
---

```{r Preamble, include = FALSE}
loc <- file.path(here::here(), "prep", "LIV")

source(file.path(here::here(), "R", "common.R"))
knitr::opts_chunk$set(message = FALSE, warning = FALSE, results = "hide")

bkgd_path <- file.path(here::here(), "ref", "goal_summaries", "LIV.Rmd")
data_path <- file.path(here::here(), "data", "LIV", assessment_yr, "LIV_data.rmd")
refs_path <- file.path(loc, "LIV_references.Rmd")
```

## 1. Background

```{r Background, child = bkgd_path, results = "asis", echo = FALSE}
```

<br/>

## 2. Data
<!-- printing kable tables not working with this approach of 'child = data_path, results = "asis"' -->

```{r Data, child = data_path, results = "asis", echo = FALSE}
```

<br/>

## 3. Prep: Wrangling & Derivations, Checks/Evaluation, Gapfilling

### 3.1 Goal Model

$$X_{LIV} = \frac{\left(RgnEmployRate / NatlEmployRate \right)_{c}}{\left(RgnEmployRate / NatlEmployRate \right)_{r}}$$

- $c$ = current year
- $r$ = reference year, which is where the region:country employment ratio was maximized within the past 5 years
- Region is the BHI region:
   - Each BHI region is composed by one or more NUTS2 regions.  
   - NUTS2 employment percentage multiplied by the by population in the 25km inland buffer associated with a BHI region.  Sum across all associated with a BHI region to get the total employed people in the BHI region.
   - Regional Employment rate = # of employment/regional population
- For country data, we have employment rate data directly


### 3.2 

<br/>

## 4. Visualizing Data Layers

<br/>

## 5. References

```{r References, child = refs_path, results = "asis", echo = FALSE}
```
