---
title: "Artisanal Fishing Opportunities (AO)"
output:
  github_document:
    toc: true
    toc_depth: 3
params:
    datasource: csv
---

<br>

```{r ao preamble, message = FALSE, warning = FALSE, include = FALSE}
loc <- here::here("prep", "AO")

source(here::here("R", "setup.R"))
knitr::opts_chunk$set(message = FALSE, warning = FALSE, results = "hide", fig.width = 9.5)
library(stringi)
bkgd_path <- here::here("supplement", "goal_summaries", "AO.Rmd")
data_path <- here::here("data", "AO", version_year, "ao_data.rmd")
refs_path <- file.path(loc, "ao_references.Rmd")

dir_B <- file.path(dirname(dir_prep), "bhi-data", "BHI 2.0")
dir_rawdata <- file.path(dir_B, "Goals", "AO")

fullpal <- c(
  RColorBrewer::brewer.pal(8, "Dark2"),
  RColorBrewer::brewer.pal(9, "Set1")
)
```

```{r ao setup spatial functions and files, warning = FALSE, include = FALSE}
## in order to assign bhi regions to data, must run this function first, to load required shapefiles:
source(here::here("R", "spatial.R"))
regions_shape(sp_dir = file.path(dirname(dir_B), "Shapefiles"))

basin_lookup <- file.path(dir_prep, "supplement", "lookup_tabs", "rgns_complete.csv") %>% 
  read_csv() %>%
  dplyr::select(bhi_id = region_id, country = eez, subbasin, rgn_nam = region_name, HELCOM_ID = helcom_id) %>%
  mutate(
    subbasin = as.character(subbasin),
    rgn_nam = as.character(rgn_nam)
  )
```


## 1. Background {-}

```{r ao prep background, child = bkgd_path, results = "asis", echo = FALSE}
```

<br/>

## 2. Data {-}

This prep document is used to generate and explore the following data layers:

- `ao_stock_status_bhi2019.csv`
- `ao_stock_slope_bhi2019.csv` 

These are saved to the `layers` folder. Saved to `data/AO/v2019/intermediate` are intermediate datasets: `ao_merged_rawdata.csv` and `ao_cleaned_cpue.csv`. All these are derived from or informed by the raw datasets from Jens Olsson (goalkeeper).

<br>

```{r ao prep data, child = data_path, results = "asis", echo = FALSE}
```

<br/>

## 3. Prep: Wrangling & Derivations, Checks/Evaluation, Gapfilling {-}

```{r load ao datasets, echo = TRUE}
## previously merged in 'data/AO/v2019/ao_data.R' script
combined_coastal_fish <- read_csv(file.path(dirname(data_path), "intermediate", "ao_merged_rawdata.csv"))
cpue <- read_csv(file.path(dirname(data_path), "intermediate", "ao_cleaned_cpue.csv"))
```

<br/>

### 3.1 Match BHI regions {-}

The raw coastal fish data must be assigned to their respective Subbasins and BHI Regions based on their latitude/longitude coordinates. 

```{r assign BHI regions to coastal fish data, echo = TRUE, message = FALSE, warning = FALSE}
# Rename some variable and subbasin name to make it match to the basin_lookup table
combined_coastal_fish1 <- combined_coastal_fish %>% 
  rename(lat = Latitude, lon = Longitude) %>% 
  filter(!is.na(lat) & !is.na(lon)) %>%  # however, by doing this, more than 73 locations will be cut off
  mutate(
    Subbasin = stri_trans_totitle(Subbasin),
    Subbasin = str_replace(Subbasin, "Åland Sea", "Aland Sea"),
    Subbasin = str_replace(Subbasin, "Archipelago Sea", "Aland Sea"),
    Subbasin = str_replace(Subbasin, "Belt Sea", "Great Belt"),
    Subbasin = str_replace(Subbasin, "Gulf Of Finland", "Gulf of Finland"),
    Subbasin = str_replace(Subbasin, "Gulf Of Riga", "Gulf of Riga"),
    Subbasin = str_replace(Subbasin, "Estern Gotland Basin", "Eastern Gotland Basin"),
    Subbasin = str_replace(Subbasin, "Mecklenburg Bight", "Bay Of Mecklenburg"),
    Subbasin = str_replace(Subbasin, "Bay Of Mecklenburg", "Bay of Mecklenburg"),
    Subbasin = str_replace(Subbasin, "Northern Baltic Sea", "Northern Baltic Proper")) %>% 
  rename(subbasin = Subbasin)
# need to check if "Archipelago Sea" (Finland) is actually Aland Sea subbasin
#detach("package:raster", unload=TRUE) 

# bhibuffer <- sf::st_read(file.path(dirname(dir_B), "Shapefiles", "BHI_shapefile_25km_buffer"))
coastalfish_rgns <- join_rgns_info(
  combined_coastal_fish1, helcomID_col = "helcom_id", country_col = "country", 
  latlon_vars = c("^lat", "^lon"), return_spatial = FALSE, 
  rgn_shps_loc = file.path(dirname(dir_B), "Shapefiles"), 
  buffer_shp = NULL
) 

```

<br/>

**Explore coastal fish data on a map**
```{r explore coastal fish data in map, results = "show", fig.width = 11, fig.height = 9}
library(leaflet)
source(here::here("R", "spatial.R"))
regions_shape()

# Remove data points out of the bhi regions
coastalfish_rgns1 <- coastalfish_rgns %>% 
  filter(monitoring.area != "Northern Limfjord" & 
           monitoring.area != "Skive Fiord and Lovns Broad") %>% 
  # assign bhi_id with na value
  mutate(BHI_ID = replace(BHI_ID, is.na(BHI_ID) & monitoring.area == "Finbo", 36),
         BHI_ID = replace(BHI_ID, is.na(BHI_ID) & monitoring.area == "Jurkalne", 24),
         BHI_ID = replace(BHI_ID, is.na(BHI_ID) & monitoring.area == "Præstø Fiord", 12))

## make basemap
basemap <- ggplot2::ggplot(rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")) + 
  geom_sf(size = 0.1, color = "burlywood", alpha = 0.4) +
  scale_x_continuous(limit = c(4, 32)) +
  scale_y_continuous(limit = c(53.5, 66)) +
  theme(panel.background = element_rect(fill = "#F8FBFC", color = "#E2EEF3"))

bhi_rgns_simple <- rmapshaper::ms_simplify(input = BHI_rgns_shp) %>% sf::st_as_sf()

basemap +
  geom_sf(data = bhi_rgns_simple, fill = "white", size = 0.2, color = "burlywood") +
  geom_sf(
    mapping = aes(colour = monitoring.area),
    data = sf::st_as_sf(
      coastalfish_rgns1, 
      coords = c("lon", "lat"), 
      crs = 4326
    ),
    size = 1.8
  ) +
  labs(x = NULL, y = NULL, colour = "Monitoring area") +
  ggtitle("Sampling Locations")

# + theme(legend.position = c(0.9, 0.1)) +
#  scale_color_manual(values = c("tomato", "aquamarine2", "gold"))

```


## 4 BHI Status score {-}

Since there will be 12 BHI regions with no score (due to missing data or no sampling there), the status score will be calculated at BHI region scale (instead of at Subbasin scale).

Step 1. Take mean score for each indicator type in each monitoring region.  

Step 2. Take mean score for each indicator at the BHI region scale. 

Step 3. Take mean of the two indicators for each BHI region.  

Step 4. Apply basin score to each BHI region  


Xao = Mean Stock Indicator Value / Reference Point

- Stock indicators = two HELCOM core indicators assessed for good environmental status (each scored between 0 and 1 by BHI region)
- Reference value = 0.6

**A status of 100 indicates target is achieved**
```{r status calculation, echo = TRUE, message = FALSE, warning = FALSE}
## Number of time series that will contribute to the mean score in each bhi rgn
basin_n_obs <- coastalfish_rgns1 %>% 
  dplyr::select(BHI_ID) %>%
  count(BHI_ID)

## Indicator mean by monitoring area
monitoring_indicator_mean <- coastalfish_rgns %>% 
  dplyr::select(
    BHI_ID, subbasin, country, monitoring.area, lat, lon, core_indicator, status.value.monitoring.location, threshold_value
    ) %>%
  group_by(BHI_ID, subbasin, country, monitoring.area, lat, lon, core_indicator, threshold_value)%>%
  summarise(mean_core_monitoring_score = mean(status.value.monitoring.location, na.rm=TRUE))%>%
  ungroup()

# Monitoring area score (mean across the two indicators)
monitoring_mean_score <- monitoring_indicator_mean %>%
  group_by(monitoring.area, BHI_ID, subbasin, country, lat, lon, threshold_value) %>%
  summarise(mean_monitoring_score = round(mean(mean_core_monitoring_score, na.rm=TRUE), 2)) %>%
  ungroup()

# join basin lookup
bhi_monitoring_mean_score <- monitoring_mean_score %>% 
  full_join(., basin_lookup, by = c("BHI_ID" = "bhi_id", "subbasin"))  

# Indicator mean by bhi region
bhi_indicator_mean <- monitoring_indicator_mean %>%
  group_by(BHI_ID, subbasin, country, core_indicator, threshold_value)%>%
  summarise(mean_core_bhi_score = mean(mean_core_monitoring_score, na.rm=TRUE)) %>%
  ungroup()

## BHI score (mean across the two indicators)
# join basin lookup
bhi_mean_score <- bhi_indicator_mean %>%  
  group_by(BHI_ID, threshold_value) %>%
  summarise(mean_bhi_score = round(mean(mean_core_bhi_score, na.rm=TRUE), 2)) %>%
  ungroup() %>% 
  full_join(., basin_lookup, by = c("BHI_ID" = "bhi_id"))

```
<br/>

### 4.1 Plot scores at each level {-}

Plot the monitoring stations and BHI regions 

#### 4.1.1 Plot monitoring area indicator mean scores

**Note that points are "jittered" in the second plot to make visible**  

**Colors in the second plot are different monitoring locations**
```{r plot scores levels of aggregation, results = "show", echo = TRUE, fig.width = 8.5, fig.height = 6}
## Plot monitoring area indicator mean scores
ggplot(monitoring_indicator_mean) + 
  geom_point(aes(monitoring.area, mean_core_monitoring_score, color = core_indicator,
                 shape = core_indicator), size = 3, alpha = 0.6) +
  geom_hline(yintercept = 0.6, color = "red") +
  theme(axis.text.x = element_text(colour="grey20", size=6, angle=90, 
                                   hjust=.5, vjust=.5, face = "plain"),
        plot.margin = unit(c(1,1,1,1), "cm"))  +
  labs(y = "Score", x = "Monitoring Areas") +
  ggtitle("Monitoring Area Indicator Mean Score")

## Plot monitoring area indicator mean scores by BHI regions
ggplot(monitoring_indicator_mean) + 
  geom_jitter(aes(BHI_ID, mean_core_monitoring_score, color = monitoring.area, 
                  shape = factor(core_indicator))) +
  guides(colour=FALSE)+
  theme(axis.text.x = element_text(colour="grey20", size=8, angle=90, 
                                   hjust=.5, vjust=.5, face = "plain"),
        plot.margin = unit(c(1,1,1,1), "cm")) +
  labs(y = "Score", x = "BHI Regions") +
  ggtitle("Monitoring Area Indicator Mean Score by BHI regions")

```

#### 4.1.2 Plot BHI indicator mean score {-}

Plot the mean bhi score for each indicator 
```{r plot basin indicator scores, results = "show", echo = TRUE, fig.width = 7.5, fig.height = 4}
## Plot bhi indicator mean scores
ggplot(bhi_indicator_mean) + 
  geom_point(aes(BHI_ID, mean_core_bhi_score, color = core_indicator, 
                 shape = core_indicator), size = 2, alpha = 0.7) +
  geom_hline(yintercept = 0.6, color = "red") +
  scale_shape_manual(values = c(6,19)) +
  theme(axis.text.x = element_text(colour = "grey20", size = 8, angle = 90, 
                                    hjust=.5, vjust = .5, face = "plain"),
         plot.margin = unit(c(1,1,1,1), "cm")) +
  labs(y = "Score", x = "BHI Regions") +
  ggtitle('Indicator Mean Score by BHI Regions')
```
<br/>

**Plot bhi mean score across indicators**
The mean of the two indicator scores was taken for each bhi region
```{r plot basin mean across indicators, results = "show", echo = TRUE, fig.width = 7.5, fig.height = 6}
## Plot bhi mean score across indicators
ggplot(bhi_mean_score) + 
  geom_point(aes(rgn_nam, mean_bhi_score), size=2) +
  geom_hline(yintercept = 0.6, color = "red") +
  theme(axis.text.x = element_text(colour = "grey20", size = 8, angle = 90, 
                                   hjust = .5, vjust = .5, face = "plain"),
        plot.margin = unit(c(1,1,1,1), "cm")) +
  labs(y = "Score", x = "BHI Regions") +
  ggtitle('Mean Score by BHI region')
```
<br/>


#### 4.1.3 Plot BHI Scores {-}

The score to regions where no sampling occurred, due to local population. 
Therefore, 12 regions have the score replaced with NA (4, 7, 8, 10, 13, 16, 17, 18, 19, 21, 22, 33).
_See this map of sampling locations provided by Jens Olsson for the above justification (note that the Finnish sites are not marked but covered the entire coastline as they are ICES regions. Coloration of the coastal waters is indication HELCOM coastal water type._  

```{r final score objects, echo = TRUE, message = FALSE, warning = FALSE}
## Calculate bhi status as: basin_mean/basin_target, constraining values to 1 as maximum score
bhi_status <- bhi_mean_score %>%
  dplyr::select(BHI_ID, mean_bhi_score, threshold_value) %>%
  mutate(
    mean_basin_score = ifelse(BHI_ID %in% c(4,7,8,10,13,16,17,18,19,21,22,33), NA, mean_bhi_score),
    dimension= "status"
    ) %>%  
  dplyr::rename(rgn_id = BHI_ID) %>%
  mutate(score = pmin(1, mean_basin_score/threshold_value),
         score = round(score * 100, 2)) %>% # keep decimal places, will round when combine eut indicators
  dplyr::select(rgn_id, dimension, score) %>% 
  filter(!is.na(rgn_id))

```
<br/>


#### 4.1.4  Plot final score objects


```{r ao final score object plotted, results = "show", echo = TRUE, fig.width = 9.5, fig.height = 6.5}
## Number of observations
##basin_n_obs    
##Plot points
plot_status <- bhi_status %>% 
  full_join(.,basin_lookup, by = c("rgn_id"= "bhi_id")) %>%
  full_join(., basin_n_obs, by= c("rgn_id" ="BHI_ID"))

# ggplot(plot_status) +
#  geom_point(aes(rgn_id, status, color = n, size = n)) +
#  ylim(0, 100) +
#  labs(y = "Score", x = "BHI regions") +
#  ggtitle("AO Status Score")

## plot the status
## BHI region plots will look the same, as BHI status are taken from subbasin means
## unless alternative calculation is used, based on BHI region means rather than subbasin means
set.seed(2)
statuspal <- colorRampPalette(fullpal)(80)[sample(1:80, size = 18)]

ggplot(plot_status) +
  geom_col(
    aes(rgn_nam, score, fill = subbasin),
    position = position_dodge(), 
    color = "black", 
    alpha = 0.6,
    size = 0.2,
    show.legend = FALSE
  ) + 
  geom_text(aes(rgn_nam, score, label = score), color = "black", size = 3) +
  coord_flip() +
  labs(x = NULL, y = "Score") +
  theme_bw() +
  theme(axis.text.y = element_text(size = 9)) +
  scale_fill_manual(values = statuspal) +
  ggtitle("AO Status by Region")

```


## 5. Trend Calculation {-}

We decided to follow the same approach as for BHI 1.0, from [Bergström et al 2016](https://www.researchgate.net/publication/288058474_Long_term_changes_in_the_status_of_coastal_fish_in_the_Baltic_Sea), but only focus on the final 10 year-time period
for the slope (2008-2018)

### 5.1 Trend calculation method {-}

STEPS:

1) Select final time period of trend assessment (2008-2018)
2) Use time series from both indicators, Key Species and Functional groups. For functional groups,
    include both cyprinid and piscivore time series.
3) For each time series:  square-root transform data, z-score, fit linear regression, extract slope
4) Within each time series group (key species, cyprinid, piscivore), take the mean slope for each group within each bhi region
5) Within each bhi region take a mean functional group indicator slope (mean of cyprinid mean and piscivore mean)
6) For each bhi region take overall mean slope - mean of key species and functional group
7) Multiple by five for future year value
8) Apply trend value for all BHI regions (except in Gulf of Finland, do not apply Finnish site value to Estonia and Russian regions)
 
**Select time period and plot data**  
```{r select time period and plot data, results = "show", echo = TRUE, fig.width = 11.5, fig.height = 6.5}
## Select years 2008-2018
cpue <- cpue %>%
  filter(Year >= 2008 & Year <= 2018) %>% 
  mutate(Year = as.character(Year)) %>% 
  right_join(basin_lookup, by = c("Subbasin" = "subbasin", "Country" = "country"))

## Plot raw data
ggplot(filter(cpue, core_indicator == "Key species")) +
  geom_point(aes(Year, cpue)) +
  facet_wrap(~monitoring_area, scales="free_y") +
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("Key Species CPUE by monitoring areas (2008-2018)")

ggplot(filter(cpue, Taxa == "Cyprinids")) +
  geom_point(aes(Year, cpue)) +
  facet_wrap(~monitoring_area, scales="free_y") +
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("Cyprinids CPUE by monitoring areas (2008-2018)")

ggplot(filter(cpue, Taxa == "Piscivores")) +
  geom_point(aes(Year, cpue)) +
  facet_wrap(~monitoring_area, scales="free_y") +
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("Piscivores CPUE by monitoring areas (2008-2018)")
```
<br/>

**Transform data**
```{r transform data, results = "show", echo = TRUE, fig.width = 11.5, fig.height = 6.5}
# square root transform and zscore each time series
cpue_trans <- cpue %>%
  mutate(cpue_sqrt = sqrt(cpue))%>% ## square root transform the data
  group_by(bhi_id, rgn_nam, monitoring_area, core_indicator, Taxa) %>%
  mutate(cpue_mean = mean(cpue_sqrt, na.rm = TRUE),
         cpue_sd = sd(cpue_sqrt, na.rm=TRUE),
         cpue_z = ((cpue_sqrt - cpue_mean)/cpue_sd)) %>% ## zscore the transformed data, each time series zscored to itself
  ungroup()

## check zscore mean should be = 0, sd should be = 1
cpue_trans %>% 
  dplyr::select(monitoring_area, core_indicator, Taxa, cpue_z) %>%
  group_by(monitoring_area, core_indicator, Taxa) %>%
  summarise(mean_check = mean(cpue_z, na.rm = TRUE),
            sd_check = sd(cpue_z, na.rm = TRUE)) %>%
  ungroup()
## means are all very different,sd are 1
  
## Plot sqrt transformed and zscored data
ggplot(filter(cpue_trans, core_indicator == "Key species")) +
  geom_point(aes(Year, cpue_z))+
  geom_hline(yintercept = 0) +
  facet_wrap(~monitoring_area, scales="free_y") +
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("Key Species CPUE Sqrt Zscore")

ggplot(filter(cpue_trans, Taxa == "Cyprinids")) +
  geom_point(aes(Year, cpue_z))+
  geom_hline(yintercept = 0) +
  facet_wrap(~monitoring_area, scales="free_y") +
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("Cyprinids CPUE Sqrt Zscore")

ggplot(filter(cpue_trans, Taxa =="Piscivores")) +
  geom_point(aes(Year, cpue_z)) +
  geom_hline(yintercept = 0) +
  facet_wrap(~monitoring_area, scales="free_y") +
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("Piscivores CPUE Sqrt Zscore")
```

**Fit LM, extract slope**
```{r fit linear model, echo = TRUE, message = FALSE, warning = FALSE}
library(broom)
cpue_mdl <- cpue_trans %>%
  dplyr::select(bhi_id, rgn_nam, monitoring_area, core_indicator, Taxa, Year, cpue_z) %>%
  mutate(Year = as.numeric(Year)) %>% 
  group_by(bhi_id, rgn_nam, monitoring_area, core_indicator, Taxa) %>%
  drop_na(cpue_z) %>% # need to drop 'NA' to make the lm work
  do(mdl = lm(cpue_z ~ Year, data = .)) %>%             # regression model to get the trend
  summarize(bhi_id = bhi_id,
            rgn_nam = rgn_nam,
            monitoring_area = monitoring_area,
            core_indicator = core_indicator,
            Taxa = Taxa,
            slope = coef(mdl)['Year'],
            r2 = as.numeric(summary(mdl)['r.squared']),
            nobs = nrow(augment(mdl))) %>%  ## augment() is from the broom library, extracts fits including data, rows of output matches number of obs used
  ungroup()

## Is the number of observations ever less than 5 for lm?
cpue_mdl %>% 
  dplyr::select(nobs) %>% 
  min()  ## yes, there is a row with 3 obs (remove that?)

```

**Plot slope extracted with number of observations**
```{r ao plot slope, results = "show", echo = TRUE, fig.width = 8.5, fig.height = 6.5}
## Plot slope extracted with number of observations
ggplot(cpue_mdl) +
  geom_point(aes(rgn_nam, slope, colour = monitoring_area, 
                 shape = core_indicator, label = nobs), size = 2) +
  geom_hline(yintercept = 0) +
  geom_text(aes(rgn_nam, slope, label = nobs), size = 2.5, alpha = 0.5) +
  theme(axis.text.x = element_text(colour = "grey20", size = 8, angle = 90,
                                   hjust = .5, vjust = .5, face = "plain")) +
  labs(x = "BHI region", y = "Slope") +
  guides(colour = "none") +
  ggtitle("lm slopes 2008-2018")

## Plot without number of observations
ggplot(cpue_mdl) +
  geom_point(aes(rgn_nam, slope, colour = monitoring_area,
                 shape = core_indicator), size = 2.5, alpha = 0.5) +
  geom_hline(yintercept = 0) +
  theme(axis.text.x = element_text(colour = "grey20", size = 8, angle = 90,
                                   hjust = .5, vjust = .5, face = "plain")) +
  labs(x = "BHI region", y = "Slope") +
  guides(colour = "none") +
  ggtitle("lm slopes per monitoring areas (2008-2018)")
```


### 5.2 Plot the slopes from all timeseries

```{r plot slopes, results = "show", echo = TRUE, fig.width = 9.5, fig.height = 6.5}
#### Plot the slopes from all timeseries
ggplot(cpue_mdl)+
  geom_hline(yintercept = 0) +
  geom_boxplot(aes(rgn_nam, slope)) +
  facet_wrap(~core_indicator + Taxa)+
  theme(axis.text.x = element_text(colour = "grey20", size = 8, angle = 90,
                                   hjust = .5, vjust = .5, face = "plain")) +
  labs(x = "BHI region", y = "Slope") +
  ggtitle("Time Series slope by Indicator type and Taxa")

## Only Functional groups
ggplot(filter(cpue_mdl, core_indicator == "Key functional group")) +
  geom_hline(yintercept = 0) +
  geom_point(aes(rgn_nam, slope, colour = monitoring_area)) +
  facet_wrap(~Taxa) +
  theme(axis.text.x = element_text(colour = "grey20", size = 8, angle = 90,
                                   hjust = .5, vjust = .5, face = "plain"),
        legend.text = element_text(colour = "grey20", size = 5),
        plot.title = element_text(colour = "grey20", size = 8)) +
  labs(x = "BHI region", y = "Slope") +
  ggtitle("Time Series slope for Functional Group by Taxa")

## Only Key Species
ggplot(filter(cpue_mdl, core_indicator == "Key species")) +
  geom_hline(yintercept = 0) +
  geom_point(aes(rgn_nam, slope, colour = monitoring_area, shape = Taxa)) +
  theme(axis.text.x = element_text(colour = "grey20", size = 8, angle = 90,
                                   hjust = .5, vjust = .5, face = "plain"),
        legend.text = element_text(colour = "grey20", size = 5),
        plot.title = element_text(colour = "grey20", size = 8)) +
  #guides(size=.5)+
  labs(x = "BHI region", y = "Slope") +
  ggtitle("Time Series slope for Key Species by Taxa")
```

### 5.3 Mean Slope group per bhi region

Within each time series group (key species, cyprinid, piscivore), take the mean slope for each group within each bhi region 
```{r mean slope by group per bhi region, results = "show", echo = TRUE, fig.width = 10.5, fig.height = 7.5}
## slope (per bhi region)
slope_bhi <- cpue_mdl %>%
  mutate(bhi_id = as.character(bhi_id)) %>% 
  group_by(bhi_id, core_indicator, Taxa) %>%
  mutate(slope_mean_bhi_taxa = mean(slope)) %>% 
#  stringi::stri_sort("bhi_id", numeric = TRUE) %>% 
  ungroup() 

## number of time series for each bhi region
bhi_n_obs_slope <- slope_bhi %>%
  dplyr::select(bhi_id)%>%
  count(bhi_id)

## Plot
ggplot(slope_bhi) + 
  geom_hline(yintercept = 0) +
  geom_boxplot(aes(rgn_nam, slope)) +
  geom_point(aes(rgn_nam, slope_mean_bhi_taxa), colour = "red") +
  facet_wrap(~core_indicator + Taxa)+
  theme(axis.text.x = element_text(colour = "grey20", size = 8, angle = 90,
                                   hjust = .5, vjust = .5, face = "plain")) +
  labs(x = "BHI region", y = "Slope") +
  ggtitle("Mean slope by bhi region and indicator taxa")
```


### 5.4 Mean slope for each indicator (Functional group, Key Spp)

Within each bhi region take a mean functional group indicator slope (mean of cyprinid mean and piscivore mean) and take mean of each Key species indicator
```{r ao mean indicator slope, results = "show", echo = TRUE, fig.width = 8.5, fig.height = 5.5}
## new object only mean value
slope <- slope_bhi %>%
  dplyr::select(bhi_id, rgn_nam, core_indicator, Taxa, slope_mean_bhi_taxa)

slope2 <- slope %>%
  group_by(bhi_id, core_indicator) %>%
  mutate(slope_mean_bhi_indicator = mean(slope_mean_bhi_taxa)) %>%
  ungroup()
# str(slope2)

# plot
ggplot(slope2) + 
  geom_hline(yintercept = 0) +
  geom_point(aes(rgn_nam, slope_mean_bhi_indicator), colour = "red") +
  facet_wrap(~core_indicator) +
  theme(axis.text.x = element_text(colour = "grey20", size = 8, angle = 90,
                                   hjust = .5, vjust = .5, face = "plain")) +
  ylim(-.3,.3) +
  labs(y = "Slope", x = "BHI region") +
  ggtitle("Mean slope by bhi region and indicator type")

```


### 5.5 BHI region mean slope

Take the mean for each bhi region across the indicator types (Functional groups, key species)
```{r ao bhi mean slope, results = "show", echo = TRUE, fig.width = 8.5, fig.height = 5.5}
## new object only bhi rgn and indicator mean
slope3 <- slope2 %>%
  dplyr::select(bhi_id, rgn_nam, core_indicator, slope_mean_bhi_indicator)

slope3 <- slope3 %>%
  group_by(bhi_id) %>%
  mutate(slope_mean_bhi = mean(slope_mean_bhi_indicator)) %>%
  ungroup()

ggplot(slope3) + 
  geom_hline(yintercept = 0) +
  geom_point(aes(rgn_nam, slope_mean_bhi), colour = "black", size = 3) +
  geom_point(aes(rgn_nam, slope_mean_bhi_indicator, shape = core_indicator), colour = "red") +
  theme(axis.text.x = element_text(colour = "grey20", size = 8, angle = 90,
                                   hjust = .5, vjust = .5, face = "plain")) +
  scale_shape_manual(values = c(1:2)) +
  ylim(-.3,.3) +
  labs(y = "Slope", x = "BHI region") +
  ggtitle("Mean slope (black dots) by bhi region")
```

#### 5.5.1 Investigate on functional groups slopes

This plot might help explaining the overall positive trend of functional groups
```{r ao functional group mean slope, results = "show", echo = TRUE, fig.width = 8.5, fig.height = 5.5}
# Investigate on functional groups
slope22 <- slope2 %>%
  group_by(bhi_id, core_indicator) %>%
  mutate(slope_mean_bhi_indicator = mean(slope_mean_bhi_taxa)) %>%
  filter(Taxa %in% c("Cyprinids", "Piscivores")) %>%
  ungroup() 
# str(slope22)

ggplot(slope22) + 
  geom_hline(yintercept = 0) +
  geom_point(aes(rgn_nam, slope_mean_bhi_indicator), colour = "black", size = 3) +
  geom_point(aes(rgn_nam, slope_mean_bhi_taxa, shape = Taxa), colour = "red") +
  theme(axis.text.x = element_text(colour = "grey20", size = 8, angle = 90,
                                   hjust = .5, vjust = .5, face = "plain")) +
  scale_shape_manual(values = c(1:2)) +
  ylim(-.3,.3) +
  labs(y = "Slope", x = "BHI region") +
  ggtitle("Functional Groups mean slope (black dots) by BHI regions")
## This plot might help explaining the overall positive trend of functional groups
```


### 5.6 Final slope object

```{r final slope object, echo = TRUE, message = FALSE, warning = FALSE}
## new object, only bhi rgn mean
slope4 <- slope3 %>%
  dplyr::select(bhi_id, slope_mean_bhi) %>% 
  distinct(.)

bhi_slope <- slope4 %>%
  dplyr::select(bhi_id, slope_mean_bhi)%>%
  dplyr::rename(rgn_id = bhi_id,
                score = slope_mean_bhi) %>%
  mutate(rgn_id = as.numeric(rgn_id),
         dimension="trend",
         score = round(score, 2)) %>% ## round to 2 decimal places
  dplyr::select(rgn_id, dimension,score)
```


#### 5.6.1. Plot final slope object

Size points by number of times series used
```{r ao plot final code object, results = "show", echo = TRUE, fig.width = 9.5, fig.height = 6.5}
##object with number of observations
plot_slope <- bhi_slope %>%
  full_join(., basin_lookup, by = c("rgn_id" = "bhi_id")) %>% 
  full_join(., basin_n_obs, by= c("rgn_id" = "BHI_ID")) %>% 
  filter(!is.na(score))

ggplot(plot_slope) +
  geom_hline(yintercept = 0) +
  geom_point(aes(rgn_nam, score, color = n, size = n)) +
  ylim(-1, 1) +
  labs(y="Slope", x = "BHI regions") +
  theme(axis.text.x = element_text(colour = "grey20", size = 8, angle = 90,
                                   hjust = .5, vjust = .5, face = "plain")) +
  ggtitle("AO Slope")

ggplot(rbind(plot_status, plot_slope)) + 
  geom_col(aes(rgn_nam, score, fill = subbasin), alpha = 0.5, show.legend = FALSE) +
  coord_flip() +
  facet_wrap(~dimension, scales = "free_x", nrow = 1) +
  labs(x = "BHI Region\n", y = "Score") +
  ggtitle("AO Status and Trend Scores by Region") +
  scale_fill_manual(values = statuspal) +
  theme(axis.text.y = element_text(size = 7))
```

**Export Data Layers**
```{r  export data layers, eval = FALSE, echo = TRUE, message = FALSE, warning = FALSE}
## Data layers for the stock component of AO

## Status Score
write_csv(bhi_status, file.path(dir_layers, 'ao_stock_status_bhi2019.csv'))

## Slope for Trend
write_csv(bhi_slope, file.path(dir_layers, 'ao_stock_slope_bhi2019.csv'))
```


## 6. Considerations for `BHI3.0` {-}

Stock was one component of the AO goal. Model could be updated with parameters representing **access** and **need** of artisanal fishing opportunities in the future. Jens Olsson and his group will compile information on current regulations for small-scale fishing in the Baltic Sea along the coast. There is an NGO called ['Low Impact Fishers of Europe (LIFE)'](https://lifeplatform.eu/) that could provide this type of information for both recreational and commercial fishing.

<br>

## 7. References {-}

```{r References, child = refs_path, results = "asis", echo = FALSE}
```
