## 2. Data

**Rainbow trout production (Denmark, Sweden and Finland) and Finfish (Denmark, Germany, Sweden and Finland)**

```{r Preamble, echo = FALSE, include = FALSE, error = FALSE}
source(file.path(here::here(), "R", "setup.R"))
library(readr)
library(sf)
# source(file.path(here::here(), "R", "data.R"))
```

### 2.1 Datasets with Sources

**Rainbow trouts mariculture**  

- **Sweden**
<!-- dataset save location BHI_share/2.0/Goals/FP/Mariculture/rb_trout/rbt_swe -->

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("PDF:", "Vattenbruk 2018"), 
  c("Location data:", "Page 7, table A")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [Jorbruksverket database](https://www.scb.se/hitta-statistik/statistik-efter-amne/jord-och-skogsbruk-fiske/vattenbruk/vattenbruk/) <br/> [Jorbruksverket](https://www.scb.se/contentassets/cef2fb103630496bb532e76c98f747e6/jo1201_2018a01_sm_jo60sm1901.pdf) Downloaded 2019-10-09 by Andrea De Cervo")
```

<br/>

- **Denmark**
<!-- dataset save location BHI_share/2.0/Goals/FP/Mariculture/rb_trout/rbt_den -->

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Faste tabeller med akvakulturststistik:", "Produktion af hovedarter fordelt på region"), 
  c("Rows:", "Regnbueørred, Saltvand")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [Ministry of Environment and Food of Denmark database](https://fiskeristyrelsen.dk/fiskeristatistik/akvakulturstatistik/) <br/> Downloaded 2019-10-11 by Andrea De Cervo")
```

<br/>

- **Finland**
<!-- dataset save location BHI_share/2.0/Goals/FP/Mariculture/rb_trout/rbt_fin -->

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Luke statistics database:", "Food fish production by species"), 
  c("Production:", "Quantity (1 000 kg)"),
  c("Area:", "Sea"),
  c("Fish:", "Rainbow trout"),
  c("Year:", "All")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [Natural Resources Institute database](http://statdb.luke.fi/PXWeb/pxweb/en/LUKE/LUKE__06%20Kala%20ja%20riista__02%20Rakenne%20ja%20tuotanto__10%20Vesiviljely/?tablelist=true&rxid=ef8a2896-d6cb-4660-a8aa-fee93bd4212f) <br/> Downloaded 2019-10-10 by Andrea De Cervo")
```

<br/>

- **Germany**
<!-- dataset save location BHI_share/2.0/Goals/FP/Mariculture/rb_trout/rbt_ger -->

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Location:", "Kiel Bay")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [Email from plant manager](tassilo@jaeger-kleinicke.de) <br/> Received 2019-10-30")
```

<br/>

**Finfish Mariculture**  
<!-- dataset save location BHI_share/2.0/Goals/FP/Mariculture/Finfish_mariculture -->

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("HELCOM Map and Data service:", "Pressures and human activities"), 
  c("Human activities:", "Finfish mariculture")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [HELCOM database](http://maps.helcom.fi/website/mapservice/?datasetID=3cfa469a-6a78-4913-b82f-57fd0e7f4dc0) <br/> (http://metadata.helcom.fi/geonetwork/srv/eng/catalog.search#/metadata/3cfa469a-6a78-4913-b82f-57fd0e7f4dc0) <br/> Downloaded 2019-10-09 by Andrea De Cervo")
```

<br/>

**All Baltic countries**
<!-- dataset save location BHI_share/2.0/Goals/FP/Mariculture/FAO_fishstat -->

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("FAO FishStatJ:", "Global aquaculture production"), 
  c("Quantity:", "1950-2017")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [FAO FishStatJ](http://www.fao.org/fishery/statistics/software/fishstatj/en#downlApp) <br/> Downloaded 2019-10-16 by Andrea De Cervo")
```

<br/>

#### Reference Point: maximum nutrient discharge (N/P)

In 2016 HELCOM Contracting Parties adopted Recommendation 37/3 on "Sustainable aquaculture in the Baltic Sea Region" (http://www.helcom.fi/Recommendations/Rec%2037-3.pdf). 

However, the only parameter that can be used as reference point for Mariculture subgoal in the Baltic Sea (that can be also found on Finnish Mariculture data provided by HELCOM) is the maximum nutrient discharge for P and N (http://www.helcom.fi/Recommendations/Rec%2025-4.pdf), where existing and new marine fish farms should not exceed the annual average of: 
- 7g of P (tot-P) 
- 50g of N (tot-N) 
per 1kg fish (living weight) produced.
[The nutrient limit values (P and N) are calculated on the basis that living fish contains 0,4% of phosphorus and 2,75% of nitrogen.]

This recommendation was adopted in 2004, but since there are not updated nutrient limit values, this will be the parameter used as reference point.

### 2.2 Centralization & Normalization

```{r load raw data, message = FALSE, warning = FALSE, echo = TRUE}
## root location of the raw data
dir_rawdata <- file.path(dir_B, "Goals", "FP", "MAR") # list.files(dir_rawdata)

## rainbow trout data
rbt_de_raw <- read_delim(file.path(dir_rawdata, "rbt_de", "rbt_de.csv"), delim = ";")
rbt_dk_raw <- read_delim(file.path(dir_rawdata, "rbt_dk", "rbt_dk.csv"), delim = ";")
rbt_dk2005raw <- read_delim(file.path(dir_rawdata, "rbt_dk", "rbt_dk_2005.csv"), delim = ";")
rbt_fi_raw <- read_delim(file.path(dir_rawdata, "rbt_fi", "rbt_fi.csv"), delim = ";")
rbt_se_raw <- read_delim(file.path(dir_rawdata, "rbt_se", "rbt_se.csv"), delim = ";")

## FAO rainbow trout data
fao_rbt_raw <- read_delim(file.path(dir_rawdata, "FAO_fishstat", "FAO_fishstat.csv"), delim = ";")

# countrycodes for FAO dataset
countrycode <- data.frame(
  area = c("Germany", "Denmark", "Estonia", "Finland","Sweden"),
  country = c("DE", "DK", "ES", "FI", "SE")
)

## HELCOM Finfish mariculture data
## this is spatial data so we need to remove the geometry column
## also need to convert many columns from factor to numeric
finfish_mariculture_raw <- sf::st_read(file.path(dir_rawdata, "Finfish_mariculture", "Finfish_mariculture.shp"))
st_geometry(finfish_mariculture_raw) <- NULL
finfish_mariculture_raw[, 8:28] <- sapply(finfish_mariculture_raw[, 8:28], as.numeric)
finfish_mariculture_raw[, 1:7] <- sapply(finfish_mariculture_raw[, 1:7], as.character)
```

#### 2.2.0 Merge datasets and calculate maximum nutrient discharge (N/P) 

```{r merge rainbow trout rawdata, echo = TRUE}
rbt_raw <- rbind(
  
  ## rbt_de rainbow trout germany
  rbt_de_raw %>% 
    dplyr::rename(farm_species = production) %>% 
    tidyr::gather(key = "area", value = "produced_tonnes", -year, -farm_species) %>%
    dplyr::mutate(
      produced_tonnes = as.numeric(produced_tonnes),
      country = "DE",
      p_kg = "",
      n_kg = "",
      ## join with max nutrient input value
      maxP_kg = 0.007*(as.numeric(produced_tonnes)*1000), 
      maxN_kg = 0.05*(as.numeric(produced_tonnes)*1000),
      source = "national"
    ) %>% 
    select(year, country, area, farm_species, produced_tonnes, p_kg, maxP_kg, n_kg, maxN_kg, source) %>% 
    arrange(country),
  
  ## rbt_dk2005 rainbow trout 2005 denmarks
  rbt_dk2005raw %>% 
    dplyr::select(-tot_wt_ton) %>% 
    dplyr::rename(
      farm_species = production,
      arhus_cty = cty1, 
      vejle_cty = cty2,
      fyns_cty = cty3,
      sønderjylland_cty = cty4,
      ribe_cty = cty5,
      ringkøbing_cty = cty6,
      viborg_cty = cty7,
      nordjyllands_cty = cty8,
      frederiksborg_cty = cty9,
      roskilde_cty = cty10,
      storstrøms_cty  = cty11,
      vestsjællands_cty = cty12
    ) %>% 
    tidyr::gather(key = "area", value = "produced_tonnes", -year, -farm_species) %>% 
    dplyr::mutate(
      country = "DK",
      p_kg = "",
      n_kg = "",
      maxP_kg = 0.007*(as.numeric(produced_tonnes)*1000), 
      maxN_kg = 0.05*(as.numeric(produced_tonnes)*1000),
      source = "national"
    ) %>% 
    arrange(country),
  
  ## rbt_dk rainbow trout denmark
  rbt_dk_raw %>% 
    dplyr::rename(
      farm_species = production,
      denmark = tot_wt_ton,
      hovedstaden = "reg1", 
      sjælland = "reg2",
      syddanmark = "reg3",
      midtjylland = "reg4"
    ) %>% 
    tidyr::gather(key = "area", value = "produced_tonnes", -year, -farm_species) %>% 
    dplyr::mutate(
      country = "DK",
      p_kg = "",
      n_kg = "",
      maxP_kg = 0.007*(as.numeric(produced_tonnes)*1000), 
      maxN_kg = 0.05*(as.numeric(produced_tonnes)*1000),
      source = "national"
    ) %>% 
    arrange(country),
  
  ## rbt_fi rainbow trout finland
  rbt_fi_raw %>% 
    dplyr::rename(
      produced_tonnes = tot_wt_tonnes, 
      farm_species = production) %>% 
    dplyr::mutate(
      country = "FI", area = "finland",
      p_kg = "",
      n_kg = "",
      maxP_kg = 0.007*(as.numeric(produced_tonnes)*1000), 
      maxN_kg = 0.05*(as.numeric(produced_tonnes)*1000),
      source = "national"
    ) %>% 
    arrange(country),
  
  ## rbt_se rainbow trout sweden
  rbt_se_raw %>% 
    dplyr::rename(
      farm_species = production,
      sweden = tot_wt_mtonnes) %>% 
    select(-ne_farm_no, -other_farm_no) %>%  # don't need these information 
    tidyr::gather(key = "area", value = "produced_tonnes", -year, -farm_species) %>% 
    mutate(
      country = "SE",
      p_kg = "",
      n_kg = "",
      maxP_kg = 0.007*(as.numeric(produced_tonnes)*1000), 
      maxN_kg = 0.05*(as.numeric(produced_tonnes)*1000),
      source = "national"
    ) %>% 
    arrange(country),
  
  ## fao_rbt fao data on rainbow trout
  fao_rbt_raw %>% 
    dplyr::rename(
      area = country,
      farm_species = species) %>% 
    filter(
      str_detect(str_to_lower(farm_species), "rainbow trout"), 
      !(environment == "Marine" & area == "Denmark")
    ) %>% 
    select(-aqua_area, -Unit, -environment, -("1950":"1971")) %>%
    tidyr::gather(key = "year", value = "produced_tonnes", -area, -farm_species) %>% 
    ## remove characters from production
    mutate(
      produced_tonnes = str_replace(produced_tonnes,"F"," "),
      produced_tonnes = as.numeric(produced_tonnes),
      p_kg = "",
      n_kg = "",
      maxP_kg = 0.007*(as.numeric(produced_tonnes)*1000), 
      maxN_kg = 0.05*(as.numeric(produced_tonnes)*1000),
      source = "fao"
    ) %>% 
    ## join with countrycode data
    left_join(countrycode, by = "area") %>% 
    arrange(country)
)

## reassign types
rbt_raw[, c(1, 5:9)] <- sapply(rbt_raw[, c(1, 5:9)], as.numeric)
```

```{r merge finfish rawdata, echo = TRUE}

## wrangling finfish data so can merge with other datasets
finfish_raw <- finfish_mariculture_raw %>%
  mutate(SUBBASIN = ifelse(is.na(SUBBASIN), COUNTY, SUBBASIN)) %>% 
  dplyr::rename(
    area = SUBBASIN,
    farm_species = PRODUCTION,
    country = COUNTRY
  ) %>% 
  ## remove the facilities on land
  ## these are entries with any of NA, on land, or On land as 'AREA'
  ## are we sure that 'NA' value for AREA means that it is on land? there are so many NAs, leaves only "SE" "DK" "FI" data...
  filter(!(is.na(AREA)|str_detect(str_to_lower(AREA), "on land"))) %>% 
  
  ## replace NA value with general mariculture name 'finfish'
  mutate(
    area = str_replace(area,"Sland Sea","Aland Sea"),
    
    farm_species = ifelse(
      is.na(farm_species), 
      "finfish",  ifelse(
        str_detect(farm_species, "Rainbow trout"),
        "Rainbow trout", farm_species
      )
    )
  )

## gather production, phosphorus and nitrogen vars by year
spreadcols <- names(finfish_raw)[1:7]
gathercols <- names(finfish_raw)[8:28]

finfish_raw <- as_tibble(finfish_raw) %>% 
  pivot_longer(cols = gathercols, names_to = "varyear", values_to = "value") %>% 
  mutate(
    year = str_extract(varyear, pattern = "[0-9]{4}"),
    var = str_extract(varyear, pattern = "[A-Z]+")
  ) %>% 
  filter(!is.na(year)) %>%  # filter out na years as these represent aggregated values, but keep NA values
  
  ## re-spread data by the three variables: production (PR), nitrogen (N), and phosphorus (P)
  tidyr::pivot_wider(id_cols = c(spreadcols, "year"), names_from = var, values_from = value) %>% 
  rename(
    produced_tonnes = PR,
    n_kg = N,
    p_kg = P
  ) %>% 
  mutate(
    maxP_kg = 0.007 * (produced_tonnes * 1000),
    maxN_kg = 0.05 * (produced_tonnes * 1000),
    source = "helcom"
  ) %>% 
  ## keep only necessary columns
  select(year, country, area, farm_species, produced_tonnes, p_kg, maxP_kg, n_kg, maxN_kg, source)
```

```{r combine into a single harmonized dataset save, echo = TRUE}
combined_rawdata <- rbind(rbt_raw, finfish_raw)
combined_rawdata[, c(1, 5:9)] <- sapply(combined_rawdata[, c(1, 5:9)], as.numeric)
# readr::write_csv(combined_rawdata, here::here("data", "MAR", version_year, "mar_full_nmerged_dataset.csv"))
```


#### 2.2.1 Standardize Units

#### 2.2.2 Rename Fields/Variables

#### 2.2.3 Save to BHI Database

<br/>

### 2.3 Initial Data Exploration

```{r}
## create datasets for the plots
plot_nutrients <- combined_rawdata %>% 
  filter(year %in% 2010:2019) %>% 
  mutate(countryarea = paste(country, area, sep = "_")) %>% 
  dplyr::select(-farm_species) %>% 
  rename(P_KG = p_kg, N_KG = n_kg) %>% 
  tidyr::pivot_longer(cols = c("P_KG", "N_KG"), names_to = "Input", values_to = "Value")

plot_nutrients_bounds <- combined_rawdata %>% 
  filter(year %in% 2010:2019) %>% 
  mutate(countryarea = paste(country, area, sep = "_")) %>% 
  dplyr::select(-farm_species, -p_kg, -n_kg) %>% 
  tidyr::pivot_longer(cols = c("maxP_kg", "maxN_kg"), names_to = "maxInput", values_to = "Value")

## plot format nutrient reference
nutrients_plot <- ggplot(data = plot_nutrients) +
  geom_errorbar(
    data = plot_nutrients_bounds, 
    aes(x = year, ymin = Value, ymax = Value, color = maxInput), 
    alpha = 0.2, 
    size = 0.5
  ) +
  geom_point(aes(x = year, y = Value, color = Input), alpha = 0.5, size = 4, shape = 18) +
  facet_grid(cols = vars(country), rows = vars(Input)) +
  scale_color_brewer(palette = "Paired") +
  theme(legend.position = "none")
```

```{r initial data exploration by national data first}
## Checking all variables before plotting them

## Germany
germany <- rbt_raw %>% filter(country == "DE")
# str(germany)
ggplot2::ggplot(germany) +
  geom_col(aes(year, produced_tonnes, fill = area), position = position_dodge()) +
  coord_flip() +
  labs(title = "German production (tons) 2009-2019")

## Denmark
denmark <- rbt_raw %>% filter(country == "DK", source == "national")
ggplot2::ggplot(denmark) +
  geom_col(aes(year, produced_tonnes, fill = area), position = position_dodge()) +
  coord_flip() +
  labs(title = "Danish production (tons) 2005-2017")
  
## Finland
finland <- rbt_raw %>% filter(country == "FI", source == "national")
ggplot2::ggplot(finland) +
  geom_col(aes(year, produced_tonnes, fill = area), position = position_dodge()) +
  coord_flip() +
  labs(title = "Finnish production (tons) 1978-2018")

## Sweden 
sweden <- rbt_raw %>% filter(country == "SE", source == "national")
ggplot2::ggplot(sweden) +
  geom_col(aes(year, produced_tonnes, fill = area), position = position_dodge()) +
  coord_flip() +
  labs(title = "Swedish production (tons) 2009-2018")

## FAO data
fao <- rbt_raw %>% filter(source == "fao")
str(fao)
ggplot2::ggplot(fao) +
  geom_col(aes(year, produced_tonnes, fill = area), position = position_dodge(), width = 2) +
  labs(title = "FAO Brackish water production (tons) 1972-2017")
```


#### 2.3.1 Compare versus Previous Years Data

#### 2.3.2 Timeseries Plots

#### 2.3.3 Map

