---
title: "Nutrients Loads Pressure Layer Data Preparation"
output:
  github_document:
    toc: true
    toc_depth: 3
params: 
    datasource: csv
---
<br>

```{r nutrient load pressure preamble prep, message = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = TRUE, results = "hide", fig.width = 9.5, fig.height = 6)
source(here::here("R", "setup.R"))
loc <- here::here("prep", "pressures", "nutrient_load")
data_path <- here::here("data", "pressures", "nutrient_load", version_year, "nutrient_load_data.rmd")
```

<br>

## 1. Background {-}

External loads of nitrogen and phosphorus into the Baltic Sea. Loads include land-based and atmospheric sources and are normalized for river flow. HELCOM Core Indicator - Inputs of Nitrogen and Phosphorus to the Baltic Sea. [6th Pollution Load Compilation, Executive Summary](https://helcom.fi/media/publications/PLC-6-Executive-Summary.pdf)  


### 1.1 Data Scale {-}

Assessment of N & P loads and the setting of MAI were done for 7 Baltic Sea basins: Baltic Proper, Kattegat, Danish Straits, Gulf of Riga, Gulf of Finland, Bothnian Sea, Bothnian Bay. These basins are larger than the HOLAS basins used define the BHI regions.  The pressure score will be calculated at the assessment scale and then all BHI regions within an assessment basin will receive the same pressure value.  


### 1.2 Reference points {-}

In discussions with Bo Gustafsson, we have made the following data rescaling decisions: 

**Current pressure** = the mean load in each basin during the most recent three year assessment period for HELCOM

**Min value** = the minimum value is when there is no pressure on the system, this is scaled to the MAI for each basin 

**Max value** = the maximum value is the highest load to each basin over the reference period used for determining MAI (1997-2003) 



<br/>

## 2. Data {-}

This prep document is used to generate and explore the following data layers:

- `po_nload_bhi2019.csv` 
- `po_pload_bhi2019.csv` 

These are saved to the `layers` folder. All these are derived from or informed by the raw datasets from HELCOM Pollution Load Compilation (PLC) reports.

<br>

```{r nutrient load pressure prep data, child = data_path, results = "asis", echo = FALSE}
```

<br/>

## 3. Prep: Wrangling & Derivations, Checks/Evaluation, Gapfilling {-}


```{r read in the nutrient load pressure data}
dir_intermediate <- file.path(dir_prep, "data", "pressures", "nutrient_load", version_year, "intermediate")

## read in N and P data
n_load <- read.csv(file.path(dir_intermediate, "basin_n_load.csv"))
n_atmos_load <- read.csv(file.path(dir_intermediate, "basin_n_atmos_load.csv"))
p_load <- read.csv(file.path(dir_intermediate, "basin_p_load.csv"))

## read in maximum allowable inputs
mai <- read.csv(file.path(dir_intermediate, "n_p_load_targets.csv"))
```


```{r merge nutrient load pressure datasets with targets}
n_load <- left_join(
  rename(n_load, n_waterborne = input_tonnes), 
  select(mai, baltic_sea_basin, maximum_allowable_input = total_nitrogen_tonnes),
  by = "baltic_sea_basin"
)
p_load <- left_join(
  p_load,
  select(mai, baltic_sea_basin, maximum_allowable_input = total_phosphorus_tonnes),
  by = "baltic_sea_basin"
)
```


```{r add the atmospheric nitrogen deposition to the waterborne inputs}
n_load <- n_atmos_load %>% 
  rename(n_atmos = input_tonnes) %>% 
  left_join(n_load, by = c("year", "baltic_sea_basin")) %>% 
  mutate(input_tonnes = n_atmos + n_waterborne)
```

<br>

### 3.1 Calculate Load Pressure Scores {-}


**Calculate average loads of Subbasins**

```{r current n pressure calculation}
n_current <- n_load %>% 
  
  ## three year assessment period
  ## take rolling means with three year window
  arrange(basin_code, year) %>% 
  group_by(basin_code) %>% 
  mutate(n_current = zoo::rollapply(input_tonnes, 3, mean, na.rm = TRUE, align = "right", fill = NA)) %>% 
  ungroup() %>% 
  filter(!is.na(n_current)) %>% 
  mutate(n_current = round(n_current))
```


```{r current p pressure calculation}
p_current <- p_load %>% 
  
  ## three year assessment period
  ## take rolling means with three year window
  arrange(basin_code, year) %>% 
  group_by(basin_code) %>% 
  mutate(p_current = zoo::rollapply(input_tonnes, 3, mean, na.rm = TRUE, align = "right", fill = NA)) %>% 
  ungroup() %>% 
  filter(!is.na(p_current)) %>% 
  mutate(p_current = round(p_current))
```


**Rescale Load pressures**

Data layer needs to be rescaled between 0 (min pressure) and 1 (max pressure).  

```{r rescale n pressure scores}
n_min <- n_load %>% 
  distinct(baltic_sea_basin, maximum_allowable_input) %>%
  dplyr::rename(n_min = maximum_allowable_input)

n_max <- n_load %>% 
  filter(year %in% 1997:2003) %>%
  select(year, baltic_sea_basin, input_tonnes) %>%
  group_by(baltic_sea_basin) %>%
  summarise(n_max = max(input_tonnes)) %>%
  ungroup()
        
## to normalize data: normalized = (x-min(x))/(max(x)-min(x))
## high pressure score corresponds to high pressure, i.e. is a bad thing
## pressure of zero indicates the target is achieved
n_current_score <- n_current %>%
  full_join(n_min, by = "baltic_sea_basin") %>%
  full_join(n_max, by = "baltic_sea_basin") %>%
  mutate(n_rescaled = (n_current - n_min)/abs(n_max - n_min)) %>%
  mutate(n_score = pmin(1, pmax(0, n_rescaled))) %>%
  mutate(n_score = round(n_score, 2)) %>% 
  select(year, baltic_sea_basin, n_score, n_current, n_min, n_max)
```

```{r rescale p pressure scores}
p_min <- p_load %>% 
  distinct(baltic_sea_basin, maximum_allowable_input) %>%
  dplyr::rename(p_min = maximum_allowable_input)

p_max <- p_load %>% 
  filter(year %in% 1997:2003) %>%
  select(year, baltic_sea_basin, input_tonnes) %>%
  group_by(baltic_sea_basin) %>%
  summarise(p_max = max(input_tonnes)) %>%
  ungroup()
        
## to normalize data: normalized = (x-min(x))/(max(x)-min(x))
p_current_score <- p_current %>%
  full_join(p_min, by = "baltic_sea_basin") %>%
  full_join(p_max, by = "baltic_sea_basin") %>%
  mutate(p_rescaled = (p_current - p_min)/abs(p_max - p_min)) %>%
  mutate(p_score = pmin(1, pmax(0, p_rescaled))) %>%
  mutate(p_score = round(p_score, 2)) %>% 
  select(year, baltic_sea_basin, p_score, p_current, p_min, p_max)
```


**Match nutrient loading values by Subbasin to BHI regions**

```{r expand n and p pressure to bhi regions}
bhi_basin_match <- read_csv(file.path(dir_prep, "supplement", "lookup_tabs", "rgns_complete.csv")) %>% 
  select(region_id, subbasin) %>% 
  left_join(
    tibble(
      subbasin = c(
        "Great Belt", "Kiel Bay", "Bay of Mecklenburg", 
        "The Sound", "Kattegat", 
        "Arkona Basin", "Bornholm Basin", "Gdansk Basin", "Northern Baltic Proper", 
        "Eastern Gotland Basin", "Western Gotland Basin", 
        "Gulf of Riga", 
        "Gulf of Finland", 
        "Aland Sea", "Bothnian Sea", 
        "The Quark", "Bothnian Bay"
      ),
      baltic_sea_basin = c(
        "Danish Straits", "Danish Straits", "Danish Straits", 
        "Kattegat", "Kattegat",
        "Baltic Proper", "Baltic Proper", "Baltic Proper", "Baltic Proper", 
        "Baltic Proper", "Baltic Proper", 
        "Gulf of Riga", 
        "Gulf of Finland", 
        "Bothnian Sea", "Bothnian Sea", 
        "Bothnian Bay", "Bothnian Bay"
      )
    )
  )

po_nload <- full_join(bhi_basin_match, n_current_score, by = "baltic_sea_basin")
po_pload <- full_join(bhi_basin_match, p_current_score, by = "baltic_sea_basin")
```

<br/>

### 3.2 Save the Nutrient Load Pressure Layers {-}

Save nutrient loading pressure data as csv in layers folder.

```{r save nutrient load pressure layers as a csvs}
write_csv(
  po_nload %>% 
    select(region_id, year, pressure_score = n_score), 
  file.path(dir_layers, "po_nload_bhi2019.csv")
)
write_csv(
  po_pload %>% 
    select(region_id, year, pressure_score = p_score), 
  file.path(dir_layers, "po_pload_bhi2019.csv")
)
```

<br>

## 4. Visualizing Data Layers {-}

### 4.1 Nutrient Load Pressure Scores by Subbasin {-}

```{r nutrient load pressure scores, results  = "show", fig.width  = 9.5}
ggplot(distinct(po_nload, year, subbasin, n_score)) +
  geom_col(aes(x = year, y = n_score)) +
  facet_wrap(~subbasin, scales = "free_y", nrow = 6) +
  labs(x = NULL, y = "Nitrogen Load, Pressure Score")

ggplot(distinct(po_pload, year, subbasin, p_score)) +
  geom_col(aes(x = year, y = p_score)) +
  facet_wrap(~subbasin, scales = "free_y", nrow = 6) +
  labs(x = NULL, y = "Phosphorus Load, Pressure Score")
```


### 4.2 Timeseries {-}

```{r timeseries plots of n and p loads, results = "show", fig.width = 9.5, fig.height = 3}
gridExtra::grid.arrange(
  ggplot(n_current) +
    geom_vline(xintercept = 1997, color = "snow", size = 2) +
    geom_vline(xintercept = 2003, color = "snow", size = 2) +
    geom_hline(aes(yintercept = maximum_allowable_input), color = "maroon") +
    geom_line(aes(year, n_current), size = 0.5) +
    geom_point(aes(year, input_tonnes)) +
    labs(
      x = NULL, y = "Input (tonnes)\n", 
      title = "Nitrogen Inputs (atmospheric and waterborne), line shows 3 year rolling mean"
    ) +
    facet_wrap(~baltic_sea_basin, scales = "free_y", nrow = 1),
  ggplot(p_current) +
    geom_vline(xintercept = 1997, color = "snow", size = 2) +
    geom_vline(xintercept = 2003, color = "snow", size = 2) +
    geom_hline(aes(yintercept = maximum_allowable_input), color = "maroon") +
    geom_line(aes(year, p_current), size = 0.5) +
    geom_point(aes(year, input_tonnes)) +
    labs(
      x = NULL, y = "Input (tonnes)\n", 
      title = "Phosphorus Inputs (only waterborne), line shows 3 year rolling mean"
    ) +
    facet_wrap(~baltic_sea_basin, scales = "free_y", nrow = 1),
  nrow = 2
)
```


### 4.3 Map of N & P pressure scores by BHI region {-}

```{r load spatial info for nutrient load presures maps}
source(here::here("R", "spatial.R"))

bhi_rgn <- st_read(file.path(dirname(dir_B), "Shapefiles", "BHI_shapefile"))
bhi_rgn_simple <- rmapshaper::ms_simplify(input = bhi_rgn) %>% 
  sf::st_as_sf() %>% 
  select(region_id = BHI_ID, rgn_nam, subbasin = Subbasin) %>% 
  mutate(region_nam = paste(rgn_nam, subbasin, sep  = ", "))

basemap <- ggplot2::ggplot(rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")) +
  geom_sf(size = 0.1, color = "burlywood", alpha = 0.4) +
  theme(panel.background = element_rect(fill = "#F8FBFC", color = "#E2EEF3")) +
  scale_x_continuous(limit = c(4, 32)) +
  scale_y_continuous(limit = c(53.5, 66)) 

cols <- c("indianred", "coral", "goldenrod1", "khaki", "lightblue", "steelblue")
```


```{r nitrogen nutrient load pressures map, results = "show", fig.width = 9.5}
basemap +
  geom_sf(
    data = left_join(
      bhi_rgn_simple,
      po_nload %>%
        select(region_id, year, n_score) %>% 
        filter(year %in% seq(1997, 2017, by = 2)),
      # %>% 
      #   tidyr::pivot_wider(names_from = "year", names_prefix = "yr", values_from = "n_score"),
      by = "region_id"
    ),
    mapping = aes(fill = n_score),
    alpha = 0.8
  ) +
  scale_fill_gradientn(
    colors = rev(cols), 
    values = c(0, 10^(-7), 0.3, 0.65, 0.8, 1), 
    limits = c(0, 1)
  ) +
  facet_wrap(~year)
```

```{r phosphorus nutrient load pressures map, results = "show", fig.width = 9.5}
basemap +
  geom_sf(
    data = left_join(
      bhi_rgn_simple,
      po_pload %>%
        select(region_id, year, p_score) %>% 
        filter(year %in% seq(1997, 2017, by = 2)),
      # %>% 
      #   tidyr::pivot_wider(names_from = "year", names_prefix = "yr", values_from = "n_score"),
      by = "region_id"
    ),
    mapping = aes(fill = p_score),
    alpha = 0.8
  ) +
  scale_fill_gradientn(
    colors = rev(cols), 
    values = c(0, 10^(-7), 0.3, 0.65, 0.8, 1), 
    limits = c(0, 1)
  ) +
  facet_wrap(~year)
```


<br>

## 5. Considerations for `BHI3.0` {-}
