---
<<<<<<< HEAD
title: "Food Provision (FP) - Fisheries (FIS) subgoal & Natural Products (NP) Goal Data Prep"
=======
title: "Wild-Caught Fisheries - Food Provision Subgoal data prep"
>>>>>>> 751dbb963601c25a35d2e65c05f27aca59b5122c
output:
  github_document:
    toc: true
    toc_depth: 3
params: 
    datasource: csv
---

```{r Preamble}
loc <- file.path(here::here(), "prep", "FIS")

<<<<<<< HEAD
source(file.path(here::here(), "R", "prep.R"))
knitr::opts_chunk$set(message = FALSE, warning = FALSE, results = "hide")

bkgd_path <- file.path(here::here(), "supplement", "goal_summaries", "FIS.Rmd")
data_path <- file.path(here::here(), "data", "FIS", 2019, "fis_np_data.rmd")
refs_path <- file.path(loc, "fis_np_references.Rmd")
=======
source(file.path(here::here(), "R", "setup.R"))
knitr::opts_chunk$set(message = FALSE, warning = FALSE, results = "hide")

bkgd_path <- file.path(here::here(), "supplement", "goal_summaries", "FIS.Rmd")
data_path <- file.path(here::here(), "data", "FIS", version_year, "fis_np_data.rmd")
refs_path <- file.path(loc, "fis_references.Rmd")
>>>>>>> 751dbb963601c25a35d2e65c05f27aca59b5122c
```

## 1. Background

```{r Background, child = bkgd_path, results = "asis", echo = FALSE}
```

<br/>

## 2. Data

```{r Data, child = data_path, results = "asis", echo = FALSE}
```

<br/>

## 3. Prep: Wrangling & Derivations, Checks/Evaluation, Gapfilling

<<<<<<< HEAD
### 3.1 Cod 
#### 3.1.1 Reorganizing/wrangling
#### 3.1.2 Evaluate flagged data & sampling patterns
#### 3.1.3 Status and trend options and calculation
#### 3.1.4 Gapfilling
#### 3.1.5 Methods discussion

### 3.2 Herring
#### 3.2.1 Reorganizing/wrangling
#### 3.2.2 Evaluate flagged data & sampling patterns
#### 3.2.3 Status and trend options and calculation
#### 3.2.4 Gapfilling
#### 3.2.5 Methods discussion

### 3.3 Sprat
#### 3.3.1 Reorganizing/wrangling
#### 3.3.2 Evaluate flagged data & sampling patterns
#### 3.3.3 Status and trend options and calculation
#### 3.3.4 Gapfilling
#### 3.3.5 Methods discussion

=======
>>>>>>> 751dbb963601c25a35d2e65c05f27aca59b5122c
<br/>

## 4. Visualizing Data Layers

<br/>

## 5. References

```{r References, child = refs_path, results = "asis", echo = FALSE}
<<<<<<< HEAD
```


## 6. Data Prep - Cod

### 6.1 Basic data cleaning and organization


#### Read in Cod data

=======
```
>>>>>>> 751dbb963601c25a35d2e65c05f27aca59b5122c
