
```{r atmos con data preamble, eval = FALSE, echo = TRUE, message = FALSE, warning = FALSE}
source(file.path(here::here(), "R", "data.R"))
source(file.path(here::here(), "R", "common.R"))
source(file.path(here::here(), "R", "spatial.R"))

## root location of the raw data
dir_B <- file.path(dirname(dir_prep), "bhi-data", "BHI 2.0")
dir_rawdata <- file.path(dir_B, "Pressure", "bottom_trawling")
```


## 2. Data {-}

### 2.1 Data sources {-}

#### 2.1.1 Fishing Effort mobile bottom-contacting gear {-}

<!-- dataset save location BHI_share/BHI 2.0/Pressure/bottom_trawling/ -->
```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Pressures and Human Activities:", "Fisheries"), 
  c("Effort (2009-2013):", "Fishing Effort mobile bottom-contacting gear (2010-2013)")
))
colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [HELCOM Map and Data Service](http://metadata.helcom.fi/geonetwork/srv/eng/catalog.search#/metadata/7986ee68-389c-4126-b36f-15407f61cbb5) <br/>") 

# select annual layers (2010-2013)
```
**There is no updated data**

<br/>

**HELCOM metadata**  
*The following text is copied directly from the [HELCOM metadata](http://maps.helcom.fi/website/getMetadata/htm/All/Fishing%20effort%20mobile%20bottom-contacting%20gear%202013.htm) file*  
"This dataset describes fishing effort (hours/c-square) for mobile bottom-contacting gear based on VMS/Log book data processed by ICES Working Group on Spatial Fisheries Data (WGSFD).  

HELCOM requires spatially explicit information on fishing activity affecting the Baltic Sea marine ecosystem for policy purposes. In order to obtain this information a joint ICES/HELCOM/OSPAR data call was issued to relevant authorities of contracting parties to deliver information on fishing activity based on VMS/Log book data. The raw data was submitted to ICES and processed to advice data products by ICES Working Group for Spatial Fisheries (WGSFD) as requested by HELCOM. Processing of the raw data requires specific resources, knowledge and guarantee of anonymity for specific vessels, thus the process was done by ICES WGSFD following Conditions for VMS data use. In 2015 ICES collated Vessel Monitoring System (VMS) and logbook data received; data from Russia were not received. ICES provided to HELCOM advice as fishing abrasion pressure maps as well as fishing effort maps.  

The correct data product citation is following: [ICES. 2015. Fishing abrasion pressure maps for mobile bottom-contacting gears in HELCOM area](http://ices.dk/sites/pub/Publication%20Reports/Data%20outputs/HELCOM_mapping_fishing_intensity_and_effort_data_outputs_2015.zip).   

**Data caveates**  
When using the data for analysis/assessments the following caveats need to be taken into consideration:The methods for identifying fishing activity from the VMS data varied between countries; therefore there may be some country-specific biases that ICES cannot evaluate. Additionally, activities other than active towing of gear may have been incorrectly identified as fishing activity. This would have the effect of overestimating the apparent fishing intensity in ports and in areas used for passage.The data for 2012 and 2013 is not directly comparable to the data of previous years in the data call (2010–2011) due to the gradual increase in VMS-enabled vessels in the range of 12–15 m. This is likely to be most relevant when examining trends in effort for inshore areas. Many countries have substantial fleets of smaller vessels that are not equipped with VMS (The fishing abrasion pressure methodology is based on very broad assumptions in terms of the area affected by abrasion. A single speed and gear width was applied across each gear category in most cases, which can lead to both underestimates and overestimates in actual surface and subsurface abrasion.  

#### 2.1.2 Fishing intensity Dredge 2016 {-}

This dataset describes fishing intensity for Dredge mobile bottom contacting gear in 2016 based on VMS/Log book data processed by ICES Working Group on Spatial Fisheries Data ( [WGSFD](http://ices.dk/community/groups/Pages/WGSFD.aspx)). 

_(For further technical information on the advice request, see http://www.ices.dk/sites/pub/Publication%20Reports/Advice/2017/Special_requests/helcom.2017.18.pdf)_

ICES secretariat collected during 2017 relevant VMS and logbook data for 2009-2016 to produce, as a technical service to HELCOM, updated spatial data layers on fishing intensity/pressure. Improved data quality control checks were implemented. Submitted data across the HELCOM area have improved in quality compared to previous data 2009-2013 published in 2015. Standardized methods were used to produce the requested data layers. The code used to create the data products is available here: https://github.com/ices-eg/wg_WGSFD

<!-- dataset save location BHI_share/BHI 2.0/Pressure/bottom_trawling/ -->
```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Pressures and Human Activities:", "Fisheries"), 
  c("Intensity (2009-2016):", "Fishing intensity Dredge 2016 v2")
))
colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [HELCOM Map and Data Service](http://metadata.helcom.fi/geonetwork/srv/eng/catalog.search#/metadata/dd24a7a8-3552-4bfd-b3a5-ce436a7be237) <br/>") 

```

_We decided to use these data, as is seems the latest compilation for the whole Baltic (this is under "Intesity" instead of "Effort")_

**Data caveates**
Several caveats, listed below, should be taken into account when considering this technical service and its data. These caveats relate to issues concerning the provision of vessel data and its interpretation, and the scale at which data are informative. 
- Data on fishing locations for vessels less than 12 m is not available and are therefore not included in the technical service. This introduces a bias in the assessment that is expected to be strongest in coastal areas. 
- VMS data from Russia were not supplied, introducing a downward bias in pressure in areas fished by Russia. The greatest bias will be in Russian waters. 
- Fishing pressure (SAR, swept area reatio) depends on the spatial resolution of the fishing pressure data. Pressure is calculated at a resolution of 0.05 × 0.05 degrees. 
- It is possible that the valuation of landings has been treated differently by different countries, potentially introducing bias. 
- Data outputs represent vessels over 15 m (2009–2011) and vessels over 12 m (2012–2015). 
- Data outputs in this technical service assume a uniform distribution of trawling within each c-square. When using the data products of this technical service it should be noted that the above assumption will apply when trawling is evaluated over longer time periods (e.g. 2012–2015). However, at shorter, yearly time scales the proportion of the sea floor trawled will be overestimated because trawling is randomly distributed at small spatial scales (Rijnsdorp et al., 1998; Ellis et al., 2014; Eigaard et al., 2016).

### 2.2 Data attributes {-}

**SubsurfSAR**: Subsurface area ratio (Swept area, subsurface =>2 cm penetration depth of the gear components) will be the parameter used to calculate the pressure for the bottom trawling (differently from the BHI 1.0, were 'fishing hours per area' was used from the fishing effort data).

#### 2.2.1 Spatial reference information {-}
*Provided in the HELCOM file*  
Spatial Reference, ArcGIS coordinate system

Type: Projected  
Geographic coordinate reference:GCS_ETRS_1989  
Projection: ETRS_1989_LAEA  

Coordinate reference details:  
Well-known identifier:3035  
X origin: -8426600  
Y origin: -9526700  
XY scale: 10000  
Z origin: 0  
Z scale: 1  
M origin: 0  
M scale: 1  
XY tolerance: 0.001  
Z tolerance: 0.001  
M tolerance: 0.001  
High precision: true  
Latest well-known identifier: 3035  
Well-known text: PROJCS["ETRS_1989_LAEA",GEOGCS["GCS_ETRS_1989",DATUM["D_ETRS_1989",SPHEROID["GRS_1980",6378137.0,298.257222101]],PRIMEM["Greenwich",0.0],UNIT["Degree",0.0174532925199433]],PROJECTION["Lambert_Azimuthal_Equal_Area"],PARAMETER["False_Easting",4321000.0],PARAMETER["False_Northing",3210000.0],PARAMETER["Central_Meridian",10.0],PARAMETER["Latitude_Of_Origin",52.0],UNIT["Meter",1.0],AUTHORITY["EPSG",3035]]  


HELCOM secretariat did the following processing to the shapefile provided by ICES:
- Conversion to ETRS89LAEA coordinate system
- Added feature and attribute "Reported" to display unreported areas.
