
```{r eco data preamble, echo = FALSE, include = FALSE, error = FALSE}
## location of the BHI raw data
source(here::here("R", "setup.R"))
dir_rawdata <- file.path(dir_B, "Goals", "LE", "ECO")
dir_intermediate <- here::here("data", "ECO", version_year, "intermediate")
```
<br>


### 2.1 Datasets with Sources {-}
<br/>

#### 2.1.1 EU Baltic countries GVA Data {-}

Data for the 8 EU Baltic countries were extracted from the [2020 EU Blue economy report](https://ec.europa.eu/maritimeaffairs/sites/maritimeaffairs/files/2020_06_blueeconomy-2020-ld_final.pdf) on July 14 2020 by Andrea De Cervo. 

In particular, detailed info on GVA% by sub-sector and activity per member state were extracted from the tables on the [Annex I](https://ec.europa.eu/maritimeaffairs/sites/maritimeaffairs/files/2020_06_blueeconomy-2020-annexes-ld-part1_final.pdf) and [Annex II](https://ec.europa.eu/maritimeaffairs/sites/maritimeaffairs/files/2020_06_blueeconomy-2020-annexes-ld-part2_final.pdf).


#### 2.1.2 Russian country level data {-}

No GVA data found for Russia.


#### 2.1.3 Reference values {-}

Target is **1.5 % annual growth** between 2009 and 2018, as envisioned in the [2020 EU Blue economy report](https://ec.europa.eu/maritimeaffairs/sites/maritimeaffairs/files/2020_06_blueeconomy-2020-ld_final.pdf). 
Status is calculated for each sector, and goal status is the average of sector status scores. 

<br>

### 2.2 Centralization & Normalization {-}

#### 2.2.1 GVA datasets {-}

The downloaded raw datasets provides an overview of the Blue Economy in the individual Member States. 
They include the established sectors, ensuring a comparable analysis across all the MS. 
In particular, these datasets shows an overview of the Blue Economy by sub-sector and activity for employment, GVA and turnover.

```{r load raw gva data for eco goal, echo = TRUE, message = FALSE, warning = FALSE}

columns <- c(
  "country",
  "Sector",
  "Sub-sector",
  "Activity",
  "Persons employed (2009)",
  "Persons employed (2018)",
  "Persons employed (delta 2018-09)",
  "Turnover (M€) (2009)",
  "Turnover (M€) (2018)",
  "Turnover (M€) (delta 2018-09)",
  "Value added at factor cost (M€) (2009)",
  "Value added at factor cost (M€) (2018)",
  "Value added at factor cost (M€) (delta 2018-09)",
  "GVA to turnover (%) (2009)",
  "GVA to turnover (%) (2018)",
  "GVA to turnover (%) (delta 2018-09)",
  "Gross operating surplus (M€) (2009)",
  "Gross operating surplus (M€) (2018)",
  "Gross operating surplus (M€) (delta 2018-09)",
  "Gross profit margin (%) (2009)",
  "Gross profit margin (%) (2018)",
  "Gross profit margin (%) (delta 2018-09)",
  "Personnel costs per employee (k€) (2009)",
  "Personnel costs per employee (k€) (2018)",
  "Personnel costs per employee (k€)  (delta 2018-09)"
) 
columns <- columns %>% 
    stringr::str_to_lower() %>% 
    stringr::str_remove(" \\(.*\\) ") %>% 
    stringr::str_replace_all(" |\\(|\\)|-", "_") %>% 
    stringr::str_remove("_$") %>% 
    stringr::str_replace_all("__", "_")

for(f in grep(".*_blueconomy.*csv", list.files(dir_rawdata), value = TRUE)){
  
  tmp <- read.csv(file.path(dir_rawdata, f), sep = ";")
  tmp[,] <- sapply(tmp[,], as.character)
  colnames(tmp) <- columns
  
  assign(
    sprintf("%s_gva", substr(f, 1, 3)),
    tmp %>% 
      mutate_at(
        setdiff(columns, c("country", "sector", "sub_sector", "activity")), 
        function(x){as.numeric(stringr::str_replace(stringr::str_remove(x, "%| |-$"), ",", "."))}
      ) %>% 
      mutate_at(
        vars(matches(".*delta.*|.*gva_to_turnover.*|.*gross_profit_margin.*")), 
        function(x){x/100}
      ) %>% 
      filter(country != "", sector != "", sub_sector!= "", activity!= "")
  )
}
```

<br>

**Clean, explore and merge GVA datasets**

```{r merging eu gva data, echo = TRUE, message = FALSE, warning = FALSE}
eco_data_combined <- rbind(den_gva, est_gva, fin_gva, ger_gva, lat_gva, lit_gva, pol_gva, swe_gva) %>% 
  filter(!(sector == "all sectors" & sub_sector == "all sub-sectors"))

#write_csv(
#  eco_data_combined,
#  file.path(dir_intermediate, "le_eco_bluegrowth_bhi2019.csv")
#)

eco_data_combined <- eco_data_combined %>% 
  dplyr::select(
    country, sector, sub_sector, activity, 
    ## only need GVA for 2009 and 2018
    value_added_at_factor_cost_2009,
    value_added_at_factor_cost_2018
  ) %>% 
  
  ## group and aggregate by sub-sector and country
  group_by(country, sector) %>% 
  summarize(
    value_added_at_factor_cost_2009 = sum(value_added_at_factor_cost_2009, na.rm = TRUE), 
    value_added_at_factor_cost_2018 = sum(value_added_at_factor_cost_2018, na.rm = TRUE)
  ) %>% 
  ungroup() %>% 
  
  ## calculate annual growth rate with cagr formula
  ## annualgrowthrate result in decimal form, multiply by 100 to get percent %
  mutate(annualgrowthrate = (value_added_at_factor_cost_2018/value_added_at_factor_cost_2009)^(1/9) - 1) %>% 
  mutate(annualgrowthrate = ifelse(is.nan(annualgrowthrate), NA, annualgrowthrate)) %>% 
  
  ## estimate GVA at years 2009-2018, 
  ## CONVERT VALUE TO SECTOR PROPORTION OF TOTAL COUNTRY MARINE GVA
  ## (so when expand to BHI regions rather than countries, isn't misleading!!!)
  mutate(year = list(c(2009:2018))) %>% 
  tidyr::unnest(cols = c(year)) %>% 
  mutate(value_added_year = value_added_at_factor_cost_2009*((annualgrowthrate+1)^(year-2009))) %>% 
  group_by(country, year) %>% 
  mutate(annual_total_marine_value_added = sum(value_added_year, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(prop_annual_value_added = value_added_year/annual_total_marine_value_added)
  
  
## add bhi regions
eco_data_bhirgns <- read_csv(file.path(dir_prep, "supplement", "lookup_tabs", "rgns_complete.csv")) %>% 
  dplyr::select(region_id, country = eez) %>% 
  mutate(year = list(c(2009:2018))) %>%
  tidyr::unnest(cols = c(year)) %>%
  full_join(eco_data_combined, by = c("country", "year")) %>% 
 
  ## convert the annual growth rate into a perentage and round to just one decimal place
  ## NOTE: NOW GROWTH RATE WILL BE IN PERCENTAGE NOT DECIMAL FORM
  mutate(annualgrowthrate = round(annualgrowthrate*100, 1))
```


**Plot of Blue Economy approximate annual growth by country**
```{r plot of blue economy approximate annual growth by country, fig.width = 9.5, fig.height = 8}
ggplot(eco_data_bhirgns %>% filter(!is.na(sector))) + 
  geom_line(aes(x = year, y = value_added_year, color = sector), size = 0.6) +
  scale_color_manual(values = colorRampPalette(RColorBrewer::brewer.pal(8, "Set2"))(30)[sample(1:30)]) +
  facet_wrap(~country, scales = "free_y") +
  labs(x = NULL, y = "Gross value added, approximated from 2009 and 2018 data, M€", color = "Sector, Blue Economy") +
  theme_bw() +
  theme(legend.position = c(0.85, 0.15), legend.background = element_rect(color = "steelblue"))
```

```{r save eco data layers, eval = FALSE, echo = TRUE, message = FALSE, warning = FALSE}
write_csv(
  select(eco_data_bhirgns, region_id, sector, year, annual_growth_rate = annualgrowthrate),
  file.path(dir_layers, "le_eco_bluegrowth_rates_bhi2019.csv")
)
write_csv(
  select(
    eco_data_bhirgns, 
    region_id, sector, year, 
    gva_sector_prop = prop_annual_value_added, 
    country_gva_estim = value_added_year
  ),
  file.path(dir_layers, "le_eco_yearly_gva_bhi2019.csv")
)
```

