
## 2. Data

```{r Preamble, echo = FALSE, include = FALSE, error = FALSE}
source(file.path(here::here(), "R", "data.R"))
```

### 2.1 Datasets with Sources
<br/>

#### Landings (for F/FMSY) and SSB (for B/BMSY) Data
<br/>

**Cod in subdivisions 22-24, western Baltic stock**  
<!-- dataset save location BHI_share/2.0/Goals/FP/Fisheries/Cod/cod_SDs22-24 -->

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c(":", ""), 
  c(":", "")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [ICES database](http://standardgraphs.ices.dk/ViewCharts.aspx?key=10446) <br/> Downloaded 2019-10-07 by Andrea De Cervo")
```
<br/>

**DATASET NAME/VARIABLE**  
<!-- dataset save location BHI_share/2.0/Goals/ -->

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c(":", ""), 
  c(":", "")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [ABC database](URL/LINK) <br/> Downloaded DATE by NAME")
```
<br/>

**DATASET NAME/VARIABLE**  
<!-- dataset save location BHI_share/2.0/Goals/ -->

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c(":", ""), 
  c(":", "")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [ABC database](URL/LINK) <br/> Downloaded DATE by NAME")
```
<br/>

**DATASET NAME/VARIABLE**  
<!-- dataset save location BHI_share/2.0/Goals/ -->

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c(":", ""), 
  c(":", "")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [ABC database](URL/LINK) <br/> Downloaded DATE by NAME")
```
<br/>

#### FMSY and BMSY

[Cod (Gadus morhua) in subdivisions 22–24, western Baltic stock (western Baltic Sea)](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/cod.27.22-24.pdf): BMSY: 21 876 ; FMSY (Ftotal 2020): 0.26  
[Cod (Gadus morhua) in subdivisions 24–32, eastern Baltic stock (eastern Baltic Sea)](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/cod.27.24-32.pdf): BMSY (Bpa): 108 035 ; FMSY: 0.3  

[Herring (Clupea harengus) in Subdivision 28.1 (Gulf of Riga)](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/her.27.28.pdf): BMSY 60 000 ; FMSY 0.32

Herring in subdivisions 20–24: BMSY 150000 , FMSY : 0.31
Herring in subdivisions 25–29 and 32, excluding the Gulf of Riga. : BMSY: 600000 ; FMSY:0.22
Herring in subdivisions 30 and 31. BMSY: 140998 ; FMSY: 0.15


[Sprat in subdivisions 22–32] : BMSY : 570000 : FMSY 0.26



### 2.2 Centralization & Normalization

#### 2.2.1 Standardize Units

#### 2.2.2 Rename Fields/Variables

#### 2.2.3 Save to BHI Database

<br/>

### 2.3 Initial Data Exploration

#### 2.3.1 Compare versus Previous Years Data

#### 2.3.2 Timeseries Plots

#### 2.3.3 Map

