
```{r Preamble, echo = FALSE, include = FALSE, error = FALSE}
source(here::here("R", "setup.R"))
# source(file.path(here::here(), "R", "data.R"))
```

### 2.1 Datasets with Sources {-}

#### 2.1.1 Landings (for F/FMSY) and SSB (for B/BMSY) Data {-}

**Cod in subdivisions 22-24, western Baltic stock**  
<!-- dataset save location BHI_share/BHI 2.0/Goals/FP/FIS/cod_SDs22_24/cod_SDs22_24.csv -->

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Species:", "Gadus morhua"), 
  c("EcoRegion (Fishstock):", "Baltic Sea"), 
  c("Assessment year:", "2019"),
  c("FishStock:", "cod.27.22-24"),
  c("Assessment Key:", "10446")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, format = "html", caption = "Source: [ICES database](http://standardgraphs.ices.dk/ViewCharts.aspx?key=10446) <br/> Downloaded 2019-10-07 by Andrea De Cervo") %>% 
  kableExtra::kable_styling(latex_option = c("hold_position"), position = "left")
```

<br>

**Cod in subdivisions 24-32, eastern Baltic stock**  
<!-- dataset save location BHI_share/BHI 2.0/Goals/FP/FIS/cod_SDs24_32/cod_SDs24_32.csv -->

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Species:", "Gadus morhua"), 
  c("EcoRegion (Fishstock):", "Baltic Sea"), 
  c("Assessment year:", "2019"),
  c("FishStock:", "cod.27.24-32"),
  c("Assessment Key:", "12941")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, format = "html", caption = "Source: [ICES database](http://standardgraphs.ices.dk/ViewCharts.aspx?key=12941) <br/> Downloaded 2019-10-07 by Andrea De Cervo") %>% 
  kableExtra::kable_styling(latex_option = c("hold_position"), position = "left")
```

<br>

**Herring in subdivisions 20-24 -Skagerrak, Kattegat and western Batic-**  
<!-- dataset save location BHI_share/BHI 2.0/Goals/FP/FIS/herring_SDs20_24/herring_SDs20_24.csv -->

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Species:", "Clupea harengus"), 
  c("EcoRegion (Fishstock):", "Baltic Sea"), 
  c("Assessment year:", "2019"),
  c("FishStock:", "her.27.20-24"),
  c("Assessment Key:", "12592")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, format = "html", caption = "Source: [ICES database](http://standardgraphs.ices.dk/ViewCharts.aspx?key=12592) <br/> Downloaded 2019-10-07 by Andrea De Cervo") %>% 
  kableExtra::kable_styling(latex_option = c("hold_position"), position = "left")
```

<br>

**Herring in subdivisions 25-29,32 -central Baltic Sea (excluding Gulf of Riga)-**  
<!-- dataset save location BHI_share/BHI 2.0/Goals/FP/FIS/herring_SDs25_29_32/herring_SDs25_29_32.csv -->

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Species:", "Clupea harengus"), 
  c("EcoRegion (Fishstock):", "Baltic Sea"), 
  c("Assessment year:", "2019"),
  c("FishStock:", "her.27.25-2932"),
  c("Assessment Key:", "10408")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, format = "html", caption = "Source: [ICES database](http://standardgraphs.ices.dk/ViewCharts.aspx?key=10408) <br/> Downloaded 2019-10-07 by Andrea De Cervo") %>% 
  kableExtra::kable_styling(latex_option = c("hold_position"), position = "left")
```

<br>

**Herring in subdivision 28.1 (Gulf of Riga)**  
<!-- dataset save location BHI_share/BHI 2.0/Goals/FP/FIS/herring_SD_28.1/herring_SD_28.1.csv -->

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Species:", "Clupea harengus"), 
  c("EcoRegion (Fishstock):", "Baltic Sea"), 
  c("Assessment year:", "2019"),
  c("FishStock:", "her.27.28"),
  c("Assessment Key:", "10404")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, format = "html", caption = "Source: [ICES database](http://standardgraphs.ices.dk/ViewCharts.aspx?key=10404) <br/> Downloaded 2019-10-08 by Andrea De Cervo") %>% 
  kableExtra::kable_styling(latex_option = c("hold_position"), position = "left")
```

<br>

**Herring in subdivisions 30-31 (Gulf of Bothnia)**  
<!-- dataset save location BHI_share/BHI 2.0/Goals/FP/FIS/herring_SDs30_31/herring_SDs30_31.csv -->

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Species:", "Clupea harengus"), 
  c("EcoRegion (Fishstock):", "Baltic Sea"), 
  c("Assessment year:", "2019"),
  c("FishStock:", "her.27.3031"),
  c("Assessment Key:", "12738")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, format = "html", caption = "Source: [ICES database](http://standardgraphs.ices.dk/ViewCharts.aspx?key=12738) <br/> Downloaded 2019-10-08 by Andrea De Cervo") %>% 
  kableExtra::kable_styling(latex_option = c("hold_position"), position = "left")
```

<br>

**Sprat in subdivisions 22-32 (Baltic Sea)**  
<!-- dataset save location BHI_share/BHI 2.0/Goals/NP/sprat_SDs22_32/sprat_SDs22_32.csv -->

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Species:", "Sprattus sprattus"), 
  c("EcoRegion (Fishstock):", "Baltic Sea"), 
  c("Assessment year:", "2019"),
  c("FishStock:", "spr.27.22-32"),
  c("Assessment Key:", "12942")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, format = "html", caption = "Source: [ICES database](http://standardgraphs.ices.dk/ViewCharts.aspx?key=12942) <br/> Downloaded 2019-10-08 by Andrea De Cervo") %>% 
  kableExtra::kable_styling(latex_option = c("hold_position"), position = "left")
```

<br>

#### 2.1.2 Cod Trawl Survey {-}

**Baltic International Trawl Survey (BITS) Cod Length and Weight - ICES subdivisions 21-29**  
<!-- dataset save location BHI_share/2.0/Goals/FP/FIS/SMALK_2019-10-22 16_34_01/SMALK_2019-10-22_16_34_01.csv -->

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Data Product:", "SMALK"),
  c("Survey:", "BITS (Baltic International Trawl Survey)"),
  c("Quarters:", "4"),
  c("Years:", "All"),
  c("Species:", "All species -> Gadus morhua")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, format = "html", caption = "Source: [ICES DATRAS database](https://datras.ices.dk/Data_products/Download/Download_Data_public.aspx) <br/> Downloaded 2019-10-22 by Eleanore Campbell") %>% 
  kableExtra::kable_styling(latex_option = c("hold_position"), position = "left")
```

<br>

**Baltic International Trawl Survey (BITS) Cod CPUE - ICES subdivisions 21-29**  
<!-- dataset save location BHI_share/2.0/Goals/FP/FIS/CPUE_per_length_per_haul_per_hour_2019-11-07_13_36_17/CPUE_per_length_per_haul_per_hour_2019-11-07_13_36_17.csv -->

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Data Product:", "CPUE per length per haul per hour"),
  c("Survey:", "BITS (Baltic International Trawl Survey)"),
  c("Quarters:", "All"),
  c("Years:", "All"),
  c("Ships:", "All"),
  c("Gear:", "All"),
  c("Areas:", "All"),
  c("Species:", "Standard species -> Gadus morhua")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, format = "html", caption = "Source: [ICES DATRAS database](https://datras.ices.dk/Data_products/Download/Download_Data_public.aspx) <br/> Downloaded 2019-11-07 by Eleanore Campbell") %>% 
  kableExtra::kable_styling(latex_option = c("hold_position"), position = "left")
```

<br>

#### 2.1.3 FMSY and BMSY Reference Points {-}

**Cod (Gadus morhua) in subdivisions 22–24, western Baltic stock (western Baltic Sea)** [Reference points](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/cod.27.22-24.pdf): 
BMSY: 21 876 ; FMSY (Ftotal 2020): 0.26  

**Cod (Gadus morhua) in subdivisions 24–32, eastern Baltic stock (eastern Baltic Sea)** [Reference points](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/cod.27.24-32.pdf): 
BMSY (Bpa): 108 035 ; FMSY: 0.3  

**Herring (Clupea harengus) in Subdivisions 20-24 (Skagerrak, Kattegat and western Baltic)** [Reference points](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/her.27.20-24.pdf): 
BMSY 150 000 ; FMSY 0.31

**Herring (Clupea harengus) in Subdivisions 25-29,32 (excluding Gulf of Riga)** [Reference points](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/her.27.25-2932.pdf): 
BMSY 600 000 ; FMSY 0.22

**Herring (Clupea harengus) in Subdivision 28.1 (Gulf of Riga)** [Reference points](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/her.27.28.pdf): 
BMSY 60 000 ; FMSY 0.32

**Herring (Clupea harengus) in Subdivisions 30 and 31 (Gulf of Bothnia)** [Reference points](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/her.27.3031.pdf): 
BMSY 140 998 ; FMSY 0.15

**Sprat (Sprattus sprattus) in Subdivisions 22-32 (Baltic Sea)** [Reference points](http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/spr.27.22-32.pdf): 
BMSY 570 000 ; FMSY 0.26


### 2.2 Centralization & Normalization {-}


Each assessed stock has its own raw datafile. Cod and herring data are saved under `Goals/FP/FIS` and sprat data are saved in the `Goals/NP` folder.

The MSY values (FMSY and BMSY for each assessed stock) are taken from the ICES advice reports, linked in the `standardgraphs.ices.dk` aspx pages. No table is saved with these values- they are entered directly into a dataframe in the code below.

```{r load fisheries raw data, message = FALSE, warning = FALSE, echo = TRUE}
## root location of the raw data
dir_rawdata <- file.path(dir_B, "Goals", "FP", "FIS") # list.files(dir_rawdata)

## Fisheries data from ICES
## if csvs are saved w/ semicolons:
# source(file.path(here::here(), "R", "semicolon_to_comma.R"))
# lapply(
#   list("cod_SDs22_24", "cod_SDs24_32",
#        "herring_SD_28.1", "herring_SDs20_24",
#        "herring_SDs25_29_32", "herring_SDs30_31"),
#   function(x){
#      fp <- file.path(dir_rawdata, x, paste0(x, "_reformat.csv"))
#      semicolon_to_comma(fp, remove_na_cols = TRUE, overwrite = TRUE)
#   }
# )
# semicolon_to_comma(file.path(dir_B,"Goals","NP","sprat_SDs22_32","sprat_SDs22_32_reformat.csv"), TRUE, TRUE)


## cod data
cod1raw <- read_csv(file.path(dir_rawdata, "cod_SDs22_24", "cod_SDs22_24_reformat.csv"))
cod2raw <- read_csv(file.path(dir_rawdata, "cod_SDs24_32", "cod_SDs24_32_reformat.csv"))

## herring data
herring1raw <- read_csv(file.path(dir_rawdata, "herring_SD_28.1", "herring_SD_28.1_reformat.csv"))
herring2raw <- read_csv(file.path(dir_rawdata, "herring_SDs20_24", "herring_SDs20_24_reformat.csv"))
herring3raw <- read_csv(file.path(dir_rawdata, "herring_SDs25_29_32", "herring_SDs25_29_32_reformat.csv"))
herring4raw <- read_csv(file.path(dir_rawdata, "herring_SDs30_31", "herring_SDs30_31_reformat.csv"))

## sprat data
sprat1raw <- read_csv(file.path(dir_B, "Goals", "NP", "sprat_SDs22_32", "sprat_SDs22_32_reformat.csv"))


## make MSY values table
## these values are obtained from ICES reports, see data/FIS/fis_np_data.rmd for more details
msy_data <- t(data.frame(
  c("cod", "22-24", "cod_SDs22_24", 21876, 0.26),
  c("cod", "24-32", "cod_SDs24_32", 108035, 0.3),
  c("herring", "28.1", "herring_SD_28.1", 60000, 0.32),
  c("herring", "20-24", "herring_SDs20_24", 150000, 0.31),
  c("herring", "25-29,32", "herring_SDs25_29_32", 600000, 0.22),
  c("herring", "30-31", "herring_SDs30_31", 140998, 0.15),
  c("sprat", "22-32", "sprat_SDs22_32", 570000, 0.26)
))
## for testing effect of using 2013 values...
# msy_2013_data <- t(data.frame(
#   c("cod", "22-24", "cod_SDs22_24", 36400, 0.26),
#   c("cod", "24-32", "cod_SDs24_32", 88200, 0.46),
#   c("herring", "28.1", "herring_SD_28.1", 60000, 0.35),
#   c("herring", "20-24", "herring_SDs20_24", 110000, 0.28),
#   c("herring", "25-29,32", "herring_SDs25_29_32", 600000, 0.26),
#   c("herring", "30-31", "herring_SDs30_31", 316000, 0.15),
#   c("sprat", "22-32", "sprat_SDs22_32", 570000, 0.29)
# ))
# msy_data <- msy_2013_data
colnames(msy_data) <- c("species", "SDs", "stockname", "BMSY", "FMSY")
rownames(msy_data) <- NULL
msy_data <- as_tibble(msy_data)
```


#### 2.2.0 Merge datasets and calculate F/FMSY and B/BMSY ratios {-}

```{r merge datasets and calculate ffmsy bbmsy ratios, echo = TRUE}
combined_rawdata <- rbind(
  ## cod
  cod1raw %>% 
    dplyr::mutate(catch_tonnes = ifelse(
      !is.na(discards_tonnes), 
      landings_tonnes + discards_tonnes,
      landings_tonnes
    )) %>%
    dplyr::select(
      year, ssb, 
      fis_mort = fishing_mortality_age3_5, 
      catch = catch_tonnes
      # landings = landings_tonnes
    ) %>%
    mutate(stockname = "cod_SDs22_24"),
  cod2raw %>% 
    dplyr::mutate(catch_tonnes = ifelse(
      !is.na(discards_tonnes), 
      landings_tonnes + discards_tonnes,
      landings_tonnes
    )) %>%
    dplyr::select(
      year, ssb, 
      fis_mort = fishing_mortality_age4_6, 
      catch = catch_tonnes
      # landings = landings_tonnes
    ) %>% 
    mutate(stockname = "cod_SDs24_32"),
  ## herring
  herring1raw %>% 
    dplyr::select(
      year, ssb = ssb_tonnes, 
      fis_mort = `F`, 
      catch = catches_tonnes # no data on discards, only catch...
    ) %>% 
    mutate(stockname = "herring_SD_28.1"),
  herring2raw %>% 
    dplyr::select(
      year, ssb = ssb_tonnes, 
      fis_mort = F_age3_6, 
      catch = catches_tonnes
    ) %>% 
    mutate(stockname = "herring_SDs20_24"),
  herring3raw %>% 
    dplyr::select(
      year, ssb = ssb_tonnes, 
      fis_mort = F_age3_6, 
      catch = catches_tonnes
    ) %>% 
    mutate(stockname = "herring_SDs25_29_32"),
  herring4raw %>% 
    dplyr::select(
      year, ssb = ssb, 
      fis_mort = F_age3_7, 
      catch = catches_tonnes
    ) %>% 
    mutate(stockname = "herring_SDs30_31"),
  ## sprat
  sprat1raw %>% 
    dplyr::select(
      year, ssb = ssb_tonnes, 
      fis_mort = F_age3_5, 
      catch = catches_tonnes
    ) %>% 
    mutate(stockname = "sprat_SDs22_32")) %>% 
  ## join with msy data
  filter(!is.na(year)) %>% 
  left_join(msy_data, by = c("stockname")) %>% 
  mutate(bbmsy = ssb/as.numeric(BMSY), ffmsy = fis_mort/as.numeric(FMSY)) %>% 
  arrange(species, stockname, year)
```

<br>

#### 2.2.1 Standardize Units: Match BHI Regions to ICES Subdivisions {-}

[Map of ICES regions.](https://www.ices.dk/marine-data/Documents/Maps/ICES-Ecoregions-hybrid-statistical-areas.png)
The map below shows the overlap between ICES subdivisions and the BHI regions:

```{r map how the ICES and BHI regions overlap, results = "show", message = FALSE, echo = TRUE, fig.width = 9.5}
source(here::here("R", "spatial.R"))
regions_shape() # loads spatial features objects

## ICES shapefile has crs EPSG:3035
## https://spatialreference.org/ref/epsg/etrs89-etrs-laea/
## transform to match BHI crs of EPSG:4326, and simplify

ices_transform <- rmapshaper::ms_simplify(input = ICES_rgns_shp) %>% 
  sf::st_as_sf() %>% 
  sf::st_transform(crs = 4326)
bhi_rgns_simple <- rmapshaper::ms_simplify(input = BHI_rgns_shp) %>% 
  sf::st_as_sf() # also simplify BHI shp for plotting

map_bhi_ices <- ggplot2::ggplot() + 
  ## baltic countries borders
  geom_sf(
    data = rnaturalearth::ne_countries(scale = "medium", returnclass = "sf") %>%
      st_crop(xmin = 0, xmax = 40, ymin = 53, ymax = 67),
    fill = "ivory", 
    size = 0.1
  ) +
  ## ICES areas
  geom_sf(
    data = ices_transform, 
    aes(fill = ICES_area), 
    color = NA,
    alpha = 0.5
  ) +
  scale_fill_manual(values = colorRampPalette(RColorBrewer::brewer.pal(9, "Set3"))(14)) +
  # scale_fill_manual(values = as.vector(pals::kelly(n = 14))) + # library(pals)
  # scale_fill_manual(values = colorRampPalette(beyonce_palette(127))(14)) + # library(beyonce)
  labs(fill = "ICES Subdivisions", x = NULL, y = NULL) +
  ## BHI regions with ID numbers
  geom_sf(data = bhi_rgns_simple, fill = NA, size = 0.15, color = "grey40") +
  scale_x_continuous(limit = c(5, 37)) +
  scale_y_continuous(limit = c(53.5, 66)) +
  theme(
    panel.background = element_rect(fill = "#F8FBFC", color = "#E2EEF3"),
    legend.position = c(0.9, 0.6)
  )
map_bhi_ices + geom_sf_text(data = bhi_rgns_simple, aes(label = BHI_ID)) 
```


A lookup table derived from the overlap between ICES subdivisions and BHI regions as shown in the map above, is used in the code below to match BHI region IDs to the raw dataset:

```{r converting ICES fisheries regions to Baltic regions, echo = TRUE}
## based on 'prep/FIS/raw/DataOrganization.R' by Melanie Frazier March 16 2016, in bhi-1.0-archive
## table for matching ICES to BHI regions
regions <- read_csv(
  here::here("supplement", "lookup_tabs", "ices_to_bhi_lookup.csv"), 
  col_types = cols()) %>% 
  mutate(ices_numeric = ifelse(ices_numeric == 3, 21, ices_numeric)) # 3a.21 overlaps BHI rgn 1 & 2

## convert ICES stock assessments to BHI regions
combined_w_rgns <- combined_rawdata %>% 
  
  ## expand ices subdivisions categories to one row per subdiv
  rowwise() %>% 
  mutate(
    ## check unique(combined_rawdata$SDs) to make sure this string parsing will work
    sd_from = as.numeric(str_split(SDs, "-|,")[[1]][1]),
    sd_to = as.numeric(str_split(SDs, "-|,")[[1]][2]),
    sd_extra = as.numeric(str_split(SDs, "-|,")[[1]][3])
  ) %>% 
  mutate(
    sd_to = ifelse(is.na(sd_to), sd_from, sd_to),
    sd_extra = ifelse(is.na(sd_extra), sd_from, sd_extra),
    incl28.2 = sd_from <= 28 & sd_to >= 28,
    incl28.1 = sd_from <= 28 & sd_to >= 28 & species != "herring"
  ) %>% 
  ## ICES regions 28.1 for Riga and 28.2 elsewhere but no 28
  mutate(SDs = list(
    unique(
      c(sd_from:sd_to, sd_extra)[c(sd_from:sd_to, sd_extra)!=28] %>% 
        c(ifelse(incl28.2, 28.2, sd_from)) %>% 
        c(ifelse(incl28.1, 28.1, sd_from))
    )
  )) %>% 
  tidyr::unnest(ices_subdiv = SDs) %>% 
  ## NOTE:
  ## using catches instead of landings, because herring landings were not reported, just catch
  ## but i dont want to go back through code and change everything to from 'landings' to 'catch'
  dplyr::select(year, species, stockname, ssb, BMSY, bbmsy, fis_mort, FMSY, ffmsy, landings = catch, ices_subdiv) %>%
  filter(ices_subdiv != 20) %>% # filter because no BHI region assigned to SD 20 and will cause NA means...
  
  ## add BHI regions info using ices_to_bhi_lookup
  left_join(dplyr::select(regions, region_id, ices_numeric, area_km2), by = c("ices_subdiv" = "ices_numeric"))

## ICES subdivs 20 and 21 are North Sea and left out of BHI1.0, but herring in 3a.21 for rgn 1 & 2
# combined_w_rgns <- combined_w_rgns %>% filter(!ices_subdiv %in% c(20, 21))
```

<br>

```{r cod condition ie fultons K from BITS surveys, echo = TRUE}
surveyBITS_cod <- read_csv(file.path(dir_rawdata, "SMALK_2019-10-22_16_34_01", "SMALK_2019-10-22_16_34_01.csv"))
# unique(surveyBITS_cod$Country) # no data for Finland, or really any areas north of North Baltic Proper

q4_fultonsK <- surveyBITS_cod %>%
  mutate(length_cm = LngtClass/10) %>% 
  dplyr::select(year = Year, ices_subdiv = Area, eez = Country, length_cm, wgt_gram = IndWgt) %>% 
  mutate(
    ## revise country EEZ codes to be consistent with codes used elsewhere by BHI
    eez = ifelse(eez == "GFR", "DEU", ifelse(eez == "DEN", "DNK", ifelse(eez == "LAT", "LVA", eez))),
    ## calculate fultons K and assign length classes
    fultonsK = wgt_gram/(length_cm^3)*100,
    length_group = floor(length_cm/10)
  ) %>% 
  filter(!is.na(fultonsK), fultonsK < 2) %>% 
  ## per Casini et al exclude lengths < 10cm and ≥60cm-- not enough data:
  filter(length_group %in% 1:5) %>% 
  mutate(
    length_group = ifelse(
      length_group == 1, "10-20cm", ifelse(
         length_group == 2, "20-30cm", ifelse(
           length_group == 3, "30-40cm", "40-60cm"
         )
      )
    )
  ) %>% 
  ## join BHI regions
  ## area 28 includes both 28.2 and gulf of Riga 28.1 
  ## http://www.ices.dk/marine-data/Documents/DATRAS/Survey_Maps_Datras.pdf
  mutate(ices_subdiv = ifelse(ices_subdiv == 28, list(28.1, 28.2), ices_subdiv)) %>% 
  tidyr::unnest(ices_subdiv = ices_subdiv) %>% 
  right_join(
    regions %>% 
      dplyr::select(region_id, eez, ices_subdiv = ices_numeric) %>% 
      filter(ices_subdiv %in% c(unique(surveyBITS_cod$Area), 28.1, 28.2)),
    by = c("ices_subdiv", "eez")
  )

# filter(q4_fultonsK, is.na(fultonsK)) # NA check
q4_fultonsK <- filter(q4_fultonsK, !is.na(fultonsK))
## will give BHI region 1 have same penalty as 2, BHI 5 same as 6, and BHI 29 same as 26
## as these regions are small, close, and/or share overlapping ICES areas; see ICES vs BHI regions map above
```


**No cod in Northern part of the Baltic Sea! Filter combined fisheries dataset to reflect this.**

```{r distributions of cod from BITS surveys CPUE, results = "show", message = FALSE, echo = TRUE, fig.width = 9.5, out.width = "120%"}
cpue_sf <- read_csv(file.path(
  dir_rawdata, 
  "CPUE_per_length_per_haul_per_hour_2019-11-07_13_36_17", 
  "CPUE_per_length_per_haul_per_hour_2019-11-07_13_36_17.csv")) %>% 
  dplyr::select(Year, DateTime, Ship, Gear, ShootLat, ShootLong, rateCPUE = CPUE_number_per_hour) %>% 
  filter(Year >= 2008) %>% 
  distinct() %>% 
  sf::st_as_sf(coords = c("ShootLong", "ShootLat"), crs = 4326)

map_bhi_ices + 
  geom_sf(aes(size = rateCPUE), color = "royalblue3", alpha = 0.02, shape = 16, data = cpue_sf) + 
  labs(size = "Number/Hour", x = NULL, y = NULL) +
  guides(fill = FALSE) + 
  theme(legend.position = c(0.15, 0.85))

combined_w_rgns <- combined_w_rgns %>% 
  filter(!(species == "cod" & region_id %in% 27:42))
```
<br>

#### 2.2.2 Save Datasets {-}

```{r save full merged dataset with raw data and calculated MSY ratios, echo = TRUE, eval = FALSE}
## full merged dataset with MSY and raw data, as well as calculated ratios
## this will be used for the shiny app among other things!
write_csv(
  combined_w_rgns %>% rename(catch = landings), 
  here::here("data", "FIS", version_year, "intermediate", "fis_full_merged_dataset.csv")
)
```

```{r save Eastern cod fultons K dataset, echo = TRUE, eval = FALSE, eval = FALSE}
## full merged dataset with MSY and raw data, as well as calculated ratios
## this will be used for the shiny app among other things!
write_csv(
  q4_fultonsK, 
  file.path(here::here(), "data", "FIS", version_year, "intermediate", "fis_cod_fultonsK.csv")
)
```
