
## 2. Data

```{r Preamble, echo = FALSE, include = FALSE, error = FALSE}
source(file.path(here::here(), "R", "data.R"))
```

### 2.1 Datasets with Sources
<br/>

#### Landings (for F/FMSY) and SSB (for B/BMSY) Data
<br/>

**Cod in subdivisions 22-24, western Baltic stock**  
<!-- dataset save location BHI_share/2.0/Goals/FP/Fisheries/Cod/cod_SDs22-24 -->

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Species:", "Gadus morhua"), 
  c("EcoRegion (Fishstock):", "Baltic Sea"), 
  c("Assessment year:", "2019"),
  c("FishStock:", "cod.27.22-24"),
  c("Assessment Key:", "10446")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [ICES database](http://standardgraphs.ices.dk/ViewCharts.aspx?key=10446) <br/> Downloaded 2019-10-07 by Andrea De Cervo")
```
<br/>

**Cod in subdivisions 24-32, eastern Baltic stock**  
<!-- dataset save location BHI_share/2.0/Goals/FP/Fisheries/Cod/cod_SDs24-32 -->

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Species:", "Gadus morhua"), 
  c("EcoRegion (Fishstock):", "Baltic Sea"), 
  c("Assessment year:", "2019"),
  c("FishStock:", "cod.27.24-32"),
  c("Assessment Key:", "12941")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [ICES database](http://standardgraphs.ices.dk/ViewCharts.aspx?key=12941) <br/> Downloaded 2019-10-07 by Andrea De Cervo")
```
<br/>

**Herring in subdivisions 20-24 -Skagerrak, Kattegat and western Batic-**  
<!-- dataset save location BHI_share/2.0/Goals/FP/Fisheries/Herring/herring_SDs20-24 -->

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Species:", "Clupea harengus"), 
  c("EcoRegion (Fishstock):", "Baltic Sea"), 
  c("Assessment year:", "2019"),
  c("FishStock:", "her.27.20-24"),
  c("Assessment Key:", "12592")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [ICES database](http://standardgraphs.ices.dk/ViewCharts.aspx?key=12592) <br/> Downloaded 2019-10-07 by Andrea De Cervo")
```
<br/>

**Herring in subdivisions 25-29,32 -central Baltic Sea (excluding Gulf of Riga)-**  
<!-- dataset save location BHI_share/2.0/Goals/FP/Fisheries/Herring/herring_SDs25-29,32 -->

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Species:", "Clupea harengus"), 
  c("EcoRegion (Fishstock):", "Baltic Sea"), 
  c("Assessment year:", "2019"),
  c("FishStock:", "her.27.25-2932"),
  c("Assessment Key:", "10408")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [ICES database](http://standardgraphs.ices.dk/ViewCharts.aspx?key=10408) <br/> Downloaded 2019-10-07 by Andrea De Cervo")
```
<br/>

**Herring in subdivision 28.1 (Gulf of Riga)**  
<!-- dataset save location BHI_share/2.0/Goals/FP/Fisheries/Herring/herring_SD_28.1 -->

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Species:", "Clupea harengus"), 
  c("EcoRegion (Fishstock):", "Baltic Sea"), 
  c("Assessment year:", "2019"),
  c("FishStock:", "her.27.28"),
  c("Assessment Key:", "10404")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [ICES database](http://standardgraphs.ices.dk/ViewCharts.aspx?key=10404) <br/> Downloaded 2019-10-08 by Andrea De Cervo")
```
<br/>

**Herring in subdivisions 30-31 (Gulf of Bothnia)**  
<!-- dataset save location BHI_share/2.0/Goals/FP/Fisheries/Herring/herring_SDs30-31 -->

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Species:", "Clupea harengus"), 
  c("EcoRegion (Fishstock):", "Baltic Sea"), 
  c("Assessment year:", "2019"),
  c("FishStock:", "her.27.3031"),
  c("Assessment Key:", "12738")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [ICES database](http://standardgraphs.ices.dk/ViewCharts.aspx?key=12738) <br/> Downloaded 2019-10-08 by Andrea De Cervo")
```
<br/>

**Sprat in subdivisions 22-32 (Baltic Sea)**  
<!-- dataset save location BHI_share/2.0/Goals/NP/Sprat/sprat_SDs22-32 -->

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Species:", ""), 
  c("EcoRegion (Fishstock):", "Baltic Sea"), 
  c("Assessment year:", "2019"),
  c("FishStock:", "cod.27.22-24"),
  c("Assessment Key:", "10446")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [ICES database](http://standardgraphs.ices.dk/ViewCharts.aspx?key=12942) <br/> Downloaded 2019-10-08 by Andrea De Cervo")
```
<br/>

#### FMSY and BMSY

[Cod (Gadus morhua) in subdivisions 22–24, western Baltic stock (western Baltic Sea)](Reference points: http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/cod.27.22-24.pdf): 
BMSY: 21 876 ; FMSY (Ftotal 2020): 0.26  

[Cod (Gadus morhua) in subdivisions 24–32, eastern Baltic stock (eastern Baltic Sea)](Reference points: http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/cod.27.24-32.pdf): 
BMSY (Bpa): 108 035 ; FMSY: 0.3  

[Herring (Clupea harengus) in Subdivisions 20-24 (Skagerrak, Kattegat and western Baltic)](Reference points: http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/her.27.20-24.pdf): 
BMSY 150 000 ; FMSY 0.31

[Herring (Clupea harengus) in Subdivisions 25-29,32 (excluding Gulf of Riga)](Reference points: http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/her.27.25-2932.pdf): 
BMSY 600 000 ; FMSY 0.22

[Herring (Clupea harengus) in Subdivision 28.1 (Gulf of Riga)](Reference points: http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/her.27.28.pdf): 
BMSY 60 000 ; FMSY 0.32

[Herring (Clupea harengus) in Subdivisions 30 and 31 (Gulf of Bothnia)](Reference points: http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/her.27.3031.pdf): 
BMSY 140 998 ; FMSY 0.15

[Sprat (Sprattus sprattus) in Subdivisions 22-32 (Baltic Sea)](Reference points: http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/spr.27.22-32.pdf): 
BMSY 570 000 ; FMSY 0.26


### 2.2 Centralization & Normalization

#### 2.2.1 Standardize Units

#### 2.2.2 Rename Fields/Variables

#### 2.2.3 Save to BHI Database

<br/>

### 2.3 Initial Data Exploration

#### 2.3.1 Compare versus Previous Years Data

#### 2.3.2 Timeseries Plots

#### 2.3.3 Map

