---
title: "Clean Water (CW) - Trash (TRA) Subgoal Data Preparation"
output:
  github_document:
    toc: true
    toc_depth: 3
params:
    datasource: csv
always_allow_html: true
---

<br>
```{r tra prep preamble, message = FALSE}
source(here::here("R", "setup.R"))
loc <- here::here("prep", "CW", "trash")
dir_intermediate <- file.path(here::here("data", "CW", "trash", version_year, "intermediate"))

bkgd_path <- here::here("supplement", "goal_summaries", "TRA.Rmd")
data_path <- here::here("data", "CW", "trash", version_year, "tra_data.rmd")
refs_path <- file.path(loc, "tra_references.Rmd")
```

## 1. Background {-}

```{r Background, child = bkgd_path, results = "asis", echo = FALSE}
```

<br/>

## 2. Data {-}

```{r Data, child = data_path, results = "asis", echo = FALSE}
```

<br/>

## 3. Prep: Wrangling & Derivations, Checks/Evaluation {-}

### 3.1 Initial Data Wrangling/Harmonizing {-}

**Load datasets**
```{r load trash data, warning = FALSE, message = FALSE, results = "hide"}
# raw jambeck data
trash_file <- read_csv(file.path(dir_intermediate, 'trash_jambeck_baltic.csv'))

## BHI regional pop
bhi_pop <- read_csv(file.path(dir_intermediate, 'NUTS2_BHI_ID_Pop_density_in_buffer.csv')) %>% 
  dplyr::select(country = rgn_nam, 
                pop_rgn = PopTot) %>% 
  filter(!country == "Russia") %>% 
  group_by(country) %>% 
  summarize(bhi_pop = round(sum(pop_rgn), 0))

## lookup tables
lookup_bhi <- read.csv(file.path(here::here(), "supplement", "lookup_tabs", 'bhi_basin_country_lookup.csv'),
                                 sep=";", stringsAsFactors = FALSE) %>% 
  dplyr::select(rgn_id =BHI_ID, 
                country = rgn_nam) %>%
  mutate(country = as.character(country))

baltic_lookup <- read_csv(file.path(here::here(), "supplement", "lookup_tabs",'country_id.csv'))

## remove Total and footnotes (NAs in Country column)
data_clean <- trash_file %>%
  filter(!is.na(Country)) %>% #head(data_raw); summary(data_raw)
  dplyr::select(
    country = Country,
    coastal_pop = `Coastal population2`,
    modeled_waste_2010_orig = `Mismanaged plastic waste in 2010\r\n [tonnes]7`,
    modeled_waste_2025_orig = `Mismanaged plastic waste in 2025\r\n [tonnes]7`) %>%
  mutate(country = str_replace_all(country, '8', '')) # remove footnotes from country names

## filter Baltic data, excluding Russia, save intermediate
baltic_all <- data_clean %>%
  filter(country %in% baltic_lookup$country_name) %>% 
  filter(!country == "Russia")
#  write.csv(baltic, '~/github/bhi/baltic2015/prep/CW/trash/intermediate/trash_jambeck_baltic.csv')

```

<br/>

### 3.2 Downweight for Germany, Denmark, Sweden 

Amount of trash proportional of coastal pop/national pop

```{r downweight trash based on pop, warning = FALSE, message = FALSE, results = "hide"}
##combine coastal and bhi pop & downweight for Germany, Denmark, Sweden 
baltic_data <- full_join(bhi_pop, baltic_all, by = "country") %>% 
  mutate(proportion = bhi_pop/coastal_pop,
         ## only downweight for Germany, Denmark, and Sweden
         proportion = ifelse(country == "Germany" | country == "Denmark" | country == "Sweden", proportion, 1), 
         modeled_waste_2010 = modeled_waste_2010_orig * proportion, 
         modeled_waste_2025 = modeled_waste_2025_orig * proportion) %>% 
  dplyr::select(country, modeled_waste_2010, modeled_waste_2025)

# baltic_pop_proportion = full_join(bhi_pop, baltic_all, by = "country") %>%
#   mutate(proportion = bhi_pop/coastal_pop) %>%
#   dplyr::select(country, bhi_pop, coastal_pop, proportion)
```

Note: Because Russia is reported as a whole country, we used the proportion of population in Kaliningrad+St Petersburg to the coastal population reported by Jambeck et al. See figure below. 

#### 3.2.1 Downweight Russia Coastal Trash 

It made sense to downweight Kaliningrad and St Petersburg by coastal population density instead of total Russia population density because of the distribution along the coasts

<!-- Then find ref point as the max of Europe, but compare 2010 v. 2025 to determine whether it's worth trying to interpolate/model to find a reference point in ~2015. Conclusion, probably not worth it, so we could just pick either 2010 or 2025 since those data are already modeled/reported by Jambeck et al.  -->


```{r data explore and prep, warning = FALSE, message = FALSE, results = "hide"}
## havea another look at the data
# baltic = read.csv('~/github/bhi/baltic2015/prep/CW/trash/intermediate/trash_jambeck_baltic.csv')
# summary(baltic) 
# summary(data_clean)

# ###### NOT NEEDED? ###############
# ## read in lookup table for European countries
# lookup = read.csv('~/github/bhi/baltic2015/prep/EUcountrynames.csv', sep=';') %>%
#   mutate(country_abb = as.character(country_abb),
#          country     = as.character(country)); # head(lookup)
# 
# ## join lookup table to trash data; remove inland NAs and Kosovo, Macedonia, Serbia (no trash data)
# europe = lookup %>%
#   left_join(data_clean, by = 'country') %>%
#   filter(!is.na(coastal_pop)) %>%
#   dplyr::select(-country_abb); # head(europe)

## downweight russia **very hacky for now**
#  4597600 # St Petersbergy: from Wikipedia https://en.wikipedia.org/wiki/Demographics_of_Saint_Petersburg
#  431902  # Kaliningrad https://en.wikipedia.org/wiki/Kaliningrad
rus_trash_pop <- data_clean %>%
  filter(country == 'Russia') %>%
  dplyr::select(coastal_pop)

rus_coastal_pop_proportion <- as.numeric((4597600+431902) / rus_trash_pop); rus_coastal_pop_proportion

rus <- data_clean %>%
  filter(country == 'Russia') %>%
  mutate(modeled_waste_2010 = modeled_waste_2010_orig * rus_coastal_pop_proportion, 
         modeled_waste_2025 = modeled_waste_2025_orig * rus_coastal_pop_proportion) %>% 
  dplyr::select(country, modeled_waste_2010, modeled_waste_2025)

## create dataframe of europe + russia
eur_rus <- rbind(
  baltic_data,
  rus) 

```

### 3.3 Calculate Mismanaged plastic waste 2015

Assuming it's a linear trend between 2010 and 2025, 2015 values are the 1/3-point between 2010 and 2025. 
```{r calculate waste 2015, echo = FALSE, message = FALSE, warning = FALSE}
modeled_waste <- eur_rus %>% 
  dplyr::select(country, modeled_waste_2010, modeled_waste_2025) %>% 
  mutate(modeled_waste_2015 = modeled_waste_2010 + 1/3 * (modeled_waste_2025 - modeled_waste_2010)) 

```

#### 3.3.1 Plot Modeled Mismanaged plastic waste 2010, 2015, and 2025

```{r plot modeled mismanaged plastic waste, results = "show", fig.width = 9.5, fig.height = 4}
modeled_waste_plot <- modeled_waste %>% 
  tidyr::gather(scenario, waste, 2:4) 

ggplot(modeled_waste_plot, aes(x = country, y = waste, fill = scenario)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
  labs(title = 'Modeled waste 2010, 2015, & 2025',
      x = 'Country', 
      y = 'Waste (tonns)')

## SAVE 2015 DATA for VISUALIZE
tra_value_data <- modeled_waste%>%
  dplyr::select(value=modeled_waste_2015,
                location =country)%>%
  mutate(unit= "tonnes",
         bhi_goal = "TRA",
         data_descrip="modeled mismanaged plastic waste 2015" )

# write.csv(tra_value_data, file.path(dir_baltic,'visualize/tra_value_data.csv'), row.names=TRUE)

```


### 3.4 Prepare data layer for Toolbox 

**Plan:** 
We will create scores for each country for 2015 related to the reference point, and then apply to all of that country's BHI regions under the assumption that the management for the country is probably equal. 

Modeled Mismanaged plastic waste data (tonnes) are reported for each country. We will make a linear model between 2010 and 2025 and choose 2015.  

Note: for BHI 2.0, would be better to recalculate their model using coastal population for each BHI region.

**Reference points**
Need a min and a max reference point (they will ultimately be inverted for CW but not as a pressure)

- **Minimum modeled trash entering the ocean = 0** (we want no trash entering the ocean)
- **Maximum modeled trash entering the ocean = highest BHI country in 2010** 

#### 3.4.1 find ref point
```{r find ref and 2015, warning = FALSE, message = FALSE, results = "hide"}
## set min and max reference points ----
## rescale to reference point -- whole European
ref_point_min = 0

## 2010 ref point
ref_point_max_2010 <- eur_rus %>%
 filter(modeled_waste_2010 == max(modeled_waste_2010))

sprintf('max trash for 2010 is %s (%s)', 
        round(ref_point_max_2010$modeled_waste_2010),
        ref_point_max_2010$country)

ref_point_max_2010 <- ref_point_max_2010$modeled_waste_2010
# "max trash for 2010 is 37566 (Russia)"

```

<!-- ### 3.5 Plot exploring ref point as max of Europe (including Russia) in 2010 v. 2025 -->

<!-- ```{r plot and save layer} -->

<!-- # compare reference points - makes minimal differences in scores, country order does not change.  Larger differences for Germany & Russia (lesser degree Poland) -->

<!-- baltic_explore2 = baltic_explore %>%  -->
<!--   dplyr::select(country, trash_score_2010, trash_score_2025) %>%  -->
<!--   gather(key = ref_point, value = score, 2:3) %>%  -->
<!--   mutate(status_score = (1-score)*100) -->

<!-- ggplot(baltic_explore2, aes(x = country, y = status_score, fill = ref_point)) + -->
<!--   geom_point(aes(color = ref_point)) +  -->
<!--   theme(axis.text.x = element_text(angle = 75, hjust = 1)) + -->
<!--   labs(title = "Trash scores comparison: \n max(2010) vs. max(2025) as reference point") -->

<!-- ## assign to BHI regions ---- -->
<!-- lookup_bhi = read.csv('~/github/bhi/baltic2015/prep/CW/trash/bhi_basin_country_lookup.csv',sep=";", stringsAsFactors = FALSE) %>% -->
<!--   dplyr::select(rgn_id =BHI_ID, -->
<!--          country = rgn_nam) %>% -->
<!--   mutate(country = as.character(country)) -->

<!-- ## join baltic_explore (baltic countries) to bhi regions, repeating the trash score for each baltic region. -->
<!-- baltic_layer = lookup_bhi %>% -->
<!--   left_join(baltic_explore, -->
<!--             by = 'country') %>% -->
<!--   dplyr::select(rgn_id, -->
<!--          score = trash_score_2010); # baltic_layer -->

<!-- ## save layer to layers folder, and register layer in layers.csv by hand -->
<!-- # write.csv(baltic_layer, '~/github/bhi/baltic2015/layers/po_trash_bhi2015.csv',row.names=FALSE) -->

<!-- ## calculate trend; maybe make a linear model between 2010 and 2025, take trend as 5 years out? -->

<!-- ``` -->

### 3.6 Explore status score 2010 and 2015 

**Status**: 2015 modeled values / ref point 2010  
**Trend**: 5 year liear trend 2010-2015

2015: Using modeled 2015 data for status but use max 2010 as a reference point. This approach was used as the final candidate for status calculation. 

```{r explore tra status 2015, results = "show", fig.width = 9.5, fig.height = 4}
## scores by country
scores_by_country <- modeled_waste %>% 
  mutate(ref_point = 37566,  # "max trash for 2010 is 37566 (Russia)"
         score_2010 = (1 - pmin(1, modeled_waste_2010/ref_point))*100, 
         score_2015 = (1 - pmin(1, modeled_waste_2015/ref_point))*100 ) %>% 
  dplyr::select(country, score_2010, score_2015)

scores_by_country_plot <- scores_by_country %>% 
  tidyr::gather(year, score, 2:3)

# plot by country
ggplot(scores_by_country_plot, aes(x = country, y = score, fill = year)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) +
  ggtitle("Trash scores by country: 2010 & 2015")

## scores by bhi region
## join baltic_explore (baltic countries) to bhi regions, repeating the trash score for each baltic region.
scores_by_bhi <- lookup_bhi %>%
  left_join(scores_by_country, 
            by = 'country') %>%
  dplyr::select(-country) 

scores_by_bhi_plot <- scores_by_bhi %>% 
  tidyr::gather(year, score, 2:3)
  
ggplot(scores_by_bhi_plot, aes(x = rgn_id, y = score, fill = year)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) +
  ggtitle("Trash scores by BHI region: 2010 & 2015")

  
# save as CW-TRA data layer
tra_status <- scores_by_bhi %>% 
  dplyr::select(rgn_id, score = score_2015)

# write_csv(tra_status, file.path(dir_layers, 'cw_tra_status_bhi2015.csv'))

```

### 3.7 Trend calculation

Based on linear model of status scores in 2010 and 2015.  
```{r tra trend calculation, warning = FALSE, message = FALSE, results = "hide"}
tra_trend <- scores_by_bhi %>% 
  tidyr::gather(year, score, 2:3) %>% 
  mutate(year = substr(year, 7, 10)) %>% 
  group_by(rgn_id) %>% 
  summarize(trend_score = round((score[year == 2015] - score[year == 2010])/5 * 0.05, 2)) %>% 
  # NOT used: could not select coefficient of year from linear model 
  # do(lmd = lm(score ~ year, data = .)) %>% 
  # mutate(trend_score = coef(lmd)[['year']])
  dplyr::select(rgn_id, score = trend_score)

# write_csv(tra_trend, file.path(dir_layers, 'cw_tra_trend_bhi2015.csv'))

```


## 4. Save as Pollution Pressures layer

```{r save as pressures layer, warning = FALSE, message = FALSE, results = "hide"}
press_scores <- tra_status %>% 
  mutate(pressure_score = 1 - score/100) %>% 
  dplyr::select(rgn_id, pressure_score)

# write_csv(press_scores, file.path(dir_layers, "po_trash_bhi2015.csv"))

```


## 5. References {-}

```{r References, child = refs_path, results = "asis", echo = FALSE}
```

