---
output:
  word_document: default
  html_document: default
---

```{r lsp preamble, message = FALSE, warning = FALSE, include = FALSE}
loc <- here::here("prep", "LSP")

source(here::here("R", "setup.R"))
knitr::opts_chunk$set(message = FALSE, warning = FALSE, results = "hide", fig.width = 9.5)

bkgd_path <- here::here("supplement", "goal_summaries", "LSP.Rmd")
data_path <- here::here("data", "LSP", version_year, "lsp_data.rmd")
refs_path <- file.path(loc, "lsp_references.Rmd")

dir_B <- file.path(dirname(dir_prep), "bhi-data", "BHI 2.0")
dir_rawdata <- file.path(dir_B, "Goals", "LSP")
dir_intermediate <- file.path(here::here("data", "LSP", version_year, "intermediate"))

fullpal <- c(
  RColorBrewer::brewer.pal(8, "Dark2"),
  RColorBrewer::brewer.pal(9, "Set1")
)
```

```{r lsp setup spatial functions and files, warning = FALSE, include = FALSE}
## in order to assign bhi regions to data, must run this function first, to load required shapefiles:
source(here::here("R", "spatial.R"))
regions_shape(sp_dir = file.path(dirname(dir_B), "Shapefiles"))

basin_lookup <- file.path(dir_prep, "supplement", "lookup_tabs", "rgns_complete.csv") %>% 
  read_csv() %>%
  dplyr::select(BHI_ID = region_id, Subbasin = subbasin, rgn_nam = region_name, 
                bhi_area_km2 = region_area_km2) %>%
  mutate(
    Subbasin = as.character(Subbasin),
    rgn_nam = as.character(rgn_nam)
  )
```


## 1. Background

```{r lsp prep background, child = bkgd_path, results = "asis", echo = FALSE}
```

<br/>

## 2. Data {-}

This prep document is used to generate and explore the following data layers:

- `lsp_status_bhi2019.csv` 
- `lsp_trend_bhi2019.csv` 

These are saved to the `layers/v2019` folder. Saved to `data/LSP/v2019/intermediate` are intermediate datasets: `mpa_bhi_data.csv`, `bhi_buffer.shp` and `bhi_transf.shp`. All these are derived from or informed by the raw datasets from HELCOM database.

<br>

```{r lsp prep data, child = data_path, results = "asis", echo = FALSE}
```

<br/>

## 3. Prep: Wrangling & Derivations, Checks/Evaluation, Gapfilling {-}


```{r load lsp datasets, echo = TRUE}
management_plans <- read.delim(
  file.path(dir_rawdata, "management_plans.csv"), 
  sep = ",",
  stringsAsFactors = FALSE)

report_sites <- read.delim(
  file.path(dir_rawdata, "report_sites.csv"), 
  sep = ";",
  stringsAsFactors = FALSE)

## previously merged in 'data/LSP/v2019/lsp_data.R' script
mpa_bhi <- sf::st_read(file.path(
  dir_intermediate, 
  "mpa_bhi.shp"
))

## previously merged in 'data/LSP/v2019/lsp_data.R' script
mpa_bhi_data <- read_csv(file.path(dirname(data_path), "intermediate", "mpa_bhi_data.csv"))
```

<br/>

In BHI 1.0, MPA regions were divided by with BHI region shapefile, to be able to calculate the total MPA area within each BHI region. (MPA area per region was saved in the prep folder _bhi-1.0-archive/baltic2015/prep/LSP/mpa_area_per_rgn.csv_ and includes information on _Area per MPA, Date established, MPA status, total MPA area per region_).

However, in April 2016, changes had been made on raw datasets, in particular the following attributes were removed: "Location_S", "AreaTot_Ha", "AreaMar_Ha", "N2K_Site", changed following attributes names: "BSPA_ID" to "MPA_ID", "BSAPLink" to "Site_link" and added attribute "Date_est". Therefore, the attribute "MPA.size.km.sup2" from the raw _report_sites_ dataset will be used as total MPA area.

### 3.1 MPA management status  {-}

Joining information from the .csv file to the shapefile data.  

``` {r combine management status with mpa area and establishment info, echo = FALSE, message=FALSE}
mgmt_status <- report_sites %>%
  mutate(status = str_replace_all(Status, " ", "_")) %>%
  dplyr::select(name_csv = HELCOM.MPA.name,
                MPA_ID = ID,
                status_csv = status,
                mpa_area_km2 = MPA.size..km.sup2..) 

# count number of unique mpa names in new mpa mgmt status csv with a total of 175 observations
mpa_name <- dplyr::select(mgmt_status, name_csv)
unique_mpa_csv_new <- base::unique(mpa_name) ## total of 177 names that are unique

# combine mpa_status with shapefile data
mpa_mgmt <- right_join(mgmt_status, mpa_bhi,  
                       by = "MPA_ID") %>%
  dplyr::rename(name_shape = Name, 
                country = rgn_nam,
                status_shape = MPA_status,
                date_est = Date_est,
                eez_area_km2 = Area_km2) %>% 
mutate(mgmt_status = ifelse(is.na(status_csv), as.character(status_shape), status_csv)) 
       # Mgmt_status = str_replace_all(mgmt_status, Managed, Designated_and_partially_managed)
             
# count number of MPAs in shape file with no matches in csv file. 
# num_mismatch_new <- dplyr::count(mpa_mgmt, !status_csv== "NA") 
# 1 rows (1 MPA) have no match from the status csv file. 
````
<br/>

### 3.2 Explore map data {-}

```{r explore mpa data in map, eval = FALSE, results = "show", fig.width = 9, fig.height = 6}
library(leaflet)
source(here::here("R", "spatial.R"))
regions_shape()

## Make basemap
basemap <- ggplot2::ggplot(rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")) + 
  geom_sf(size = 0.1, color = "burlywood", alpha = 0.4) +
  scale_x_continuous(limit = c(4, 32)) +
  scale_y_continuous(limit = c(53.5, 66)) +
  theme(panel.background = element_rect(fill = "#F8FBFC", color = "#E2EEF3"))

# bhi_rgns_simple <- rmapshaper::ms_simplify(input = BHI_rgns_shp) %>% sf::st_as_sf()

## previously merged in 'data/LSP/v2019/lsp_data.R' script
bhi_buff <- sf::st_read(file.path(
  dir_intermediate, 
  "bhi_buffer.shp"
))

## previously merged in 'data/LSP/v2019/lsp_data.R' script
bhi_transf <- sf::st_read(file.path(
  dir_intermediate, 
  "bhi_transf.shp"
))

mpa_plot <- basemap +
  geom_sf(data = bhi_transf, fill = "white", size = 0.2, color = "burlywood") +
  geom_sf(
    mapping = aes(colour = MPA_status),
    data = sf::st_as_sf(
      mpa_bhi, 
      coords = c("lon", "lat"), 
      crs = 3035
    ),
    size = 1.8
  ) +
  labs(x = NULL, y = NULL, colour = "MPA Status") +
  ggtitle("MPA Status in the Baltic Sea")
# + theme(legend.position = c(0.9, 0.1)) +
#  scale_color_manual(values = c("tomato", "aquamarine2", "gold"))

```
<br/>

## 4. Status calculation {-}

Following a similar method as for the BHI 1.0, the model assesses the area of MPAs in each *bhi region* in relation to its EEZs, and their management status. 

Status is calculated by *bhi region* (which is also associated with that respective country).

STEPS:

1) Calculate the reference point: 10% total eez per BHI region (instead of per country, in order to have a fair representation of both countries and their respective basins)

2) Calculate cumulative mpa_area * weight, per BHI region per year [cumulative_sum (weight * mpa_area_km2)]

3) Calculate status by BHI region per year [cumulative_sum (weight * mpa_area_km2) / ref_point]

4) Calculate status by BHI region

_Xlsp_country = sum(w_i * MPA area)_m / Reference_pt_bhiregions_  

- Numerator is the sum over all MPAs within a bhi region's EEZ of the MPA area weighted by the management status.  
- w_i = value between 0 -1   

Weight (w_i) is based upon management status, which is broken down to three categories and weighted on a 0-1 scale:

0.1 = designated
0.4 = designated and partly managed
1.0 = designated and managed

**Reference_pt_bhiregion** = 10% of the area in a bhi region's EEZ is designated as an MPA and is 100% managed = 10% area bhi region's EEZ  


### 4.1 Calculate reference point {-}

```{r status calculation, ECHO = FALSE, message = FALSE}
mgmt_weight <- data.frame(mgmt_status = c(
  "Designated", "Designated_and_partly_managed", "Designated_and_managed"),
  weight = c(0.1, 0.4, 1))
                          
mpa_mgmt_with_wt = full_join(mgmt_weight, mpa_mgmt, 
                              by = 'mgmt_status') %>% 
  mutate(rgn_nam = paste(Subbasin,",", country))

### Calculate the reference point: 10% total eez per bhi region 
eez_data <- mpa_mgmt_with_wt %>%
  dplyr::select(MPA_ID, BHI_ID, country, eez_area_km2) %>%
#  filter(!duplicated(MPA_ID)) %>% #remove duplicated rows of the same MPA (but also MPA shared by two bhi_rgn)
  filter(!duplicated(BHI_ID)) %>%
  group_by(BHI_ID) %>%
  summarize(total_eez_km2 = sum(eez_area_km2)) %>% 
  ungroup() %>%
  mutate(ref_eez_km2 = 0.1 * total_eez_km2) %>%
  dplyr::select( -total_eez_km2)

### Calculate cumulative mpa_area*weight, per bhi region per year:
### cumulative_sum (weight * mpa_area_km2)
cum_mpa_area_wt <- mpa_mgmt_with_wt %>%
  dplyr::select(MPA_ID, BHI_ID, weight, country, rgn_nam, mpa_area_km2, year = Year_est) %>%
  filter(!is.na(year)) %>% 
  group_by(BHI_ID, year) %>%
  dplyr::summarize(sum_mpa_wt = sum(mpa_area_km2 * weight)) %>%
  arrange(year) %>%
  mutate(cum_mpa_wt = cumsum(sum_mpa_wt)) %>%
  ungroup()

### Calculate status per bhi region per year: 
### cumulative_sum (weight * mpa_area_km2) / ref_point
status_per_year_by_bhiregion <- full_join(
  cum_mpa_area_wt, eez_data, by = 'BHI_ID') %>% 
  group_by(BHI_ID, year) %>%
  summarize(status = round(max(-1, pmin(1, cum_mpa_wt/ref_eez_km2)) * 100, 1)) %>%
  ungroup() %>%
  arrange(BHI_ID, year)

### Calculate status per bhi region
status_by_bhiregion <- status_per_year_by_bhiregion %>%
  group_by(rgn_id = BHI_ID) %>%
  filter(year == max(year)) %>% # there are also differences in MPA status among years
  dplyr::select(-year, -BHI_ID)

lsp_last_year <- status_per_year_by_bhiregion %>%
  group_by(BHI_ID)%>%
  summarise(last_year = last(year))
# bhi_rgn have different last year data

knitr::kable(
  lsp_last_year %>% 
    group_by(last_year) %>%
    summarize(`Num. BHI Regions with Year` = n(), rgn_nam = paste(rgn_nam, collapse = ", ")) %>% 
    rename(`Last Year ` = last_year))

r.status <- mpa_mgmt_with_wt %>%
  filter(!duplicated(BHI_ID)) %>%
  dplyr::select(rgn_id = BHI_ID, 
                country, rgn_nam) %>%
  full_join(status_by_bhiregion, 
            by = 'rgn_id') %>% 
  dplyr::select(rgn_id, 
                score = status) %>% 
  mutate(rgn_id = as.integer(rgn_id)) %>% 
  mutate(dimension= "status")

# lsp_status <- r.status_helcom %>%
#  full_join(r.status_bhi, 
#            by = 'rgn_id')

# Save status layer
write_csv(r.status, file.path(dir_layers, 'lsp_status_bhi2019.csv'))

```
<br/>


#### 4.1.1 Alternative calculation of sum area {-}

It seems like the management weighs quite heavily in the score that even if there are lots of MPAs on the map, they are not worth much unless they are at least partly managed. Therefore, apply the managemnet designation weighting before doing proportion of area. 

What the score would be before applying  the 'penalties' (the  different  management categories based on whether the MPA is managed)

```{r}
# Calculate the proportion
alt <- mpa_mgmt_with_wt %>%
  group_by(BHI_ID) %>%
  mutate(
    total_mpa_area_in_bhirgn = sum(mpa_area_km2),
    area_refpt = 0.1 * eez_area_km2
  ) %>% 
  ungroup() %>% 
  ## if the total MPA area exceeds the 10% mark,
  ## don't want sum of weights for weighted sum to exceed 1
  ## so need to use ratio of MPA area to Sum of MPA areas in given BHI region
  ## but want to penalize for not reaching 10% mark
  ## so use MPA area vs Reference Area (10%) in all other cases
  mutate(mpa_prop = ifelse(
    total_mpa_area_in_bhirgn >= area_refpt,
    mpa_area_km2/total_mpa_area_in_bhirgn,
    mpa_area_km2/area_refpt
  )) %>% 
  group_by(BHI_ID) %>% 
  mutate(
    ## check_area_ref_relationship should be less than 1 for BHI regions where 10% ref area as MPA is not achieved
    ## but should never exceed 1!
    ## (if you were to calc mpa_prop above without ifelse, i.e. mpa_prop = mpa_area/area_refpt, would get many cases above 1
    check_area_ref_relationship = sum(mpa_prop),
    status_score = sum(mpa_prop*weight))

### Calculate status per year by bhi region
status_per_year_by_bhiregion_alt <- alt %>% 
  group_by(BHI_ID, year = Year_est) %>%
  summarize(status = round(max(-1, pmin(1, status_score)) * 100, 1)) %>%
  ungroup() %>%
  arrange(BHI_ID, year)

### Calculate status per bhi region
status_by_bhiregion_alt <- status_per_year_by_bhiregion_alt %>%
  group_by(rgn_id = BHI_ID) %>%
  filter(year == max(year)) %>% # there are also differences in MPA status among years
  dplyr::select(-year, -BHI_ID)

lsp_last_year <- status_per_year_by_bhiregion_alt %>%
  group_by(BHI_ID)%>%
  summarise(last_year = last(year))
# bhi_rgn have different last year data

knitr::kable(
  lsp_last_year %>% 
    group_by(last_year) %>%
    summarize(`Num. BHI Regions with Year` = n(), rgn_nam = paste(rgn_nam, collapse = ", ")) %>% 
    rename(`Last Year ` = last_year))

r.status_alt <- alt %>%
  filter(!duplicated(BHI_ID)) %>%
  dplyr::select(rgn_id = BHI_ID, 
                country, rgn_nam) %>%
  full_join(status_by_bhiregion_alt, 
            by = 'rgn_id') %>% 
  dplyr::select(rgn_id, 
                score = status) %>% 
  mutate(dimension= "status",
         status = score/100)

# Save status layer
write_csv(r.status_alt, file.path(dir_layers, 'lsp_status_bhi2019.csv'))

```


#### 4.1.2 Plot alternative status {-}

```{r}
## plot the status
plot_status_alt <- r.status_alt %>% 
  full_join(.,basin_lookup, by = c("rgn_id"= "BHI_ID")) 

set.seed(2)
statuspal <- colorRampPalette(fullpal)(80)[sample(1:80, size = 18)]

ggplot(plot_status_alt) +
  geom_col(
    aes(rgn_nam, score, fill = Subbasin),
    position = position_dodge(), 
    color = "black", 
    alpha = 0.6,
    size = 0.2,
    show.legend = FALSE
  ) + 
  geom_text(aes(rgn_nam, score, label = score), color = "black", size = 3) +
  coord_flip() +
  labs(x = NULL, y = "Score") +
  theme_bw() +
  theme(axis.text.y = element_text(size = 9)) +
  scale_fill_manual(values = statuspal) +
  ggtitle("LSP Status by BHI Region")
```

#### 4.1.3 Plot status map {-}

```{r}
# Make a map with score and mpa status categories
basemap <- ggplot2::ggplot(rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")) +
  geom_sf(size = 0.1, color = "burlywood", alpha = 0.4) +
  theme(panel.background = element_rect(fill = "#F8FBFC", color = "#E2EEF3")) +
  scale_x_continuous(limit = c(4, 32)) +
  scale_y_continuous(limit = c(53.5, 66)) 

cols <- c("indianred", "coral", "goldenrod1", "khaki", "lightblue", "steelblue")

bhi_rgn <- st_read(file.path(dirname(dir_B), "Shapefiles", "BHI_shapefile"))

bhi_rgn_simple <- rmapshaper::ms_simplify(input = bhi_rgn) %>% 
  sf::st_as_sf() %>% 
  mutate(rgn_nam = stringr::str_replace(rgn_nam, ", ",  " , "))

MPA_shp <- rmapshaper::ms_simplify(input = mpa_bhi) %>% 
  sf::st_as_sf()

## try joining data before plotting to make sure join works
plotsf <- left_join(
   bhi_rgn_simple,
   r.status_alt,
   by = c("BHI_ID" = "rgn_id")
)

statusmap <- basemap +
  geom_sf(
    data = plotsf,
    mapping = aes(fill = score)
  ) +
  scale_fill_gradientn(
    colors = (cols),
    values = c(0, 0.1, 0.3, 0.65, 0.8, 1),
    limits = c(0, 100)
  )

statusmap + geom_sf(data = MPA_shp, mapping = aes(color = MPA_status), fill = NA, alpha = 0.4)
```


### 4.2 Plot status per country {-}

```{r plot status by bhi region, ECHO = FALSE, message=FALSE}
plot_status <- ggplot(status_by_bhiregion, aes(x = rgn_id, y = status)) +
 geom_bar(stat = 'identity') +
 theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
 labs(title = 'LSP status by BHI region',
      x = 'BHI Region', 
      y = 'LSP status')

print(plot_status)
```
<br/>

#### 4.2.3 Plot number of MPAs per bhi regions {-}

```{r plot number of MPAs per country, ECHO = FALSE, message=FALSE}
## plot the number of MPAs per bhi regions
num_mpa_per_bhi_rgn = mpa_mgmt %>%
  group_by(BHI_ID) %>%
  summarize(count = length(!duplicated(name_shape))) %>%
  ungroup %>% 
  full_join(basin_lookup, by = "BHI_ID")

mpa_per_bhi_rgn_plot <- ggplot(num_mpa_per_bhi_rgn, aes(x = rgn_nam, y = count)) +
 geom_bar(stat = 'identity') +
 geom_text(aes(label = sprintf('n = %s', count), y = count), 
           size = 3, 
           angle = 90, hjust = 0, color = 'grey30') +
 theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
 labs(title = 'Number of MPAs per BHI region',
      x = 'BHI regions', 
      y = 'Number of MPAs')

```

#### 4.2.4 Plot the number of MPAs per bhi region by management level {-}

```{r plot number of MPAs per country by mgmt levels, ECHO = FALSE, message = FALSE}

## plot number of MPAs per bhi rgn by mgmt levels 
mgmt_weight_alt <- data.frame(mgmt_status = c("Designated", "Designated_and_partly_managed", "Designated_and_managed"),
                              weight = c(0.1, 0.4, 1)) 

num_mpa_per_bhi_rgn_mgmt <- mpa_mgmt_with_wt %>%
  group_by(BHI_ID, weight) %>%
  summarize(count = length(!duplicated(name_shape))) %>%
  ungroup %>% 
  full_join(mgmt_weight_alt, by = "weight") %>% 
  full_join(basin_lookup, by = "BHI_ID")

num_bhi_rgn_mgmt_plot <- ggplot(num_mpa_per_bhi_rgn_mgmt, aes(x = rgn_nam, y = count, fill = mgmt_status)) +
 geom_bar(stat = 'identity') +
 geom_text(aes(label = sprintf('n = %s', count), y = count), 
           size = 3, 
           angle = 90, hjust = 0, color = 'grey30') +
 theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
 labs(title = 'Number of MPAs by BHI region and management level',
      x = 'BHI region', 
      y = 'Number of MPAs',
      fill = 'Management Level')

```

#### 4.2.5 Plot total MPA area vs. 10% of the EEZ 

In the plot below, the red line represents where total MPA area equals that of the reference point (ie. 10% EEZ). Any point above that line means that the country has exceeded the reference point of designated MPA area and therefore will have a status score of 100, and any point below the line will have a score lower than 100.  

```{r compare total MPA area with ref EEZ, ECHo = FALSE, message = FALSE}
mpa_vs_eez <- mpa_mgmt_with_wt %>%
  dplyr::select(MPA_ID, BHI_ID, country, mpa_area_km2, year = Year_est) %>%
#  filter(!is.na(year), 
#         !duplicated(MPA_ID)) %>%  # the same MPA could exist in different BHI_IDs 
  group_by(BHI_ID, year)  %>%
  summarize(total_mpa_per_year = sum(mpa_area_km2)) %>%
  ungroup() %>%
  group_by(BHI_ID) %>%
  mutate(total_mpa = cumsum(total_mpa_per_year)) %>%
  filter(year == max(year)) %>%
  full_join(dplyr::select(eez_data, BHI_ID, ref_eez_km2), 
            by = 'BHI_ID') %>%
  dplyr::select( -total_mpa_per_year, -year) %>% 
  full_join(basin_lookup, by = "BHI_ID")
  
mpa_vs_eez_plot <- ggplot(mpa_vs_eez, aes(x = ref_eez_km2, y = total_mpa)) +
  geom_point(aes(color = BHI_ID), size = 3) +
  geom_abline(slope = 1, intercept = 0, color = 'red') +
  geom_text(aes(label = BHI_ID), nudge_y = -300) +
  theme(legend.position = 'none') +
  coord_cartesian(xlim = c(0, 15000), ylim = c(0, 15000)) +
  labs(title = 'Total MPA vs. Reference EEZ area by BHI region',
       x = '10% EEZ (km2)', 
       y = 'Total MPA (km2)')

print( mpa_vs_eez_plot)

```


#### 4.2.6 Plot the area of MPA by management level vs. 10% EEZ

Even when a BHI region has many MPAs that are just designated but not managed, it can still achieve a score of 100, is it because the area of MPAs that are "managed" (ie. higher weight) is relatively larger compared to that of the "designated"" (ie. lower weight), and thus have a higher _weight*area_ score?  

The plot below compares the total MPA area of each management level with the reference point (10% EEZ). 
The black dots represent 10% EEZ. 

```{r MPA area vs management level}
area_vs_mgmt_lvl <- mpa_mgmt_with_wt %>%
  dplyr::select(MPA_ID, BHI_ID, country, mpa_area_km2, year = Year_est, mgmt_status) %>%
#  filter(!is.na(year), 
#         !duplicated(BSPA_ID)) %>%  # the same MPA could exist in different BHI_IDs
  group_by(BHI_ID, mgmt_status)  %>%
  summarize(total_mpa_area = sum(mpa_area_km2)) %>%
  full_join(eez_data, by = "BHI_ID") %>% 
  full_join(basin_lookup, by = "BHI_ID")

area_vs_mgmt_lvl_plot <- ggplot(area_vs_mgmt_lvl, aes(x = rgn_nam, y = total_mpa_area, fill = mgmt_status)) +
 geom_bar(stat = 'identity') +
 geom_point(aes(x = rgn_nam, y = ref_eez_km2, size = ref_eez_km2), col = "black", show.legend = NA) +
 theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
 labs(title = 'MPA area of each management level vs. 10% EEZ per BHI region',
      x = 'BHI region', 
      y = 'MPA area (km2)',
      fill = 'Management Level',
      size = '10% EEZ')

print(area_vs_mgmt_lvl_plot)

```


### 4.3  Plot final score objects {-}

**Plot Status**
```{r final score object plotted}
## plot the status
plot_status <- r.status %>% 
  full_join(.,basin_lookup, by = c("rgn_id"= "BHI_ID")) 

set.seed(2)
statuspal <- colorRampPalette(fullpal)(80)[sample(1:80, size = 18)]

ggplot(plot_status) +
  geom_col(
    aes(rgn_nam, score, fill = Subbasin),
    position = position_dodge(), 
    color = "black", 
    alpha = 0.6,
    size = 0.2,
    show.legend = FALSE
  ) + 
  geom_text(aes(rgn_nam, score, label = score), color = "black", size = 3) +
  coord_flip() +
  labs(x = NULL, y = "Score") +
  theme_bw() +
  theme(axis.text.y = element_text(size = 9)) +
  scale_fill_manual(values = statuspal) +
  ggtitle("LSP Status by BHI Region")

```
<br/>

This status score looks more similar to the previous assessment. 
Will use this for the trend calculation.


## 6. Trend calculation {-}

### 6.1 Linear trend of status scores {-}

A linear trend was drawn on the status scores of the past five years. 

```{r trend}
trend_data <- status_per_year_by_bhiregion %>% 
  filter(!is.na(year)) %>% # one russian MPA didn't have info on established year
  mutate(year = as.integer(year)) %>%
  tidyr::complete(year = full_seq(year, 1), nesting(BHI_ID)) %>%
  group_by(BHI_ID) %>%
  fill(status) %>%
  mutate(status = ifelse(is.na(status), 0, status)) %>%
  ungroup()

r.trend <- trend_data %>%
  filter(year > (max(year) - 5)) %>% # most recent 5 years of data (2013- 2018)
  group_by(BHI_ID) %>%
  do(dlm = lm(status ~ year, data = .)) %>%
  mutate(score = round(pmax(-1, pmin(1, coef(dlm)[['year']]* 5/100)), 1)) %>%
  dplyr::select(BHI_ID, score)%>% 
  full_join(basin_lookup, by = "BHI_ID")   

r.trend_by_bhi <- dplyr::select(mpa_mgmt_with_wt, BHI_ID, country) %>% 
  filter(!duplicated(BHI_ID)) %>% 
  full_join(r.trend,
            by = "BHI_ID") %>% 
  dplyr::select(BHI_ID, 
                score) %>% 
  mutate(dimension = "trend") 

# Save trend layer
write_csv(r.trend_by_bhi, file.path(dir_layers, 'lsp_trend_bhi2019.csv'))  

### plot r.status
trend_plot <- ggplot(r.trend, aes(x = rgn_nam, y = score)) +
 geom_bar(stat = 'identity') +
 theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
 labs(title = 'Trend score by BHI region based on status scores',
      x = 'BHI Region', 
      y = 'Trend score')

```


### 6.2 Plot final slope object {-}

```{r plot final code object}
plot_slope <- r.trend_by_bhi %>%
  full_join(., basin_lookup, by = "BHI_ID") %>% 
  rename(rgn_id = BHI_ID)

ggplot(plot_slope) +
  geom_hline(yintercept = 0) +
  geom_point(aes(rgn_id, score)) +
  ylim(-1, 1) +
  labs(y="Slope", x = "BHI regions") +
  ggtitle("LSP Slope")

ggplot(rbind(plot_status, plot_slope)) + 
  geom_col(aes(rgn_nam, score, fill = Subbasin), alpha = 0.5, show.legend = FALSE) +
  coord_flip() +
  facet_wrap(~dimension, scales = "free_x", nrow = 1) +
  labs(x = "BHI Region\n", y = "Score") +
  ggtitle("LSP Status and Trend Scores by Region") +
  scale_fill_manual(values = statuspal) +
  theme(axis.text.y = element_text(size = 7))
```


## 7. Considerations for `BHI3.0` {-}

<br>

## 8. References {-}

```{r References, child = refs_path, results = "asis", echo = FALSE}
```

