---
output:
  word_document: default
  html_document: default
---

```{r lsp data preamble, echo = FALSE, include = FALSE, error = FALSE}
source(here::here("R", "data.R"))
source(here::here("R", "spatial.R"))

## root location of the raw data
dir_A <- file.path(dirname(dir_prep), "bhi-data")
dir_B <- file.path(dirname(dir_prep), "bhi-data", "BHI 2.0")
dir_rawdata <- file.path(dir_B, "Goals", "LSP")
dir_intermediate <- file.path(here::here("data", "LSP", version_year, "intermediate"))
  
```

### 2.1 Datasets with Sources {-}
<br/>

#### 2.1.1 MPA Data {-}

**HELCOM MPAs**

This dataset contains borders of the HELCOM MPAs (former Baltic Sea Protected Areas (BSPAs). 
The dataset has been compiled from data submitted by HELCOM Contracting Parties. It includes the borders of designated HELCOM MPAs stored in the [HELCOM Marine Protected Areas database](http://mpas.helcom.fi/apex/f?p=103:1::::::).
The designation is based on the HELCOM Recommendation 15/5 (1994). 

<!-- dataset save location BHI_share/BHI 2.0/Goals/LSP/ -->
```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Biodiversity:", "Protected areas"), 
  c("Protected areas:", "HELCOM MPAs")
))
colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [HELCOM Map and Data Service](http://metadata.helcom.fi/geonetwork/srv/eng/catalog.search#/metadata/d27df8c0-de86-4d13-a06d-35a8f50b16fa) <br/> Downloaded 23 June 2020 by Andrea De Cervo") 
```

<br/>

**MPAs**

Data were downloaded from the *MPAs* tab on the HELCOM MPA database.  
These MPA status levels depend upon the existence of implemented management plans for the MPA areas.
Also, MPA status is based on self-reporting. If countries differ in their definitions of "managed" or inflate their MPA status, we cannot account for those biases in the data. 

<!-- dataset save location BHI_share/BHI 2.0/Goals/LSP/ -->
```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("MPAs", "Actions"), 
  c("Download", ".csv")
))
colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [HELCOM Marine Protected Areas database](http://mpas.helcom.fi/apex/f?p=103:5::::::) <br/> Downloaded 23 June 2020 by Andrea De Cervo") 
```

<br/>

**Management Plans**

The status of the management plans associated with each MPA were downloaded from HELCOM MPA database, under the *Management Plans* tab. There are three levels of management plan status that can be assigned to each MPA: 
*No plan*, *In development*, *Implemented*.  

A challenge is that each MPA can have multiple management plans associated with it. 
There is no limit to the number of plans not an ability to assess their relative importance. 
Different management plans for the same MPA can have different levels of implementation.  

<!-- dataset save location BHI_share/BHI 2.0/Goals/LSP/ -->
```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Management plans", "Actions"), 
  c("Download", ".csv")
))
colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [HELCOM Marine Protected Areas database](http://mpas.helcom.fi/apex/f?p=103:5::::::) <br/> Downloaded 23 June 2020 by Andrea De Cervo") 
```

<br/>

#### 2.1.2 Reference values {-}

The Convention on Biodiversity's target is for 10% of a country's EEZs is designated as MPAs, and are fully managed. For more information see [Target 11 - Technical Rationale](https://www.cbd.int/sp/targets/rationale/target-11/).

For the BHI 2.0, the designation of at least 10% of the BHI region area as MPAs with a full implemented management plan, in order to give a fair representation of spatial coverage to the country and its respective basin. 
Management status are broken down to three categories and weighted on a 0-1 scale:

0.1 = designated
0.4 = designated and partly managed
1.0 = designated and managed


<br>

### 2.2 Centralization & Normalization {-}

#### 2.2.1 MPAs datasets {-}

Get them in the same coordinate reference system for the Baltic. 
The MPA file is in the LAEA coordinate reference system, for details see [here](http://spatialreference.org/ref/epsg/etrs89-etrs-laea/). 

**Read in MPA and BHI regions shapefiles**
```{r read in shapefiles, ECHO = FALSE, message=FALSE}
library(rgdal)  # install.packages('rgdal')
library(raster) # install.packages('raster')
library(rgeos)  # install.packages('rgeos')
library(sp)
## inspect BHI regions ----
# bhi <- sf::st_read(file.path(dir_A, "Shapefiles", "BHI_shapefile", "BHI_shapefile.shp"))
bhi <- rgdal::readOGR(dsn = path.expand(file.path(dir_A, "Shapefiles", "BHI_shapefile")),
                      layer = 'BHI_shapefile') 
# bhi # coord. ref. : +proj=longlat
## explore/plot
# plot(bhi) #takes ~15 seconds
# bhi@data
head(bhi@data)

# helcom_mpas <- sf::st_read(file.path(dir_rawdata, "HELCOM_MPAs", "HELCOM_MPAs.shp"))
helcom_mpas <- rgdal::readOGR(dsn = path.expand(file.path(dir_rawdata, "HELCOM_MPAs")),
                      layer = 'HELCOM_MPAs')

# mpa # coord. ref. : +proj=laea

## plot
# plot(helcom_mpas) # quick
# head(helcom_mpas@data)

## transform bhi to mpa coordinate reference system (CRS) ----
## LAEA: http://spatialreference.org/ref/epsg/etrs89-etrs-laea/
## https://www.nceas.ucsb.edu/~frazier/RSpatialGuides/OverviewCoordinateReferenceSystems.pdf
bhi_transf <- spTransform(bhi, helcom_mpas@proj4string)
# bhi # coord. ref. : +proj=laea

## save intersected shapefile
writeOGR(bhi_transf, dsn = path.expand(dir_intermediate), layer = 'bhi_transf', driver = 'ESRI Shapefile', overwrite = TRUE) 

## plot on top of each other
plot(bhi_transf, col = 'light blue', border = "grey", main = "BHI regions and MPAs intercept"); plot(helcom_mpas, col = 'blue', add = TRUE); legend('bottomright', c("BHI regions", "BHI borders", "MPA"), lty = c(1,1), lwd = c(2.5, 2.5, 2.5), col = c( "light blue", "grey", "blue"), text.font = 1, box.lty = 0 )
```

#### 2.2.2 Intersect BHI and HELCOM_MPA polygons {-}

MPA regions were divided by with BHI region shapefile, and thus we were able to calculate the total MPA area within each BHI region. MPA area per region is saved in the prep folder (`mpa_area_per_rgn.csv`). 

The .csv file includes information: Area per MPA, Date established, MPA status, total MPA area per region.  

```{r intersect BHI and MPA, echo=FALSE, message=FALSE, cache=TRUE}
# bhi <- gSimplify(bhi, tol = 0.00001)
# helcom_mpas <- gSimplify(helcom_mpas, tol = 0.00001)

# Need to 'buffer' both bhi and mpa shapefiles before intersecting
bhi_buff <- gBuffer(bhi_transf, byid=TRUE, width=0)
helcom_mpas_buff <- gBuffer(helcom_mpas, byid=TRUE, width=0)

mpa_bhi_buff <- raster::intersect(helcom_mpas_buff, bhi_buff)
# system.time({mpa_bhi <- raster::intersect(helcom_mpas, bhi_transf)})
## system.time({mpa_bhi <- raster::intersect(mpa, bhi)}) # large file size and took hours to run. bhi ~ 45.5 Mb, mpa ~ 8.7 Mb.

## save intersected shapefile
writeOGR(mpa_bhi_buff, dsn = path.expand(dir_intermediate), layer = 'mpa_bhi', driver = 'ESRI Shapefile', overwrite = TRUE) 

plot(mpa_bhi, col ="red", border = "black")
head(mpa_bhi_buff@data)

# Clean mpa_bhi data and save it to the intermediate folder
mpa_bhi_data <- mpa_bhi_buff@data %>% 
  rename(name_shape = Name, 
         country = rgn_nam,
         status_shape = MPA_status,
         date_est = Date_est,
         eez_area_km2 = Area_km2)

## save mpa_bhi_data file to prep/LSP folder 
write_csv(mpa_bhi_data, file.path(dir_intermediate, 'mpa_bhi_data.csv'))

```

