
## 3. Pressure Model
Will create two separate pressure layers (PCB 153 and PCDDF) which can either be combined into a single pressure layer or only one can be used.  

### 3.1 Current conditions 
Value in most recent year. 
PCDDF = 2012  
PCB 153 = 2013  

### 3.2 Rescaling  
Min value = 0

Max value = maximum value across all basin 1990 - max year  




## 4. Data layer preparation

```{r setup}

## Libraries
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(RMySQL)
library(stringr)
library(tools)
library(rprojroot) # install.packages('rprojroot')

## source common libraries, directories, functions, etc
source('~/github/bhi/baltic2015/prep/common.r')

## rprojroot
root <- rprojroot::is_rstudio_project


## make_path() function to 
make_path <- function(...) rprojroot::find_root_file(..., criterion = is_rstudio_project)



dir_layers = make_path('baltic2015/layers') # replaces  file.path(dir_baltic, 'layers')


# root$find_file("README.md")
# 
# root$find_file("ao_need_gl2014.csv")
# 
# root <- find_root_file("install_ohicore.r", 
# 
# withr::with_dir(
#   root_file("DESCRIPTION"))



dir_atmos_con    = file.path(dir_prep,'pressures/atmos_contaminants')


## add a README.md to the prep directory with the rawgit.com url for viewing on GitHub
create_readme(dir_atmos_con, 'atmos_con_prep.rmd') 
```

### 4.1 Read in and organize data

#### 4.1.1 Read in data

```{r read in data}
pcb153 = read.csv(file.path( dir_atmos_con, "data_database/pcb153.csv"), stringsAsFactors = FALSE)
str(pcb153)

pcddf = read.csv(file.path( dir_atmos_con, "data_database/pcddf.csv"), stringsAsFactors = FALSE)
str(pcddf)

atmos_basin_lookup = read.csv(file.path( dir_atmos_con, "data_database/atmos_loading_basin_lookup.csv"), stringsAsFactors = FALSE)
str(atmos_basin_lookup)


bhi_lookup = read.csv(file.path( dir_atmos_con, "bhi_basin_country_lookup.csv"),sep=";", stringsAsFactors = FALSE) %>%
              select(BHI_ID, Subbasin)%>%
              dplyr::rename(rgn_id = BHI_ID,
                            holas_basin = Subbasin)
str(bhi_lookup)
```

#### 4.1.2 Transform data to long format
```{r transform to long format}
pcb153_long = pcb153 %>%
              gather(year, value, -basin_loading,-basin_abb,-data_type,-substance,-unit) %>%
              mutate(year = substr(year,2,5),
                     year = as.numeric(year),
                     value = as.numeric(value))
              

pcddf_long = pcddf %>%
              gather(year, value, -basin_loading,-basin_abb,-data_type,-substance,-unit) %>%
              mutate(year = substr(year,2,5),
                     year = as.numeric(year),
                     value = as.numeric(value))
              
```

#### 4.1.3 Join to HOLAS basin names  
Data are joined to HOLAS basin names because these are what are associated with BHI regions. This is a slightly indirect approach, could have made a look up table to associated the major basins from the deposition data directly with the BHI regions, but this should have the same result.  
```{r join to holas basin names}
pcb153_long1 = pcb153_long %>%
               full_join(., atmos_basin_lookup,by="basin_loading")

pcddf_long1 = pcddf_long %>%
               full_join(., atmos_basin_lookup,by="basin_loading")
```

#### 4.1.4 Plot data by Major basin provided
```{r plot raw data by major basin provided}
ggplot(pcb153_long1)+
  geom_point(aes(year,value))+
  facet_wrap(~basin_loading)+
  ylab("PCB 153 kg/year")+
  theme(axis.text.x = element_text(colour="grey20", size=8, angle=90, 
                                  hjust=.5, vjust=.5, face = "plain"),
          strip.text.x = element_text(size = 6))+
  ggtitle("PCB153 Atmospheric Deposition by Major Basin")

ggplot(pcddf_long1)+
  geom_point(aes(year,value,colour=data_type))+
  facet_wrap(~basin_loading)+
  ylab("PCDD/F g TEQ / year")+
  theme(axis.text.x = element_text(colour="grey20", size=8, angle=90, 
                                  hjust=.5, vjust=.5, face = "plain"),
          strip.text.x = element_text(size = 6))+
  ggtitle("PCDD/F Atmospheric Deposition by Major Basin")



```


#### 4.1.5 Plot data by HOLAS basin

```{r plot raw data by holas basin}
ggplot(pcb153_long1)+
  geom_point(aes(year,value))+
  facet_wrap(~holas_basin)+
  ylab("PCB 153 kg/year")+
  theme(axis.text.x = element_text(colour="grey20", size=8, angle=90, 
                                  hjust=.5, vjust=.5, face = "plain"),
          strip.text.x = element_text(size = 6))+
  ggtitle("PCB153 Atmospheric Deposition by HOLAS Basin")

ggplot(pcddf_long1)+
  geom_point(aes(year,value,colour=data_type))+
  facet_wrap(~holas_basin)+
  ylab("PCDD/F g TEQ / year")+
  theme(axis.text.x = element_text(colour="grey20", size=8, angle=90, 
                                  hjust=.5, vjust=.5, face = "plain"),
          strip.text.x = element_text(size = 6))+
  ggtitle("PCDD/F Atmospheric Deposition HOLAS Basin")



```

#### 4.1.6 Select PCDFF scenario data
```{r select pcddf scenario data}


pcddf_long2 = pcddf_long1 %>%
              filter(data_type == "scenario with adjusted emissions")
dim(pcddf_long1); dim(pcddf_long2)

```


### 4.2 Current value

#### 4.2.1 Get most recent year for current value
```{r get current year value}
max_yr_pcb153 = pcb153_long1 %>%
                select(year)%>%
                max()%>%
                as.numeric()



max_yr_pcddf = pcddf_long2 %>%
                select(year)%>%
                max()%>%
                as.numeric()
```

#### 4.2.2 Get current pressure data
```{r get current pressure data values}

current_pcb153 = pcb153_long1 %>%
                 filter(year == max_yr_pcb153) %>%
                 select(holas_basin, value, unit, year)



current_pcddf =  pcddf_long2%>%
                 filter(year == max_yr_pcddf) %>%
                 select(holas_basin, value, unit, year)
```

#### 4.2.3 Plot current year data
```{r plot current year data}
ggplot(current_pcb153)+
  geom_point(aes(holas_basin, value), size = 2.5)+
  ylab("PCB 153 kg/year")+
  ylim(0,8)+
   theme(axis.text.x = element_text(colour="grey20", size=8, angle=90, 
                                  hjust=.5, vjust=.5, face = "plain"),
          strip.text.x = element_text(size = 6))+
  ggtitle("PCB 153 Atmospheric Deposition in 2013")


ggplot(current_pcddf)+
geom_point(aes(holas_basin, value), size = 2.5)+
ylab("PCDD/F g TEQ / year")+
ylim(0,150)+
 theme(axis.text.x = element_text(colour="grey20", size=8, angle=90, 
                                hjust=.5, vjust=.5, face = "plain"),
        strip.text.x = element_text(size = 6))+
ggtitle("PCDD/F Atmospheric Deposition in 2012")



```


### 4.3 Rescale data  

#### 4.3.1 Maximum value
Select maximum value across all basins within each time series. 
PCB 153:  Baltic Proper in 1990, 21.1 kg/year  
PCDD/F: Baltic Proper in 1990, 320 g TEQ / year  
```{r select max values}
max_pcb153 = pcb153_long1 %>%
             select(value)%>%
             max()%>%
             as.numeric()
max_pcb153

## which year and basin 
pcb153_long1 %>% filter(value == max_pcb153) %>% select(basin_loading, year, unit,value)%>% distinct()


max_pcddf = pcddf_long2 %>%
             select(value)%>%
             max()%>%
             as.numeric()
max_pcddf

## which year and basin 
pcddf_long2 %>% filter(value == max_pcddf) %>% select(basin_loading, year, unit,value)%>% distinct()


```

#### 4.3.2 Minimum value
Minimum value is 0, the value at which there is no pressure 
```{r select min values}
min_pcb153 =0

min_pcddf = 0




```

#### 4.3.3 Join current, max, and min values
```{r join current max and min values}
pcb153_rescale = current_pcb153%>%
                 mutate(min = min_pcb153, 
                        max= max_pcb153) %>% 
                 dplyr:: rename(current = value)

pcddf_rescale = current_pcddf%>%
                 mutate(min = min_pcddf, 
                        max= max_pcddf)%>%
                  dplyr:: rename(current = value)


```

#### 4.3.4 Plot current, max and min
```{r plot current max and min}

ggplot(gather(pcb153_rescale, type, concentration, -holas_basin,-unit,-year))+
  geom_point(aes(holas_basin, concentration, colour = type, shape=type), size= 2.5)+
  ylab("PCB 153 kg/year")+
   theme(axis.text.x = element_text(colour="grey20", size=8, angle=90, 
                                hjust=.5, vjust=.5, face = "plain"))+
  ggtitle("PCB 153 minimum, maximum, and current pressure value")




ggplot(gather(pcddf_rescale, type, concentration, -holas_basin,-unit,-year))+
  geom_point(aes(holas_basin, concentration, colour = type, shape=type), size= 2.5)+
  ylab("PCDD/F g TEQ / year")+
   theme(axis.text.x = element_text(colour="grey20", size=8, angle=90, 
                                hjust=.5, vjust=.5, face = "plain"))+
  ggtitle("PCDD/F minimum, maximum, and current pressure value")

```


#### 4.3.5 Rescale data
```{r rescale data}

pcb153_rescale1 = pcb153_rescale %>%
                  mutate(pcb153_normalize = (current - min)/(max - min))



pcddf_rescale1 = pcddf_rescale %>%
                  mutate(pcddf_normalize = (current - min)/(max - min))
```

#### 4.3.6 Plot rescaled pressure layer
```{r plot rescales pressure layer}
ggplot(pcb153_rescale1)+
  geom_point(aes(holas_basin,pcb153_normalize), size= 2.5)+
  ylab("Pressure Value")+
  ylim(0,1)+
   theme(axis.text.x = element_text(colour="grey20", size=8, angle=90, 
                                hjust=.5, vjust=.5, face = "plain"))+
  ggtitle("PCB 153 pressure value")




ggplot(pcddf_rescale1)+
  geom_point(aes(holas_basin, pcddf_normalize), size= 2.5) +
 ylab("Pressure Value")+
  ylim(0,1)+
   theme(axis.text.x = element_text(colour="grey20", size=8, angle=90, 
                                hjust=.5, vjust=.5, face = "plain"))+
  ggtitle("PCDD/F Pressure value")



```

#### 4.3.7 Plot and compare two contaminants rescaled pressure layers  
**PCB153 is in the open red dots, and PCDDF is in the filled black dots**  
Higher pressure from PCDDF, but magnitude of difference varies among basins
```{r plot and compare pcb153 and pcddf}

plot_compare = full_join(
                select(pcb153_rescale1, holas_basin, pcb153_normalize),
                select(pcddf_rescale1, holas_basin, pcddf_normalize),
                by = "holas_basin")

ggplot(plot_compare)+
  geom_point(aes(holas_basin,pcb153_normalize), size= 2.5, colour = "red", shape = 1)+
  geom_point(aes(holas_basin, pcddf_normalize), size= 2.5, colour = "black") +
  ylab("Pressure Value")+
  ylim(0,1)+
   theme(axis.text.x = element_text(colour="grey20", size=8, angle=90, 
                                hjust=.5, vjust=.5, face = "plain"))+
  ggtitle("Compare PCB153 and PCDD/F Pressure values")



```

### 4.4 Pressure layer for BHI regions

#### 4.4.1 Apply basin values to BHI regions
```{r apply basin values to BHI regions}

pcb153_rgn = pcb153_rescale1 %>%
            full_join(., bhi_lookup, by = "holas_basin") %>%
            select(rgn_id, pcb153_normalize)%>%
            arrange(rgn_id)

pcddf_rgn = pcddf_rescale1 %>%
            full_join(., bhi_lookup, by = "holas_basin") %>%
            select(rgn_id, pcddf_normalize)%>%
            arrange(rgn_id)

```

#### 4.4.2 Plot Pressure layer by BHI regions
```{r Plot Pressure layer by BHI regions}
ggplot(pcb153_rgn)+
  geom_point(aes(rgn_id,pcb153_normalize), size= 2.5)+
  ylab("Pressure Value")+
  ylim(0,1)+
   theme(axis.text.x = element_text(colour="grey20", size=8, angle=90, 
                                hjust=.5, vjust=.5, face = "plain"))+
  ggtitle("PCB 153 pressure value by BHI regions")




ggplot(pcddf_rgn)+
  geom_point(aes(rgn_id, pcddf_normalize), size= 2.5) +
 ylab("Pressure Value")+
  ylim(0,1)+
   theme(axis.text.x = element_text(colour="grey20", size=8, angle=90, 
                                hjust=.5, vjust=.5, face = "plain"))+
  ggtitle("PCDD/F Pressure value by BHI regions")

```


### 4.5 Prepare and save object for layers  

#### 4.5.1 Prepare objects
```{r prepare object for layers}
po_atmos_pcb153 = pcb153_rgn %>%
                  dplyr::rename(pressure_score = pcb153_normalize)


po_atmos_pcddf = pcddf_rgn %>%
                  dplyr::rename(pressure_score = pcddf_normalize)
```

#### 4.5.2 Save objects to layers
```{r save objects to layers}
write.csv(po_atmos_pcb153,file.path(dir_layers,"po_atmos_pcb153_bhi2015.csv"), row.names = FALSE)

write.csv(po_atmos_pcddf,file.path(dir_layers,"po_atmos_pcddf_bhi2015.csv"), row.names = FALSE)
```

