#Preparation of Munitions pressure data layer 

These data need to be explored before can be determined to be a useful pressure layer.  

```{r setup}

## Libraries
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(RMySQL)
library(stringr)
library(tools)
library(rprojroot) # install.packages('rprojroot')

## root location of the raw data
source(here::here("R", "setup.R"))
source(here::here("R", "spatial.R"))

dir_A <- file.path(dirname(dir_prep), "bhi-data")
dir_B <- file.path(dirname(dir_prep), "bhi-data", "BHI 2.0")
dir_rawdata <- file.path(dir_B, "Pressure", "munitions")
dir_intermediate <- file.path(here::here("data", "pressures", "bottom_trawling", version_year, "intermediate"))

bhi_basin_lookup <- read_delim(file.path(dir_prep, "supplement", "lookup_tabs", "bhi_basin_country_lookup.csv"), delim = ";") %>% 
  rename(subbasin = Subbasin) 
```

## 1. Background
Legacy of munitions dumping in the Baltic Sea.

## 2. Data

###2.1 Data source

Multiple shapefiles from HELCOM 

<!-- dataset save location BHI_share/BHI 2.0/Pressure/munitions/ -->
```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Pressures and Human Activities:", "Dumped chemical munitions")
))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [HELCOM Map and Data Service](http://maps.helcom.fi/website/mapservice/) downloaded on 26 August 2020 by Andrea De Cervo <br/>") 

```


### 2.2 Data layers

**Different data layers hae different formats: areas (polygons), points, tracks.**
Unclear how best to combine.  

#### 2.2.1 Reported encounters with chemical warfare materials 1961 to 2012 

[This dataset](http://metadata.helcom.fi/geonetwork/srv/eng/catalog.search#/metadata/e54e0cc7-c646-4b82-87bf-9a1132712ae7) contains information about reported encounters of sea dumped chemical munitions in the Baltic Sea from 1961 to 2012. For further details see Annex 7.2 of the HELCOM 2013 report: Chemical Munitions Dumped in the Baltic Sea - Report of the ad hoc Expert Group to Update and Review the Existing Information on Dumped Chemical Munitions in the Baltic Sea (HELCOM MUNI). This dataset includes incidents of encounters with chemical warfare agents dumped in the Baltic Sea. The position of encounters is given in decimal degrees (WGS84), where the information is available (encounters lacking this geographic information are also included in the dataset, although they lack coordinates). 

The weight of the warfare agent payload (net weight) was estimated by visual inspection. Disposal on land or the relocation to emergency disposal sites due to potential explosive hazards is indicated, where the information is available. 

Data sources include incidents reported by HELCOM countries since the establishment of the HELCOM reporting system in 1994, fishing and shoring incidents with chemical warfare materials in Poland(since 1952-1983) and Denmark (1968-1984); the 1991 SACSA report and HELCOM 1994.

Further, information about encounters of chemical warfare material by the Nord Stream pipeline project have been included in the dataast.

For more details, see the [HELCOM report](http://www.helcom.fi/Documents/Ministerial2013/Associated%20documents/Background/Dumped%20chemical%20munitions%20in%20the%20Baltic%20Sea.pdf) on Dumped chemical munitions in the Baltic Sea (HELCOM 2013).

References:

HELCOM (1994): Report on chemical munitions dumped in the Baltic Sea –report to the 16th Meeting of Helsinki Commission, 8- 11 March 1994, from the Ad Hoc Working Group on Dumped Chemical Munition (HELCOM CHEMU). Danish Environmental Protection Agency (1994).

SACSA (1991): Dumping of war gas ammunition –documentation of the Helsinki Commission. Document SACSA 18/8/1-E, presented at the 18thmeeting of the Standing Advisory Committee for Scientific Advicetranscribed from SACSA 1991.

#### 2.2.2 Emergency relocation areas for netted sea dumped chemical warfare material

[This dataset](http://metadata.helcom.fi/geonetwork/srv/eng/catalog.search#/metadata/353c558c-08f2-441a-9a5e-89e81cf653ac) depicts two areas in the Baltic Sea that have been assigned for the emergency disposal of netted warfare materials too unsafe to be brought and handled ashore. Two relocation areas have been assigned by the Bornholm-located Danish Navy Maritime Surveillance Centre South in the vicinity of the Bornholm dumpsite for the emergency disposal of netted warfare materials too unsafe to be brought and handled ashore. The two areas with a diameter of 0.5 nautical miles and an area of about 270 ha each are located at 55°4.7’N, 15°14.2’E and 55°2.4’N, 14°37.9’E (WGS84). 

For more information, see: HELCOM 2013. Chemical Munitions Dumped in the Baltic Sea. Report of the ad hoc Expert Group to Update and Review the Existing Information on Dumped Chemical Munitions in the Baltic Sea (HELCOM MUNI)

#### 2.2.3 Chemical weapons dumpsites in the Baltic Sea

Areas that were formerly chosed and designated for dumping munitions at sea. 
[The dataset](http://metadata.helcom.fi/geonetwork/srv/eng/catalog.search#/metadata/b55b508d-3c40-484c-8c7f-38869a8df368) was made based on information compiled by the HELCOM ad hoc Expert Group to Update and Review the Existing Information on Dumped Chemical Munitions in the Baltic Sea (HELCOM MUNI). 

The areas displayed indicate areas formerly chosen and designated for dumping munitions at sea, e.g, by instructing the ships' captains to dump in sight of a landmark or moored buoy. In addition the dataset contains information about a suspected dumpsite in the Gdansk deep. 

It should also be noted that dumping of munitions was also occasionally conducted outside of the known and designated dumping areas.

The use of these terms is not necessarily reflected in the descriptions contained in sea charts. While the area south of Little Belt was used as a dumping area, for example, in contemporary sea charts it is only described as ‘foul (explosives)’ and not as ‘explosives dumping ground'.

#### 2.2.4 Chemical munition transport routes to dumpsites

Designated and suspected alternative routes along which chemical munitions were transported to the designated dumping areas in the Baltic Sea after World War II. 
[This dataset](http://metadata.helcom.fi/geonetwork/srv/eng/catalog.search#/metadata/5253b733-74de-4ed6-9151-720468e0d5f9) illustrates the designated and suspected alternative routes along which chemical munitions were transported to the designated dumping areas in the Baltic Sea after World War II. 

#### 2.2.5 Suspected en route dumping areas  

Areas where en-route dumping of chemical munitions is suspected to have occured while they were being transported to the designated dumpsite south of Little Belt. The southern entrance to the Little Belt is an area that was used by German troops and refugees for munitions disposal during the last days of WWII. 
[This dataset](http://metadata.helcom.fi/geonetwork/srv/eng/catalog.search#/metadata/c59001a4-97c3-4dc5-8e5e-5ee03d8438b0) illustrates that areas where, according to historical accounts, en-route dumping of chemical munitions is likely to have taken place on the way to the official designated dumping area. 

#### 2.2.6 Areas where sea dumped chemical warfare materials have been encountered

[This dataset](http://metadata.helcom.fi/geonetwork/srv/eng/catalog.search#/metadata/5724455d-4589-4b00-b255-c1989742a4ed) gives an overview of the main areas in the Baltic Sea where sea dumped chemical warfare materials have been reported to be found. This dataset gives an overview of the main areas in the Baltic Sea where sea dumped chemical warfare materials have been found. The areas are drawn based on reported encounters where fishermenhave netted chemical warfare materials. 

For more information, see the HELCOM 2013 MUNI report on dumped chemical munitions in the Baltic Sea. 

#### 2.2.7 Mines sunk in the World War II - Risk areas

[This dataset](http://metadata.helcom.fi/geonetwork/srv/eng/catalog.search#/metadata/d424a749-6dba-4c54-89b1-abbfc3c5be53) shows sea bottom risk areas for mines sunk in the World War II. The big areas in Danish and German areas as well as in the Gdansk Bay are British flight mine areas.

This dataset was created by the HELCOM Expert Group on Environmental Risks of Hazardous Submerged Objects (SUBMERGED). SUBMERGED works to compile and assess information about all kinds of hazardous objects and assess the associated risks. The dataset was provided by Gunnar Möller (Mine Warfare Data Center (C MWDC), 4th Naval Warfare Flottilla, Berga, Sweden) for the HELCOM Maritime Assessment published in 2018.

### 2.3 Data exploration Needed
Need to plot evaluate information associated with each data layer available.  

## 3. Pressure model
Option: total area in in BHI region associated with one or more munitions impact.

### 3. 1 Current conditions
Percent of BHI area with one or more munitions impact.  

### 3.2 Rescaling between 0 to 1

min value = 0

max value = BHI region with greatest percent munitions impact  


## Layer prep

```{r TODO, echo=FALSE, message=FALSE}
## read in data...

## plot data...

```
