---
title: "Carbon Storage (CS) Goal Data Preparation"
output:
  github_document:
    toc: true
    toc_depth: 3
params:
    datasource: csv
---

<br>

```{r cs preamble, message = FALSE, warning = FALSE, include = FALSE}
loc <- here::here("prep", "CS")

source(here::here("R", "setup.R"))
knitr::opts_chunk$set(message = FALSE, warning = FALSE, results = "hide", fig.width = 9.5)

bkgd_path <- here::here("supplement", "goal_summaries", "CS.Rmd")
data_path <- here::here("data", "CS", version_year, "cs_data.rmd")
refs_path <- file.path(loc, "cs_references.Rmd")

dir_B <- file.path(dirname(dir_prep), "bhi-data", "BHI 2.0")
dir_intermediate <- file.path(here::here("data", "CS", version_year, "intermediate"))
dir_rawdata <- file.path(dir_B, "Goals", "CS")

fullpal <- c(
  RColorBrewer::brewer.pal(8, "Dark2"),
  RColorBrewer::brewer.pal(9, "Set1")
)
```

```{r cs setup spatial functions and files, warning = FALSE, include = FALSE}
## in order to assign bhi regions to data, must run this function first, to load required shapefiles:
source(here::here("R", "spatial.R"))
regions_shape(sp_dir = file.path(dirname(dir_B), "Shapefiles"))

basin_lookup <- file.path(dir_prep, "supplement", "lookup_tabs", "rgns_complete.csv") %>% 
  read_csv() %>%
  dplyr::select(bhi_id = region_id, subbasin, rgn_nam = region_name, HELCOM_ID = helcom_id) %>%
  mutate(
    subbasin = as.character(subbasin),
    rgn_nam = as.character(rgn_nam)
  )
```


## 1. Background {-}

```{r cs prep background, child = bkgd_path, results = "asis", echo = FALSE}
```

<br/>

## 2. Data {-}

This prep document is used to generate and explore the following data layers:

- `cs_status_bhi2019.csv` 

These are saved to the `layers` folder. Saved to `data/CS/v2019/intermediate` are intermediate datasets: `bhi_transf.shp` and `cs_bhi_intersect.shp`. All these are derived from HELCOM Map and Data Service.

<br>

```{r cs prep data, child = data_path, results = "asis", echo = FALSE}
```

<br/>

## 3. Prep: Wrangling & Derivations, Checks/Evaluation, Gapfilling {-}

```{r load cs datasets, echo = TRUE}
## previously created in 'data/CS/v2019/ao_data.R' script
cs_rawdata <- sf::st_read(file.path(
  dir_intermediate, "cs_bhi_intersect.shp"
))
st_geometry(cs_rawdata) <- NULL
```

<br/>

## 4. Status

Zostera new shape file explore: 

- 10km grid cell
- field description: 

0 = no observations
1 = present before year 1995 or in 1995
2 = present after year 1995
3 = present before and after year 1995

Expert opinion (Christoffer Boström) suggested that no growth of Zostera sp naturally occurs in these BHI ID regions: 12, 15, 17, 19, 21, 22, 23, 24, 37, 38, 39, 40, 41, 42, and were left as NA in this analysis. 

**Status**
```{r cs status calculation, results = "show", echo = TRUE, fig.width = 9.5, fig.height = 5.5}
cs_status <- cs_rawdata %>% 
  dplyr::select(rgn_id = BHI_ID, country = rgn_nam, zostera = Z_marina) %>% 
  mutate(helcom_score = ifelse(zostera == 2, 100, zostera/3*100)) %>% # if zostera = 2, set score to 100
  group_by(rgn_id) %>% 
  summarize(score = mean(helcom_score)) %>%
  mutate(score = round(score, 2)) %>% 
  ungroup

### identify regions with no eelgrass growth potential and set as NA
no_eelgrass <-c(12, 15, 17, 19, 21, 22, 23, 24, 37, 38, 39, 40, 41, 42)

cs_status_2 <- cs_status %>% 
  mutate(score = ifelse(rgn_id %in% no_eelgrass, NA, score)) %>% 
  dplyr::select(rgn_id, 
                score) %>% 
  mutate(dimension = "status")

# Write status as csv file
write_csv(cs_status_2, file.path(dir_layers, 'cs_status_bhi2019.csv'))

## Plot Status
cs_plot <- ggplot(cs_status_2) +
  geom_bar(aes(rgn_id, score), stat = 'identity') +
  labs(title = "CS score based on HELCOM Zostera data",
       x = "BHI region",
       y = "Score") 
  
plot_status <- cs_status_2 %>% 
  full_join(.,basin_lookup, by = c("rgn_id"= "bhi_id"))

set.seed(2)
statuspal <- colorRampPalette(fullpal)(80)[sample(1:80, size = 18)]

ggplot(plot_status) +
  geom_col(
    aes(rgn_nam, score, fill = subbasin),
    position = position_dodge(), 
    color = "black", 
    alpha = 0.6,
    size = 0.2,
    show.legend = FALSE
  ) + 
  geom_text(aes(rgn_nam, score, label = score), color = "black", size = 3) +
  coord_flip() +
  labs(x = NULL, y = "Score") +
  theme_bw() +
  theme(axis.text.y = element_text(size = 9)) +
  scale_fill_manual(values = statuspal) +
  ggtitle("CS Status by Region")
```

<br/>


## 5. Trend

**Trend**
Trend will be NA for all regions. 


## 6. References

```{r References, child = refs_path, results = "asis", echo = FALSE}
```
