
```{r ao data preamble, echo = FALSE, include = FALSE, error = FALSE}
source(here::here("R", "setup.R"))
source(here::here("R", "data.R"))
source(here::here("R", "spatial.R"))
library(tidyr)
## root location of the raw data
dir_B <- file.path(dirname(dir_prep), "bhi-data", "BHI 2.0")
dir_rawdata <- file.path(dir_B, "Goals", "AO")
```

### 2.1 Datasets with Sources {-}
<br/>

#### 2.1.1 Coastal fish Data {-}

Data were received by Jens Olsson via email (18 March 2020). 
Dare from monitoring locations (described in the HELCOM core indicators). 
Finnish data are fisheries data from ICES assessment regions (ICES 29-32).


### 2.2 Centralization & Normalization {-}

#### 2.2.1 Coastal fish datasets {-}

Since coastal fish data ('coastal_fish_data_2018.csv') are updated until 2018, but the latest status assessment (HOLAS II) only includes data until 2016, we have decided to use the data included in HOLAS II (until 2016) for this first BHI 2.0 calculations ('key_species_status.csv', 'piscivores_status.csv', 'cyprinids_status.csv'). Then, once the most updated data will be included into the next "intermediate" status assessment, we will also include the newest data.

```{r load raw data, echo = TRUE, message = FALSE, warning = FALSE}
# library(readr)
# load all raw data (containing the newest data)
coastal_fish_raw <- read.csv(
  file.path(dir_rawdata, "coastal_fish_data_2018.csv"), 
  sep = ";")

# key species status data (until 2016)
key_species_status_raw <- read.csv(
  file.path(dir_rawdata, "key_species_status.csv"),
  sep = ";")

# functional groups (piscivores and cyprinids) status data (until 2016)
piscivores_status_raw <- read.csv(
  file.path(dir_rawdata, "piscivores_status.csv"), 
  sep = ";")

cyprinids_status_raw <- read.csv(
  file.path(dir_rawdata, "cyprinids_status.csv"), 
  sep = ";")

coastal_areas <- sf::st_read(file.path(
  dir_rawdata, "Coastal fish monitoring areas",
  "CoastalFishAreas.shp"
)) # CRS: 3035
```

**Explore and clean coastal fish data for trend calculation**
```{r cleaning coastal fish raw data, echo = TRUE, message = FALSE, warning = FALSE}
coastal_fish_cpue <- coastal_fish_raw %>% 
  dplyr::rename(
    monitoring_area = Area, 
    coastal_water_type = Assessment.unit,
    core_indicator = Indicator,
    cpue = Value
  ) %>% 
   dplyr::mutate(
      monitoring_area = str_replace(monitoring_area, "≈", "A"), ## need to clean the areas' names!
      monitoring_area = str_replace(monitoring_area, "Area south of Zealand (SmÂlandsfarvandet)", "Area south of Zealand (Smalandsfarvandet)"),
      monitoring_area = str_replace(monitoring_area, "Askˆ", "Asko"),
      monitoring_area = str_replace(monitoring_area, "Gaviksfj‰rden", "Gaviksfjaerden"),
      monitoring_area = str_replace(monitoring_area, "Holmˆn", "Holmoen"),
      monitoring_area = str_replace(monitoring_area, "Kinnb‰cksfj‰rden", "Kinnbaecksfjaerden"),
      monitoring_area = str_replace(monitoring_area, "RÂneÂ", "Ranea"),
      monitoring_area = str_replace(monitoring_area, "LÂngvindsfj‰rden", "Langvindsfjaerden"),
      monitoring_area = str_replace(monitoring_area, "Kv‰dˆfj‰rden", "Kvaedoefjaerden"),
      monitoring_area = str_replace(monitoring_area, "LÂngvindsfj‰rden", "Langvindsfjaerden"),
      monitoring_area = str_replace(monitoring_area, "Rectangle 23", "ICES SD 23"),
      monitoring_area = str_replace(monitoring_area, "Rectangle 28", "ICES SD 28"),
      monitoring_area = str_replace(monitoring_area, "PrÊst¯ Fiord", "Praestoe Fiord"),
      Subbasin = str_replace(Subbasin, "≈", "A"),
      Subbasin = str_replace(Subbasin, "Arkona basin", "Arkona Basin"),
      Subbasin = str_replace(Subbasin, "Bornholm basin", "Bornholm Basin"),
      cpue = str_replace(cpue, ",", "."),
      cpue = str_replace(cpue, "-", ""),
      cpue = as.numeric(cpue),
      core_indicator = ifelse(Functional.group == "Cyprinids" & core_indicator  == "Key species", "Key functional group", core_indicator),
      Taxa = if_else(
        Species == "-", Functional.group, Species)) %>% 
  dplyr::select(-Species, -Functional.group)
```
<br>

```{r save cleaned cpue data, eval = FALSE}
readr::write_csv(
  coastal_fish_cpue, 
  here::here("data", "AO", version_year, "intermediate", "ao_cleaned_cpue.csv")
)
```
<br>


**Clean, explore and merge key species, cyprinids and piscivores status data**
```{r merging coastal fish raw data, echo = TRUE, message = FALSE, warning = FALSE}
coastal_fish_status <- rbind(
  key_species_status_raw %>% 
    dplyr::rename(
      Latitude = Coordinates,
      Longitude = X,
      Taxa = Identity.of.key.species,
      Monitoring.area = Monitoring.location,
      Subbasin = Sub.basin.name,
      coastal_water_type = Coastal.area.name..assessment.unit.,
      period = Time.period.assessed
    ) %>% 
    dplyr::mutate(
      Latitude = str_replace(Latitude, ",", "."),
      Latitude = str_replace(Latitude, "Area not point", ""),
      Longitude = str_replace(Longitude, ",", "."),
      ## assign lat-lon for ICES area to allow map plotting 
      # (assign location in center of Finnish coasts for ICES areas (23,28,29-32)
      Latitude = ifelse(Monitoring.area == "Finnish ICES SD 31", 64.207281,
                 ifelse(Monitoring.area == "Finnish ICES SD 30", 61.679815,
                 ifelse(Monitoring.area == "Finnish ICES SD 29", 59.705518,
                 ifelse(Monitoring.area == "Finnish ICES SD 32", 60.108130,
                 ifelse(Monitoring.area == "Finnish ICES rect 23", 63.505817, 
                 ifelse(Monitoring.area == "Finnish ICES rect 28", 63.505817, Latitude)))))),
      Longitude = ifelse(Monitoring.area == "Finnish ICES SD 31", 23.511161,
                  ifelse(Monitoring.area == "Finnish ICES SD 30", 21.269112,
                  ifelse(Monitoring.area == "Finnish ICES SD 29", 21.264879,
                  ifelse(Monitoring.area == "Finnish ICES SD 32", 25.772410,
                  ifelse(Monitoring.area == "Finnish ICES rect 23", 21.351243, 
                  ifelse(Monitoring.area == "Finnish ICES rect 28", 21.351243, Longitude)))))),
      Latitude = as.numeric(Latitude),
      Longitude = as.numeric(Longitude),
      Status.value.monitoring.location = str_replace(Status.value.monitoring.location, ",", "."),
      Status.value.monitoring.location = as.numeric(Status.value.monitoring.location),
      Monitoring.area = str_replace(Monitoring.area, "Kvädöfjärden, autumn", "Kvädöfjärden"),
      Monitoring.area = str_replace(Monitoring.area, "Kvädöfjärden, perch", "Kvädöfjärden"),
      threshold_value = 0.6,
      core_indicator = "key species"),
  cyprinids_status_raw %>% 
    dplyr::rename(
      Latitude = Coordinates,
      Longitude = X,
      Monitoring.area = Monitoring.location,
      Subbasin = Sub.basin.name,
      coastal_water_type = Coastal.area.name..assessment.unit.,
      period = Time.period.assessed
    ) %>% 
    dplyr::select(-Identity.of.indicator) %>% 
    dplyr::mutate(
      Latitude = str_replace(Latitude, ",", "."),
      Latitude = str_replace(Latitude, "Area not point", ""),
      Longitude = str_replace(Longitude, ",", "."),
       Latitude = ifelse(Monitoring.area == "Finnish ICES SD 31", 64.207281,
                 ifelse(Monitoring.area == "Finnish ICES SD 30", 61.679815,
                 ifelse(Monitoring.area == "Finnish ICES SD 29", 59.705518,
                 ifelse(Monitoring.area == "Finnish ICES SD 32", 60.108130,
                 ifelse(Monitoring.area == "Finnish ICES rect 23", 63.505817, 
                 ifelse(Monitoring.area == "Finnish ICES rect 28", 63.505817, Latitude)))))),
      Longitude = ifelse(Monitoring.area == "Finnish ICES SD 31", 23.511161,
                  ifelse(Monitoring.area == "Finnish ICES SD 30", 21.269112,
                  ifelse(Monitoring.area == "Finnish ICES SD 29", 21.264879,
                  ifelse(Monitoring.area == "Finnish ICES SD 32", 25.772410,
                  ifelse(Monitoring.area == "Finnish ICES rect 23", 21.351243, 
                  ifelse(Monitoring.area == "Finnish ICES rect 28", 21.351243, Longitude)))))),
      Latitude = as.numeric(Latitude),
      Longitude = as.numeric(Longitude),
      Status.value.monitoring.location = str_replace(Status.value.monitoring.location, ",", "."),
      Status.value.monitoring.location = as.numeric(Status.value.monitoring.location),
      threshold_value = 0.6,
      Taxa =  "cyprinids",
      core_indicator = "functional group"),
  piscivores_status_raw %>% 
    dplyr::rename(
      Latitude = Coordinates,
      Longitude = X,
      Subbasin = Sub.basin.name,
      coastal_water_type = Coastal.area.name..assessment.unit.,
      period = Time.period.assessed
    ) %>% 
    # separate_rows(Taxa, sep =",\\s+") %>% # serapate different taxa from single rows
    dplyr::select(-Identity.of.key.piscivores) %>% 
    dplyr::mutate(
      Latitude = str_replace(Latitude, ",", "."),
      Latitude = str_replace(Latitude, "Area not point", ""),
      Longitude = str_replace(Longitude, "Åland Sea", "."),
      Longitude = str_replace(Longitude, ",", "."),
       Latitude = ifelse(Monitoring.area == "Finnish ICES SD 31", 64.207281,
                 ifelse(Monitoring.area == "Finnish ICES SD 30", 61.679815,
                 ifelse(Monitoring.area == "Finnish ICES SD 29", 59.705518,
                 ifelse(Monitoring.area == "Finnish ICES SD 32", 60.108130,
                 ifelse(Monitoring.area == "Finnish ICES rect 23", 63.505817, 
                 ifelse(Monitoring.area == "Finnish ICES rect 28", 63.505817, Latitude)))))),
      Longitude = ifelse(Monitoring.area == "Finnish ICES SD 31", 23.511161,
                  ifelse(Monitoring.area == "Finnish ICES SD 30", 21.269112,
                  ifelse(Monitoring.area == "Finnish ICES SD 29", 21.264879,
                  ifelse(Monitoring.area == "Finnish ICES SD 32", 25.772410,
                  ifelse(Monitoring.area == "Finnish ICES rect 23", 21.351243, 
                  ifelse(Monitoring.area == "Finnish ICES rect 28", 21.351243, Longitude)))))),
      Latitude = as.numeric(Latitude),
      Longitude = as.numeric(Longitude),
      Status.value.monitoring.location = str_replace(Status.value.monitoring.location, ",", "."),
      Status.value.monitoring.location = as.numeric(Status.value.monitoring.location),
      threshold_value = 0.6,
      Taxa =  "piscivores",
      core_indicator = "functional group"))
  
```
<br>

```{r save merged data, eval = FALSE, echo = TRUE, message = FALSE, warning = FALSE}
readr::write_csv(
  coastal_fish_status, 
  here::here("data", "AO", version_year, "intermediate", "ao_merged_rawdata.csv")
)
```
<br>


### 2.3 Initial Data Exploration {-}

#### 2.3.1 Check unique indicators per monitoring location

1. Is more than one key species monitored at a given locations?  **Yes, in Kvädöfjärden both flounder (autumn) and perch (summer)**
2. Is more than one functional group monitored?  **Yes, except in Muskö (Sweden) where no assessment of cyprinids/mesopredators is available. However, in Danish areas, only key species (flounder) are assessed as there are no reliable data for either piscivores or cyprinids/mesopredators**  
3. Are both key species and functional groups monitored at all locations? **No**

- 11 monitoring areas without Functional groups (all Danish areas)
- 1 monitoring area with two Key_spp status: *Kvädöfjärden*  

**Number of indicators by monitoring location**
```{r unique indicators monitoring region, results = "show", echo = TRUE, fig.width = 9.5, fig.height = 6}
## how many core indicators, how many taxa
indicator_count <- coastal_fish_status %>%
  select(
    Monitoring.area, core_indicator, Taxa, Status.value.monitoring.location
  ) %>%
  group_by(Monitoring.area) %>%
  summarise(unique_indicator = length(unique(core_indicator)),
            unique_taxa_func = length(unique(Taxa))
  ) %>%
  ungroup()
# indicator_count %>% print(n=40)

## which stations have only 1 core indicator type (either have only functional or only key spp)
one_indicator <- indicator_count %>% 
  filter(unique_indicator == 1)
# one_indicator %>% print(n=14)

## number of taxa organized by location and core indicator type
indicator_taxa_count <- coastal_fish_status %>% 
  select(
    Monitoring.area, core_indicator, Taxa
    ) %>%
  group_by(Monitoring.area, core_indicator) %>%
  summarise(unique_taxa_func = length(unique(Taxa))) %>%
  ungroup()
# indicator_taxa_count %>% print(n=66)

## which locations are missing an indicator type?
indicator_taxa_count %>% 
  filter(Monitoring.area %in% one_indicator$Monitoring.area)

ggplot(indicator_taxa_count) + 
  geom_point(aes(Monitoring.area, unique_taxa_func)) +
  facet_wrap(~core_indicator) +
  labs(y = "Taxa obs", x = "Monitoring Areas") +
  ggtitle("Number of indicators by monitoring location") +
  theme(axis.text.x = element_text(colour="grey20", size=6, angle=90, 
                                   hjust=.5, vjust=.5, face = "plain"),
        plot.margin = unit(c(1,1,1,1), "cm")) 

## Number of NAs by each indicator
indicator_taxa_count %>%  
  spread(core_indicator, unique_taxa_func) %>% 
  dplyr::rename(key_species = `key species`) %>%
  dplyr::rename(functional_group = `functional group`) %>%
  summarise(Func_na = sum(is.na(functional_group)),                                                                               KeySpp_na = sum(is.na(key_species)))
```
<br>


#### 2.3.2 Compare status values across taxa {-}

```{r show data, results = "show", echo = TRUE, fig.width = 7.5, fig.height = 6}
# make coastal_fish_scores in long format then plot
temp_long <- coastal_fish_status %>% 
  select(
    Monitoring.area, core_indicator, Taxa, Status.value.monitoring.location
    ) %>%
  group_by(
    Monitoring.area, core_indicator, Taxa
    ) %>% 
  ungroup()

# Coastal Fish Status Value (by monitoring area)
ggplot(coastal_fish_status) + 
  geom_point(
    aes(Monitoring.area, Status.value.monitoring.location, 
        colour = Taxa, shape = Taxa)) +
  facet_wrap(~core_indicator) +
  theme(axis.text.x = element_text(colour= "grey20", 
                                   size=7, angle=90, hjust=.5, vjust=.5, face="plain"),
        plot.margin = unit(c(1,1,1,1), "cm")) +
  labs(x = "Monitoring Area", y = "Status Value") +
  ggtitle("Coastal Fish Status Values (by monitoring area)")

# Coastal Fish Status Value (by regions)
ggplot(coastal_fish_status) + 
  geom_point(
    aes(coastal_water_type, Status.value.monitoring.location, 
        colour = Taxa, shape = Taxa)) +
  facet_wrap(~ core_indicator) +
  theme(axis.text.x = element_text(colour= "grey20", 
                                   size=7, angle=90, hjust=.5, vjust=.5, face="plain"),
        plot.margin = unit(c(1,1,1,1), "cm")) +
  labs(x = "Regions", y = "Status Value") +
  ggtitle("Coastal Fish Status Values (by regions)")

```
