---
output:
  word_document: default
  html_document: default
---

```{r eut data preamble, eval = FALSE, echo = TRUE, message = FALSE, warning = FALSE}
source(here::here("R", "setup.R"))
source(here::here("R", "data.R"))
source(here::here("R", "spatial.R"))

## root location of the raw data
dir_B <- file.path(dirname(dir_prep), "bhi-data", "BHI 2.0")
dir_rawdata <- file.path(dir_B, "Goals", "AO")
```

The Artisanal Fishing Opportunities goal, often also called small-scale fishing, provides a critical source of food, nutrition, poverty alleviation and livelihood opportunities for many people around the world, in particular in developing nations. 
This goal measures whether people who need to fish on a small, local scale have the opportunity to do so. 
It has three sub-components: stock, access, and need. 
A score of 100 means the country or region is meeting the needs of artisanal fishermen or communities by implementing institutional supports, providing access to near-shore water, and maintaining the health of targeted species.

For the BHI, the focus is on the **stock** sub-component and will use this as a proxy for the entire goal.

Two core indicators are included:

- abundance of coastal fish key functional groups (CPUE of cyprinids/mesopredators and CPUE of piscivores)
- abundance of key coastal fish species (CPUE of perch, cod or flounder, depending on the sub-basin)


### 2.1 Datasets with Sources {-}
<br/>

#### 2.1.1 Coastal fish Data {-}

Data were received by Jens Olsson via email (18 March 2020). 
Dare from monitoring locations (described in the HELCOM core indicators). Finnish data are fisheries data from ICES assessment regions (ICES 29-32).


### 2.1.2 Reference values {-}

Previously, indicator status was scored per monitoring area using the assessment protocol for coastal fish (baseline or trend-based approach) => yielding whether or not the status per monitoring area is **GES** or **subGES**. If a monitoring station received a "sub-GES" assessment, it was given a qualifier as "low" or "high".  
Environmental status assessments provided by Jens Olsson (SLU). See [HELCOM FISH-PRO II](http://www.helcom.fi/helcom-at-work/projects/fish-pro/) now in progress for the [HELCOM FISH-PRO III](https://helcom.fi/helcom-at-work/projects/fish-pro-iii/) 

Now, to match the [BEAT-tool as developed within HOLAS II](http://stateofthebalticsea.helcom.fi/wp-content/uploads/2017/07/HELCOM_The_integrated_assessment_of_biodiversity_supplementary_report_first_version_2017.pdf), indicator status was normalized in categorial classes:
- Nearest lower value of 0.125 = 0.125
- Nearest lower value of 0.375 = 0.375
- Nearest lower value of 0.6 = 0.6
- Nearest lower value of 0.875 = 0.875
where **threshold value = 0.6**. 
Threshold values for the status assessments are identified based on site-specific time-series data for each indicator. (Site specific values are used, as coastal fish generally have local population structures, limited migration, and show local responses to environmental change) _See also powerpoint provided by Jens Olsson of BHI shared folder_


### 2.2 Centralization & Normalization {-}

#### 2.2.1 Coastal fish datasets {-}

Since coastal fish data ('coastal_fish_data_2018.csv') are updated until 2018, but the latest status assessment (HOLAS II) only includes data until 2016, we have decided to use the data included in HOLAS II (until 2016) for this first BHI 2.0 calculations ('key_species_status.csv', 'piscivores_status.csv', 'cyprinids_status.csv'). Then, once the most updated data will be included into the next "intermediate" status assessment, we will also include the newest data.

```{r load raw data, eval = FALSE, echo = TRUE, message = FALSE, warning = FALSE}
# library(readr)
# load all raw data (containing the newest data)
coastal_fish_raw <- read.csv(
  file.path(dir_rawdata, "coastal_fish_data_2018.csv"), 
  sep = ";")

# key species status data (until 2016)
key_species_status_raw <- read.csv(
  file.path(dir_rawdata, "key_species_status.csv"),
  sep = ";")

# functional groups (piscivores and cyprinids) status data (until 2016)
piscivores_status_raw <- read.csv(
  file.path(dir_rawdata, "piscivores_status.csv"), 
  sep = ";")

cyprinids_status_raw <- read.csv(
  file.path(dir_rawdata, "cyprinids_status.csv"), 
  sep = ";")
```

**Explore and clean coastal fish data for trend calculation**
```{r cleaning coastal fish raw data, eval = FALSE, echo = TRUE, message = FALSE, warning = FALSE}
coastal_fish_cpue <- coastal_fish_raw %>% 
  dplyr::rename(
    monitoring_area = Area, 
    coastal_water_type = Assessment.unit,
    core_indicator = Indicator,
    cpue = Value
  ) %>% 
   dplyr::mutate(
      monitoring_area = str_replace(monitoring_area, "≈", "A"), ## need to clean the areas' names!
      monitoring_area = str_replace(monitoring_area, "Area south of Zealand (SmÂlandsfarvandet)", "Area south of Zealand (Smalandsfarvandet)"),
      monitoring_area = str_replace(monitoring_area, "Askˆ", "Asko"),
      monitoring_area = str_replace(monitoring_area, "Gaviksfj‰rden", "Gaviksfjaerden"),
      monitoring_area = str_replace(monitoring_area, "Holmˆn", "Holmoen"),
      monitoring_area = str_replace(monitoring_area, "Kinnb‰cksfj‰rden", "Kinnbaecksfjaerden"),
      monitoring_area = str_replace(monitoring_area, "RÂneÂ", "Ranea"),
      monitoring_area = str_replace(monitoring_area, "LÂngvindsfj‰rden", "Langvindsfjaerden"),
      monitoring_area = str_replace(monitoring_area, "Kv‰dˆfj‰rden", "Kvaedoefjaerden"),
      monitoring_area = str_replace(monitoring_area, "LÂngvindsfj‰rden", "Langvindsfjaerden"),
      monitoring_area = str_replace(monitoring_area, "Rectangle 23", "ICES SD 23"),
      monitoring_area = str_replace(monitoring_area, "Rectangle 28", "ICES SD 28"),
      monitoring_area = str_replace(monitoring_area, "PrÊst¯ Fiord", "Praestoe Fiord"),
      Subbasin = str_replace(Subbasin, "≈", "A"),
      Subbasin = str_replace(Subbasin, "Arkona basin", "Arkona Basin"),
      Subbasin = str_replace(Subbasin, "Bornholm basin", "Bornholm Basin"),
      cpue = str_replace(cpue, ",", "."),
      cpue = str_replace(cpue, "-", ""),
      cpue = as.numeric(cpue),
      Taxa = if_else(
        Species == "-", Functional.group, Species)) %>% 
  select(-Species, -Functional.group)
```
<br>

```{r save cleaned cpue data, eval = FALSE}
readr::write_csv(
  coastal_fish_cpue, 
  here::here("data", "AO", version_year, "intermediate", "ao_cleaned_cpue.csv")
)
```
<br>


**Clean, explore and merge key species, cyprinids and piscivores status data**
```{r merging coastal fish raw data, eval = FALSE, echo = TRUE, message = FALSE, warning = FALSE}
coastal_fish_status <- rbind(
  key_species_status_raw %>% 
    dplyr::rename(
      Latitude = Coordinates,
      Longitude = X,
      Taxa = Identity.of.key.species,
      Monitoring.area = Monitoring.location,
      Subbasin = Sub.basin.name,
      coastal_water_type = Coastal.area.name..assessment.unit.,
      period = Time.period.assessed
    ) %>% 
    dplyr::mutate(
      Latitude = str_replace(Latitude, ",", "."),
      Latitude = str_replace(Latitude, "Area not point", ""),
      Longitude = str_replace(Longitude, ",", "."),
      ## assign lat-lon for ICES area to allow map plotting 
      # (assign location in center of Finnish coasts for ICES areas (23,28,29-32)
      Latitude = ifelse(Monitoring.area == "Finnish ICES SD 31", 64.207281,
                 ifelse(Monitoring.area == "Finnish ICES SD 30", 61.679815,
                 ifelse(Monitoring.area == "Finnish ICES SD 29", 59.705518,
                 ifelse(Monitoring.area == "Finnish ICES SD 32", 60.108130,
                 ifelse(Monitoring.area == "Finnish ICES rect 23", 63.505817, 
                 ifelse(Monitoring.area == "Finnish ICES rect 28", 63.505817, Latitude)))))),
      Longitude = ifelse(Monitoring.area == "Finnish ICES SD 31", 23.511161,
                  ifelse(Monitoring.area == "Finnish ICES SD 30", 21.269112,
                  ifelse(Monitoring.area == "Finnish ICES SD 29", 21.264879,
                  ifelse(Monitoring.area == "Finnish ICES SD 32", 25.772410,
                  ifelse(Monitoring.area == "Finnish ICES rect 23", 21.351243, 
                  ifelse(Monitoring.area == "Finnish ICES rect 28", 21.351243, Longitude)))))),
      Latitude = as.numeric(Latitude),
      Longitude = as.numeric(Longitude),
      Status.value.monitoring.location = str_replace(Status.value.monitoring.location, ",", "."),
      Status.value.monitoring.location = as.numeric(Status.value.monitoring.location),
      Monitoring.area = str_replace(Monitoring.area, "Kvädöfjärden, autumn", "Kvädöfjärden"),
      Monitoring.area = str_replace(Monitoring.area, "Kvädöfjärden, perch", "Kvädöfjärden"),
      threshold_value = 0.6,
      core_indicator = "key species"),
  cyprinids_status_raw %>% 
    dplyr::rename(
      Latitude = Coordinates,
      Longitude = X,
      Monitoring.area = Monitoring.location,
      Subbasin = Sub.basin.name,
      coastal_water_type = Coastal.area.name..assessment.unit.,
      period = Time.period.assessed
    ) %>% 
    select(-Identity.of.indicator) %>% 
    dplyr::mutate(
      Latitude = str_replace(Latitude, ",", "."),
      Latitude = str_replace(Latitude, "Area not point", ""),
      Longitude = str_replace(Longitude, ",", "."),
       Latitude = ifelse(Monitoring.area == "Finnish ICES SD 31", 64.207281,
                 ifelse(Monitoring.area == "Finnish ICES SD 30", 61.679815,
                 ifelse(Monitoring.area == "Finnish ICES SD 29", 59.705518,
                 ifelse(Monitoring.area == "Finnish ICES SD 32", 60.108130,
                 ifelse(Monitoring.area == "Finnish ICES rect 23", 63.505817, 
                 ifelse(Monitoring.area == "Finnish ICES rect 28", 63.505817, Latitude)))))),
      Longitude = ifelse(Monitoring.area == "Finnish ICES SD 31", 23.511161,
                  ifelse(Monitoring.area == "Finnish ICES SD 30", 21.269112,
                  ifelse(Monitoring.area == "Finnish ICES SD 29", 21.264879,
                  ifelse(Monitoring.area == "Finnish ICES SD 32", 25.772410,
                  ifelse(Monitoring.area == "Finnish ICES rect 23", 21.351243, 
                  ifelse(Monitoring.area == "Finnish ICES rect 28", 21.351243, Longitude)))))),
      Latitude = as.numeric(Latitude),
      Longitude = as.numeric(Longitude),
      Status.value.monitoring.location = str_replace(Status.value.monitoring.location, ",", "."),
      Status.value.monitoring.location = as.numeric(Status.value.monitoring.location),
      threshold_value = 0.6,
      Taxa =  "cyprinids",
      core_indicator = "functional group"),
  piscivores_status_raw %>% 
    dplyr::rename(
      Latitude = Coordinates,
      Longitude = X,
      Subbasin = Sub.basin.name,
      coastal_water_type = Coastal.area.name..assessment.unit.,
      period = Time.period.assessed
    ) %>% 
    # separate_rows(Taxa, sep =",\\s+") %>% # serapate different taxa from single rows
    select(-Identity.of.key.piscivores) %>% 
    dplyr::mutate(
      Latitude = str_replace(Latitude, ",", "."),
      Latitude = str_replace(Latitude, "Area not point", ""),
      Longitude = str_replace(Longitude, "Åland Sea", "."),
       Latitude = ifelse(Monitoring.area == "Finnish ICES SD 31", 64.207281,
                 ifelse(Monitoring.area == "Finnish ICES SD 30", 61.679815,
                 ifelse(Monitoring.area == "Finnish ICES SD 29", 59.705518,
                 ifelse(Monitoring.area == "Finnish ICES SD 32", 60.108130,
                 ifelse(Monitoring.area == "Finnish ICES rect 23", 63.505817, 
                 ifelse(Monitoring.area == "Finnish ICES rect 28", 63.505817, Latitude)))))),
      Longitude = ifelse(Monitoring.area == "Finnish ICES SD 31", 23.511161,
                  ifelse(Monitoring.area == "Finnish ICES SD 30", 21.269112,
                  ifelse(Monitoring.area == "Finnish ICES SD 29", 21.264879,
                  ifelse(Monitoring.area == "Finnish ICES SD 32", 25.772410,
                  ifelse(Monitoring.area == "Finnish ICES rect 23", 21.351243, 
                  ifelse(Monitoring.area == "Finnish ICES rect 28", 21.351243, Longitude)))))),
      Latitude = as.numeric(Latitude),
      Longitude = as.numeric(Longitude),
      Status.value.monitoring.location = str_replace(Status.value.monitoring.location, ",", "."),
      Status.value.monitoring.location = as.numeric(Status.value.monitoring.location),
      threshold_value = 0.6,
      Taxa =  "piscivores",
      core_indicator = "functional group"))
  
```
<br>

```{r save merged data, eval = FALSE}
readr::write_csv(
  coastal_fish_status, 
  here::here("data", "AO", version_year, "intermediate", "ao_merged_rawdata.csv")
)
```
<br>


### 2.3 Initial Data Exploration {-}

#### 2.3.1 Check unique indicators per monitoring location

1. Is more than one key species monitored at a given locations?  **Yes, in Kvädöfjärden both flounder (autumn) and perch (summer)**
2. Is more than one functional group monitored?  **Yes, except in Muskö (Sweden) where no assessment of cyprinids/mesopredators is available. However, in Danish areas, only key species (flounder) are assessed as there are no reliable data for either piscivores or cyprinids/mesopredators**  
3. Are both key species and functional groups monitored at all locations? **No**

- 11 monitoring areas without Functional groups (all Danish areas)
- 1 monitoring area with two Key_spp status: *Kvädöfjärden*  

**Number of indicators by monitoring location**
```{r unique indicators monitoring region}
## how many core indicators, how many taxa
indicator_count <- coastal_fish_status %>%
  select(
    Monitoring.area, core_indicator, Taxa, Status.value.monitoring.location
  ) %>%
  group_by(Monitoring.area) %>%
  summarise(unique_indicator = length(unique(core_indicator)),
            unique_taxa_func = length(unique(Taxa))
  ) %>%
  ungroup()
# indicator_count %>% print(n=40)

## which stations have only 1 core indicator type (either have only functional or only key spp)
one_indicator <- indicator_count %>% 
  filter(unique_indicator == 1)
# one_indicator %>% print(n=14)

## number of taxa organized by location and core indicator type
indicator_taxa_count <- coastal_fish_status %>% 
  select(
    Monitoring.area, core_indicator, Taxa
    ) %>%
  group_by(Monitoring.area, core_indicator) %>%
  summarise(unique_taxa_func = length(unique(Taxa))) %>%
  ungroup()
# indicator_taxa_count %>% print(n=66)

## which locations are missing an indicator type?
indicator_taxa_count %>% 
  filter(Monitoring.area %in% one_indicator$Monitoring.area)

ggplot(indicator_taxa_count) + 
  geom_point(aes(Monitoring.area, unique_taxa_func)) +
  facet_wrap(~core_indicator) +
  labs(y = "Taxa obs", x = "Monitoring Areas") +
  theme(axis.text.x = element_text(colour="grey20", size=6, angle=90, 
                                   hjust=.5, vjust=.5, face = "plain"),
        plot.margin = unit(c(1,1,1,1), "cm")) 

## Number of NAs by each indicator
indicator_taxa_count %>%  
  spread(core_indicator, unique_taxa_func) %>% 
  dplyr::rename(key_species = `key species`) %>%
  dplyr::rename(functional_group = `functional group`) %>%
  summarise(Func_na = sum(is.na(functional_group)),                                                                               KeySpp_na = sum(is.na(key_species)))
```
<br>


#### 2.3.2 Compare versus Previous Years Data {-}

#### 2.3.3 Timeseries Plots {-}

```{r show data, results = "show", eval = FALSE, echo = TRUE, fig.width = 7.5, fig.height = 4}
# make coastal_fish_scores in long format then plot
temp_long <- coastal_fish_status %>% 
  select(
    Monitoring.area, core_indicator, Taxa, Status.value.monitoring.location
    ) %>%
  group_by(
    Monitoring.area, core_indicator, Taxa
    ) %>% 
  ungroup()

# Coastal Fish Status Value (by monitoring area)
ggplot(coastal_fish_status) + 
  geom_point(
    aes(Monitoring.area, Status.value.monitoring.location, 
        colour = Taxa, shape = Taxa)) +
  facet_wrap(~core_indicator) +
  theme(axis.text.x = element_text(colour= "grey20", 
                                   size=7, angle=90, hjust=.5, vjust=.5, face="plain"),
        plot.margin = unit(c(1,1,1,1), "cm")) +
  labs(x = "Monitoring Area", y = "Status Value") +
  ggtitle("Coastal Fish Status Value (by monitoring area)")

# Coastal Fish Status Value (by subbasins)
ggplot(coastal_fish_status) + 
  geom_point(
    aes(Subbasin, Status.value.monitoring.location, 
        colour = Taxa, shape = Taxa)) +
  facet_wrap(~ core_indicator) +
  theme(axis.text.x = element_text(colour= "grey20", 
                                   size=7, angle=90, hjust=.5, vjust=.5, face="plain"),
        plot.margin = unit(c(1,1,1,1), "cm")) +
  labs(x = "Subbasins", y = "Status Value") +
  ggtitle("Coastal Fish Status Value (by Subbasins)")

```
