
```{r eut data preamble, echo = FALSE, include = FALSE, error = FALSE}
source(here::here("R", "data.R"))
source(here::here("R", "spatial.R"))

## root location of the raw data
dir_B <- file.path(dirname(dir_prep), "bhi-data", "BHI 2.0")
dir_rawdata <- file.path(dir_B, "Goals", "AO")
```

The Artisanal Fishing Opportunities goal, often also called small-scale fishing, provides a critical source of food, nutrition, poverty alleviation and livelihood opportunities for many people around the world, in particular in developing nations. 
This goal measures whether people who need to fish on a small, local scale have the opportunity to do so. 
It has three sub-components: stock, access, and need. 
A score of 100 means the country or region is meeting the needs of artisanal fishermen or communities by implementing institutional supports, providing access to near-shore water, and maintaining the health of targeted species.

For the BHI, the focus is on the **stock** sub-component and will use this as a proxy for the entire goal.

Two core indicators are included:

- abundance of coastal fish key functional groups (CPUE of cyprinids/mesopredators and CPUE of piscivores)

- abundance of key coastal fish species (CPUE of perch, cod or flounder)

The status assessment protocol is found here.

### 2.1 Datasets with Sources {-}
<br/>

### 2.2 Centralization & Normalization {-}

```{r load raw data, message = FALSE, echo = TRUE, warning = FALSE, results = "hide"}
## root location of the raw data
dir_rawdata <- file.path(dir_B, "Goals", "GOAL/OR/SUBGOAL")
```

#### 2.2.1 Rename Fields/Variables {-}

#### 2.2.2 Standardize Units {-}

#### 2.2.3 Save to BHI Database {-}

<br/>

### 2.3 Initial Data Exploration {-}

#### 2.3.1 Compare versus Previous Years Data {-}

#### 2.3.2 Timeseries Plots {-}

```{r CODE CHUNK WITH FIGURE OR GRAPH, results = "show", message = FALSE, echo = TRUE, fig.width = 9.5, fig.height = 4.5}
```

#### 2.3.3 Map {-}

