---
title: "fp_prep.rmd"
output:
  github_document:
    toc: true
    toc_depth: 3
params: 
    datasource: csv
---

# File to set up layers for the overall Food Provisioning (FP) Goal



```{r setup}

## Libraries
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(RMySQL)
library(stringr)
library(tools)
library(rprojroot) # install.packages('rprojroot')


## source common libraries, directories, functions, etc
source('~/github/bhi/baltic2015/prep/common.r')

## rprojroot
root <- rprojroot::is_rstudio_project


## make_path() function to 
make_path <- function(...) rprojroot::find_root_file(..., criterion = is_rstudio_project)



dir_layers = make_path('baltic2015/layers') # replaces  file.path(dir_baltic, 'layers')


# root$find_file("README.md")
# 
# root$find_file("ao_need_gl2014.csv")
# 
# root <- find_root_file("install_ohicore.r", 
# 
# withr::with_dir(
#   root_file("DESCRIPTION"))



dir_fp   = file.path(dir_prep,'FP')




## add a README.md to the prep directory with the rawgit.com url for viewing on GitHub
create_readme(dir_fp, 'fp_prep.rmd') 
```



## 1. Layer prep for FIS & MAR weight

### 1.1 Read in data

#### 1.1.1 MAR data
Read in the tonnes of MAR production per BHI area from the layers file. 
```{r read in MAR data}
## read in data
mar_data = read.csv(file.path(dir_layers, "mar_harvest_tonnes_bhi2015.csv"))
str(mar_data)

```


#### 1.1.2 FIS data
```{r read in FIS data}

fis_data = read.csv(file.path(dir_layers,'fis_landings_bhi2015.csv'),stringsAsFactors = FALSE)
str(fis_data)

```


### 1.2 Organize data to total tonnes by subgoal per year
```{r organize data}
mar_data = mar_data %>%
           select(rgn_id,year, tonnes) %>%
           arrange(rgn_id,year) %>%
           mutate(subgoal= 'MAR')


fis_data = fis_data %>%
           select(rgn_id, year, landings)%>%
           group_by(rgn_id,year) %>%
           summarise(tonnes = sum(landings, na.rm=TRUE))%>%
           ungroup()%>%
           arrange(rgn_id,year) %>%
           mutate(subgoal = 'FIS')

str(fis_data)
```

### 1.3 Join MAR and FIS data
```{r join mar and fis}
fp_data = bind_rows(mar_data,fis_data)

str(fp_data)
head(fp_data)
tail(fp_data)
```

### 1.4 Spread FP data and Get total tonnes and Fraction from FIS
```{r spread FP data}
fp_data = fp_data %>%
          spread(subgoal, tonnes) %>%
          mutate(FIS = ifelse(is.na(FIS),0,FIS),
                 MAR = ifelse(is.na(MAR),0,MAR))%>%  ## replace NA with zero, no production from that subgoal
          mutate(tot_tonnes = FIS + MAR,
                 fis_prop = FIS / tot_tonnes)
          
  
```

### 1.4 Plot totals and proportion FIS
```{r plot tonnes and proportion FIS}
ggplot(fp_data)+
    geom_point(aes(year,MAR), colour="green",size=1)+
    geom_point(aes(year,FIS), colour="blue", size=1)+
  facet_wrap(~rgn_id)+
  ylab("tonnes")+
  theme(axis.text.x = element_text(colour="grey20", size=6, angle=90, 
                                    hjust=.5, vjust=.5, face = "plain"),
        axis.text.y = element_text(size=6))+
  ggtitle("Mariculture Production (green) and Fishery Landings(blue)")


ggplot(fp_data)+
    geom_point(aes(year,fis_prop), colour="black",size=1)+
    facet_wrap(~rgn_id)+
    ylab("proportion tonnes")+
  theme(axis.text.x = element_text(colour="grey20", size=6, angle=90, 
                                    hjust=.5, vjust=.5, face = "plain"))+
    ggtitle("Proportion of Total Fishery Harvest from wild-capture fisheries (FIS)")

```

### 1.5 Take FIS prop mean across years by BHI region
Only take mean for the most recent 10 years (not MAR data prior and FIS has a very long time series). 2003-2014
```{r mean across years}

fp_data1 = fp_data %>%
           select(rgn_id, year, fis_prop) %>%
           filter(year>2002 & year < 2015) %>%
           group_by(rgn_id) %>%
           summarise(mean_fis_prop = mean(fis_prop))%>%
           ungroup()

```

### 1.6 Plot mean FIS prop 2003-2014
```{r Plot mean FIS prop 2003-2014}

ggplot(fp_data1)+
    geom_point(aes(rgn_id,mean_fis_prop), colour="black",size=2.5)+
    ylab("proportion tonnes")+
  theme(axis.text.x = element_text(colour="grey20", size=6, angle=90, 
                                    hjust=.5, vjust=.5, face = "plain"))+
    ggtitle("Mean Proportion of Total Fishery Harvest from wild-capture fisheries (FIS) 2003-2014")


```


### 1.7 Prepare data layer
```{r prepare prop fis data layer}

fp_wildcaught_weight = fp_data1 %>%
                       select(rgn_id, mean_fis_prop)%>%
                       dplyr:: rename(w_fis = mean_fis_prop)%>%
                       arrange(rgn_id)

```

### 1.8 Save layer
```{r save fp_wildcaught_weight data layer}

write.csv(fp_wildcaught_weight, file.path(dir_layers, 'fp_wildcaught_weight_bhi2015.csv'), row.names = FALSE)
```

