---
title: "Livelihoods and Economies (LE) - Economies (ECO) Subgoal Data Preparation"
output:
  html_document:
    toc: true
    toc_depth: 4
    code_folding: hide
---

<br>
<br>

```{r eco preamble prep, message = FALSE}
loc <- here::here("prep", "ECO")

source(here::here("R", "setup.R"))
knitr::opts_chunk$set(message = FALSE, warning = FALSE, results = "hide")

bkgd_path <- here::here("supplement", "goal_summaries", "ECO.Rmd")
data_path <- here::here("data", "ECO", version_year, "eco_data.rmd")
refs_path <- file.path(loc, "eco_references.Rmd")
# file.exists(c(loc, bkgd_path, data_path, refs_path))
```

## 1. Background

```{r eco background, child = bkgd_path, results = "asis", echo = FALSE}
```

<br/>

## 2. Data

This prep document is used to generate and explore the following data layers:

- `le_eco_bluegrowth_rates_bhi2019.csv` 
- `le_eco_yearly_gva_bhi2019.csv`

These are saved to the `layers` folder. Saved to `data/ECO/v2019/intermediate` is intermediate dataset: `le_eco_bluegrowth_bhi2019.csv`. All these are derived from or informed by the raw datasets from the EU Blue Growth 2020 Report.

<br>

<!-- ## 2. Data --- header in the child  document -->
```{r eco data, child = data_path, results = "asis", echo = FALSE}
```

<br/>

## 3. Prep: Wrangling & Derivations, Checks/Evaluation, Gapfilling

<br/>

## 4. Visualizing Data Layers

### 4.1 Map

```{r marine sectors economic growth rates, eval = FALSE, results = "show", fig.width = 9.5, fig.height = 8.5}
# source(here::here("R", "spatial.R"))
# regions_shape()
# bhirgn_simple <- rmapshaper::ms_simplify(input = BHI_rgns_shp) %>% sf::st_as_sf()

mapsf <- left_join(
  bhirgn_simple %>% 
    select(region_id = BHI_ID, Subbasin, country = rgn_nam),
  eco_data_bhirgns %>% 
    filter(year == 2018) %>% 
    select(region_id, country, sector, annualgrowthrate),
    # mutate(sector = stringr::str_replace_all(stringr::str_to_lower(sector), " ", "_")) %>% 
    # tidyr::pivot_wider(names_from = sector, values_from = annualgrowthrate),
  by = c("region_id", "country")
)

ggplot2::ggplot(rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")) +
  geom_sf(size = 0.1, color = "burlywood", alpha = 0.4) +
  theme(
    panel.background = element_rect(fill = "#F8FBFC", color = "#E2EEF3"), 
    legend.position = c(0.9, 0.15),
    legend.background = element_rect(color = "lavender")
  ) +
  scale_x_continuous(limit = c(4, 32)) +
  scale_y_continuous(limit = c(53.5, 66)) +
  geom_sf(data = filter(mapsf, !is.na(annualgrowthrate)), aes(fill = annualgrowthrate)) +
  labs(fill = "Annual Growth Rate (%)") +
  facet_wrap(~sector) +
  scale_fill_gradientn(
    colors = c("coral", "khaki", "cadetblue", "lightblue", "lavender", "ivory"),
    limits = c(-15, 40)
  )
```

<br>

## 5. Considerations for `BHI3.0`

Will need to find new data sources for GVA in next BHI Assessments, if Blue Growth Reports are discontinued. Eurostat has SBS trade indicators, which together with other sources form the basis for the Blue Growth Reports. These data could be explored as a replacement for GVA values extracted from the Blue Growth Reports. This option (using eurostat data) may be made easier by using the Eurostat R package.

Use [Eurostat API for Administrative Units](https://ec.europa.eu/eurostat/web/gisco/geodata/reference-data/administrative-units-statistical-units) and [R package 'eurostat: Tools for Eurostat Open Data'](https://cran.r-project.org/web/packages/eurostat/index.html)

<br>

## 6. References

```{r eco references, child = refs_path, results = "asis", echo = FALSE}
```
