---
title: "Clean Water (CW) - Eutrophication (EUT) Subgoal Data Preparation"
output:
  github_document:
    toc: true
    toc_depth: 3
params:
    datasource: csv
---

<br>

```{r eut preamble, message = FALSE, warning = FALSE, include = FALSE}
loc <- here::here("prep", "CW", "eutrophication")

source(here::here("R", "setup.R"))
knitr::opts_chunk$set(message = FALSE, warning = FALSE, results = "hide", fig.width = 9.5)

glossary <- here::here("supplement", "glossary.Rmd")
bkgd_path <- here::here("supplement", "goal_summaries", "EUT.Rmd")
data_path <- here::here("data", "CW", "eutrophication", version_year, "eut_data.rmd")
refs_path <- file.path(loc, "eut_references.Rmd")

dir_layers <- here::here("layers")

dir_B <- file.path(dirname(dir_prep), "bhi-data", "BHI 2.0")
dir_rawdata <- file.path(dir_B, "Goals", "CW", "EUT")

fullpal <- c(
  RColorBrewer::brewer.pal(8, "Dark2"),
  RColorBrewer::brewer.pal(9, "Set1")
)
```

```{r eut setup spatial functions and files, warning = FALSE, include = FALSE}
## in order to assign bhi regions to data, must run this function first, to load required shapefiles:
source(here::here("R", "spatial.R"))
regions_shape(sp_dir = file.path(dirname(dir_B), "Shapefiles"))

## import offshore helcom codes,  used for Secchi and Chlorophyll a
offshore_codes <- sf::st_read(file.path(
  dir_B, "Goals", "CW", "EUT", 
  "HELCOM_subbasins_with_coastal_and_offshore_division_2018",
  "HELCOM_subbasins_with_coastal_and_offshore_division_2018.shp"
))
offshore_codes <- offshore_codes %>% 
  filter(country == "Opensea") %>% 
  rename(Subbasin = level_2, helcom_coastal_code = coastal) %>% 
  select(-level_3, -country)  

## import coastal helcom codes, used for Chlorophyll a
coastal_codes <- sf::st_read(file.path(
  dir_B, "Goals", "CW", "EUT", 
  "HELCOM_subbasins_with_coastal_WFD_waterbodies_or_watertypes_2018",
  "HELCOM_subbasins_with_coastal_WFD_waterbodies_or_watertypes_2018.shp"
))

basin_lookup <- file.path(dir_prep, "supplement", "lookup_tabs", "rgns_complete.csv") %>% 
  read_csv() %>%
  select(bhi_id = region_id, subbasin, rgn_nam = region_name, HELCOM_ID = helcom_id) %>%
  mutate(
    subbasin = as.character(subbasin),
    rgn_nam = as.character(rgn_nam)
  )
```

## 1. Background {-}

```{r eut prep background, child = glossary, results = "asis", echo = FALSE}
```

```{r eut prep background, child = bkgd_path, results = "asis", echo = FALSE}
```

<br>

## 2. Data {-}

This prep document is used to generate and explore the following data layers:

- `cw_eut_secchi_bhi2019.csv`
- `cw_eut_chla_bhi2019.csv`
- `cw_eut_din_bhi2019.csv`
- `cw_eut_dip_bhi2019.csv`
- `cw_eut_oxydebt_bhi2019.csv`

These are saved to the `layers` folder. Saved to `data/CW/eutrophication/v2019/intermediate` are intermediate datasets: `secchi_merged_rawdata.csv`,`chla_merged_rawdata.csv`, `ctd_merged_rawdata.csv`, and `ox_merged_rawdata.csv`. All these are derived from or informed by the raw datasets from ICES, SMHI and Baltic Nest database.

<br>

```{r eut prep data, child = data_path, results = "asis", echo = FALSE}
```

<br>

## 3. Prep: Wrangling & Derivations, Checks/Evaluation, Gapfilling {-}

```{r load eut datasets, echo = TRUE}
## secchi, oxygen, and chlorophyll a datasets, 
## previously merged in 'data/CW/eutrophication/eut_data.R' script
combined_secchi_rawdata <- read_csv(file.path(dirname(data_path), "intermediate", "secchi_merged_rawdata.csv"))
combined_ox_rawdata <- read_csv(file.path(dirname(data_path), "intermediate", "ox_merged_rawdata.csv"))

chla_rawdata <- read_csv(file.path(dirname(data_path), "intermediate", "chla_rawdata.csv"))
# tmp <- problems(chla_rawdata) # to check parsing errors
# unique(tmp$col) # only problem with parsing cruise column, doesn't matter, not used in the analysis

## nutrients dataset
nutrients_rawdata <- read_csv(file.path(dir_rawdata, "NestData", "nestdata_winter_nutrients.csv"))

## eutrophication thresholds data
## Note: this table exists in its original form on the bhi internal shared server 
## read.delim(file.path(dir_rawdata, "eut_threshold_values.csv"), sep = ";")
## obtained originally from (?)
## with columns:  
## basin = str_trim(str_replace_all(Assessment.units,"_",""), side = "right"),
## oxyg_debt = as.numeric(str_replace(O2..mg.l.1., "N", " ")),
## winter_DIN = DIN.._mol.l.1., winter_DIP = DIP.._mol.l.1.,
## summer_chla = Chla.._g.l.1., summer_secchi = Water.clarity..m. 
eut_thresholds <- read_csv(file.path(loc, "eut_threshold_values.csv"))

chla_coastal_thresholds <- read_csv(file.path(loc, "chla_coastal_thresholds.csv"))
din_coastal_thresholds <- read_csv(file.path(loc, "din_coastal_thresholds.csv"))
dip_coastal_thresholds <- read_csv(file.path(loc, "dip_coastal_thresholds.csv"))
```
<br/>

**The basin-specific offshore thresholds used in calculating the five Eutrophication indicators**

```{r eutrophication thresholds, results = "show", fig.width = 9.5}
tab <- eut_thresholds
newnames <- colnames(tab) %>% 
  str_replace("_", " ") %>% 
  str_to_title() %>% 
  str_replace("D[a-z]{2}$", str_to_upper(str_extract(., "D[a-z]{2}$")))
colnames(tab) <- newnames
knitr::kable(tab)
```
<br>

**The basin-specific coastal thresholds** _(later excluded from the analyses)_

Since only [Latvia, Poland, and Sweden have threshold values](https://helcom.fi/media/core%20indicators/Dissolved-inorganic-nitrogen-DIN-HELCOM-core-indicator-2018.pdf) for dissoved inorganic nutrients, we only assess offshore components for these indicators, despite the significant impact of coastal nutrients on overall Baltic Health. Nutrient inflows, however, are accounted for elsewhere in the index as pressures e.g. on Biodiversity. 

Both offshore and coastal data are included in calculation of the Chlorophyll a indicator, with different and respective offshore and coastal threshold values. 

```{r eutrophication coastal thresholds, results = "hide"}
tab <- chla_coastal_thresholds %>% 
  rename(summer_chla = threshold, summer_chla_units = units) %>% 
  full_join(
    din_coastal_thresholds %>% 
      rename(winter_DIN = threshold, winter_DIN_units = units) ,
    by = "code"
  ) %>% 
  full_join(
    dip_coastal_thresholds %>% 
      rename(winter_DIP = threshold, winter_DIP_units = units),
    by = "code"
  )
# newnames <- colnames(tab) %>% 
#   str_replace_all("_", " ") %>% 
#   str_to_title() %>% 
#   str_replace("D[a-z]{2}$", str_to_upper(str_extract(., "D[a-z]{2}$"))) %>% 
#   str_replace("D[a-z]{2} Units$", paste(str_to_upper(str_extract(., "D[a-z]{2}")), "Units"))
# colnames(tab) <- newnames
# knitr::kable(tab)

## only have coastal nutrient thresholds for Latvia, Poland, Sweden..
```
<br>

```{r make and save full eutrophication indicators targets table, eval = FALSE}
thresholds_combined <- bind_rows(
  eut_thresholds %>% 
    rename(subbasin = basin) %>% 
    left_join(distinct(basin_lookup, subbasin, HELCOM_ID), by = "subbasin") %>% 
    rename(helcom_id = HELCOM_ID) %>% 
    select(-subbasin) %>% 
    mutate(
      summer_secchi_units =  "meters",
      summer_chla_units = "ug_per_l",
      winter_DIN_units = "umol_per_l",
      winter_DIP_units = "umol_per_l"
    ),
  tab %>% 
    rename(helcom_id = code) %>% 
    mutate(oxyg_debt = NA, summer_secchi = NA, summer_secchi_units = NA)
)

## convertions for uniform units per indicator...
## https://ocean.ices.dk/tools/UnitConversion.aspx
## not sure what to do when have EQR units!!!
thresholds_combined <- thresholds_combined %>% 
  mutate(
    winter_DIN = ifelse(winter_DIN_units == "mg_per_l", winter_DIN*1000, winter_DIN),
    winter_DIN = ifelse(winter_DIN_units == "EQR", NA, winter_DIN),
    winter_DIN_units = "umol_per_l",
    winter_DIP = ifelse(winter_DIP_units == "mg_per_l", winter_DIP*1000, winter_DIP),
    winter_DIP = ifelse(winter_DIP_units == "EQR", NA, winter_DIP),
    winter_DIP_units = "umol_per_l",
  ) %>% 
  select(helcom_id, summer_secchi, summer_chla, winter_DIN, winter_DIP, oxyg_debt) %>% 
  tidyr::pivot_longer(2:6, names_to = "indicator", values_to = "value")

write_csv(
  thresholds_combined, 
  file.path(dir_layers, sprintf("cw_eut_targets_bhi%s.csv", assess_year))
)
```


### 3.1 Secchi Indicator {-}

#### 3.1.1 Match BHI regions {-}

The raw Secchi data (consisting  of merged ICES and SMHI data) must be assigned to their respective Subbasins (with HELCOM Coastal/Offshore division codes) and BHI Regions based on their latitude/longitude coordinates. Then the data are subsetted to assure only summer months (June-September) and years 2000 onward are included.

```{r assign BHI regions to Secchi data, echo = TRUE, message = FALSE, warning = FALSE}
## before assigning bhi regions to secchi data, run this function first, to load required shapefiles:
# source(here::here("R", "spatial.R"))
# regions_shape(sp_dir = file.path(dirname(dir_B), "Shapefiles"))

combined_secchi_rawdata <- combined_secchi_rawdata %>% 
  rename(lat = Latitude, lon = Longitude)

## assign helcom offshore codes
secchi_rgns_w_offshore <- combined_secchi_rawdata %>% 
  sf::st_as_sf(
    crs = 4326,
    coords = c("lon","lat"),
    remove = FALSE
  ) %>%  
  sf::st_join(sf::st_transform(offshore_codes, 4326)) %>% 
  sf::st_drop_geometry() %>% 
  select(-Subbasin, -HELCOM_ID)

## assign bhi basins to secchi data 
secchi_rgns_shp <- join_rgns_info(
  secchi_rgns_w_offshore, helcomID_col = "helcom_id", country_col = "country", 
  latlon_vars = c("^lat", "^lon"), return_spatial = FALSE, 
  rgn_shps_loc = file.path(dirname(dir_B), "Shapefiles"), 
  buffer_shp = NULL
) 
secchi_rgns <- secchi_rgns_shp %>% 
  ## select summer months (June-September) and Years >= 2000  
  filter(month == "6" | month == "7" | month == "8" | month == "9") %>% 
  filter(year >= 2000) %>% 
  mutate(supplier = stringr::str_to_upper(supplier)) %>% 
  mutate(coastal = ifelse(is.na(helcom_coastal_code), "Coastal", "Offshore"))
```

<br>

**Remove coastal observations**

For the Secchi depth indicator, only offshore data points will be used. 
These are identified using [HELCOM Coastal and Offshore divisions (2018 version)](https://maps.helcom.fi/arcgis/rest/services/MADS/Sea_environmental_monitoring/MapServer/4).

```{r remove coastal data points, results = "hide", echo = TRUE, fig.width = 16, fig.height = 8}
## Select only data with coastal code "0"  
dim(secchi_rgns) # 45479   

## Do any observations have NA for coast_code
secchi_rgns %>% 
  filter(is.na(helcom_coastal_code) & is.na(BHI_ID)) %>% 
  dim() # 4803  

secchi_rgns %>% 
  filter(is.na(helcom_coastal_code) & !is.na(BHI_ID)) %>% 
  dim() # 32661

## 617 observations with no coast_code or BHI_ID
loc_no_coastcode_nobhi <- secchi_rgns %>% 
  filter(is.na(helcom_coastal_code) & is.na(BHI_ID))%>%
  dplyr::select(lat, lon)%>%
  distinct()

## check locations
# library(ggmap)
# map <- get_map(location = c(8.5, 53, 32, 67.5))
# no_coastal_or_bhi_id <- ggmap(map) + 
#   geom_point(aes(x = lon, y = lat), data = loc_no_coastcode_nobhi, size = 2) +
#   labs(x = NULL, y = NULL)
## these locations are very coastal or outside of the Baltic Sea

## observations with NA for the coast_code but have BHI_ID
loc_no_coastcode_bhi <- secchi_rgns %>% 
  filter(is.na(helcom_coastal_code) & !is.na(BHI_ID)) %>% 
  dplyr::select(lat, lon) %>%
  distinct()
# no_coastal_id <- ggmap(map) + 
#   geom_point(aes(x = lon, y = lat), data = loc_no_coastcode_bhi, size = 2, alpha = 0.3) +
#   labs(x = NULL, y = NULL)
## many coastal location

## Filter data that are only offshore, coast_code == 0
secchi_offshore <- filter(secchi_rgns, helcom_coastal_code == 0) 
dim(secchi_offshore) # 8015
# secchi_offshore_spatial <- ggmap(map) + 
#   geom_point(aes(x = lon, y = lat), data = secchi_offshore, colour = "tomato", size = 2, alpha = 0.3) +
#   labs(x = NULL, y = NULL)
## This is a substantial reduction in the number of observations

# gridExtra::grid.arrange(no_coastal_or_bhi_id, no_coastal_id, secchi_offshore_spatial, nrow = 1)
```

<br>

#### 3.1.2 Visualize Secchi Depth Data {-}

**Spatial Visualization of Secchi data**

The four plots generated in this section show temporal trends in offshore Secchi depth (right) and monthly distributions (left) by BHI regions (top) or subbasins (bottom), with data source indicated by color. Pink points correspond to ICES data, and turquoise points to SMHI data. 

```{r secchi timeseries yearly and monthly, results = "show", echo = TRUE, fig.width = 11, fig.height = 15}
plotdf <- left_join(
  secchi_offshore, 
  data.frame(month = c(6,7,8,9), Month = c("Jun", "Jul", "Aug", "Sep"))
)
plotdf$Month <- factor(plotdf$Month, levels = c("Jun", "Jul", "Aug", "Sep"))

## Plot by BHI regions and year
plot_secchi_bhi_ts <- ggplot(secchi_offshore) + 
  geom_point(
    aes(date, secchi_m, color = supplier), 
    size = 1, alpha = 0.3, show.legend = FALSE
  ) +
  facet_wrap(
      c("Subbasin", "rgn_nam"), 
      labeller = label_wrap_gen(width = 35, multi_line = FALSE),
      ncol = 4
  ) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5), 
    strip.text.x = element_text(size = 6)
  ) +
  labs(x = NULL, y = "Secchi depth (m)", color = "Data Source") +
  ggtitle("Yearly Secchi by BHI regions")

## Month boxplots, by BHI region
plot_secchi_bhi_box <- ggplot(plotdf) + 
  geom_boxplot(
    aes(Month, secchi_m, fill = supplier), 
    size = 0.2, outlier.size = 0.3, alpha = 0.6, show.legend = FALSE
  ) +
  facet_wrap(
      c("BHI_ID", "Subbasin", "rgn_nam"), 
      labeller = label_wrap_gen(width = 35, multi_line = FALSE),
      ncol = 4
  ) +
  labs(x = NULL, y = "Secchi depth (m)", fill = "Data Source") +
  theme(strip.text.x = element_text(size = 6)) +
  ggtitle("Monthly Secchi by BHI regions")

## Plot by subbasins and year
plot_secchi_basin_ts <- ggplot(secchi_offshore) + 
  geom_point(
    aes(date, secchi_m, color = supplier),
    size = 1, alpha = 0.3, show.legend = FALSE
  ) +
  # geom_point(aes(year, secchi_m), size = 0.3, alpha = 0.4) +
  facet_wrap(~Subbasin, ncol = 4) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5), 
    strip.text.x = element_text(size = 6)
  ) +
  labs(x = NULL, y = "Secchi depth (m)", color = "Data Source") +
  ggtitle("Yearly Secchi by Subbasins")

## Month boxplots, by Subbasin region
plot_secchi_basin_box <- ggplot(plotdf) + 
  geom_boxplot(
    aes(Month, secchi_m, fill = supplier), 
    size = 0.2, outlier.size = 0.3, alpha = 0.6, show.legend = FALSE
  ) +
  facet_wrap(~Subbasin, ncol = 4) +
  labs(x = NULL, y = "Secchi depth (m)", fill = "Data Source") +
  theme(strip.text.x = element_text(size = 6)) +
  ggtitle("Monthly Secchi by Subbasins")

gridExtra::grid.arrange(
  plot_secchi_basin_ts, plot_secchi_basin_box, 
  plot_secchi_bhi_ts, plot_secchi_bhi_box,
  nrow = 2,
  heights = c(1, 2)
)
```

<br/>

#### 3.1.3 Mean Secchi depth Calculation {-}

Mean Secchi depths are calculated per HELCOM sub-basin or BHI region, using only offshore Secchi depth measurements. Means are first calculated within each summer month, then those monthly means are averaged to get an overall summer mean. Variances (standard deviations squared) are also calculated within each month, then averaged to get an average variance across summer months for the subbasin or BHI region. This is presented as a standard deviation (square root of variance across months) on the plots below.

Along with the mean Secchi calculation, we calculate a five-year moving average for each year in the dataset. The moving average (or 'rolling mean') is an average of summer means from the current and previous four years. 

```{r calculate monthly mean secchi with target, results = "hide"}
for(r in c("Subbasin", "BHI_ID")){
  
  ## Calculate mean monthly value for each summer month
  ## using only offshore data for secchi indicator
  secchi_monthly_means <- secchi_offshore %>% 
    group_by(year, month, Subbasin, !!!syms(r)) %>%
    summarise(
      mean_secchi = mean(secchi_m, na.rm = TRUE),
      secchi_sd = sd(secchi_m, na.rm = TRUE),
      n_pts = n()
    ) %>%
    mutate(secchi_sd = ifelse(is.na(secchi_sd), 0, secchi_sd))
  
  # filter(secchi_monthly_means, is.na(mean_secchi)|is.nan(mean_secchi))
  ## what prop. of data where mean is based on <4 data points?
  # nrow(filter(secchi_monthly_means, n_pts <= 3))/nrow(secchi_monthly_means)
  
  ## Calculate summer mean secchi (subbasin)
  secchi_summer_months_mean <- secchi_monthly_means %>% 
    group_by(year, Subbasin, !!!syms(r)) %>%
    summarise(
      mean_secchi = mean(mean_secchi, na.rm = TRUE),
      secchi_sd = sqrt(mean((secchi_sd^2), na.rm = TRUE)),
      n_pts = sum(n_pts)
      
      # standard_err = secchi_sd/sqrt(n_pts)
    ) 
  
  ## calculate min and max, and rolling 5 year mean
  secchi_summer_months_mean <- secchi_summer_months_mean %>% 
    arrange(!!!syms(r), year) %>% 
    group_by(!!!syms(r)) %>%
    mutate(ma5yr = zoo::rollapply(mean_secchi, 5, mean, na.rm = TRUE, align = "right", fill = NA)) %>% 
    mutate(secchi_min = mean_secchi-secchi_sd, secchi_max = mean_secchi+secchi_sd) %>% 
    ungroup() # in mean calculation some months have NA, ignore for that year's calculation
  
  ## join with basin-specific eutrophication secchi targets
  secchi_summer_months_mean <- eut_thresholds %>% 
    select(Subbasin = basin, target_secchi = summer_secchi) %>% 
    right_join(secchi_summer_months_mean, by = "Subbasin")
  assign(paste0(str_to_lower(r), "_target_secchi"), secchi_summer_months_mean)
  
  ## Plot summer mean secchi
  secchimeansplot <- ggplot(secchi_summer_months_mean, aes(year, mean_secchi, group = 1)) + 
    geom_errorbar(
      aes(x = year, ymin = secchi_min, ymax = secchi_max), 
      alpha = 0.6, size = 0.3, width = 0.3
    ) +
    geom_point(size = 0.3) +
    # geom_line(aes(year, ma5yr), size = 0.2) +
    geom_line(size = 0.2) +
    ## helcom target value
    geom_line(aes(y = target_secchi), colour = "tomato", size = 0.4) +
    facet_wrap(vars(!!!syms(r)), ncol = 5) +
    scale_y_continuous(limits = c(0, 10)) +
    labs(x = NULL, y = "Secchi depth (m)") +
    ggtitle(sprintf("Mean Summer (June-Sept) Secchi by %s", r)) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
  assign(paste0("secchimeansplot", str_to_lower(r)), secchimeansplot)
}
```

<br>

**Plot summer Secchi with target values indicated**

The following shows annual mean summer Secchi depths in meters. Horizontal lines are HELCOM target values, as summarized in the table at the top of section 3.

```{r summer secchi with targets, results = "show", echo = TRUE, fig.width = 9.5, fig.height = 10}
## adding onto/calling plots made in code chunk above
gridExtra::grid.arrange(
  secchimeansplotsubbasin + 
    ggtitle("Mean Summer Secchi by Subbasin (June-Sept, Offshore) against Target Values"),
  secchimeansplotbhi_id + 
    ggtitle("Mean Summer Secchi by BHI Region (June-Sept, Offshore) against Target Values"),
  nrow = 2, 
  heights = c(1,2)
)
```

<br>

#### 3.1.4 Status Calculation {-}

**Explore years in data available for status calculation**

Only five basins have data for 2019, which will be used as the status year, nine basins have data for 2018 and three basins for 2017 and for these basins, their most recent year's data is used for current status calculations.

```{r what is the last year in secchi subbasin means dataframe, results = "show"}
## get the last year of non-NA data
secchi_last_year <- subbasin_target_secchi %>%
  filter(!is.na(ma5yr)) %>%
  group_by(Subbasin)%>%
  summarise(last_year = last(year))

knitr::kable(
  secchi_last_year %>% 
    group_by(last_year) %>%
    summarize(`Num. Subbasins with Year` = n(), Subbasins = paste(Subbasin, collapse = ", ")) %>% 
    rename(`Last Year ` = last_year)
)
## which are not in 2019
# filter(secchi_last_year, last_year < 2019)
```
<br>

**Status calculation with raw (non-modeled) mean summer Secchi by basin** 

Status is calculated on the basin level and then applied to all regions. It is calculated here for additional visualization and evaluation, but the final method of status calculation is also completed by the EUT function in `functions.R` of the relevant bhi repository assessment folder.

Status for the Secchi indicator is calculated as `basin summer mean` divided by `basin target`, with the value capped at 1. The plot shows BHI region scores (uniform across regions within the same basin, as the data is aggregated by basin for the summer means calculation), and also indicates which data year the status is based on.

```{r constants for eut secchi status and trend calculations, echo = TRUE}
## Define constants for status and trend calculations
min_year <- 2009 # earliest year to use as a start for regr_length timeseries
regr_length <- 10 # number of years to use for regression
future_year <- 5 # the year at which we want the likely future status
min_regr_length <- 5 # min actual number of years with data to use for regression
```

```{r eut secchi indicator status calculation, results = "show", echo = TRUE, fig.width = 9.5, fig.height = 5.5}
## Calculate basin status as: basin_mean/basin_target, constraining values to 1 as maximum score
secchi_status_allyrs <- subbasin_target_secchi %>%
  filter(year >= min_year) %>% 
  rename(subbasin = Subbasin) %>%
  # mutate(status = pmin(1, mean_secchi/target_secchi)) 
  ## or use five year rolling means (moving average) rather than annual mean?
  mutate(status = pmin(1, ma5yr/target_secchi))
secchi_status <- group_by(secchi_status_allyrs, subbasin)

## Alternative status calculation: Status calculated using BHI region means
# secchi_status_allyrs <- bhi_id_target_secchi %>%
#   rename(subbasin = Subbasin, bhi_id = BHI_ID) %>%
#   mutate(status = pmin(1, mean_secchi/target_secchi))
# secchi_status <- group_by(secchi_status_allyrs, bhi_id)

## select last year of data for status in each basin or bhi region
## this means status year differs by basin!
## joining to basin lookup expands from subbasins to one row per BHI region
secchi_status <- secchi_status %>%
  filter(year == max(year)) %>%
  mutate(status = round(status*100, 2)) %>% # keep decimal places, will round when combine eut indicators
  ungroup() %>%
  dplyr::select(score = status, !!!syms(intersect(c("subbasin", "bhi_id"), names(secchi_status)))) %>% 
  mutate(dimension = "status", indicator = "secchi") %>% 
  left_join(basin_lookup) %>%
  mutate(region = as.factor(rgn_nam))

## plot the status
## BHI region plots will look the same, as BHI status are taken from subbasin means
## unless alternative calculation is used, based on BHI region means rather than subbasin means
set.seed(2)
statuspal <- colorRampPalette(fullpal)(80)[sample(1:80, size = 18)]

ggplot(secchi_status %>% left_join(secchi_last_year, by = c("subbasin" = "Subbasin"))) +
  geom_col(
    aes(region, score, fill = subbasin),
    position = position_dodge(), 
    color = "grey", 
    alpha = 0.6,
    size = 0.2,
    show.legend = FALSE
  ) + 
  geom_text(aes(region, score, label = last_year), color = "slategray", size = 3) +
  coord_flip() +
  labs(x = NULL, y = NULL) +
  theme_bw() +
  theme(axis.text.y = element_text(size = 9), plot.caption = element_text(hjust = 0)) +
  scale_fill_manual(values = statuspal) +
  labs(
    title = "Secchi Indicator Status by Region",
    caption = paste(
      "Summer mean Secchi Depth divided by basin target,",
      "scaled between 0-100 via multiplication by 100.\n",
      "The status score reflects the state of the goal in the Sub-basin,",
      "between worst-possible and 'Good Environmental Status'.\n",
      "A status of 100 indicates target (GES) is achieved."
    )
  )
```
<br>

#### 3.1.5 Trend Calculation {-}

In our approach here, we use a 10 year period to calculate the trend with a minimum of 5 data points. In most cases, the BHI framework uses a 5 year period for the trend, but as Secchi is a slow response variable, we use a longer time period. The trend value is the slope of the linear regression (mean offshore summer Secchi depth per year) multiplied by the year of future interest (5 years from status year) and this value is constrained to between -1 and 1.

The trend plots show Secchi depth measurements by basin, with a red line indicating where the linear model predicts mean Secchi depths will be five years in the future, and a blue line indicating current assessment year summer mean. Vertical lines frame the 5-year time frame over which trend is projected.

```{r eut secchi indicator trend calculation, results = "hide", echo = TRUE, fig.width = 9.5}
trendyrs <- min_year:(min_year + regr_length)

## linear model from the basin mean secchi values ----

secchi_lm_estim <- secchi_status_allyrs %>%
  distinct(subbasin, year, mean_secchi) %>%
  filter(year %in% trendyrs) %>% 
  group_by(subbasin) %>%
  ## calculate trend only if:
  ## have at least X yrs data (min_regr_length) in the last Y yrs (regr_length) of time series
  do(tail(., n = regr_length)) %>%
  do(data.frame(enough_data = ifelse(sum(!is.na(.$mean_secchi)) >= min_regr_length, TRUE, FALSE))) %>%
  right_join(secchi_status_allyrs %>% filter(year %in% trendyrs), by  = "subbasin") %>%
  group_by(subbasin, enough_data, target_secchi) %>%
  do(trend_mdl = lm(mean_secchi ~ year, data = .))

## trend plots ----

trendplotdf <- secchi_offshore %>%
  rename(subbasin = Subbasin) %>%
  left_join(secchi_status_allyrs) %>% 
  filter(year %in% trendyrs) %>% 
  left_join(
    mutate(
      secchi_lm_estim,
      current = predict(trend_mdl, data.frame(year = max(trendyrs))),
      prj = predict(trend_mdl, data.frame(year = max(trendyrs) + future_year))
    )
  )
ggplot(trendplotdf) +
  geom_point(aes(x = year, y = secchi_m), size = 0.2, alpha = 0.2) +
  geom_point(aes(x = year, y = mean_secchi), size = 0.6) +
  geom_vline(xintercept = max(trendyrs), size = 0.2) +
  geom_vline(xintercept = max(trendyrs) + future_year, size = 0.2) +
  geom_hline(aes(yintercept = prj), color = "tomato") +
  geom_hline(aes(yintercept = current), color = "lightsteelblue") +
  facet_wrap(~subbasin) +
  labs(x = NULL, y = NULL) +
  ggtitle("Offshore Secchi with 5 Years in Future and Current Estimates (Red and Blue lines Respectively)")

## trend score dataframe ----

## assign basin trend to BHI regions
secchi_trend <- secchi_lm_estim %>% 
    mutate(
      expectedchange5yrs = coef(trend_mdl)["year"]*future_year,
      ## divide expected 5 year change in secchi by respective basin targets
      ## interpretation of trend is roughly expected (percent) change in status
      trend_score = max(-1, min(1, expectedchange5yrs/target_secchi))
    ) %>%
  right_join(basin_lookup, by = "subbasin") %>% 
  mutate(
    score = round(ifelse(enough_data, trend_score, NA), 3),
    region = rgn_nam,
    dimension = "trend",
    indicator = "secchi"
  ) %>% 
  select(score, subbasin, dimension, indicator, bhi_id, rgn_nam, HELCOM_ID, region)
```
<br>

#### 3.1.5 Status and Trend Timeseries Plots {-}

**Basin status**

As a ratio between summer mean and basin target, the sub-basin status is initially a value between 0 and 1. This is later rescaled to the 0 to 100 range via multiplication by 100. The status is calculated for each (data) year between 2000 and 2019.

```{r timeseries plot of basin eut secchi status, results = "show", echo = TRUE, fig.width = 9.5}
plotdf <- secchi_status_allyrs %>% 
  distinct(subbasin, year, status, n_pts) %>% 
  mutate(year = as.factor(year))
ggplot(plotdf) + 
  geom_col(aes(year, status, fill = n_pts), alpha = 0.7) +
  facet_wrap(~subbasin) +
  scale_fill_viridis_c() +
  geom_hline(yintercept = 1) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(x = NULL, y = "Secchi Indicator Status\n", fill = "Num. Contributing of Points") +
  ggtitle("Yearly Secchi Status by Subbasin")
```

<br>

**Plot region Status and Trend values**

Status values can range from 0-100 -- this is the status for the *most recent* set of 5 years. In most cases this is 2015-2019.

```{r secchi status and trend for regions plot, results = "show", echo = TRUE, fig.width = 9.5, fig.height = 5.5}
ggplot(rbind(secchi_status, secchi_trend)) + 
  geom_col(aes(region, score, fill = subbasin), alpha = 0.5, show.legend = FALSE) +
  coord_flip() +
  facet_wrap(~dimension, scales = "free_x", nrow = 1) +
  labs(x = "BHI Region\n", y = NULL) +
  ggtitle("Secchi Status and Trend Scores by Region") +
  scale_fill_manual(values = statuspal) +
  theme(axis.text.y = element_text(size = 7))
```

<br>

#### 3.1.6 Save Secchi depth eutrophication layer {-}

```{r save secchi eut layer, eval = FALSE}
write_csv(
  secchi_offshore %>% 
    select(
      region_id = BHI_ID,
      lat, lon, 
      month, year,
      secchi_depth = secchi_m
    ), 
  file.path(dir_layers, sprintf("cw_eut_secchi_bhi%s.csv", assess_year))
)
```

<br>

### 3.2 Chlorophyll a Indicator {-}

#### 3.2.1 Match BHI regions {-}

The raw chlorophyll a data (consisting of merged ICES, SMHI, and Baltic Nest data) must be assigned to their respective Sub-basins (along with HELCOM Coastal WFD /Offshore division codes) and BHI Regions based on their latitude/longitude coordinates. Then the data are subsetted to assure only summer months (June-September), years 2000 onward, and measurements from the surface layer (0 - 10 m depth) are included.

```{r assign BHI regions to Chl a data, echo = TRUE, message = FALSE, warning = FALSE}
## assign helcom offshore codes
# nrow(filter(chla_rawdata, is.na(lat) | is.na(lon))) # first check if any lat/lon have NAs
chla_rgns_w_offshore_code <- chla_rawdata %>% 
  filter(!is.na(lat) & !is.na(lon)) %>%
  sf::st_as_sf(
    crs = 4326,
    coords = c("lon", "lat"),
    remove = FALSE
  ) %>%  
  sf::st_join(sf::st_transform(offshore_codes, 4326)) %>% 
  select(-HELCOM_ID, -Subbasin)

## spatial join also the WFD water bodies
chla_rgns_w_offshore_code <- chla_rgns_w_offshore_code %>% 
  sf::st_join(
    coastal_codes %>% 
      sf::st_transform(4326) %>% 
      rename(area_km2_wfd = Area_km2, code = HELCOM_ID) %>% 
      select(area_km2_wfd, code)
  ) %>% 
  sf::st_drop_geometry() %>% 
  filter(!is.na(lat) & !is.na(lon))
  

## assign bhi basins to chla data 
chla_rgns_shp <- join_rgns_info(
  chla_rgns_w_offshore_code, helcomID_col = "helcom_id", country_col = "country", 
  latlon_vars = c("^lat", "^lon"), return_spatial = FALSE, 
  rgn_shps_loc = file.path(dirname(dir_B), "Shapefile"), buffer_shp = NULL
)

## select summer months (June-September), surface layer (0 - 10 m depth), Years >= 2000
chla_rgns <- chla_rgns_shp %>% 
  mutate(month = stringr::str_pad(month, 2, "left", "0")) %>% 
  filter(month == "06" | month == "07" | month == "08" | month == "09") %>% 
  filter(depth_m <= 10) %>% 
  filter(year >= 2000) %>% 
  mutate(supplier = stringr::str_to_upper(supplier)) %>% 
  mutate(coastal = ifelse(is.na(helcom_coastal_code), "Coastal", "Offshore")) %>% 
  ## some cases where code SEA-010 is matched with Northern Baltic Proper, 
  ## by our spatial join should be code SEA-012,
  ## happens because our polygons are different than HELCOM's
  mutate(code = as.character(code)) %>% 
  mutate(area_km2_wfd = ifelse(
    code == "SEA-010" & Subbasin == "Northern Baltic Proper", 
    unique(filter(chla_rgns_shp, code == "SEA-012")$area_km2_wfd), 
    area_km2_wfd
  )) %>% 
  mutate(code = ifelse(code == "SEA-010" & Subbasin == "Northern Baltic Proper", "SEA-012", code))

## Gdansk and The Quark have only coastal, no offshore data points
distinct(chla_rgns, Subbasin, HELCOM_ID, coastal) %>% 
  group_by(HELCOM_ID) %>% 
  mutate(n = n()) %>% 
  arrange(n, Subbasin)

## not all coastal regions have data... 76 of 257 do not have data
length(levels(coastal_codes$HELCOM_ID))
coastalareas_nodata <- setdiff(unique(coastal_codes$HELCOM_ID), as.character(unique(chla_rgns$code)))
length(coastalareas_nodata)

## 17 of 257 are offshore, 240 are coastal
length(levels(coastal_codes$HELCOM_ID)[grep("SEA-[0-9]{3}", levels(coastal_codes$HELCOM_ID))])

## 224 of the 240 coastal areas have corresponding chla coastal thresholds, 16 do not
length(unique(chla_coastal_thresholds$code))
coastal_nothresh <- setdiff(
  levels(coastal_codes$HELCOM_ID)[grep("SEA-[0-9]{3}", levels(coastal_codes$HELCOM_ID), invert = TRUE)], 
  unique(chla_coastal_thresholds$code)
)

## coastal areas with no data, but which do have threshold
coastalareas_nodata[!coastalareas_nodata %in% coastal_nothresh]

## coastal areas with data but no thresholds
levels(coastal_codes$HELCOM_ID)[levels(coastal_codes$HELCOM_ID) %in% coastal_nothresh]

```

<br>

**Checking coastal and offshore observations**

For the chlorophyll a indicator, coastal and offshore data points must be associated with different thresholds. Coastal versus offshore points are distinguished using [HELCOM Coastal and Offshore divisions (2018 version)](https://maps.helcom.fi/arcgis/rest/services/MADS/Sea_environmental_monitoring/MapServer/4), and each category (coastal and offshore) gives important information regarding the status of Eutrophication in the Baltic Sea. However, not every coastal area has sufficient data and/or a threshold value, and therefore coastal data must be excluded at this time from our analysis.

```{r basemap eut prep to be used in spatial plotting, echo = TRUE}
## make basemap
basemap <- ggplot2::ggplot(rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")) + 
  geom_sf(size = 0.1, color = "burlywood", alpha = 0.4) +
  scale_x_continuous(limit = c(4, 32)) +
  scale_y_continuous(limit = c(53.5, 66)) +
  theme(panel.background = element_rect(fill = "#F8FBFC", color = "#E2EEF3"))
```

```{r check chla data spatial info, results = "hide"}
pal <- colorRampPalette(RColorBrewer::brewer.pal(12, "Set3"))(257)[sample(1:257)]
coastal_simple <- rmapshaper::ms_simplify(input = coastal_codes) %>% sf::st_as_sf()
## note: DEN-034, DEN-035, DEN-040 are eliminated by simplification of polygons for plotting
## coastal_simple cannot be used for analysis; is only used for plotting

## coastal regions which are not accounted for i.e. no data matched to these
coastalareas_nodata_map <- basemap +
  geom_sf(
    mapping = aes(fill = Name),
    data = filter(coastal_simple, HELCOM_ID %in% coastalareas_nodata),
    color = "snow", size = 0.05,
    show.legend = FALSE
  ) +
  scale_fill_manual(values = pal)


## checking chla_rgns observations, coastal/offshore locations
nrow(chla_rgns) # 38386

## observations with no coast_code and no BHI_ID
## distinct locations
loc_no_coastcode_nobhi <- chla_rgns %>% 
  filter(is.na(helcom_coastal_code) & is.na(BHI_ID))%>%
  dplyr::select(lat, lon)%>%
  distinct() # 588

## these locations are only coastal or outside of the Baltic Sea
coastal_noid <- basemap +
  geom_sf(
    data = sf::st_as_sf(
      loc_no_coastcode_nobhi, 
      coords = c("lon", "lat"), 
      crs = 4326
    ),
    size = 1.8, alpha = 0.2
  )

## any coastal observations have NA for bhi id / how many with non-NA bhi id
nrow(filter(chla_rgns, coastal == "Coastal" & is.na(BHI_ID))) # 9658
nrow(filter(chla_rgns, coastal == "Coastal" & !is.na(BHI_ID))) # 23636


## observations with NA for the coast_code but having BHI_ID
## distinct locations
loc_bhi_no_coastcode <- chla_rgns %>% 
  filter(is.na(helcom_coastal_code) & !is.na(BHI_ID)) %>% 
  dplyr::select(lat, lon) %>%
  distinct() # 1506

## no coastal code, all coastal stations
coastal_with_id <- basemap +
  geom_sf(
    data = sf::st_as_sf(
      loc_bhi_no_coastcode, 
      coords = c("lon", "lat"), 
      crs = 4326
    ),
    size = 1.8, alpha = 0.2
  )

## filter data that are only offshore, coast_code == 0 and having BHI_ID
chla_offshore <- filter(chla_rgns, helcom_coastal_code == 0, !is.na(BHI_ID))
nrow(chla_offshore) # 5087
chla_offshore_spatial <- basemap +
  geom_sf(
    mapping = aes(colour = supplier),
    data = sf::st_as_sf(
      chla_offshore, 
      coords = c("lon", "lat"), 
      crs = 4326
    ),
    size = 1.8, alpha = 0.2
  ) +
  labs(x = NULL, y = NULL, colour = "Data Source") +
  theme(legend.position = c(0.9, 0.1)) +
  scale_color_manual(values = c("tomato", "aquamarine2", "gold"))
## taking only offshore data is a substantial reduction in the number of observations, 
## especially in the northern Baltic 


## Check coastal observations
chla_coastal <- filter(chla_rgns, is.na(helcom_coastal_code) & !is.na(BHI_ID))
nrow(chla_coastal) # 23636
chla_coast_spatial <- basemap +
  geom_sf(
    mapping = aes(colour = supplier),
    data = sf::st_as_sf(
      chla_coastal, 
      coords = c("lon", "lat"), 
      crs = 4326
    ),
    size = 1.8, alpha = 0.2
  ) +
  labs(x = NULL, y = NULL, colour = "Data Source") +
  theme(legend.position = c(0.9, 0.1)) +
  scale_color_manual(values = c("tomato", "aquamarine2", "gold"))

## Look at the complete dataset, including both coastal and offshore locations
chla_whole <- filter(chla_rgns, !is.na(BHI_ID))
nrow(chla_whole) # 28723

plot_chla_whole <- sf::st_as_sf(
  mutate(chla_whole, code_na = is.na(code)),
  coords = c("lon", "lat"), 
  crs = 4326
)

chla_whole_spatial_codeareas <- basemap +
  geom_sf(
    mapping = aes(fill = Name),
    data = coastal_simple,
    alpha = 0.7, color = "snow", size = 0.05,
    show.legend = FALSE
  ) +
  geom_sf(
    mapping = aes(shape = code_na),
    data = filter(plot_chla_whole, code_na),
    size = 4, shape = 3,
    show.legend = FALSE
  ) +
  scale_fill_manual(values = pal)
## missing codes for a few points in the inner coastal areas
## estuarine areas near Göteborg, Västervik, Rostock, Haapsalu, Luleå

chla_missing_data_or_targets <- basemap +
  ## areas with no data
  geom_sf(
    mapping = aes(fill = Name),
    data = filter(coastal_simple, HELCOM_ID %in% coastalareas_nodata),
    alpha = 0.7, color = "snow", size = 0.05,
    show.legend = FALSE
  ) +
  ## areas with no threshold
  geom_sf(
    data = filter(coastal_simple, HELCOM_ID %in% coastal_nothresh),
    fill = NA, color = "black", size = 0.3, 
    show.legend = FALSE
  ) +
  ## overlay data points locations
  # geom_sf(
  #   data = plot_chla_whole,
  #   size = 0.01, alpha = 0.5, color = "red"
  # ) +
  labs(x = NULL, y = NULL) +
  theme(legend.position = c(0.9, 0.1)) +
  scale_fill_manual(values = pal)
```

```{r chla offshore coastal data maps, results = "show", echo = FALSE, fig.width = 12, fig.height = 10}
gridExtra::grid.arrange(
  chla_offshore_spatial + ggtitle("Offshore Chl a. Data"), 
  chla_coast_spatial + ggtitle("Coastal Chl a. Data"), 
  chla_whole_spatial_codeareas + ggtitle("Chl a. Data missing Coastal Code, overlying WFD regions"),
  chla_missing_data_or_targets + ggtitle("No Data (filled polygons) and/or No Threshold (outlined polygons)"),
  nrow = 2
)

# Comment from Vivi: Strange how little offshore ICES chla, why is it completely missing in the north? It looks also that NEST data is missing some countries (Eg. Finland /SYKE) which should be there.
```

<br/>

#### 3.2.2 Visualize Chrorophyll a Data {-}

**Visualization of Chl a data (whole Baltic i.e. both Coastal and Offshore)**

The four plots generated in this section show temporal trends (top) and monthly distributions (bottom) by subbasin, for Offshore (right) and Coastal (left) chlorophyll a, with data source indicated by color. Pink points correspond to ICES data, turquoise points to Nest data, and gold to SMHI data. 

```{r chl a timeseries yearly and monthly, results = "hide", echo = TRUE}
plotinfo <- list(
  list("chla_offshore", c("BHI_ID", "Subbasin", "rgn_nam"), "Chlorophyll a (Offshore) by BHI Region", "bhi"),
  list("chla_offshore", "Subbasin", "Chlorophyll a (Offshore) by Subbasin", "basin")
)

## expand cases for equal number of offshore and coastal plots, so visual comparison is easier 
expandcases <- tidyr::expand(chla_whole, tidyr::nesting(Subbasin, BHI_ID, year, month, coastal))
lapply(
  plotinfo,
  function(x){
    ## 1. evaluate data by source
    plottimeseries <- ggplot(full_join(get(x[[1]]), expandcases)) +
      geom_point(
        aes(year, chla_ug_l, color = supplier), 
        size = 0.5, alpha = 0.5, show.legend = FALSE
      ) +
      labs(x = "Year", y = "Chl a (ug/L)", color = "Data Source", title = x[[3]]) +
      scale_color_manual(values = c("tomato", "aquamarine2", "gold"))
    
    ## 2. evaluate data by by coastal/offshore
    # plottimeseries <- ggplot(chla_whole) + 
    #   geom_point(aes(year, chla_ug_l, color = coastal), size = 0.5, alpha = 0.5) +
    #   labs(x = "Year", y = "Chl a (ug/L)", color = "Coastal/Offshore", title = x[[3]]) +
    #   scale_color_manual(values = c("lightsteelblue", "sandybrown")) 
    
    plottimeseries <- plottimeseries + 
      facet_wrap(
        x[[2]], 
        labeller = label_wrap_gen(width = 35, multi_line = FALSE), 
        ncol = 3
      ) 
    assign(sprintf("plot_%s_%s_ts", x[[1]], x[[4]]), plottimeseries, envir = .GlobalEnv)
    createdplots <- sprintf("plot_%s_%s_ts", x[[1]], x[[4]])
  }
)
lapply(
  plotinfo,
  function(x){
    ## 1. evaluate data by source
    plotdf <- left_join(
      get(x[[1]]) %>% 
        full_join(expandcases) %>% 
        mutate(month = as.factor(month)),
      data.frame(month = c("06","07","08","09"), Month = c("Jun","Jul","Aug","Sep"))
    )
    plotdf$Month <- factor(plotdf$Month, levels = c("Jun","Jul","Aug","Sep"))
    
    plotboxplot <- ggplot(plotdf) +
      geom_boxplot(
        aes(Month, chla_ug_l, fill = supplier), 
        size = 0.2, outlier.size = 0.3, alpha = 0.4, show.legend = FALSE
      ) +
      labs(x = NULL, y = "Chl a (ug/L)", fill = "Data Source", title = x[[3]]) +
      scale_fill_manual(values = c("tomato", "aquamarine2", "gold"))
    
    ## 2. evaluate data by by coastal/offshore
    # plotboxplot <- ggplot(mutate(chla_whole, month = as.factor(month))) +
    #   geom_boxplot(aes(month, chla_ug_l, fill = coastal), size = 0.2, outlier.size = 0.3, alpha = 0.4) +
    #   labs(x = "Month", y = "Chl a (ug/L)", fill = "Coastal/Offshore", title = x[[3]]) +
    #   scale_fill_manual(values = c("lightsteelblue", "sandybrown")) + 
    #   ggtitle("Chlorophyll a Monthly")
    
    plotboxplot <- plotboxplot +
      facet_wrap(
        x[[2]], 
        labeller = label_wrap_gen(width = 35, multi_line = FALSE), 
        ncol = 3
      )
    assign(sprintf("plot_%s_%s_box", x[[1]], x[[4]]), plotboxplot, envir = .GlobalEnv)
    createdplots <- sprintf("plot_%s_%s_box", x[[1]], x[[4]])
  }
)
```

```{r coastal and offshore chla plots by basin, results = "show", echo = TRUE, fig.width = 12, fig.height = 7}
# gridExtra::grid.arrange(
#   plot_chla_offshore_bhi_ts, plot_chla_coastal_bhi_ts,
#   plot_chla_offshore_bhi_box, plot_chla_coastal_bhi_box, 
#   nrow = 2
# )
gridExtra::grid.arrange(
  plot_chla_offshore_basin_ts,
  plot_chla_offshore_basin_box,
  nrow = 1
)
```

<br>

#### 3.2.3 Mean Chlorophyll a Calculation

_Chlorophyll a_ concentration averages are calculated per basin or BHI region, using offshore measurements (coastal measurements were added in a separate category just for the analyses). Means are calculated for each coastal and offshore categories, first within each summer month. Then the monthly means are averaged across to get an overall summer mean in each the coastal and offshore category. Variances (standard deviations squared) are also calculated within each month, then averaged to get an average variance across summer months for the subbasin or BHI region and coastal/offshore category. This is presented as a standard deviation (square root of variance across months) on the plots below.

Along with the mean calculation, we calculate five-year moving averages. The moving average is an average of summer means from the current and previous four years. 

```{r calculate monthly and summer mean chl a with target, results = "hide"}
## Calculate mean monthly value for each summer month
## For coastal and offshore separately
chla_monthly_means <- chla_whole %>% 
  filter(!is.na(code)) %>% 
  ## now group and take means
  group_by(year, month, coastal, Subbasin, code, area_km2_wfd) %>%
  summarise(
    mean_chla = mean(chla_ug_l, na.rm = TRUE),
    chla_sd = sd(chla_ug_l, na.rm = TRUE),
    n_pts = n()
  ) %>% 
  mutate(chla_sd = ifelse(is.na(chla_sd), 0, chla_sd))

## Calculate summer mean chl a
## each year will have multiple means, for coastal areas and an offshore (open-sea) subbasin
## https://stats.stackexchange.com/questions/25848/how-to-sum-a-standard-deviation
chla_summer_months_mean <- chla_monthly_means %>% 
  group_by(year, coastal, Subbasin, code, area_km2_wfd) %>%
  summarise(
    mean_chla = mean(mean_chla, na.rm = TRUE), 
    chla_sd = sqrt(mean((chla_sd^2), na.rm = TRUE)),
    n_pts = sum(n_pts)
  ) %>% 
  ungroup() %>% 
  mutate(code = as.character(code))

## calculate min and max, and rolling 5 year mean
chla_summer_months_mean <- chla_summer_months_mean %>% 
  arrange(code, coastal, year) %>% 
  group_by(code, Subbasin, coastal) %>%
  mutate(ma5yr = zoo::rollapply(mean_chla, 5, mean, na.rm = TRUE, align = "right", fill = NA)) %>% 
  mutate(chla_min = mean_chla-chla_sd, chla_max = mean_chla+chla_sd)


## DIFFERENT TARGETS FOR COASTAL (WFD AREAS) AND OFFSHORE
wfd_area_target_chla <- file.path(dir_prep, "supplement", "lookup_tabs", "rgns_complete.csv") %>% 
  read_csv() %>% 
  dplyr::select(basin = subbasin, code = helcom_id) %>% 
  distinct() %>% 
  left_join(eut_thresholds, by = "basin") %>% 
  dplyr::select(code, target_chla = summer_chla) %>% 
  mutate(units = "ug_per_l") %>% 
  bind_rows(rename(chla_coastal_thresholds, target_chla = threshold)) %>% 
  ## join with basin-specific eutrophication chlorophyll a targets
  right_join(chla_summer_months_mean, by = "code")


## Plot summer mean secchi
chla_meansplot <- wfd_area_target_chla %>% 
  ggplot(aes(year, mean_chla, group = code)) +  
  geom_errorbar(
    aes(x = year, ymin = chla_min, ymax = chla_max), 
    alpha = 0.6, size = 0.3, color = "snow", width = 0.3,
    show.legend = FALSE
  ) +
  geom_point(size = 0.3, color = "snow", show.legend = FALSE) +
  geom_line(aes(year, ma5yr), size = 0.2, color = "snow", show.legend = FALSE) + 
  geom_point(
    data = filter(wfd_area_target_chla, str_detect(code, "^SEA-")),
    size = 0.7
  ) +
  geom_line(
    data = filter(wfd_area_target_chla, str_detect(code, "^SEA-")),
    aes(year, ma5yr), 
    size = 0.3
  ) + 
  facet_wrap(~Subbasin, scales = "free", ncol = 4) +
  scale_y_continuous(limits = c(0, NA)) +
  labs(x = NULL, y = "Chlorophyll a (ug/L)") +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5),
    strip.text.y = element_text(size = 6)
  ) +
  theme_dark() 
```
<br>

**Plot summer chlorophyll a with target values indicated**

The following plot shows annual mean summer chlorophyll a concentrations (ug/l). Horizontal lines are HELCOM target values, as summarized in the table at the top of section 3. With one row per subbasin, the three columns show Coastal, Offshore, and Combined summer mean chlorophyll a concentrations. 

```{r summer chl a with target, results = "show", echo = TRUE, fig.width = 9.5, fig.height = 8}
chla_meansplot + ggtitle(sprintf("Mean Summer (June-Sept) Chlorophyll a by WFD Area, organized by Subbasin"))
```

<br>


#### 3.2.3 Status Calculations: Coastal & Offshore {-}

**Explore years in data available for status calculation**

Only two basins have data for 2019, which will be used as the status year, while thirteen have data for 2018 and two for 2016. For these basins, their most recent year's data is used for current status calculations.

We use the summer mean rather than the 5 year moving average because, when excluding coastal data, the most recent year for Kiel Bay and The Sound would become 2010, and the following would have no values in our compiled dataset to base status calculation on: Bothnian Bay, Bothnian Sea, The Quark, Aland Sea, Gdansk Basin.

Using annual summer means and only offshore data however, we are still missing The Quark and Gdansk Basin.

```{r what is the last year for chla, results = "show"}
## get the last year of non-NA data
chla_last_year <- wfd_area_target_chla %>%
  filter(!is.na(mean_chla), str_detect(code, "^SEA-")) %>%
  group_by(code, Subbasin) %>%
  summarise(last_year = last(year))

## which are not in 2019
# filter(chla_last_year, last_year < 2019)
```

<br>

**Status calculation with mean summer Chla by basin, from in situ measurements** 

Status is calculated on the basin level and then applied to all regions. It is calculated in data prep for additional visualization and evaluation, but the final method of status calculation is also completed by the EUT function in functions.R of the relevant `bhi` repository assessment folder.

Status for the _chlorophyll a_ indicator is calculated as `basin target` divided by `basin summer means`, with the value capped at 1. We calculate status for each Coastal, Offshore, and Combined categories for evaluation, but select the Combined version for the indicator status. The plot shows BHI region scores (uniform across regions within the same basin, as the data is aggregated by basin for the summer means calculation), and also indicates which data year the status is based on.

```{r constants for eut chla status and trend calculations, echo = TRUE}
## Define constants for status and trend calculations
min_year <- 2009 # earliest year to use as a start for regr_length timeseries
regr_length <- 10 # number of years to use for regression
future_year <- 5 # the year at which we want the likely future status
min_regr_length <- 5 # min actual number of years with data to use for regression
```

```{r eut chla status calculation, results = "show", echo = TRUE, fig.width = 9.5, fig.height = 5.5}
## calculate for each offshore and coastal separately, then combine
## quick plot, distributions of means
ggplot(wfd_area_target_chla) + 
  geom_histogram(aes(mean_chla), bins = 80, fill = NA, color = "grey") + 
  facet_grid(cols = vars(coastal), scales = "free_x") + 
  labs(x = "Mean Chlorphyll a Concentration", y = NULL)

## Calculate basin status as: basin_target/basin_mean, constraining values to 1 as maximum score
## ratio inverted compared to secchi because levels of chl a above target are undesireable
chla_status_allyrs <- wfd_area_target_chla %>%
  rename(subbasin = Subbasin) %>% 
  mutate(status = pmin(1, target_chla/mean_chla))
  ## or use five year rolling means (moving average) rather than annual mean?
  # mutate(status = pmin(1, target_chla/ma5yr))
chla_status <- group_by(chla_status_allyrs, coastal, subbasin, code)

## select last year of data for status in each basin, this means status year differs by basin
## joining to basin lookup expands from subbasins to one row per BHI region
chla_status <- chla_status %>% 
  filter(year == max(year)) %>% 
  mutate(status = round(status*100, 2)) %>% # keep decimal places, will round when combine eut indicators
  ungroup() %>% 
  dplyr::select(score = status, code, coastal, subbasin) %>% 
  mutate(dimension = "status", indicator = "chla") %>% 
  left_join(basin_lookup) %>%
  mutate(region = as.factor(rgn_nam))

## plot the status
## BHI region plots will look the same, as BHI status are taken from subbasin means
## unless alternative calculation is used, based on BHI region means rather than subbasin means
set.seed(2)
statuspal <- colorRampPalette(fullpal)(80)[sample(1:80, size = 18)]

chlaplotdf <- chla_status %>% 
  filter(coastal == "Offshore") %>% 
  left_join(chla_last_year, by = c("subbasin" = "Subbasin", "code"))

ggplot(chlaplotdf) + 
  geom_col(
    aes(region, score, fill = subbasin),
    position = position_dodge(), 
    color = "grey", 
    alpha = 0.6,
    size = 0.2,
    show.legend = FALSE
  ) + 
  geom_text(aes(region, score, label = last_year), color = "slategray", size = 3) +
  coord_flip() +
  facet_grid(cols = vars(coastal)) +
  labs(x = NULL, y = NULL) +
  theme_bw() +
  theme(axis.text.y = element_text(size = 9)) +
  scale_fill_manual(values = statuspal) +
  ggtitle("Chlorophyll a Status by Region, for Coastal, Offshore, and Combined data")

chla_status <- chla_status %>% 
  filter(coastal == "Offshore") %>% 
  select(-coastal, -code)
```

#### 3.2.4 Trend Calculations: Coastal & Offshore {-}

In our approach here, we use a 10 year period to calculate the trend with a minimum of 5 data points. In most cases, the BHI framework uses a 5 year period for the trend, but as _chlorophyll a_ is a slower response variable, we use a longer time period. The trend value is the slope of the linear regression (mean summer chlorophyll vs year) multiplied by the year of future interest (5 years from status year) and this value is constrained to between -1 and 1.

The linear regressions look very different when combining offshore and coastal categories of data, e.g. for Western Gotland both offshore and coastal categories separately have slight positive trends (increasing chlorophyll a), but a linear regression using both offshore/coastal combined would have a negative trend (decreasing chlorophyll a). For this reason the _chlorophyll a_ indicator trend dimension is calculated as an average of trend scores calculated for coastal and offshore chlorophyll a.

The trend plots show chlorophyll measurements by basin, with a red line indicating where the linear model predicts mean Secchi depths will be five years in the future, and a blue line indicating current assessment year summer mean. Vertical lines frame the five-year time frame over which the trend is projected.

```{r eut chl a trend calculation, results = "show", warning = FALSE, fig.width = 9, fig.height = 10}
trendyrs <- min_year:(min_year + regr_length)

## linear model from the basin mean chla values ----

chla_lm_estim <- chla_status_allyrs %>%
  filter(year %in% trendyrs) %>% 
  distinct(subbasin, code, coastal, year, mean_chla) %>%
  group_by(subbasin, code, coastal) %>%
  ## calculate trend only if:
  ## have at least X yrs data (min_regr_length) in the last Y yrs (regr_length) of time series
  do(tail(., n = regr_length)) %>%
  do(data.frame(enough_data = ifelse(sum(!is.na(.$mean_chla)) >= min_regr_length, TRUE, FALSE))) %>%
  right_join(chla_status_allyrs %>% filter(year %in% trendyrs), by = c("subbasin", "code", "coastal")) %>%
  group_by(subbasin, code, coastal, enough_data, target_chla) %>%
  do(trend_mdl = lm(mean_chla ~ year, data = .))

## trend plots ----

trendplotdf <- chla_whole %>%
  rename(subbasin = Subbasin) %>%
  left_join(chla_status_allyrs) %>% 
  filter(year %in% trendyrs) %>% 
  left_join(
    mutate(
      chla_lm_estim,
      current = predict(trend_mdl, data.frame(year = max(trendyrs))),
      prj = predict(trend_mdl, data.frame(year = max(trendyrs) + future_year))
    )
  ) %>% 
  filter(coastal %in% c("Offshore"))

maketrendplot <- function(trendplotdf){
  trendplot <- ggplot(trendplotdf) +
    geom_point(aes(x = year, y = chla_ug_l), size = 0.2, alpha = 0.2) +
    geom_point(aes(x = year, y = mean_chla), size = 0.6) +
    geom_vline(xintercept = max(trendyrs), size = 0.2) +
    geom_vline(xintercept = max(trendyrs) + future_year, size = 0.2) +
    geom_hline(aes(yintercept = prj), color = "tomato") +
    geom_hline(aes(yintercept = current), color = "lightsteelblue") +
    facet_wrap(~subbasin, scales = "free_y", ncol = 4) +
    labs(x = NULL, y = NULL)
  return(trendplot)
}
plottitle <- "Chlorophyll a with 5 Years in Future and Current Estimates (Red and Blue lines Respectively)"
maketrendplot(filter(trendplotdf, coastal == "Offshore")) + ggtitle(paste("Offshore", plottitle))

## trend score dataframe ----

## assign basin trend to BHI regions
chla_trend <- chla_lm_estim %>% 
  mutate(
    expectedchange5yrs = coef(trend_mdl)["year"]*future_year,
    ## divide expected 5 year change in chla by respective basin target
    ## i.e get as proportion relative to target (target is status score of 100, so equivalent scales)
    ## then interpretation of trend is roughly expected (percent) change in status
    trend_score = max(-1, min(1, -expectedchange5yrs/target_chla))
  ) %>%
  ## only using offshore data for now:
  filter(coastal == "Offshore", enough_data) %>%
  group_by(subbasin, target_chla) %>% 
  summarize(trend_score = mean(trend_score, na.rm = TRUE)) %>% 
  ungroup() %>% 
  right_join(basin_lookup, by = "subbasin") %>% 
  mutate(
    score = round(trend_score, 3),
    region = rgn_nam,
    dimension = "trend",
    indicator = "chla"
  ) %>% 
  select(score, subbasin, dimension, indicator, bhi_id, rgn_nam, HELCOM_ID, region)
```

#### 3.2.5 Status and Trend Timeseries Plots {-}

**Basin status**

Basin status is initially a value between 0 and 1. Calculated for each year between 2000 and 2020.

```{r timeseries plot of basin eut chla status, results = "show", echo = TRUE, fig.width = 9.5}
plotdf <- chla_status_allyrs %>% 
  filter(coastal == "Offshore") %>% 
  distinct(subbasin, year, status, n_pts)
ggplot(plotdf) + 
  geom_col(aes(year, status, fill = n_pts), alpha = 0.7) +
  facet_wrap(~subbasin) +
  scale_fill_viridis_c() +
  geom_hline(yintercept = 1) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  labs(x = NULL, y = "Chl a Indicator Status\n", fill = "Num. Contributing of Points") +
  ggtitle("Yearly Chlorophyll a Status by Subbasin")
```

<br>


**Plot region Status and Trend values**

Status values can range from 0-100 -- this is the status for the *most recent* year. In most cases this is 2019. Current trends for Åland, Bothnian  Bay, Bothnian Sea, and Kiel Bay are based only on coastal data trends, as our offshore data were not sufficient for fitting trends.

```{r chlorphyl a status and trend for regions plot, results = "show", echo = TRUE, fig.width = 9.5, fig.height = 5.5}
ggplot(rbind(chla_status, chla_trend)) + 
  geom_col(aes(region, score, fill = subbasin), alpha = 0.5, show.legend = FALSE) +
  coord_flip() +
  facet_wrap(~dimension, scales = "free_x", nrow = 1) +
  labs(x = "BHI Region\n", y = NULL) +
  ggtitle("Chlorophyll a Status and Trend Scores by Region") +
  scale_fill_manual(values = statuspal) +
  theme(axis.text.y = element_text(size = 9))
```

<br>

#### 3.2.6 Save chlorophyll a eutrophication layer {-}

```{r save chla eut layer, eval = FALSE}
write_csv(
  chla_whole %>% 
    filter(!is.na(code)) %>% 
    select(
      region_id = BHI_ID,
      helcom_id = code,
      lat, lon, 
      month, year,
      chla_conc = chla_ug_l
    ), 
  file.path(dir_layers, sprintf("cw_eut_chla_bhi%s.csv", assess_year))
)
```

<br>

### 3.3 Nutrients Indicator

The nutrients indicator measures winter (December-February) dissolved inorganic nitrogen (DIN: nitrate, nitrite, and ammonium) and winter dissolved inorganic phosphorus (DIP) concentrations in the surface layer (0 - 10 m depth).

#### 3.3.1 Match BHI regions

The raw nutrients data obtained from Baltic Nest include nitrate and nitrite (measured separately or combined), ammonia, and phosphate. Nitrate, nitrite, and ammonia are summed for DIN, and only phosphate is used for DIP.

These points are assigned to their respective Subbasins (with HELCOM Coastal/Offshore division codes) and BHI Regions based on their latitude/longitude coordinates. Then the data are subsetted to assure only winter months (December-February) and years 2000 onward are included. A column 'winter of year' is created, to group e.g. January and February 2006 with December 2005 rather than with December 2006.

```{r filter and assign BHI regions to winter nutrients data, echo = TRUE, message = FALSE, warning = FALSE}
nutrients_rawdata <- nutrients_rawdata %>% 
  mutate(nitrate_and_nitrite = ifelse(NO23N == 0, NO3N + NO2N, NO23N)) %>% 
  rename(
    nitrate = NO3N, nitrite = NO2N, ammonia = NH4N, phosphate = PO4P,
    latitude = LATITUDE, longitude = LONGITUDE, depth_m = OBSDEP, 
    date = OBSDATE, cruise = SHIP, time = OBSTIME
  ) %>%
  mutate(
    date = as.Date(date), 
    year = format(date, "%Y"), 
    month = format(date, "%m")
  ) %>% 
  select(-ID, -NO23N, -TOTN) %>% 
  mutate(din = nitrate_and_nitrite + ammonia, dip = phosphate)

## check for duplicates
## double entries, but measurements taken at different times
nutrients_rawdata %>%
  group_by(
    cruise, date, year, month, 
    latitude, longitude, depth_m,
    nitrate_and_nitrite, ammonia, phosphate
  ) %>%
  mutate(n = n()) %>%
  filter(n > 1)
  
## assign helcom offshore codes
# nrow(filter(nutrients_rawdata, is.na(latitude) | is.na(longitude))) # first check if any lat/lon have NAs
nut_rgns_w_offshore_code <- nutrients_rawdata %>% 
  filter(!is.na(latitude) & !is.na(longitude)) %>%
  sf::st_as_sf(
    crs = 4326,
    coords = c("longitude", "latitude"),
    remove = FALSE
  ) %>%  
  sf::st_join(sf::st_transform(offshore_codes, 4326))

## spatial join also the WFD water bodies
nut_rgns_w_offshore_code <- nut_rgns_w_offshore_code %>% 
  sf::st_join(
    coastal_codes %>% 
      sf::st_transform(4326) %>% 
      rename(area_km2_wfd = Area_km2, code = HELCOM_ID) %>% 
      select(area_km2_wfd, code)
  ) %>% 
  sf::st_drop_geometry() %>% 
  filter(!is.na(latitude) & !is.na(longitude))

## assign bhi basins to nutrients data 
buffer_sf <- st_read(
  file.path(dirname(dir_B), "Shapefiles", "BHI_shapefile_25km_buffer"),
  "BHI_shapefile_25km_buffer"
)
nut_rgns_shp <- join_rgns_info(
  nut_rgns_w_offshore_code, 
  rgn_shps_loc = file.path(dirname(dir_B), "Shapefiles")
  # buffer_shp = buffer_sf
)

## confirm only winter months (December-February) are included
## surface layer (0 - 10 m depth) and years >= 2000
## also need to account for fact that winter spans across new year
nut_rgns <- nut_rgns_shp %>% 
  mutate(month = stringr::str_pad(month, 2, "left", "0")) %>% 
  filter(month == "12" | month == "01" | month == "02") %>% 
  filter(depth_m <= 10) %>% 
  mutate(year = as.numeric(year)) %>% 
  mutate(winterofyear = ifelse(month == 12, year, year-1)) %>% 
  filter(winterofyear >= 2000)

## reshape some for next steps
nut_rgns <- nut_rgns %>% 
  mutate(coastal = ifelse(is.na(helcom_coastal_code), "Coastal", "Offshore")) %>% 
  select(
    latitude, longitude, depth_m,
    Subbasin, BHI_ID, code, coastal, 
    area_km2_wfd,
    winterofyear, year, month, time, 
    din, dip
  ) %>% 
  tidyr::pivot_longer(c("din", "dip"), names_to = "nutrient", values_to = "value") %>% 
  mutate(nutrient = stringr::str_to_upper(nutrient)) %>% 
  
  ## some cases where code SEA-010 is matched with Northern Baltic Proper, 
  ## by our spatial join should be code SEA-012,
  ## happens because our polygons are different than HELCOM's
  mutate(code = as.character(code)) %>% 
  mutate(area_km2_wfd = ifelse(
    code == "SEA-010" & Subbasin == "Northern Baltic Proper", 
    unique(filter(nut_rgns, code == "SEA-012")$area_km2_wfd), 
    area_km2_wfd
  )) %>% 
  mutate(code = ifelse(code == "SEA-010" & Subbasin == "Northern Baltic Proper", "SEA-012", code)) %>% 
  mutate(area_km2_wfd = ifelse(
    code == "SEA-001" & Subbasin == "The Sound", 
    unique(filter(nut_rgns, code == "SEA-003")$area_km2_wfd), 
    area_km2_wfd
  )) %>% 
  mutate(code = ifelse(code == "SEA-001" & Subbasin == "The Sound", "SEA-003", code)) 
```
<br>


#### 3.3.2 Visualize Nutrients Data {-}

**Spatial Visualization of DIN and DIP data, Offshore vs Coastal Measurements**

The map shows locations of offshore (pink) and coastal (green) measurements. The two plots show temporal trends in offshore and coastal nutrients (right) and monthly distributions (left) by subbasins. Pink points correspond to offshore data, and green points to coastal data in both the map and the plots. 

```{r eut nutrients map of raw data points, results = "show", fig.width = 9, fig.height = 6}
bhi_rgns_simple <- rmapshaper::ms_simplify(input = BHI_rgns_shp) %>% sf::st_as_sf()

nutrients_obs_mapdf <- nut_rgns %>% 
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>% 
  filter(winterofyear %in% 2010:2019) %>% 
  mutate(no_bhi_id = is.na(BHI_ID))

basemap +
  geom_sf(data = bhi_rgns_simple, fill = "white", size = 0.2, color = "burlywood") +
  # geom_sf(data = nutrients_obs_mapdf, aes(color = no_bhi_id), size = 1.8, alpha = 0.7, show.legend = FALSE) +
  geom_sf(data = nutrients_obs_mapdf, aes(color = coastal), size = 0.7, alpha = 0.2) +
  scale_color_manual(values = c("darkseagreen", "salmon")) +
  labs(color = "Offshore or Coastal Measurement") +
  theme_bw() +
  ggtitle("Offshore and Coastal Nutrient Measurements") +
  theme(legend.position = c(0.25, 0.9), legend.background = element_rect(fill = "ghostwhite"))
```

```{r nutrients data check coastal versus offshore, results = "show", fig.width = 11, fig.height = 8}
## make dataframe to use in plotting
expandcases <- tidyr::expand(nut_rgns, tidyr::nesting(Subbasin, BHI_ID, winterofyear, month, nutrient))
plotdf <- nut_rgns %>% 
  full_join(expandcases) %>%
  filter(!(is.na(Subbasin) & is.na(nutrient)), !is.na(Subbasin)) %>% 
  left_join(data.frame(month = c("12","01","02"), Month = c("Dec","Jan","Feb")))
plotdf$Month <- factor(plotdf$Month, levels = c("Dec","Jan","Feb"))

for(r in c("Subbasin", "BHI_ID")){
  for(n in c("DIN", "DIP")){
    ## timeseries plots
    plottimeseries <- ggplot(filter(plotdf, nutrient == n)) +
      geom_point(
        aes(winterofyear, value, color = coastal), 
        size = 0.5, alpha = 0.6
      ) +
      labs(x = "\nWinter of Year", y = NULL, color = "Coastal/Offshore") +
      ggtitle(paste(n, "vs Year")) +
      scale_color_manual(values = c("darkseagreen", "salmon")) +
      facet_wrap(r, scales = "free_y", ncol = 3) +
      theme(axis.text.x = element_text(size = 7, angle = 90), legend.position = "bottom")
    assign(sprintf("plot_%s_%s_ts", str_to_lower(n), str_to_lower(r)), plottimeseries, envir = .GlobalEnv)
    
    ## boxplots comparing coastal/offshore by region or subbasin
    plotboxplot <- ggplot(filter(plotdf, nutrient == n)) +
      geom_boxplot(
        aes(Month, value, fill = coastal),
        size = 0.2, outlier.size = 0.3, alpha = 0.4, show.legend = FALSE
      ) +
      labs(x = NULL, y = NULL, fill = "Coastal/Offshore") +
      ggtitle(paste("Monthly", n)) +
      scale_fill_manual(values = c("darkseagreen", "salmon")) +
      facet_wrap(r, scales = "free_y", ncol = 3)
    assign(sprintf("plot_%s_%s_box", str_to_lower(n), str_to_lower(r)), plotboxplot, envir = .GlobalEnv)
  }
}

gridExtra::grid.arrange(plot_din_subbasin_ts, plot_dip_subbasin_ts, nrow = 1)
```

<br>

#### 3.3.3 Mean Nutrient Calculations

Mean nutrient concentrations are calculated from coastal and offshore measurements grouped by open-sea sub-basins and WFD coastal spatial units. Means are calculated for each coastal or offshore zone, first within each winter month. Then monthly means are averaged to get an overall winter mean in each the coastal or offshore zone. Variances (standard deviations squared) are also calculated within each month, then averaged to get an average variance across winter months for the subbasin and coastal/offshore category. This is presented as a standard deviation (square root of variance across months) on the plots below.

Along with the mean calculation, we calculate five-year moving averages. The moving average (or 'rolling mean') is an average of winter means from the current and previous four winters. 

However, the final scores for the nutrient indicators only consider the offshore means, as the coastal thresholds available for both [DIN](https://helcom.fi/media/core%20indicators/Dissolved-inorganic-nitrogen-DIN-HELCOM-core-indicator-2018.pdf) and [DIP](https://helcom.fi/media/core%20indicators/Dissolved-inorganic-phosphorus-DIP-HELCOM-core-indicator-2018.pdf) include only Sweden, Latvia, and Poland.

```{r calculate monthly mean nutrients with target, results = "hide"}
## check that each subbasin is actually properly matched with a single code
# nut_rgns %>% 
#   ungroup() %>% 
#   select(Subbasin, code) %>% 
#   distinct() %>% 
#   filter(str_detect(code, "SEA")) %>%
#   group_by(Subbasin) %>% 
#   count()

## Calculate DIN and DIP mean monthly values for each winter month
## For coastal and offshore separately
nutrient_monthly_means <- nut_rgns %>% 
  filter(!is.na(Subbasin), !is.na(BHI_ID), !is.na(code)) %>% 
  ## now group and take means
  group_by(winterofyear, month, coastal, nutrient, Subbasin, code, area_km2_wfd) %>%
  summarise(
    nut_mean = mean(value, na.rm = TRUE),
    nut_sd = sd(value, na.rm = TRUE),
    n_pts = n()
  ) %>% 
  mutate(nut_sd = ifelse(is.na(nut_sd), 0, nut_sd))

## Calculate winter DIN and DIP means
## each year will have multiple means, for coastal areas and an offshore (open-sea) subbasin
nut_winter_months_mean <- nutrient_monthly_means %>% 
  group_by(winterofyear, coastal, nutrient, Subbasin, code, area_km2_wfd) %>%
  summarise(
    nut_mean = mean(nut_mean, na.rm = TRUE), 
    nut_sd = sqrt(mean((nut_sd^2), na.rm = TRUE)),
    n_pts = sum(n_pts)
  )

## calculate min and max, and rolling 5 year mean
nut_winter_months_mean <- nut_winter_months_mean %>% 
  arrange(Subbasin, code, nutrient, coastal, winterofyear) %>% 
  group_by(Subbasin, code, nutrient, coastal) %>%
  mutate(ma5yr = zoo::rollapply(nut_mean, 5, mean, na.rm = TRUE, align = "right", fill = NA)) %>% 
  mutate(nut_min = nut_mean-nut_sd, nut_max = nut_mean+nut_sd)


## join with basin-specific eutrophication chlorophyll a targets
wfd_area_target_nut <- file.path(dir_prep, "supplement", "lookup_tabs", "rgns_complete.csv") %>% 
  read_csv() %>% 
  dplyr::select(basin = subbasin, code = helcom_id) %>% 
  distinct() %>% 
  left_join(eut_thresholds, by = "basin") %>% 
  dplyr::select(code, DIN = winter_DIN, DIP = winter_DIP) %>% 
  mutate(units = "umol_per_l") %>% 
  bind_rows(left_join(
      rename(din_coastal_thresholds, DIN = threshold),
      rename(dip_coastal_thresholds, DIP = threshold)
  )) %>% 
  ## join with basin-specific eutrophication nutrient a targets
  tidyr::pivot_longer(c("DIN", "DIP"), names_to = "nutrient", values_to = "target") %>% 
  right_join(nut_winter_months_mean, by = c("code", "nutrient"))


## Plot winter mean nutrients
for(n in c("DIN", "DIP")){
  meansplot <- ggplot(filter(wfd_area_target_nut, nutrient == n), aes(winterofyear, nut_mean, group = code)) + 
    geom_errorbar(
      aes(x = winterofyear, ymin = nut_min, ymax = nut_max), 
      alpha = 0.4, size = 0.3, color = "snow", width = 0.3,
      show.legend = FALSE
    ) +
    geom_point(size = 0.3, alpha = 0.7, color = "snow", show.legend = FALSE) +
    geom_line(size = 0.2, alpha = 0.7, color = "snow", show.legend = FALSE) + 
    geom_point(
      data = filter(wfd_area_target_nut, str_detect(code, "^SEA-"), nutrient == n),
      size = 0.7
    ) +
    geom_line(
      data = filter(wfd_area_target_nut, str_detect(code, "^SEA-"), nutrient == n),
      size = 0.3
    ) + 
    facet_wrap(~Subbasin, scales = "free", ncol = 4) +
    scale_y_continuous(limits = c(0, NA)) +
    labs(x = NULL, y = NULL) +
    theme(
      axis.text.x = element_text(angle = 90, vjust = 0.5),
      strip.text.y = element_text(size = 6)
    ) +
    theme_dark() 
  assign(sprintf("%s_%s_meansplot", str_to_lower(n), "subbasin"), meansplot)
}
```

**Plot winter nutrients with target values indicated**

Horizontal lines are HELCOM target values 

```{r winter nutrients with target, results = "show", echo = TRUE, fig.width = 9.5, fig.height = 7}
din_subbasin_meansplot + ggtitle("Mean Winter Dissolved Inorganic Nitrogen (Dec-Feb) against Target values")
dip_subbasin_meansplot + ggtitle("Mean Winter Dissolved Inorganic Phosphorus (Dec-Feb) against Target values")
```

<br>

#### 3.3.4 Status Calculations: Coastal and Offshore {-}

**Explore years in data available for status calculation**

For both DIN and DIP, no basins have data for 2019, which is the assessment year; after excluding coastal spatial units, thirteen basins have data for 2018 and for three (Åland, Gulf of Riga, The Quark) the most recent year is 2017. Most recent 'offshore' data for The Sound is 2010. For each basin, the most recent year's data in that region is used for _current status_ calculations.

```{r what is the last year for nutrients, results = "show"}
## get the last year of non-NA DIN data
din_last_year <- wfd_area_target_nut %>%
  filter(!is.na(ma5yr), str_detect(code, "^SEA-"), nutrient == "DIN") %>%
  group_by(code, Subbasin) %>%
  summarise(last_year = last(winterofyear)) %>% 
  mutate(indicator = "din") 
## which are not in 2019
# filter(din_last_year, last_year < 2019)

## get the last year of non-NA DIP data
dip_last_year <- wfd_area_target_nut %>%
  filter(!is.na(ma5yr), str_detect(code, "^SEA-"), nutrient == "DIP") %>%
  group_by(code, Subbasin) %>%
  summarise(last_year = last(winterofyear)) %>% 
  mutate(indicator = "dip")
## which are not in 2019
# filter(chla_last_year, `Last Year` < 2019)

knitr::kable(
  bind_rows(
    din_last_year %>% 
      group_by(last_year, indicator) %>%
      summarize(`Num. subbasins with year` = n(), Subbasins = paste(Subbasin, collapse = ", ")),
    dip_last_year %>% 
      group_by(last_year, indicator) %>%
      summarize(`Num. subbasins with year` = n(), Subbasins = paste(Subbasin, collapse = ", "))
  ) %>% 
  mutate(indicator = str_to_upper(indicator)) %>% 
  rename(Indicator = indicator, `Last Year` = last_year)
)
```

<br>

**Status calculation with raw (non-modeled) mean winter nutrients by basin** 

Status is calculated on the basin level and then applied to all regions. It is calculated in data prep for additional visualization and evaluation, but the final method of status calculation is also completed by the EUT function in functions.R of the relevant `bhi` repository assessment folder.

Status for each the _DIN_ and _DIP_ indicators is calculated as `basin target` divided by `basin winter means`, with the value capped at 1. We calculate status using only Offshore data for evaluation, and using the 5 year moving average. The plot shows BHI region scores (uniform across regions within the same basin, as the data is aggregated by basin for the winter means calculation), and also indicates which data year the status is based on.

```{r constants for eut DIN and DIP status and trend calculations, echo = TRUE}
## Define constants for status and trend calculations
min_year <- 2009 # earliest year to use as a start for regr_length timeseries
regr_length <- 10 # number of years to use for regression
future_year <- 5 # the year at which we want the likely future status
min_regr_length <- 5 # min actual number of years with data to use for regression
```

```{r eut nutrients status calculation, results = "show", echo = TRUE, fig.width = 9.5, fig.height = 7}
## distributions of means for offshore and coastal categories
ggplot(wfd_area_target_nut) + 
  geom_histogram(aes(nut_mean), bins = 100, fill = NA, color = "grey") + 
  facet_grid(rows = vars(coastal), cols = vars(nutrient), scales = "free_x") + 
  labs(x = "Mean Nutrient Concentrations", y = NULL)

## Calculate basin status as: basin_target/basin_mean, constraining values to 1 as maximum score
nut_status_allyrs <- wfd_area_target_nut %>%
  rename(subbasin = Subbasin) %>% 
  # mutate(status = pmin(1, target/nut_mean))
  ## or use five year rolling means (moving average) rather than annual mean?
  mutate(status = pmin(1, target/ma5yr))
nut_status <- group_by(nut_status_allyrs, nutrient, coastal, subbasin, code)

## select last year of data for status in each subbasin, this means status year differs by basin
## joining to subbasin lookup expands from subbasins to one row per BHI region 
nut_status <- nut_status %>% 
  filter(winterofyear == max(winterofyear)) %>% 
  mutate(status = round(status*100, 2)) %>% # keep decimal places, will round when combine eut indicators
  ungroup() %>% 
  mutate(
    dimension = "status", 
    indicator = ifelse(nutrient == "DIP", "dip", "din")
  ) %>% 
  dplyr::select(score = status, indicator, dimension, coastal, subbasin, code) %>% 
  left_join(basin_lookup) %>%
  mutate(region = as.factor(rgn_nam))

## plot the status
## BHI region plots will look the same, as BHI status are taken from subbasin means
## unless alternative calculation is used, based on BHI region means rather than subbasin means
set.seed(2)
statuspal <- colorRampPalette(fullpal)(80)[sample(1:80, size = 18)]

plotdf <- nut_status %>% 
  filter(coastal == "Offshore") %>% 
  left_join(
    bind_rows(din_last_year, dip_last_year),
    by = c("subbasin" = "Subbasin", "indicator", "code")
  ) %>% 
  mutate(indicator = ifelse(indicator == "dip", "DIP", "DIN"))

ggplot(plotdf) +
  geom_col(
    aes(region, score, fill = subbasin),
    position = position_dodge(), 
    color = "grey", 
    alpha = 0.6,
    size = 0.2,
    show.legend = FALSE
  ) + 
  geom_text(aes(region, score, label = last_year), color = "slategray", size = 3) +
  coord_flip() +
  facet_grid(cols = vars(coastal), rows = vars(indicator)) +
  labs(x = NULL, y = NULL) +
  theme_bw() +
  theme(axis.text.y = element_text(size = 7)) +
  scale_fill_manual(values = statuspal) +
  ggtitle("DIN and DIP indicators Status by Region, for Coastal, Offshore, and Combined data")

nut_status <- nut_status %>% 
  filter(coastal == "Offshore") %>% 
  select(-coastal, -code)
```

<br>

#### 3.3.4 Trend Calculations: Coastal & Offshore {-}

In our approach here, we use a 10 year period to calculate the trend with a minimum of 5 data points. In most cases, the BHI framework uses a 5 year period for the trend, but as dissolved inorganic nutrients are slower response variables, we use a longer time period. The trend value is the slope of the linear regression (mean winter dissolved inorganic nutrients vs year) multiplied by the year of future interest (5 years from status year) and this value is constrained to between -1 and 1.

We use the same approach as with the Chlorophyll indicator, calculating trend as an average of trend scores calculated for coastal and offshore DIN or DIP.

The trend plots show (inorganic) nutrient measurements by basin, with a red line indicating where the linear model predicts mean Secchi depths will be five years in the future, and a blue line indicating current assessment year summer mean. Vertical lines frame the five-year time frame over which trend is projected.

**Nutrients with 5 Years in Future and Current Estimates (Red and Blue lines Respectively)**

```{r eut nutrients indicators trend calculation, results = "hide", echo = TRUE, fig.width = 9.5, fig.height = 10}
trendyrs <- min_year:(min_year + regr_length)

## linear model from the basin mean nutrient values ----

nut_lm_estim <- nut_status_allyrs %>%
  filter(winterofyear %in% trendyrs) %>% 
  distinct(subbasin, code, coastal, nutrient, winterofyear, nut_mean) %>%
  group_by(subbasin, code, coastal, nutrient) %>%
  ## calculate trend only if:
  ## have at least X yrs data (min_regr_length) in the last Y yrs (regr_length) of time series
  do(tail(., n = regr_length)) %>%
  do(data.frame(enough_data = ifelse(sum(!is.na(.$nut_mean)) >= min_regr_length, TRUE, FALSE))) %>%
  right_join(nut_status_allyrs, by = c("subbasin", "code", "coastal", "nutrient")) %>%
  group_by(subbasin, code, coastal, nutrient, enough_data, target) %>%
  do(trend_mdl = lm(nut_mean ~ winterofyear, data = .))

## trend plots ----

trendplotdf <- nut_rgns %>%
  filter(!is.na(Subbasin)) %>% 
  rename(subbasin = Subbasin) %>%
  left_join(nut_status_allyrs) %>% 
  filter(winterofyear %in% trendyrs) %>% 
  left_join(
    mutate(
      nut_lm_estim,
      current = predict(trend_mdl, data.frame(winterofyear = max(trendyrs))),
      prj = predict(trend_mdl, data.frame(winterofyear = max(trendyrs) + future_year))
    )
  ) %>% 
  filter(coastal %in% c("Offshore"))

maketrendplot <- function(trendplotdf){
  trendplot <- ggplot(trendplotdf) +
    geom_point(aes(x = winterofyear, y = value), size = 0.2, alpha = 0.2) +
    geom_point(aes(x = winterofyear, y = nut_mean), size = 0.6) +
    geom_vline(xintercept = max(trendyrs), size = 0.2) +
    geom_vline(xintercept = max(trendyrs) + future_year, size = 0.2) +
    geom_hline(aes(yintercept = prj), color = "tomato") +
    geom_hline(aes(yintercept = current), color = "lightsteelblue") +
    facet_wrap(~subbasin, scales = "free_y", ncol = 4) +
    labs(x = NULL, y = NULL)
  return(trendplot)
}
gridExtra::grid.arrange(
  maketrendplot(filter(trendplotdf, coastal == "Offshore", nutrient == "DIN")) + ggtitle("Offshore DIN"),
  maketrendplot(filter(trendplotdf, coastal == "Offshore", nutrient == "DIP"))+ ggtitle("Offshore DIP"),
  nrow = 2
)
```


```{r eut nutrients indicators trend dataframe, results = "hide", echo = TRUE}
## trend score dataframe ----

## assign basin trend to BHI regions
nut_trend <- nut_lm_estim %>% 
  mutate(
    expectedchange5yrs = coef(trend_mdl)["winterofyear"]*future_year,
    ## divide expected 5 year change in DIN or DIP by respective basin target
    ## i.e get as proportion relative to target (target is status score of 100, so equivalent scales)
    ## then interpretation of trend is roughly expected (percent) change in status
    trend_score = max(-1, min(1, -expectedchange5yrs/target))
  ) %>% 
  ungroup() %>% 
  ## only using offshore (open-sea) data for now:
  filter(coastal == "Offshore", enough_data) %>%
  group_by(subbasin, target, nutrient) %>% 
  summarize(trend_score = mean(trend_score, na.rm = TRUE)) %>% 
  ungroup() %>% 
  right_join(basin_lookup, by = "subbasin") %>% 
  mutate(
    score = round(trend_score, 3),
    region = rgn_nam,
    dimension = "trend",
    indicator = str_to_lower(nutrient)
  ) %>% 
  select(score, subbasin, dimension, indicator, bhi_id, rgn_nam, HELCOM_ID, region)
```

<br>


#### 3.3.5 Status and Trend Timeseries Plots {-}

**Basin status**

Basin status is initially a value between 0 and 1. Calculated for each year between 2000 and 2019.

```{r timeseries plot eut nutrients status, results = "show", echo = TRUE, fig.width = 9.5, fig.height = 9}
plotdf <- nut_status_allyrs %>% 
  filter(coastal == "Offshore") %>% 
  distinct(subbasin, winterofyear, nutrient, status, n_pts)

gridExtra::grid.arrange(
  ggplot(filter(plotdf, nutrient == "DIN")) + 
    geom_col(aes(winterofyear, status, fill = n_pts), alpha = 0.7) +
    facet_wrap(~subbasin) +
    scale_fill_viridis_c() +
    geom_hline(yintercept = 1) +
    labs(x = NULL, y = "DIN Indicator Status\n", fill = "Num. Contributing of Points") +
    ggtitle("Yearly DIN Status by Subbasin"),
  ggplot(filter(plotdf, nutrient == "DIP")) + 
    geom_col(aes(winterofyear, status, fill = n_pts), alpha = 0.7) +
    facet_wrap(~subbasin) +
    scale_fill_viridis_c() +
    geom_hline(yintercept = 1) +
    labs(x = NULL, y = "DIP Indicator Status\n", fill = "Num. Contributing of Points") +
    ggtitle("Yearly DIP Status by Subbasin"),
  nrow = 2
)
```

<br>

**Plot region Status and Trend values**

Status values can range from 0-100 -- this is the status for the *most recent* year. In most cases this is 2019. Note: Kiel Bay trend for both DIN and DIP is based only on offshore data as coastal data were not sufficient to fit a trend.

```{r nutrient status and trend for regions plot, results = "show", echo = TRUE, fig.width = 9.5, fig.height = 7}
rbind(nut_status, nut_trend) %>% 
  mutate(indicator = stringr::str_to_upper(indicator)) %>% 
  filter(!is.na(indicator)) %>% 
  ggplot() + 
  geom_col(aes(region, score, fill = subbasin), alpha = 0.5, show.legend = FALSE) +
  coord_flip() +
  facet_grid(cols = vars(dimension), rows = vars(indicator), scales = "free_x") +
  labs(
    x = "BHI Region\n", 
    y = NULL,
    title = "DIN and DIP Status and Trend Scores by Region",
    caption = "Negative trend values correspond to increasing nutrient concentrations."
  ) +
  scale_fill_manual(values = statuspal) +
  theme(axis.text.y = element_text(size = 7), plot.caption = element_text(hjust = 0))
```

<br>

#### 3.3.6 Save dissolved inorganic nutrients eutrophication layers {-}

```{r save din and dip eut layers, eval = FALSE}
write_csv(
  nut_rgns %>% 
    filter(!is.na(code), nutrient == "DIN") %>% 
    select(
      region_id = BHI_ID,
      helcom_id = code,
      latitude, longitude,
      month, year,
      din_conc = value
    ), 
  file.path(dir_layers, sprintf("cw_eut_din_bhi%s.csv", assess_year))
)
write_csv(
  nut_rgns %>% 
    filter(!is.na(code), nutrient == "DIP") %>% 
    select(
      region_id = BHI_ID,
      helcom_id = code,
      latitude, longitude,
      month, year,
      dip_conc = value
    ), 
  file.path(dir_layers, sprintf("cw_eut_dip_bhi%s.csv", assess_year))
)
```

<br>

### 3.4 Oxygen Debt {-}

For details on data preparation for oxygen debt layer, see [this github repository](https://github.com/OHI-Baltic/HEAT), forked from ICES github page for the HEAT assessment tool for fitting oxygen debt profiles to oxygen data measurements. Oxygen data measurements used were obtained from ICES and merged as documented in section 2.2.4.

```{r read in oxy debt data from HEAT repository, results = "hide", message = FALSE}
heat_output <- "https://raw.githubusercontent.com/OHI-Baltic/HEAT/master/analysis/output/"
oxydebt <- read_csv(paste0(heat_output,"OxygenDebt/uncorrected_indicator_table_by_year_2000_2019.csv"))
```

#### 3.4.1 Match BHI regions {-}

```{r assign BHI regions to oxygen debt data, echo = TRUE, message = FALSE, warning = FALSE}
oxydebt <- oxydebt %>% 
  select(-AssessmentUnitID) %>% 
  mutate(AssessmentUnitName = case_when(
    AssessmentUnitName == "Gulf of Gdansk" ~ "Gdansk Basin",
    str_detect(AssessmentUnitName, "land Sea") ~ "Aland Sea",
    TRUE ~ AssessmentUnitName
  )) %>% 
  rename(subbasin = AssessmentUnitName, year = Year)

## only Bornholm Basin and Baltic Proper have target values, the other subbasins are too shallow
## also, the sub-basins within Bornholm Basin and Baltic Proper have uniform values within each year...
oxydebt_rgns <- oxydebt %>% 
  full_join(basin_lookup) %>% 
  left_join(select(eut_thresholds, subbasin = basin, oxydebt_target = oxyg_debt))
```

<br>

#### 3.4.2 Status Calculaion {-}

```{r constants for eut oxy debt status and trend calculations, echo = TRUE}
## Define constants for status and trend calculations
min_year <- 2009 # earliest year to use as a start for regr_length timeseries
regr_length <- 10 # number of years to use for regression
future_year <- 5 # the year at which we want the likely future status
min_regr_length <- 5 # min actual number of years with data to use for regression
```


```{r eut oxy debt status calculation, results = "show", echo = TRUE, fig.width = 9.5, fig.height = 5.5}
## Calculate basin status as: basin_mean/basin_target, constraining values to 1 as maximum score
oxydebt_status_allyrs <- oxydebt_rgns %>%
  filter(year >= min_year) %>% 
  mutate(status = pmin(1, oxydebt_target/ES)) %>% 
  group_by(subbasin)

## select last year of data for status in each subbasin
oxydebt_status <- oxydebt_status_allyrs %>%
  filter(year == max(year)) %>%
  ungroup() %>%
  dplyr::select(score = status, subbasin, bhi_id, rgn_nam, HELCOM_ID) %>%
  mutate(dimension = "status", indicator = "oxydebt") %>% 
  mutate(region = as.factor(rgn_nam))

## plot the status
set.seed(2)
statuspal <- colorRampPalette(fullpal)(80)[sample(1:80, size = 18)]

ggplot(oxydebt_status) +
  geom_col(
    aes(region, score, fill = subbasin),
    position = position_dodge(), 
    color = "grey", 
    alpha = 0.6,
    size = 0.2,
    show.legend = FALSE
  ) + 
  coord_flip() +
  labs(x = NULL, y = NULL) +
  theme_bw() +
  theme(axis.text.y = element_text(size = 9)) +
  scale_fill_manual(values = statuspal) +
  scale_y_continuous(limits = c(0, 1)) +
  ggtitle("Oxygen Debt Indicator Status by Region")
```
<br>

#### 3.4.3 Trend Calculation {-}

#### 3.4.4 Status and Trend Timeseries Plots {-}

**Basin status**

Basin status is initially a value between 0 and 1. Calculated for each year between 2000 and 2019.

**Plot region Status and Trend values**

Status values can range from 0-100 -- this is the status for the *most recent* year. In most cases this is 2019. Note: Kiel Bay trend for both DIN and DIP is based only on offshore data as coastal data were not sufficient to fit a trend.

<br>

#### 3.3.6 Save oxygen debt eutrophication layer {-}

```{r save oxygen debt eut layer, eval = FALSE}
write_csv(
  oxydebt_rgns %>% 
    select(
      region_id = bhi_id,
      helcom_id = HELCOM_ID,
      year,
      oxygendebt = ES
    ), 
  file.path(dir_layers, sprintf("cw_eut_oxydebt_bhi%s.csv", assess_year))
)
```

<br>

## 4. Visualizing Data Layers {-}

### 4.1 Status and Trend Maps by Subbasin {-}

```{r color palettes for eut status and trend maps}
statusmapcols <- c("indianred", "coral", "goldenrod1", "khaki", "lightblue", "steelblue")
trendmapcols <- c("indianred", "coral", "white", "lightblue", "steelblue")
```

<br>

**Secchi Status and Trend Maps**

```{r secchi status and trend maps, results = "show", echo = TRUE, fig.width = 9.5, fig.height = 4}
mapdf_secchi <- left_join(
  bhi_rgns_simple %>% 
    mutate(Subbasin = as.character(Subbasin)),
  rbind(secchi_status, secchi_trend) %>% 
    select(Subbasin = subbasin, BHI_ID = bhi_id, score, dimension, region),
  by = c("Subbasin", "BHI_ID")
)
status_secchi_map <- basemap + 
  geom_sf(data = filter(mapdf_secchi, dimension == "status"), aes(fill = score), size = 0.1, alpha = 0.7) +
  scale_fill_gradientn(colors = statusmapcols, limits = c(0, 100), na.value = "gainsboro") +
  theme(
    legend.background = element_rect(color = "grey"),
    legend.position = c(0.15, 0.75)
  ) +
  labs(fill = "Score", title = "Secchi Status Map")

trend_secchi_map <- basemap + 
  geom_sf(data = filter(mapdf_secchi, dimension == "trend"), aes(fill = score), size = 0.1, alpha = 0.7) +
  scale_fill_gradientn(colors = trendmapcols, limits = c(-0.5, 0.5), na.value = "gainsboro") +
  theme(
    legend.background = element_rect(color = "grey"),
    legend.position = c(0.15, 0.75)
  ) +
  labs(fill = "Score", title = "Secchi Trend Map")

gridExtra::grid.arrange(status_secchi_map, trend_secchi_map, nrow = 1)
```

<br>

**Chl a Status and Trend Maps**

```{r chla status and trend maps, results = "show", echo = TRUE, fig.width = 9.5, fig.height = 4}
mapdf_chla <- left_join(
  bhi_rgns_simple %>% 
    mutate(Subbasin = as.character(Subbasin)),
  rbind(chla_status, chla_trend) %>% 
    select(Subbasin = subbasin, BHI_ID = bhi_id, score, dimension, region),
  by = c("Subbasin", "BHI_ID")
)
status_chla_map <- basemap + 
  geom_sf(data = filter(mapdf_chla, dimension == "status"), aes(fill = score), size = 0.1, alpha = 0.7) +
  scale_fill_gradientn(colors = statusmapcols, limits = c(0, 100), na.value = "gainsboro") +
  theme(
    legend.background = element_rect(color = "grey"),
    legend.position = c(0.15, 0.75)
  ) +
  labs(fill = "Score", title = "Chlorophyll a Status Map")

trend_chla_map <- basemap + 
  geom_sf(data = filter(mapdf_chla, dimension == "trend"), aes(fill = score), size = 0.1, alpha = 0.7) +
  scale_fill_gradientn(colors = trendmapcols, limits = c(-0.5, 0.5), na.value = "gainsboro") +
  theme(
    legend.background = element_rect(color = "grey"),
    legend.position = c(0.15, 0.75)
  ) +
  labs(fill = "Score", title = "Chlorophyll a Trend Map")

gridExtra::grid.arrange(status_chla_map, trend_chla_map, nrow = 1)
```
<br>

**DIN and DIP Status and Trend Maps**

```{r make eut nutrients status and trend maps, results = "hide", echo = TRUE}
mapdf_nut <- left_join(
  bhi_rgns_simple %>% 
    mutate(Subbasin = as.character(Subbasin)),
  rbind(nut_status, nut_trend) %>% 
    select(Subbasin = subbasin, BHI_ID = bhi_id, score, dimension, indicator, region),
  by = c("Subbasin", "BHI_ID")
)
lapply(
  list(
    list("status", "din", "statusmapcols"), list("trend", "din", "trendmapcols"),
    list("status", "dip", "statusmapcols"), list("trend", "dip", "trendmapcols")
  ),
  function(x){
    if(x[[1]] == "status"){lims = c(0, 100)} else {lims = c(-0.5, 0.5)}
    assign(
      sprintf("%s_%s_map", x[[1]], x[[2]]), 
      basemap + 
        geom_sf(
          data = filter(mapdf_nut, dimension == x[[1]], indicator == x[[2]]), 
          aes(fill = score), 
          size = 0.1, alpha = 0.7
        ) +
        scale_fill_gradientn(
          colors = get(x[[3]]), 
          limits = lims, 
          na.value = "gainsboro"
        ) +
        theme(
          legend.background = element_rect(color = "grey"),
          legend.position = c(0.15, 0.75)
        ) +
        labs(fill = "Score", title = paste(str_to_upper(x[[2]]), str_to_title(x[[1]]), "Map")), 
      envir = .GlobalEnv
    )
    createdplots <- sprintf("%s_%s_map", x[[1]], x[[2]])
  }
)
```

```{r eut nutrients status and trend maps, results = "show", echo = TRUE, fig.width = 9.5, fig.height = 8.5}
gridExtra::grid.arrange(
  status_din_map, trend_din_map, 
  status_dip_map, trend_dip_map, 
  nrow = 2
)
```


<br>


## 5. Considerations for `BHI3.0` {-}

**Regarding coastal thresholds for nutrients**: Could use the nationally set nutrient indicators and their targets in the coastal areas. In some cases that could be DIN&DIP, in other cases total nutrients; in some cases for the winter period, sometimes summer, sometimes all-year. 

<br>

## 6. References {-}

```{r References, child = refs_path, results = "asis", echo = FALSE}
```
<br>
