---
title: "Clean Water (CW) - Eutrophication (EUT) Subgoal Data Preparation"
output:
  html_document:
    toc: true
    toc_depth: 3
    code_folding: hide
---

<br>

```{r preamble prep including spatial functions and files, message = FALSE, warning = FALSE, include = FALSE}
loc <- here::here("prep", "CW", "eutrophication")

source(here::here("R", "setup.R"))
knitr::opts_chunk$set(message = FALSE, warning = FALSE, results = "hide", fig.width = 9.5)

bkgd_path <- here::here("supplement", "goal_summaries", "EUT.Rmd")
data_path <- here::here("data", "CW", "eutrophication", version_year, "eut_data.rmd")
refs_path <- file.path(loc, "eut_references.Rmd")
dir_lookup <- file.path(dir_prep, "supplement", "lookup_tabs")

basin_lookup_table <-  read.csv(file.path(dir_lookup, "bhi_basin_country_lookup.csv"), sep=";")
basin_lookup <- basin_lookup_table %>% 
  dplyr::select(bhi_id = BHI_ID, subbasin = Subbasin) %>% 
  mutate(subbasin = as.character(subbasin))

source(here::here("R", "spatial.R"))
regions_shape()
buffer_sf <- st_read(
  file.path(dirname(dir_B), "Shapefiles", "BHI_shapefile_25km_buffer"), 
  "BHI_shapefile_25km_buffer"
)
bhi_rgns_simple <- rmapshaper::ms_simplify(input = BHI_rgns_shp) %>% sf::st_as_sf()
```

## 1. Background

```{r Background, child = bkgd_path, results = "asis", echo = FALSE}
```

<br/>

## 2. Data

This prep document is used to generate and explore the following data layers:

- `cw_eut_status_bhi2019.csv` 
- `cw_eut_trend_bhi2019.csv` 

These are saved to the `layers/v2019` folder. Saved to `data/GOAL/v2019/intermediate` are intermediate datasets: `secchi_merged_rawdata.csv`, `ox_merged_rawdata.csv` and `DATASETINTEMED3.csv`. All these are derived from or informed by the raw datasets from ICES, SMHI and Baltic Nest database.

<br>

<!-- ## 2. Data --- header in the child  document -->
```{r Data, child = data_path, results = "asis", echo = FALSE}
```

<br/>

## 3. Prep: Wrangling & Derivations, Checks/Evaluation, Gapfilling

```{r load datasets, echo = FALSE, include = FALSE}
combined_secchi_rawdata <- read_csv(file.path(dirname(data_path), "intermediate", "secchi_merged_rawdata.csv"))
combined_ox_rawdata <- read_csv(file.path(dirname(data_path), "intermediate", "ox_merged_rawdata.csv"))
```

### 3.1 Secchi Indicator

#### 3.1.1 Match BHI regions

```{r assign BHI regions to Secchi data, echo = TRUE, message = FALSE, warning = FALSE}
# before assigning bhi basins to secchi data, run this function first:
regions_shape(sp_dir = file.path(dirname(dir_B), "Shapefiles"))
 
# assign bhi basins to secchi data 
secchi_rgns_shp <- join_rgns_info(
  combined_secchi_rawdata, helcomID_col = "helcom_id", country_col = "country", 
                           latlon_vars = c("^lat", "^lon"), return_spatial = FALSE, 
                           rgn_shps_loc = file.path(dirname(dir_B), "Shapefile"), buffer_shp = NULL) %>% 
  # select summer months (June-September) and Years >= 2000  
  filter(month == "06" | month == "07" | month == "08" | month == "09") %>% 
  filter(year >=2000) %>% 
  rename(country = rgn_nam,
         ctry_key = rgn_key,
         subbasin = Subbasin,
         helcom_id = HELCOM_ID,
         bhi_id = BHI_ID,
         ices_area = ICES_area)

##secchi_rgns_shp <- join_rgns_info(
  ##secchi, helcomID_col = "helcom_id", country_col = "country", 
    ##                       latlon_vars = c("^lat", "^lon"), return_spatial = FALSE, 
      ##                     regions_shape(sp_dir = file.path(dirname(dir_B), "Shapefiles")), buffer_shp = NULL)

```

**Spatial Visualization of Secchi data**
```{r}
# Plot by bhi regions
ggplot(secchi_rgns_shp) + 
  geom_point(aes(year, secchi_m), size = 0.3, alpha = 0.4) +
  facet_grid(cols = vars(country), rows = vars(bhi_id))

ggplot(secchi_rgns_shp) + 
  geom_point(aes(year, secchi_m, colour = supplier), size = 0.3, alpha = 0.4) +
  facet_wrap(~bhi_id, scales ="free_y")

ggplot(secchi_rgns_shp) + 
  geom_point(aes(month, secchi_m, colour = supplier), size = 0.3, alpha = 0.4) +
  facet_wrap(~bhi_id, scales ="free_y")

# Plot by subbasins
ggplot(secchi_rgns_shp) + 
  geom_point(aes(year, secchi_m), size = 0.3, alpha = 0.4) +
  facet_grid(cols = vars(subbasin), rows = vars(country))

ggplot(secchi_rgns_shp) + 
  geom_point(aes(year, secchi_m, colour = supplier), size = 0.3, alpha = 0.4) +
  facet_wrap(~subbasin, scales ="free_y")

ggplot(secchi_rgns_shp) + 
  geom_point(aes(month, secchi_m, colour = supplier), size = 0.3, alpha = 0.4) +
  facet_wrap(~subbasin, scales ="free_y")

```

<br/>

#### 3.1.2 Mean secchi Calculation

**Calculate mean monthly value for each summer month (subbasin)** 
basin monthly mean = mean of all samples within month and subbasin
```{r calculate monthly mean secchi}
sub_mean_months <- secchi_rgns_shp %>% 
  dplyr::select(year, month, subbasin, secchi_m)%>%
  group_by(year, month, subbasin) %>%
  summarise(mean_secchi = round(mean(secchi_m, na.rm=TRUE), 1)) %>%
  ungroup()

## Plot mean monthly value by subbasin
ggplot(sub_mean_months) + 
  geom_point(aes(year, mean_secchi, colour = factor(month))) +
  geom_line(aes(year, mean_secchi, colour = factor (month))) +
  facet_wrap (~subbasin) +
  scale_y_continuous(limits = c(0,10)) +
  ggtitle("Mean Monthly Secchi")

## There might be some outlier to check (TO-DO)
```


**Calculate summer mean secchi (subbasin)**
basin summer mean = mean of basin monthly mean values
```{r calculate mean summer secchi}
sub_mean_months_summer <- sub_mean_months %>% 
  dplyr::select(year, subbasin, mean_secchi) %>%
  group_by(year, subbasin) %>%
  summarise(mean_secchi = round(mean(mean_secchi,na.rm=TRUE),1)) %>%
  ungroup()  #in mean calculation all some months to have NA, ignore for that years calculation

# Plot summer mean secchi
ggplot(sub_mean_months_summer) + 
  geom_point(aes(year, mean_secchi)) +
  geom_line(aes(year, mean_secchi)) +
  facet_wrap(~subbasin)+
  scale_y_continuous(limits = c(0,10)) +
  ggtitle("Mean Summer Secchi")

## Only few outliers to check (TO-DO)
```

**Calculate mean monthly value for each summer month (bhi basins)** 
basin monthly mean = mean of all samples within month and bhi basins
```{r calculate monthly mean secchi}
bhi_mean_months <- secchi_rgns_shp %>% 
  dplyr::select(year, month, bhi_id, secchi_m)%>%
  group_by(year, month, bhi_id) %>%
  summarise(mean_secchi = round(mean(secchi_m, na.rm=TRUE), 1)) %>%
  ungroup()

## Plot mean monthly value by bhi basin
ggplot(bhi_mean_months) + 
  geom_point(aes(year, mean_secchi, colour = factor(month)), size = 1.2, alpha = 0.4) +
  geom_line(aes(year, mean_secchi, colour = factor(month)), alpha = 0.3) +
  facet_wrap (~bhi_id) +
  scale_y_continuous(limits = c(0,10)) +
  ggtitle("Mean Monthly Secchi")

## There might be some outlier to check (TO-DO)
```

**Calculate summer mean secchi (bhi basins)**
basin summer mean = mean of basin monthly mean values
```{r calculate mean summer secchi}
bhi_mean_months_summer <- bhi_mean_months %>% 
  dplyr::select(year, bhi_id, mean_secchi) %>%
  group_by(year, bhi_id) %>%
  summarise(mean_secchi = round(mean(mean_secchi,na.rm=TRUE),1)) %>%
  ungroup()  #in mean calculation all some months to have NA, ignore for that years calculation

# Plot summer mean secchi
ggplot(bhi_mean_months_summer) + 
  geom_point(aes(year, mean_secchi)) +
  geom_line(aes(year, mean_secchi)) +
  facet_wrap(~bhi_id)+
  scale_y_continuous(limits = c(0,10)) +
  ggtitle("Mean Summer Secchi")

## There might be some outlier to check (TO-DO)
```

**Plot summer secchi with target values indicated**
Horizontal lines are HELCOM target values 
```{r summer secchi with target }
target <- left_join(sub_mean_months_summer, target_secchi, by = "subbasin")
# head(secchi_target)

ggplot(target) + 
  geom_point(aes(year, mean_secchi)) +
  geom_line(aes(year, target_secchi)) +
  facet_wrap(~subbasin) +
  scale_y_continuous(limits = c(0,10)) +
  ggtitle("Mean Summer Secchi against Target values")
```

#### 3.1.3 Status Calculation

##### 3.1.3.1 Explore status years

All basins have data for 2019, which will be used as the status year, except for seven basins (which have 2018): 
- Aland Sea, Bay of Mecklenburg, Gdansk Basin, Great Belt, Gulf of Finland, Gulf of Riga, and Kiel Bay.

For these seven basins, their most recent year's data is used for current status calculations. 

```{r what is the last year, results = "hide"}
## get the last year of non-NA data
last_year = target %>%
  filter(!is.na(mean_secchi)) %>%
  group_by(subbasin) %>%
  summarise(last_year = last(year)) %>%
  print(n=15)

## which are not in 2019
last_year %>% filter(last_year < 2019)
```

##### 3.1.3.2 Calculate status for each subbasin

**Status calculation with raw (non-modeled) mean summer secchi by subbasin** 

Status must be calculated in data prep because it's calculated on the basin level and then applied to all regions.  

```{r Status calculation}
## Define constants for status calculation
min_year = 2000        # earliest year to use as a start for regr_length timeseries
##data already filtered for 
regr_length = 10       # number of years to use for regression
future_year = 5        # the year at which we want the likely future status
min_regr_length = 5    # min actual number of years with data to use for regression.

## Basin data with target
target

## Calculate basin status = basin_mean/basin_target
basin_status <- target %>%
  mutate(., status =  pmin(1, mean_secchi/target_secchi)) %>%
  dplyr::select(subbasin, year, status) 

## Assign basin status to BHI regions
bhi_status <- basin_status %>%
  group_by(subbasin) %>%
  filter(year == max(year)) %>% #select last year of data for status in each basin (this means status year differs by basin)
  mutate(status = round(status*100)) %>% #status is whole number 0-100
  ungroup() %>%
  left_join(basin_lookup, by = "subbasin") %>% #join bhi regions to basins
  mutate(dimension = 'status') %>%
  dplyr::select(bhi_id, dimension, score=status)

```
<br/>


##### 3.1.3.3 Calculate Trend 

```{r trend calculation}
## Calculate basin trend
basin_trend <- basin_status %>%
  group_by(subbasin) %>%
  do(tail(. , n = regr_length)) %>%  # calculate trend only if there is at least X years of data (min_regr_length) in the last Y years of time serie (regr_length)
  do({if(sum(!is.na(.$status)) >= min_regr_length)
    data.frame(trend_score = 
                 max(-1, min(1, coef(lm(status ~ year, .))['year'] * future_year)))
    else data.frame(trend_score = NA)}) %>%
  ungroup() 

## Assign basin trend to BHI regions
bhi_trend <- left_join(basin_lookup, basin_trend, by = "subbasin") %>%
  mutate(score = round(trend_score,2),
         dimension = "trend")%>%
  dplyr::select(bhi_id, dimension, score )
```


##### 3.1.3.4 Plot Status and Trend

**Plot basin status**
Basin status is initially a value between 0 - 1.  Calculated for each year between 2000 and 2019.  
```{r plot basin status}
ggplot(basin_status) + 
  geom_point((aes(year, status))) +
  facet_wrap(~subbasin) +
  theme(axis.text.x = element_text(colour="grey20", size=8, angle=90, 
                                   hjust=.5, vjust=.5, face = "plain"))
```

**Plot BHI region status and trend values**
Status values can range from 0-100 -- this is the status for the *most recent* year. In most cases this is 2019. No status or trend for 5 or 6 - these are The Sound, which had no data. 
```{r bhi status and trend plot}
ggplot(rbind(bhi_status, bhi_trend)) + 
  geom_point(aes(bhi_id, score), size=2) +
  facet_wrap(~dimension, scales = "free_y") +
  xlab("BHI region")
```

NOTE: Look at status transformation from BHI 1.0!!!

### 3.2 Oxygen Debt Indicator

#### 3.2.1 Match BHI regions

**Spatial visualization of Oxygen data**
```{r map showing datasets overlaps, results = "show", message = FALSE, echo = TRUE, fig.height= 25, fig.width=11}
# assign secchi data to bhi basins
oxyg_rgns_shp <- join_rgns_info(
  class(combined_ox_rawdata), helcomID_col = "helcom_id", country_col = "country", 
                           latlon_vars = c("^lat", "^lon"), return_spatial = FALSE, 
                           rgn_shps_loc = file.path(dirname(dir_B), "Shapefile"), buffer_shp = NULL)

# Plot by bhi regions
ggplot(oxyg_rgns_shp) + 
  geom_point(aes(year, oxygen_ml_l), size = 0.3, alpha = 0.4) +
  facet_grid(cols = vars(country), rows = vars(bhi_id))

ggplot(oxyg_rgns_shp) + 
  geom_point(aes(year, oxygen_ml_l), size = 0.3, alpha = 0.4) +
  facet_wrap(~bhi_id, scales ="free_y")

ggplot(oxyg_rgns_shp) + 
  geom_point(aes(month, oxygen_ml_l), size = 0.3, alpha = 0.4) +
  facet_wrap(~bhi_id, scales ="free_y")

# Plot by subbasins
ggplot(oxyg_rgns_shp) + 
  geom_point(aes(year, oxygen_ml_l), size = 0.3, alpha = 0.4) +
  facet_grid(cols = vars(subbasin), rows = vars(country))

ggplot(oxyg_rgns_shp) + 
  geom_point(aes(year, oxygen_ml_l, colour = country), size = 0.7) +
  facet_wrap(~Subbasin, scales ="free_y") 

ggplot(oxyg_rgns_shp) + 
  geom_point(aes(month, oxygen_ml_l, colour = country), size = 0.7) +
  facet_wrap(~Subbasin, scales ="free_y") 
  
# Plot by country
ggplot(oxyg_rgns_shp) + 
  geom_point(aes(year, oxygen_ml_l), size = 0.7) +
  facet_wrap(~country, scales ="free_y")

```


## 4. Visualizing Data Layers

### 4.1 Map

```{r SOME MAAP, results = "show", message = FALSE, echo = TRUE, fig.width = 9.5}
library(leaflet)
source(here::here("R", "spatial.R"))
regions_shape()

## join datalayers with simplified spatial data for mapping
plotshp <- rmapshaper::ms_simplify(input = BHI_rgns_shp) %>% 
  sf::st_as_sf() %>% 
  dplyr::select(rgn_nam, rgn_key, Subbasin, HELCOM_ID, region_id = BHI_ID, Area_km2) %>% 
  left_join(
    DATALAYERS,
    by = "region_id"
  )
  
pal <- leaflet::colorNumeric("RdYlBu", log10(seq(1, 5, 0.01)), "#fcfcfd", reverse = TRUE)

## create map
leaflet(data = plotshp) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>%
  setView(18, 59, zoom = 5) %>% 
  addMapPane("popup", zIndex = 450) %>% 
  
  addPolygons(
    stroke = TRUE, opacity = 0.3, weight = 1, fillOpacity = 0.95,
    fillColor = ~palProd(DATALAYER),
    group = "GROUPNAME"
  ) %>% 
  
  ## layers controls, popup layer, and formatting
  addLayersControl(
    baseGroups = c(
      BASEGROUPS
    ),
    options = layersControlOptions(collapsed = FALSE)
  ) %>% 
  addPolygons(
    popup = paste(
      "<h5><strong>", "Region:", "</strong>",
      plotshp$Name, "</h5>",
      "<h5><strong>", "BHI Region ID:", "</strong>",
      plotshp$region_id, "</h5>"
    ),
    fillOpacity = 0,
    stroke = FALSE,
    options = pathOptions(pane = "popup")
  )
```

<br>

## 5. Considerations for `BHI3.0`

<br>

## 6. References

```{r References, child = refs_path, results = "asis", echo = FALSE}
```

