
## 2. Data

```{r preamble data, echo = FALSE, include = FALSE, error = FALSE}
source(here::here("R", "setup.R"))
```

### 2.1 Datasets with Sources
<br/>

**Plastic Marine Pollution Global Dataset**  
<!-- dataset save location BHI_share/BHI 2.0/Goals/ -->

Dataset posted on 03.05.2014, 13:14 by Marcus Eriksen. *This is a global dataset of 1571 locations where surface manta tows were conducted.  Samples were divided into 4 size categories. Weights and particle counts were recoreded for each category.* Dataset supplement to paper [Plastic Pollution in the World's Oceans: More than 5 Trillion Plastic Pieces Weighing over 250,000 Tons Afloat at Sea](https://doi.org/10.1371/journal.pone.0111913) by Marcus Eriksen et al.


```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Sheet1"), 
  c("All columns"),
  c("Remove column P and O (Comments) and save as .csv")))

colnames(tab) <- c("Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [Plastic Marine Pollution Global Dataset, shared via figshare](http://dx.doi.org/10.6084/m9.figshare.1015289) <br/> Downloaded 10 March, 2020 by Ellie Campbell")
```

---

<br/>


### 2.2 Centralization & Normalization

```{r load raw data, message = FALSE, echo = TRUE, warning = FALSE, results = "hide"}
## root location of the raw data
dir_rawdata <- file.path(dir_B, "Goals", "CW", "TRA")
marinepastic_raw <- readr::read_csv(file.path(dir_rawdata, "Eriksen_PlasticMarine", "marineplastic_eriksen.csv"))
```

#### 2.2.1 Rename Fields/Variables

#### 2.2.2 Standardize Units

#### 2.2.3 Save to BHI Database

<br/>

### 2.3 Initial Data Exploration

#### 2.3.1 Compare versus Previous Years Data

#### 2.3.2 Timeseries Plots

```{r CODE CHUNK WITH FIGURE OR GRAPH, results = "show", message = FALSE, echo = TRUE, fig.width = 9.5, fig.height = 4.5}
```

#### 2.3.3 Map

