
```{r tra data preamble, echo = FALSE, include = FALSE, error = FALSE}
source(here::here("R", "data.R"))
source(here::here("R", "spatial.R"))
source(here::here("R", "setup.R"))

## root location of the raw data
dir_B <- file.path(dirname(dir_prep), "bhi-data", "BHI 2.0")
dir_rawdata <- file.path(dir_B, "Goals", "CW", "TRA")
dir_intermediate <- file.path(here::here("data", "CW", "trash", version_year, "intermediate"))
```

## 2. Data

### 2.1 Datasets with Sources {-}

#### 2.1.1 Marine Litter {-}

**Modelled Marine Litter and Trash Data** 

Data for litter/trash from [Plastic waste inputs from land into the ocean(Jambeck et al 2015)](http://science.sciencemag.org/content/347/6223/768.full.pdf+html)

Raw data were stored in `bhi-1.0-archive/baltic2015/prep/CW/trash/raw` and now also stored in 

```{r write jamberk trash data, eval = FALSE}
write_csv(
  NUTS2_BHI_ID_Pop_density_in_buffer, 
  file.path(dir_intermediate, "NUTS2_BHI_ID_Pop_density_in_buffer.csv")
)

write_csv(
  X1260352_SupportingFile_Suppl_Excel_seq1_v1, 
  file.path(dir_intermediate, "trash_jambeck_baltic.csv")
)
```

The Jambeck et al method includes:

1. the mass of waste generated per capita annually; 
2. the percentage of waste that is plastic; and 
3. the percentage of plastic waste that is mismanaged and, therefore, has the potential to enter the ocean as marine debris (12) (data S1). By applying a range of conversion rates from mismanaged waste to marine debris, we estimated the mass of plastic waste entering the ocean from each country in 2010, used population growth data (13) to project the increase in mass to 2025, and predicted growth in the percentage of waste that is plastic.

Footnotes from .xls file column headers 
1 - Based upon 2010 Gross National Income
2 - Based upon a 50 km coastal buffer created in GIS with global population densities
3 - Bold data were taken directly from World Bank estimates
4 - Bold data were taken directly from World Bank estimates 
5 - Using a model developed for this study (see Supplemental Information)
6 - From U.S. national litter study
7 - Calculated values
8 - Economic status changed from 2005 to 2010; waste generation rate and %plastic were assigned using average values for the 2010 economic category  

**Litter Exchange Data = (HH and LT) – continuously updated**
<!-- dataset save location BHI_share/BHI 2.0/Goals/CW/TRA/DATRAS -->

For the Baltic Sea, data are delivered by Denmark, Estonia, Germany, Latvia, Lithuania, Poland, Russia and Sweden and are related mainly to BITS. In the Baltic Sea, the gears TVS and TVL are mostly used. Even if the coverage by year and country is quite good, some data gaps are still remaining (for details and vocab see [DATRAS vocabulary](https://vocab.ices.dk/?theme=1)).

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Data products:", "Litter exchange data"), 
  c("Survey:", "BITS (Baltic International Trawl Survey)"),
  c("Quarter(s):", "All"),
  c("Year(s):", "All")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [ICES database (DATRAS)](https://datras.ices.dk/Data_products/Download/Download_Data_public.aspx) <br/> Downloaded 8 May 2020 by Andrea De Cervo")
```

---

<br/>

**Litter Assessment Output (OSPAR) = (merged HH, LT, and GIS-info) – updated once a year** 
<!-- dataset save location BHI_share/BHI 2.0/Goals/CW/TRA/DATRAS -->

For the Baltic Sea, data are delivered by Denmark, Estonia, Germany, Latvia, Lithuania, Poland, Russia and Sweden and are related mainly to BITS. In the Baltic Sea, the gears TVS and TVL are mostly used. Even if the coverage by year and country is quite good, some data gaps are still remaining (for details and vocab see [DATRAS vocabulary](https://vocab.ices.dk/?theme=1)).

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Data products:", "Litter Assessment Output"), 
  c("Survey:", "BITS (Baltic International Trawl Survey)"),
  c("Quarter(s):", "All"),
  c("Year(s):", "All")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [ICES database (DATRAS)](https://datras.ices.dk/Data_products/Download/Download_Data_public.aspx) <br/> Downloaded 8 May 2020 by Andrea De Cervo")
```


**Plastic Marine Pollution Global Dataset**  
<!-- dataset save location BHI_share/BHI 2.0/Goals/CW/TRA/Eriksen_PlasticMarine -->

Dataset posted on 03.05.2014, 13:14 by Marcus Eriksen. *This is a global dataset of 1571 locations where surface manta tows were conducted.  Samples were divided into 4 size categories. Weights and particle counts were recoreded for each category.* Dataset supplement to paper [Plastic Pollution in the World's Oceans: More than 5 Trillion Plastic Pieces Weighing over 250,000 Tons Afloat at Sea](https://doi.org/10.1371/journal.pone.0111913) by Marcus Eriksen et al.


```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Sheet1"), 
  c("All columns"),
  c("Remove column P and O (Comments) and save as .csv")))

colnames(tab) <- c("Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [Plastic Marine Pollution Global Dataset, shared via figshare](http://dx.doi.org/10.6084/m9.figshare.1015289) <br/> Downloaded 10 March, 2020 by Ellie Campbell")
```

<br/>

**Overview of monitoring methods in the Baltic Sea**

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Beach litter", "Marlin/ OSPAR", "Poland", "years"), 
  c("Floating litter", "Net curtains / from ships", "Estonia", "years"),
  c("Seafloor litter", "BITS (Baltic International Trawl Survey)", "Denmark (?)", "years"),
  c("Microplastics", "", "Finland", "years")))

colnames(tab) <- c("Litter monitoring", "Method", "Lead Country", "Years")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [Blastic](https://www.sei.org/wp-content/uploads/2017/12/blastic-overview-of-available-monitoring-methods-final.pdf) <br/> [HELCOM Meeting notes](https://portal.helcom.fi/workspaces/MARINE%20LITTER-92/Beach%20litter%20pre%20core%20indicator/Memo%20on-line%20meeting%20HELCOM%20EN-ML_27_1_2020.pdf)Downloaded 18 May 2020 by Andrea De Cervo")
```

<br>

### 2.2 Centralization & Normalization {-}

**Marine litter data**
```{r load raw data, message = FALSE, echo = TRUE, warning = FALSE, results = "hide"}
## root location of the raw data
dir_rawdata <- file.path(dir_B, "Goals", "CW", "TRA")
marinepastic_raw <- readr::read_csv(file.path(dir_rawdata, "Eriksen_PlasticMarine", "marineplastic_eriksen.csv"))
litterdata_raw <- readr::read_csv(file.path(dir_rawdata, "DATRAS", "Litter Exchange Data", "litter_exchange_data.csv"))
litterassess_raw <- readr::read_csv(file.path(dir_rawdata, "DATRAS", "Litter Assessment Output", "litter_assessment_output.csv"))
```

