---
title: "Mariculture - Food Provision Subgoal"
output:
  html_document:
    toc: true
    toc_depth: 4
---


```{r preamble prep, message = FALSE}
loc <- here::here("prep", "MAR")

source(here::here("R", "setup.R"))
source(here::here("R", "prep.R"))
knitr::opts_chunk$set(message = FALSE, warning = FALSE, results = "hide", fig.width = 9.5)

bkgd_path <- here::here("supplement", "goal_summaries", "MAR.Rmd")
data_path <- here::here("data", "MAR", version_year, "mar_data.rmd")
refs_path <- file.path(loc, "mar_references.Rmd")
```

## 1. Background

```{r Background, child = bkgd_path, results = "asis", echo = FALSE}
```

<br/>

## 2. Data

**Rainbow trout production (Denmark, Sweden and Finland) and Finfish (Denmark, Germany, Sweden and Finland)**

This prep document is used to generate and explore the following data layers:

- `mar_harvest_bhi2019.csv` 
- `mar_harvest_bhi2019.csv` 
- `fis_ffmsy_bhi2019.csv` 

These are saved to the `layers/v2019` folder. Saved to `data/MAR/v2019/intermediate` is two intermediate datasets: `mar_full_merged_dataset.csv` and `mar_gapfilling.csv`. All these are derived from or informed by the raw datasets from HELCOM, national databases, and FAO.

<br>

<!-- ## 2. Data --- header in the child  document -->
```{r Data, child = data_path, results = "asis", echo = FALSE}
```

<br/>

## 3. Prep: Wrangling & Derivations, Checks/Evaluation, Gapfilling

### 3.1 Reorganizing/wrangling

This section prepares production and spatial data for MAR. 

<br>

#### 3.1.1 Load datasets

<!-- Swedish aquaculture facilities --> 


## 4. Visualizing Data Layers

<br/>

## 5. References

```{r References, child = refs_path, results = "asis", echo = FALSE}
```
