
```{r mar preamble prep, message = FALSE}
loc <- here::here("prep", "MAR")

source(here::here("R", "setup.R"))
knitr::opts_chunk$set(message = FALSE, warning = FALSE, results = "hide", fig.width = 9.5)

bkgd_path <- here::here("supplement", "goal_summaries", "MAR.Rmd")
data_path <- here::here("data", "MAR", version_year, "mar_data.rmd")
refs_path <- file.path(loc, "mar_references.Rmd")
```

<br>

## 1. Background {-}

```{r mar background, child = bkgd_path, results = "asis", echo = FALSE}
```

<br/>

## 2. Data {-}

**Rainbow trout production (Denmark, Sweden and Finland) and Finfish (Denmark, Germany, Sweden and Finland)**

This prep document is used to generate and explore the following data layers:

- `mar_harvest_bhi2019.csv` 
- `mar_harvest_bhi2019.csv` 
- `fis_ffmsy_bhi2019.csv` 

These are saved to the `layers/v2019` folder. Saved to `data/MAR/v2019/intermediate` are two intermediate datasets: `mar_full_merged_dataset.csv` and `mar_gapfilling.csv`. All these are derived from or informed by the raw datasets from HELCOM, national databases, and FAO.

<br>

<!-- ## 2. Data --- header in the child  document -->
```{r mar data, child = data_path, results = "asis", echo = FALSE}
```

<br/>

## 3. Prep: Wrangling & Derivations, Checks/Evaluation, Gapfilling {-}

```{r load mar raw datasets, echo = FALSE, include = FALSE}
combined_rawdata <- read_csv(file.path(dirname(data_path), "intermediate", "mar_merged_rawdata.csv"))
combined_w_rgns <- read_csv(file.path(dirname(data_path), "intermediate", "mar_combined_w_rgns.csv"))
lgl2dbl <- c("n_kg", "p_kg", "split_N", "split_P")
combined_w_rgns[,lgl2dbl] <- sapply(combined_w_rgns[, lgl2dbl], as.numeric) 
```

### 3.1 Reorganizing/wrangling {-}

### 3.2 Imputation/Gapfilling of merged datasets {-}

#### 3.2.1 Investigate estimates with linear models {-}

```{r investigate nutrient estimates with linear models, echo = TRUE}
## gapfill using raw merged data, 
## not data split and assigned to BHI regions!
plotdf <- combined_rawdata %>% 
  filter(!is.na(p_kg) & !is.na(n_kg)) %>% 
  filter(produced_tonnes > 20) %>% 
  ## to use predict function later based on lm obj, need these names in models
  dplyr::rename(split_production = produced_tonnes, split_N = n_kg, split_P = p_kg)

## check relationships between production and nutrients
## quite linear, so probably ok to gapfill using this relationship...
## note: remember 'split' values here are actually original data renamed!
p_mdl <- lm(split_P ~ 0 + split_production, data = plotdf)
n_mdl <- lm(split_N ~ 0 + split_production, data = plotdf) 
# summary(p_mdl)
# summary(n_mdl)

## production estimates with linear models
## better without phosphorus, among other things because of negative coefficient...
prod_mdl <- lm(split_production ~ 0 + split_N, data = plotdf) 
# summary(prod_mdl)
```

```{r nutrient and production estimates, results = "show", message = FALSE, echo = TRUE, fig.width = 9.5, fig.height = 4}
ggplot(plotdf) +
  geom_abline(slope = coef(p_mdl)[["split_production"]], intercept = 0, color = "grey") +
  geom_point(aes(x = split_production, y = split_P), color = "blue", alpha = 0.75) +
  labs(x = "Production", y = "Phosphorus Use\n")

ggplot(plotdf) +
  geom_abline(slope = coef(prod_mdl)[["split_N"]], intercept = 0, color = "grey") +
  geom_point(aes(x = split_N, y = split_production), color = "blue", alpha = 0.75) +
  labs(x = "Nitrogen Use\n", y = "Production")
```

<br>

#### 3.2.2 Gapfill and save MAR layers {-}

```{r mar gapfilling and save datasets, echo = TRUE}
## create dataset with nutrient ESTIMATES (kg) from COMBINED data
mar_gapfilling <- combined_w_rgns %>% 

  ## identify where to gapfill
  mutate(
    gapfill_nutrients = (is.na(split_P)|is.na(split_N)) & !is.na(split_production),
    gapfill_prod = is.na(split_production) & !is.na(split_N) & !is.na(split_P)
  ) %>% 
  
  ## estimates based on production or nutrient use
  cbind(
    estim_p_kg = predict(p_mdl, combined_w_rgns),
    estim_n_kg = predict(n_mdl, combined_w_rgns),
    estim_prod = predict(prod_mdl, combined_w_rgns)
  ) %>% 
  
  ## gapfilling nutrients based on production and vice versa...
  ## production from regression w N and P, so cannot gapfill if either are NA
  mutate(
    split_N_gf = ifelse(is.na(split_N), estim_n_kg, split_N),
    split_P_gf = ifelse(is.na(split_P), estim_p_kg, split_P),
    split_prod_gf = ifelse(gapfill_prod, estim_prod, split_production),
    gapfilled = gapfill_nutrients|gapfill_prod
  ) %>%
  select(
    year, country, region_id, farm_species, 
    gapfilled, gapfill_nutrients, gapfill_prod, gapfill_nutrients,
    estim_n_kg, estim_p_kg, estim_prod, 
    split_N_gf, split_P_gf, split_prod_gf
  ) %>% 
  mutate(year = as.character(year)) %>% 
  mutate(year = as.numeric(year))

mar_datalayers <- mar_gapfilling %>% 
  ## get summarized values by BHI region
  group_by(year, region_id) %>% 
  summarize(
    n_kg = sum(split_N_gf, na.rm = TRUE),
    p_kg = sum(split_P_gf, na.rm = TRUE),
    produced_tonnes = sum(split_prod_gf, na.rm = TRUE),
    pct_prod_gf = sum(gapfill_prod)/n(),
    pct_nutrient_gf = sum(gapfill_nutrients)/n(),
    pct_gf = sum(gapfilled)/n()
  ) %>% 
  ungroup() %>% 
  ## nutrient ratios
  mutate(
    ## calculate ratios of nutrients used to fish produced
    ## grams per kg fish, unit conversion factors cancel
    n_ratio = ifelse(produced_tonnes == 0, 0, n_kg/produced_tonnes),
    p_ratio = ifelse(produced_tonnes == 0, 0, p_kg/produced_tonnes)
  )
```

```{r save mar intermediate datasets, echo = TRUE, eval = FALSE}
## save gapfilled dataset and dataset recording where gapfilling occurred
write_csv(mar_gapfilling, here::here("data", "MAR", version_year, "intermediate", "mar_gapfilling.csv"))

write_csv(
  mar_datalayers %>% 
    select(year, region_id, produced_tonnes),
  file.path(dir_layers, version_year, sprintf("mar_harvest_bhi%s.csv", assess_year))
)
write_csv(
  mar_datalayers %>% 
    select(year, region_id, n_kg, p_kg, n_ratio, p_ratio),
  file.path(dir_layers, version_year, sprintf("mar_nutrients_bhi%s.csv", assess_year))
)
```

---

<br>

## 4. Visualizing Data Layers {-}

<br/>

### 4.1 Nutrient Use {-}

As noted above, the HELCOM recommnedation is 50gN and 7kgP per kg of fish. The plots below explore differences between recorded and recommended nutrient use. The histogram plots show the distributions of differences with max allowable nutrients minus actual or estimated, so negative values are in excess of the recommendation.

```{r nutrients allowances minus nutrients values, echo = TRUE}
## using gapfilled data
plotdf <- mar_datalayers %>% 
  left_join(
    read_csv(here::here("supplement", "lookup_tabs", "rgns_complete.csv")) %>% 
      select(region_id, country = region_name) %>% 
      mutate(country = str_replace(country, "[A-Za-z ]+, ", "")) %>% 
      distinct(),
    by = "region_id"
  ) %>% 
  mutate(
    region_id = as.factor(region_id), 
    diffsN = 50*produced_tonnes - n_kg, 
    diffsP = 7*produced_tonnes - p_kg,
    diffsN_Ratio = 50 - n_ratio, 
    diffsP_Ratio = 7 - p_ratio
  ) %>% 
  rename(Year = year, Production = produced_tonnes)
  
histdf <- plotdf %>% 
  select(Year, country, region_id, diffsN, diffsP, diffsN_Ratio, diffsP_Ratio, Production) %>% 
  tidyr::pivot_longer(cols = c(diffsN, diffsP))

## distributions differences: max allowable nutrients minus actual or estimated
## negative values are in excess of the allowance
hist_all <- ggplot(histdf) +
  geom_histogram(aes(value), color = "blue", fill = NA, binwidth = 800) +
  facet_wrap(~ name, nrow = 2) +
  labs(x = NULL, y = NULL)

hist_lessthan0 <- ggplot(filter(histdf, value < 0)) +
  geom_histogram(aes(value), color = "blue", fill = NA) +
  facet_wrap(~ name, nrow = 2) +
  labs(x = NULL, y = NULL)

## siginificant difference between more recent and older N and P values?
# wilcox.test(value ~ timegroup, diffs_timewise)

histN_yrs <- ggplot(histdf %>% 
    filter(name == "diffsN", value < 5000) %>% 
    mutate(timegroup = ifelse(Year < 2008, "N before 2008", "N after 2008"))) + 
  geom_histogram(aes(value), color = "blue", fill = NA, binwidth = 50) +
  facet_wrap(~ timegroup, nrow = 2, scales = "free_y") +
  labs(x = NULL, y = NULL)

histP_yrs <- ggplot(histdf %>% 
    filter(name == "diffsP", value < 5000) %>% 
    mutate(timegroup = ifelse(Year < 2016, "P before 2008", "P after 2008"))) + 
  geom_histogram(aes(value), color = "blue", fill = NA, binwidth = 50) +
  facet_wrap(~ timegroup, nrow = 2, scales = "free_y") +
  labs(x = NULL, y = NULL)
```


**Nutrient Use: Recommended minus Reported Use (g/kg Fish)**

```{r mar nutrients histograms, results = "show", message = FALSE, echo = TRUE, fig.width = 9.5, fig.height = 4}
gridExtra::grid.arrange(hist_all, hist_lessthan0, histN_yrs, histP_yrs, nrow = 1)
```

---

<br>

### 4.2 Timeseries Plots {-}

```{r explore mar data by country, results = "show", message = FALSE, echo = TRUE, fig.width = 9.5}
## time series of production and nutrient use by country
timeseriesdf <- plotdf %>% 
  group_by(Year, country) %>% 
  mutate(
    Nmax = sum(50*Production, na.rm = TRUE), 
    Pmax = sum(7*Production, na.rm = TRUE)
  ) %>% 
  ungroup() %>% 
  rename(Nitrogen = n_kg, Phosphorus = p_kg)

ggplot(timeseriesdf, aes(Year, Production, fill = region_id)) +
  geom_area(position = "stack", alpha =  0.5, show.legend = FALSE) + 
  facet_wrap(~ country, ncol = 1, scales = "free_y") + 
  labs(y = "Mariculture production (tonnes, gapfilled)\n")

nutrientTimeseries <- ggplot() +
  geom_area(
    mapping = aes(x = Year, y = Value, fill = RegionID),
    data = timeseriesdf %>% 
      tidyr::pivot_longer(cols = c(Nitrogen, Phosphorus, Nmax, Pmax, Production)) %>% 
      mutate(Value = round(value), RegionID = region_id) %>% 
      filter(name %in% c("Nitrogen", "Phosphorus"), Year > 1975),
    position = "stack", 
    alpha =  0.5, 
    show.legend = FALSE
  ) + 
  geom_line(
    mapping = aes(x = Year, y = Value),
    data = timeseriesdf %>% 
      tidyr::pivot_longer(cols = c(Nitrogen, Phosphorus, Nmax, Pmax, Production)) %>% 
      mutate(Value = round(value)) %>% 
      filter(name %in% c("Nmax", "Pmax"), Year > 1975) %>% 
      mutate(name = ifelse(name == "Nmax", "Nitrogen", "Phosphorus")),
    inherit.aes = FALSE,
    size = 0.4,
    color = "grey",
    show.legend = FALSE
  ) + 
  facet_grid(rows = vars(country), cols = vars(name), scales = "free_y") + 
  labs(y = "Nutrient Use (kg, gapfilled)\n")
# nutrientTimeseries_plotly <- plotly::ggplotly(nutrientTimeseries)
# nutrientTimeseries_plotly$x$layout$showlegend <- FALSE
# nutrientTimeseries_plotly
nutrientTimeseries
```

---

<br/>

### 4.3 Map of Mariculture Data {-}

Below is a map with the `MAR` goal data. Included in the map are data on production (tonnes), nutrient use (grams per kg fish produced), percentage of data gapfilled for (1) production and (2) nutrient use.

```{r mar data map, results = "show", message = FALSE, echo = TRUE, fig.width = 9.5}
library(leaflet)

plotshp <- rmapshaper::ms_simplify(input = BHI_rgns_shp) %>% 
  sf::st_as_sf() %>% 
  dplyr::select(rgn_nam, rgn_key, Subbasin, HELCOM_ID, region_id = BHI_ID, Area_km2) %>% 
  left_join(
    mar_datalayers %>%
      filter(year == 2017) %>% 
      select(region_id, produced_tonnes, n_ratio, p_ratio, pct_prod_gf, pct_nutrient_gf),
    by = "region_id"
  ) %>% 
  ## because production data were split among regions
  ## based on the assumption that each unit area was equally likely to contain given mariculture facility
  ## should present this data as average production per area...
  mutate(produced_tonnes = produced_tonnes/Area_km2) %>% 
  mutate(Name = paste(Subbasin, rgn_nam, sep = ", "))

## create palettes based on min max of data values
palProd <- leaflet::colorNumeric("RdYlBu", log10(seq(1, 5, 0.01)), "#fcfcfd", reverse = TRUE)
palN <- leaflet::colorNumeric("RdYlBu", seq(0, 60, 0.5), "#fcfcfd", reverse = TRUE)
palP <- leaflet::colorNumeric("RdYlBu", seq(0, 8.4, 0.05), "#fcfcfd", reverse = TRUE)
pal_gf <- leaflet::colorNumeric("Greys", seq(0, 1, 0.01))

## create maps
leaflet(data = plotshp) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>%
  setView(18, 59, zoom = 5) %>% 
  addMapPane("popup", zIndex = 450) %>% 
  
  ## production data
  addPolygons(
    stroke = TRUE, opacity = 0.3, weight = 1, fillOpacity = 0.95,
    fillColor = ~palProd(produced_tonnes),
    group = "Production (tonnes)"
  ) %>% 
  
  ## nutrients use data
  addPolygons(
    stroke = TRUE, opacity = 0.3, weight = 1, fillOpacity = 0.95,
    fillColor = ~palN(n_ratio),
    group = "Nitrogen use ratio (grams/kg fish)"
  ) %>% 
  addPolygons(
    stroke = TRUE, opacity = 0.3, weight = 1, fillOpacity = 0.95,
    fillColor = ~palP(p_ratio),
    group = "Phosphorus use ratio (grams/kg fish)"
  ) %>% 
  
  ## layers controls, popup layer, and formatting
  addLayersControl(
    baseGroups = c(
      "Production (tonnes/per km2)", 
      "Nitrogen use ratio (grams/kg fish)", 
      "Phosphorus use ratio (grams/kg fish)"
      # "Production data gapfilled (proportion)",
      # "Nutrient data gapfilled (proportion)"
    ),
    options = layersControlOptions(collapsed = FALSE)
  ) %>% 
  addPolygons(
    popup = paste(
      "<h5><strong>", "Region:", "</strong>",
      plotshp$Name, "</h5>",
      "<h5><strong>", "BHI Region ID:", "</strong>",
      plotshp$region_id, "</h5>",
      "<h5><strong>", "Production:", "</strong>",
      round(plotshp$produced_tonnes, 5), "</h5>",
      "<h5><strong>", "Nitrogen:", "</strong>",
      round(plotshp$n_ratio, 2), "</h5>",
      "<h5><strong>", "Phosphorus:", "</strong>",
      round(plotshp$p_ratio, 2), "</h5>",
      "<h5><strong>", "Gapfill Prop., Nutrients:", "</strong>",
      plotshp$pct_nutrient_gf, "</h5>",
      "<h5><strong>", "Gapfill Prop., Production:", "</strong>",
      plotshp$pct_prod_gf, "</h5>"
    ),
    fillOpacity = 0,
    stroke = FALSE,
    options = pathOptions(pane = "popup")
  )




```

<br>

### 4.4 Gapfilling percentages by Country and Region {-}

```{r mar data gapfilled percentages by country and region id, results = "show", message = FALSE, echo = TRUE, fig.width = 9.5, fig.height = 4}
percentagegf <- ggplot(
  mar_datalayers %>% 
    left_join(
      read_csv(here::here("supplement", "lookup_tabs", "rgns_complete.csv")) %>% 
        select(region_id, country = region_name) %>% 
        mutate(country = str_replace(country, "[A-Za-z ]+, ", "")) %>% 
        distinct(),
      by = "region_id"
    ) %>% 
    filter(year > 2008) %>% 
    mutate(region_id = as.factor(region_id)) %>% 
    mutate(year = as.character(year)),
  aes(fill = region_id, x = year, y = pct_prod_gf) # also visualize: pct_gf, pct_nutrient_gf
)
percentagegf +
  geom_col(position = position_dodge(), alpha = 0.5, show.legend = FALSE) + 
  facet_wrap(~ country) + 
  labs(x = "Year", y = "Percentage gapfilled production data\n", fill = "BHI Region ID")
```

---

<br>

## 5. Considerations for `BHI3.0` {-}

If more adequate data are available, mariculture sites may be matched to BHI regions more precisely using spatial files (shapefiles). In particular, regarding Sweden, there are reports (e.g. [page 8, table 9](https://www.scb.se/contentassets/8964475d3a8b4aa88af32e4fcbc248eb/jo1201_2014a01_sm_jo60sm1501.pdf)) where information about rainbow trout production in "other coasts" is split into two areas: sounth-eastern coast and south-western coast. However, these data are only available until 2014.  

<br>

## 6. References {-}

```{r mar references, child = refs_path, results = "asis", echo = FALSE}
```

<br>
