---
title: "Food Provision (FP) - Mariculture (MAR) Subgoal Data Preparation"
output:
  github_document:
    toc: true
    toc_depth: 3
params: 
    datasource: csv
always_allow_html: true
---


```{r mar preamble prep, message = FALSE}
loc <- here::here("prep", "MAR")

source(here::here("R", "setup.R"))
source(here::here("R", "spatial.R"))
regions_shape(sp_dir = file.path(dirname(dir_B), "Shapefiles"))
knitr::opts_chunk$set(message = FALSE, warning = FALSE, results = "hide", fig.width = 9.5)

bkgd_path <- here::here("supplement", "goal_summaries", "MAR.Rmd")
data_path <- here::here("data", "MAR", version_year, "mar_data.rmd")
refs_path <- file.path(loc, "mar_references.Rmd")
```

<br>

## 1. Background {-}

```{r mar background, child = bkgd_path, results = "asis", echo = FALSE}
```

<br/>

## 2. Data {-}

**Rainbow trout production (Denmark, Sweden and Finland) and Finfish (Denmark, Germany, Sweden and Finland)**

<br>

```{r mar prep data, child = data_path, results = "asis", echo = FALSE}
```

<br/>


## 3. Considerations for `BHI3.0` {-}

- If more adequate data are available, mariculture sites may be matched to BHI regions more precisely using spatial files (shapefiles). In particular, regarding Sweden, there are reports (e.g. [page 8, table 9](https://www.scb.se/contentassets/8964475d3a8b4aa88af32e4fcbc248eb/jo1201_2014a01_sm_jo60sm1501.pdf)) where information about rainbow trout production in "other coasts" is split into two areas: sounth-eastern coast and south-western coast. However, these data are only available until 2014. 

- Also, collect more consistent information and data on mariculture and its sustainable production.

<br>

## 4. References {-}

```{r mar references, child = refs_path, results = "asis", echo = FALSE}
```

<br>
