---
title: "Sense of Place (SP) - Iconic Species (ICO) Subgoal Data Preparation"
output:
  github_document:
    toc: true
    toc_depth: 3
params:
    datasource: csv
---

<br>

```{r preamble prep, message = FALSE, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = TRUE, results = "show", fig.width = 9.5, fig.height = 6)
source(here::here("R", "setup.R"))

loc <- here::here("prep", "ICO")
bkgd_path <- here::here("supplement", "goal_summaries", "ICO.Rmd")
data_path <- here::here("data", "ICO", version_year, "ico_data.rmd")
refs_path <- file.path(loc, "ico_references.Rmd")
```

## 1. Background

```{r Background, child = bkgd_path, results = "asis", echo = FALSE}
```

<br/>

## 2. Data

This prep document is used to generate and explore the following data layers:

- `sp_ico_assessments_bhi2019.csv`

These are saved to the layers folder. These are based on the `bd_spp_all_assessments.csv` dataset saved to the `data/ICO/v2019/intermediate` folder during the Biodiversity species subgoal data prep (see `prep/ICO/v2019/spp_prep.Rmd` for details). All these are derived from or informed by the raw datasets from datasets from the HELCOM Biodiversity database:  Observations data, Red list information data, Species data.

<br>

```{r ico species data, child = data_path, results = "asis", echo = FALSE}
```

<br/>

<br>

## 3. Prep: Wrangling & Derivations, Checks/Evaluation, Gapfilling

```{r creating dataframe with missing fish species info}
subbasins <- c(
  "Kattegat", "Great Belt", "The Sound", "Bay of Mecklenburg", "Kiel Bay",
  "Arkona Basin", "Bornholm Basin", "Western Gotland Basin", "Eastern Gotland Basin",
  "Gdansk Basin", "Northern Baltic Proper", "Gulf of Riga", "Gulf of Finland",
  "Aland Sea", "Bothnian Sea", "The Quark", "Bothnian Bay"
)

## create a dataframe with the info on fish which were missing from the biodiversity goal 'observations' data...
## we need these fish species here even more than for the BD goal, because half the iconic species list are fish...
## not finding specific ranges for pike and trout and also they are coastal/inland not in opensea, so will exclude here
missing_fis <- tibble(
  scientific_name = c(
    "Clupea harengus", "Salmo salar", "Gadus morhua",
    "Clupea harengus", "Salmo salar", "Gadus morhua",
    "Gadus morhua (Kattegat subpopulation)",
    "Gadus morhua (Western Baltic subpopulation)",
    "Gadus morhua (Eastern Baltic subpopulation)",
    # "Esox lucius", "Salmo trutta", 
    "Sprattus sprattus"
  ),
  Subbasin = c(
    list(subbasins[1:17]), list(subbasins[c(15:17, 8)]), list(subbasins[1:12]), 
    list(subbasins[1:17]), list(subbasins[c(15:17, 8)]), list(subbasins[1:12]), 
    list(subbasins[1]), 
    list(subbasins[2:6]), 
    list(subbasins[6:12]),
    # list(subbasins[13:17]), list(subbasins[1:3]), 
    list(subbasins[1:11])
  ),
  no_rec_this_dec = rep(FALSE, 10),
  ## cod has 2007 and a 2013 assessment
  assessment_year_red_list = c(rep(2007, 3), rep(2013, 7))
)

missing_fis <- tidyr::unnest(missing_fis, cols = c(Subbasin))

## still missing ringed seal and European/Eurasian otter...
# missing_mammals <- tidyr::unnest(missing_mammals, cols = c(Subbasin))
```


```{r read in species data saved from biodiversity spp prep and the iconic species list}
## species data with assessment (iucn) categories
sppdata <- rbind(
  read_csv(here::here("data", "ICO", "v2019", "intermediate", "bd_spp_all_assessments.csv")) %>% 
    mutate(scientific_name = str_replace(scientific_name, "Phoca vitulina", "Phoca vitulina vitulina")),
  read_csv(here::here("data", "ICO", "v2019", "intermediate", "fis_redlist.csv")) %>% 
    mutate(scientific_name = ifelse(
      scientific_name == "Gadus morhua" & !is.na(remarks), 
      sprintf("%s (%s)", scientific_name, remarks), 
      scientific_name)
    ) %>% 
    select(scientific_name, species_group, red_list_category, assessment_year_red_list) %>% 
    right_join(missing_fis, by = c("scientific_name", "assessment_year_red_list"))
)

## looking at redlist dataframe, as used in the prep/SPP/v2019/spp_prep...
## Pusa hispida botnica has category VU - Vulnerable, assessed in 2013
ringed_seal <- tibble(
  scientific_name = "Pusa hispida",
  species_group = "Mammals",
  Subbasin = list(subbasins[11:17]),
  no_rec_this_dec = FALSE,
  assessment_year_red_list = 2013,
  red_list_category = "VU - Vulnerable"
)
ringed_seal <- tidyr::unnest(ringed_seal, cols = c(Subbasin))

sppdata_w_ringedseal <- rbind(sppdata, ringed_seal)


## list of 'iconic species' execept birds
ico_spp_list <- read_csv(file.path(loc, "ico_species_selection_fish_mammals.csv"))
ico_spp_list <- rbind(
  filter(ico_spp_list, common_name != "harbour porpoise"),
  ico_spp_list %>% 
    filter(common_name == "harbour porpoise") %>% 
    mutate(latin_name = list(c(
      "Phocoena phocoena (Baltic Sea Sub-population)",
      "Phocoena phocoena (Western Baltic Sea subpopulation)"
    ))) %>% 
    tidyr::unnest(cols = latin_name)
)

# knitr::kable(rename(ico_spp_list, `Common Name` = common_name, `Latin Name` = latin_name), format = "html") %>% 
#   kableExtra::kable_styling(latex_option = c("hold_position"), position = "left")

tab <- ico_spp_list
colnames(tab) <- c("Common Name", "Latin Name")
rownames(tab) <- NULL
htmlTable::htmlTable(tab, align = "lr", col.rgroup = c("none", "#8D909E"))
```


### 3.1 Filtering to keep only Iconic Species

```{r select ico spp}
ico_spp_data <- left_join(ico_spp_list, sppdata_w_ringedseal,  by = c("latin_name" = "scientific_name"))
## still missing mammals: ringed seal (pusa hispida) and european otter (lutra lutra)...
```


### 3.2 Add birds: white-tailed sea eagle & common eider

Discussion with Baltic Sea Centre and SRC (Nov 2016) concluded that 2 bird species as being seen as very important iconic species: white-tailed sea eagle (Haliaeetus albicilla) and common eider (Somateria mollissima). It would be important to add them to the iconic species list and in the ICO calculation. The sea -eagle need to be added manually as it is not in the species redlist from HELCOM but here we could use the status LC from the IUCN criteria based on the Europe IUCN redlist (http://www.iucnredlist.org/details/22695137/0). The common eider is already in the red list as VU.

```{r add iconic birds, message = FALSE, warning = FALSE}
## iconic birds in Baltic Sea include sea eagle and common eider
## adding both these species for all Subbasins
sea_eagle_data <- data_frame(Subbasin = subbasins) %>% 
  mutate(
    common_name = "sea eagle", 
    latin_name = "Haliaeetus albicilla", 
    species_group = "Birds", 
    red_list_category  = "LC - Least Concern", 
    no_rec_this_dec = FALSE,
    assessment_year_red_list = 2013
  )
common_eider_data <- data_frame(Subbasin = subbasins) %>% 
  mutate(
    common_name = "common eider", 
    latin_name = "Somateria mollissima", 
    species_group = "Birds", 
    red_list_category  = "VU - Vulnerable", 
    no_rec_this_dec = FALSE,
    assessment_year_red_list = 2013
  )

ico_spp_w_birds <- rbind(ico_spp_data, sea_eagle_data, common_eider_data)
```


```{r saving ICO data layers}
rgns_complete <- readr::read_csv(here::here("supplement", "lookup_tabs", "rgns_complete.csv"))
  
## Write as sp_ico to layers folder
ico_spp_w_birds %>% 
  right_join(rgns_complete, by = c("Subbasin" = "subbasin")) %>% 
  select(region_id, assessment_year_red_list, scientific_name = latin_name, species_group, red_list_category) %>% 
  write.csv(here::here("layers/sp_ico_assessments_bhi2019.csv"), row.names = FALSE)

## save 'no records this decade' info also as a ICO layer, so can use for penalty/scaling?
# ico_spp_w_birds %>% 
#   right_join(rgns_complete, by = c("Subbasin" = "subbasin")) %>% 
#   filter(no_rec_this_dec == TRUE) %>% 
#   select(region_id, assessment_year_red_list, scientific_name = latin_name) %>% 
#   write.csv(here::here("layers/sp_ico_norec_bhi2019.csv"), row.names = FALSE)
```

<br/>

## 4. Visualizing Data Layers

### 4.1 Iconic Species by Subbasin

```{r plotting presence of different iconic species by subbasin}
ico_spp_w_birds %>% 
  filter(!is.na(Subbasin), assessment_year_red_list == 2013) %>% 
  tidyr::complete(latin_name, Subbasin) %>% 
  mutate(presence = ifelse(is.na(red_list_category), 0, 1)) %>% 
  ggplot() +
  geom_bar(
    aes(latin_name, presence, fill = red_list_category),
    show.legend = FALSE,
    color = NA, 
    stat = "identity"
  ) +
  scale_fill_manual(values = c("indianred", "slategray2", "lightsalmon", "grey")) +
  facet_grid(cols = vars(Subbasin)) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks = element_blank(),
    strip.text.x = element_text(size = 8, angle = 90, hjust = 0)
  ) +
  coord_flip() +
  labs(title = "Presence/Absence of ICO species by basin", y = NULL, x = NULL)
```

<br>

## 5. Considerations for `BHI3.0`

**Species ranges:** One limitation of the observation data used is larger uncertainties in spatial ranges of rare species. Estimation of rare species could be improved to more confidently represent distributions of species around the Baltic Sea; one way to do this would be using IUCN species range maps to establish species occurrence in relation to their spatial habitat area.

**Relation to other assessments:** Improve the link between the BHI and the future Biodiversity assessments by IPBES and use the UN Ocean Biodiversity Information System (OBIS) as national and regional assessments will be performed and linked to IPBES in the future.

**Data consideration:** Lack of quantitative information about the level of effort involved in obtaining the species observations data, or background environmental conditions corresponding to the data points, precludes useful interpretation from observation frequencies of species; rigorous assessment of the historical conditions of all species collectively would require this data which is not readily.

<br>

## 6. References

```{r References, child = refs_path, results = "asis", echo = FALSE}
```
