
## 2. Data

```{r preamble data, echo = FALSE, include = FALSE, error = FALSE}
source(here::here("R", "setup.R"))
source(here::here("R", "spatial.R"))
```

### 2.1 Datasets with Sources
<br/>

**Species Presence, Observations and Red List Assessments**  
<!-- dataset save location BHI_share/BHI 2.0/Goals/BD/baltic_species/baltic_species.shp -->

To obtain data in shapefile format: downloaded, unzipped, opened (observations_final.gdb, observation_points layer) in QGIS, exported to shapefile

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Tab:", "Download"), 
  c("Dataset:", "Observations data ESRI Geodatabase (ZIP)")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [HELCOM Biodiversity Database](http://maps.helcom.fi/website/Biodiversity/) <br/> Downloaded 26 February by Ellie")
```

*These data are obtained from the HELCOM species database, compiled from original data sources: Shark SMHI, Species Observations System (Artportalen), MVM, and Entomological Collections (NHRS) from GBIF.*

<br>

**Red list (IUCN) information, about Species and their Assessment at HELCOM**  
<!-- dataset save location BHI_share/BHI 2.0/Goals/BD/species_with_redlist_category/species_with_redlist_category.csv -->
```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Tab:", "Download"), 
  c("Dataset:", "Red list information data CSV")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [HELCOM Biodiversity Database](http://maps.helcom.fi/website/Biodiversity/) <br/> Downloaded 12 March by Ellie")
```

---

<br/>


### 2.2 Centralization & Normalization

```{r load raw data, message = FALSE, echo = TRUE, warning = FALSE, results = "hide"}
## root location of the raw data
dir_rawdata <- file.path(dir_B, "Goals", "BD")
```

#### 2.2.1 Rename Fields/Variables

#### 2.2.2 Standardize Units

**Remove spatial 'depth' dimension, and assign BHI Region IDs**

```{r remove z spatial dimension and assign BHI region ids}
## read in raw data
## need to rename columns, as names were shortened when exported to shapefile...
spp_raw <- sf::st_read(file.path(dir_rawdata, "raw_species_observations")) %>% 
   dplyr::rename(
    scientific_name = scientific, 
    accepted_name = accepted_n, 
    species_group = species_gr, 
    sampling_group = sampling_g,
    red_list_assessments = red_list_a,
    year_collected = year_colle
  )
buffer_sf <- st_read(
  file.path(dirname(dir_B), "Shapefiles", "BHI_shapefile_25km_buffer"), 
  "BHI_shapefile_25km_buffer"
)
spp_w_rgns <- join_rgns_info(
  sf::st_zm(spp_raw),
  latlon_vars = c("decimal_la", "decimal_lo"),
  return_spatial = TRUE,
  rgn_shps_loc = file.path(dirname(dir_B), "Shapefiles"),
  buffer_shp = buffer_sf
)
```

**Reorganize into Presence/Absence data by BHI Region and Species**

```{r presence absence species data}
species_info <- spp_w_rgns %>% 
  sf::st_drop_geometry() %>% 
  select(
    species_id, 
    scientific_name,
    taxonomica, 
    accepted_name, 
    synonyms, 
    species_group, 
    red_list_assessments
  ) %>% 
  distinct()

spp_presence <- spp_w_rgns %>% 
  sf::st_drop_geometry() %>% 
  select(BHI_ID, species_id, year_collected, red_list_assessments) %>% 
  filter(year_collected > 2016) %>% 
  mutate(red_list_assessments = as.numeric(c)) %>% 
  group_by(BHI_ID, species_id) %>% 
  summarize(redlist_mean = mean(red_list_assessments, na.rm = TRUE)) %>% 
  ungroup() %>% 
  tidyr::pivot_wider(id_cols = species_id, names_from = BHI_ID, values_from = redlist_mean) %>% 
  left_join(species_info, by = "species_id") %>% 
  distinct()
```


#### 2.2.3 Save to BHI Database

```{r save species data}
if(!file.exists(file.path(dir_rawdata, "baltic_species"))){
  dir.create(file.path(dir_rawdata, "baltic_species"))
}
sf::st_write(spp_w_rgns, file.path(dir_rawdata, "baltic_species", "baltic_species.shp"))
```

<br/>

### 2.3 Initial Data Exploration

#### 2.3.1 Compare versus Previous Years Data

#### 2.3.2 Timeseries Plots

```{r CODE CHUNK WITH FIGURE OR GRAPH, results = "show", message = FALSE, echo = TRUE, fig.width = 9.5, fig.height = 4.5}
```

#### 2.3.3 Map


