
```{r spp preamble data, echo = FALSE, include = FALSE, error = FALSE}
source(here::here("R", "setup.R"))
source(here::here("R", "spatial.R"))
```

### 2.1 Datasets with Sources {-}
<br/>

#### 2.1.1 Species Presence, Observations and Red List Assessments {-}

To obtain data in shapefile format: downloaded, unzipped, opened (observations_final.gdb, observation_points layer) in QGIS, exported to shapefile. *These data are obtained from the HELCOM species database, compiled from original data sources: Shark SMHI, Species Observations System (Artportalen), MVM, and Entomological Collections (NHRS) from GBIF.*

<!-- dataset save location BHI_share/BHI 2.0/Goals/BD/baltic_species/baltic_species.shp -->
```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Tab:", "Download"), 
  c("Dataset:", "Observations data ESRI Geodatabase (ZIP)")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [HELCOM Biodiversity Database](http://maps.helcom.fi/website/Biodiversity/) <br/> Downloaded 9 March by Chester")
```

<br>

#### 2.1.2 Red list (IUCN) information, about Species and their Assessment at HELCOM {-}

<!-- dataset save location BHI_share/BHI 2.0/Goals/BD/species_with_redlist_category/species_with_redlist_category.csv -->
```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Tab:", "Download"), 
  c("Dataset:", "Red list information data CSV")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [HELCOM Biodiversity Database](http://maps.helcom.fi/website/Biodiversity/) <br/> Downloaded 12 March by Ellie")
```

#### 2.1.3 Species ID, with taxanomic information for each species in HELCOM database {-}

<!-- dataset save location BHI_share/BHI 2.0/Goals/BD/species_taxa.csv -->

```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Tab:", "Download"), 
  c("Dataset:", "Species data CSV")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [HELCOM Biodiversity Database](http://maps.helcom.fi/website/Biodiversity/) <br/> Downloaded 4 April by Chester")
```

---

<br/>


### 2.2 Adding Region Information to Dataset {-}

```{r spp load raw data, echo = TRUE, message = FALSE, warning = FALSE, results = "hide"}
## root location of the raw data
dir_rawdata <- file.path(dir_B, "Goals", "BD")
```


**Remove spatial 'depth' dimension, and assign BHI Region IDs**

Data points in the raw species observation dataset have latitude/longitude information. We join BHI region and Subbassin information based on the given latitude/longitude. The code chunk below alsoo renames some columns.

```{r remove z spatial dimension and assign BHI region ids, warning = FALSE, message = FALSE}
## create dataset with regions joined
## if already created don't need to recreate... take a long time...
if(!file.exists(file.path(dir_rawdata, "species_obs_w_rgns.csv"))){
  
  ## read in raw data
  ## need to rename columns, as names were shortened when exported to shapefile...
  if(!file.exists(file.path(dir_rawdata, "raw_species_observations.csv"))){
    spp_raw <- sf::st_read(file.path(dir_rawdata, "raw_species_observations")) %>%
      dplyr::rename(
        scientific_name = scientific, 
        accepted_name = accepted_n, 
        species_group = species_gr, 
        sampling_group = sampling_g,
        red_list_assessments = red_list_a,
        year_collected = year_colle
      ) %>% 
      st_drop_geometry()
  } else {
    spp_raw <- data.table::fread(file.path(dir_rawdata, "raw_species_observations.csv"))
  }
  
  buffer_sf <- st_read(
    file.path(dirname(dir_B), "Shapefiles", "BHI_shapefile_25km_buffer"), 
    "BHI_shapefile_25km_buffer"
  )
  
  ## join just spatial info and IDs with BHI IDs, the rejoin 
  ## (smaller dataframes are more manageable!)
  rgns <- join_rgns_info(
    spp_raw %>% select(OBJECTID, decimal_la, decimal_lo),
    latlon_vars = c("decimal_la", "decimal_lo"),
    return_spatial = FALSE,
    rgn_shps_loc = file.path(dirname(dir_B), "Shapefiles"),
    buffer_shp = buffer_sf
  )
  spp_w_rgns <- left_join(
    rename(rgns, OBJECTID = objectid), 
    spp_raw,
    by = c("OBJECTID", "decimal_la", "decimal_lo")
  )
  
  write_csv(spp_w_rgns, file.path(dir_rawdata, "species_obs_w_rgns.csv"))
}
```

<br/>
