
```{r liv data preamble, echo = FALSE, include = FALSE, error = FALSE}
source(here::here("R", "setup.R"))
source(here::here("R", "spatial.R"))
dir_rawdata <- file.path(dir_B, "Goals", "LE", "LIV")
```


### 2.1 Datasets with Sources {-}
<br/>

#### 2.1.1 EuroStat Data {-}

**Employment rates by sex, age and NUTS 2 regions % (lfst_r_lfe2emprt)**
<!-- dataset save location BHI_share/BHI 2.0/Goals/LE/LIV/Employment_NUTS2/lfst_r_lfe2emprt_2019_w_nutscodes -->

```{r employment rates by sex age and nuts2 regions, echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("TIME:", "1999-2019"), 
  c("AGE:", "Y15-74 (code) From 15 to 64 years (label)"),
  c("GEO:", "NUTS2 (Filtering - Nuts level) SHOW: CODES"), # need NUTS codes to match to BHI regions
  c("SEX:", "Total"),
  c("Unit of measure:", "Percentage")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [eurostat database](https://appsso.eurostat.ec.europa.eu/nui/show.do?dataset=lfst_r_lfe2emprt&lang=en) <br/> Downloaded 3 August, 2020 by Ellie Campbell")
```
<br/>

The source for the regional labour market information down to NUTS level 2 is the EU Labour Force Survey (EU-LFS). This is a quarterly household sample survey conducted in all Member States of the EU and in EFTA and Candidate countries.   

The EU-LFS survey follows the definitions and recommendations of the International Labour Organisation (ILO). To achieve further harmonization, the Member States also adhere to common principles when formulating questionnaires. The LFS' target population is made up of all persons in private households aged 15 and over. For more information see the EU Labour Force Survey (lfsi_esms, see paragraph 21.1.).  

The EU-LFS is designed to give accurate quarterly information at national level as well as annual information at NUTS 2 regional level and the compilation of these figures is well specified in the regulation. Microdata including the NUTS 2 level codes are provided by all the participating countries with a good degree of geographical comparability, which allows the production and dissemination of a complete set of comparable indicators for this territorial level.  

- b 	break in time series  
- c 	confidential  
- d 	definition differs, see metadata  
- e 	estimated  
- f 	forecast  
- i 	see metadata (phased out)  
- n 	not significant  
- p 	provisional  
- r 	revised  
- s 	Eurostat estimate (phased out)  
- u 	low reliability  
- z 	not applicable 

<br/>

**Population change - Demographic balance and crude rates at regional level (NUTS 2), Filter to NUTS2 Level**
<!-- dataset save location BHI_share/2.0/Goals/LE/LIV/Population_Jan1/demo_r_gind3_2019_w_nutscodes -->

```{r population change demographics nuts2 regions, echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("TIME (Years):", "2000-2019"), 
  c("GEO:", "NUTS2 (Filtering - Nuts level) SHOW: CODES"), # need NUTS codes to match to BHI regions!
  c("INDIC_DE (Demographic Indicator):", "Population on 1 January - total (JAN)")
))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [eurostat database](https://appsso.eurostat.ec.europa.eu/nui/show.do?dataset=demo_r_gind3&lang=en) <br/> Downloaded 4 August, 2020 by Ellie Campbell")
```

<br/>

Summary explanation about indicator: Eurostat aims at collecting from the EU-28's Member States' data on population on 31st December, which is further published as 1 January of the following year. The recommended definition is the 'usual resident population' and represents the number of inhabitants of a given area on 31st December. However, the population transmitted by the countries can also be either based on data from the most recent census adjusted by the components of population change produced since the last census, either based on population registers.

<br/>

**Employment and population (subset to data for Russia)**  
<!-- dataset save location BHI_share/2.0/Goals/LE/LIV/Pop_Employ_Russia_2019 -->

```{r employment and population for russia, echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Years:", "1990-2019"), 
  c("NA_ITEM:", "Total population national concept (POP_NC), Employees domestic concept (SAL_DC), Total employment domestic concept (EMP_DC)"), 
  c("GEO:", "RU	Russia"),
  c("Unit:", "Thousand persons (THS_PER)")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [eurostat database](http://appsso.eurostat.ec.europa.eu/nui/show.do?dataset=naida_10_pe&lang=en) <br/> Downloaded 21 July, 2020 by Andrea De Cervo")
```
<br/>


**Population change - Demographic balance and crude rates at regional level (NUTS 3), Filter to NUTS3 Level**
<!-- dataset save location BHI_share/2.0/Goals/LE/LIV/Population_NUTS3/demo_r_pjangrp3_w_nutscodes -->

```{r population change demographics nuts3 regions, echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("TIME (Years):", "2014-2019"), 
  c("GEO:", "NUTS3 (Filtering - Nuts level) SHOW: CODES") # need NUTS codes to match to BHI regions!
))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [eurostat database](http://appsso.eurostat.ec.europa.eu/nui/show.do?dataset=demo_r_pjangrp3&lang=en) <br/> Downloaded 11 September, 2020 by Andrea De Cervo")
```

<br/>

**Employment rate** 
<!-- dataset save location BHI_share/2.0/Goals/LE/LIV/Employment_NUTS3 -->

Data were extracted from the [EU Blue Economy report](https://ec.europa.eu/maritimeaffairs/sites/maritimeaffairs/files/2020_06_blueeconomy-2020-annexes-ld-part1_final.pdf) for each Baltic country.


#### 2.1.2 HYDE Database {-}

**Population density data**  
<!-- dataset save location BHI_share/2.0/Goals/LE/LIV/GriddedPop_2016AD -->
<!-- ftp://ftp.pbl.nl/hyde/ -->

```{r population density data from hyde database, echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Proceed to the FTP-server to", "download the datasets"), 
  c("Connect as", "Guest"),
  c("Index:", "hyde/hyde3.2/2017_beta_release/001/zip/"), 
  c("Name:", "2016AD_pop.zip")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [HYDE database, FTP-server](https://themasites.pbl.nl/tridion/en/themasites/hyde/download/index-2.html) <br/> Downloaded 21 July, 2020 by Andrea De Cervo")
```
<br/>

Year of data = 2005. Data were a 5' resolution. Population density within a 25km buffer from the coast will be used.

<br/>


### 2.2 Centralization & Normalization {-}


Summary of steps/analysis in this document:

**Wrangling of Spatial Assessment Units**
- intersect BHI 25km buffer (coastal area) with NUTS2 (reporting areas in Eurostat employment and population data) 
- review spatial units used in the analysis

**Clean Employment Rates Data**
- check and remove dataflags
- clean and visualize NUTS2 employment data
- explore reporting NUTS2 regions, check whether need spatial corrections based on time-period/diff. versions of NUTS2

**Clean/Create Coastal Population Data**
- clean and visualize NUTS2 population data
- corrections for Finland data-- different regions used than included in the raw data...
- corrections for Lithuania...
- calculate population in buffer-NUTS2 polygons as average pop per km2 times buffer-NUTS2 polygon area

**Derive and save Numbers Employed per BHI Region & Country**
- calculate number employed (in polygons from BHI buffer-NUTS2 intersection) as the population in the polygon times the relevant NUTS2 employment rate
- sum number employed in BHI buffer-NUTS2 polygons across each BHI region, to get coastal employed per BHI region
- sum population in BHI buffer-NUTS2 polygons across each BHI region, to get coastal population per BHI region
- divide coastal employed per BHI region by coastal population per BHI region, to get BHI region coastal employment rate
- clean and visualize derived coastal population and employment data by BHI region


#### 2.2.1 Spatial Data {-}

[Eurostat NUTS regions shapefiles can be found here](https://ec.europa.eu/eurostat/web/gisco/geodata/reference-data/administrative-units-statistical-units/nuts). The NUTS 2016, 2013, 2010, 2006, and 2003 were downloaded at 1:10 million resolution (NUTS_RG_10M_year_4326_LEVL_2or3 files within the zipped downloads are the polygons in espg 4326 projection).

```{r nuts2 polygons, echo = TRUE, message = FALSE, warning = FALSE}
##  NUTS2 shapefiles
## using level 2 but also have level 3 NUTS if those are better?
nutsshpfiles <- grep(
  "NUTS_RG_10M_[0-9]{4}_4326_LEVL_2/.*\\.shp", 
  list.files(file.path(dirname(dir_B), "Shapefiles", "all_NUTS_shapefiles"), recursive = TRUE),
  value = TRUE
)
for(f in nutsshpfiles){
  ## gives GDAL Errors / CPL_read_ogr warnings, but reads in the shapefile...
  shp <- sf::st_read(
    dsn = file.path(dirname(dir_B), "Shapefiles", "all_NUTS_shapefiles", stringr::str_extract(f, "ref.*LEVL_2(?=/)")),
    stringsAsFactors = FALSE
  )
  shp_baltic <- shp %>% filter(CNTR_CODE %in% c("SE", "FI", "EE", "LV", "LT", "PL", "DE", "DK"))
  assign(sprintf("nuts2_%s_shp", stringr::str_extract(f, "(?<=10M_)[0-9]{4}")), shp_baltic)
}

## 25km inland buffer for BHI regions
buffer_25km_inland <- st_read(
  file.path(dirname(dir_B), "Shapefiles", "BHI_shapefile_25km_buffer"),
  "BHI_shapefile_25km_buffer"
)
buffer_25km_inland <- cbind(
  buffer_25km_inland,
  buffer_area_km2 = as.numeric(sf::st_area(buffer_25km_inland)/(1000^2))
)


## create some spatial objects for making maps later
buffer_25k_simple <- rmapshaper::ms_simplify(input = buffer_25km_inland) %>% sf::st_as_sf()

basemap <- ggplot2::ggplot(rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")) +
  geom_sf(size = 0.1, color = "burlywood", alpha = 0.4) +
  theme(panel.background = element_rect(fill = "#F8FBFC", color = "#E2EEF3")) +
  scale_x_continuous(limit = c(4, 32)) +
  scale_y_continuous(limit = c(53.5, 66)) 
```

<br>

#### 2.2.2 Clean Employment Rates Data: standardize units, rename fields/variables {-}

```{r read in NUTS2 regional employment data, message = FALSE}
## regional employment per NUTS2 assessemnt areas, raw data from eurostat
regional_employ_data <- file.path(
    dir_rawdata, "Employment_NUTS2", 
    "lfst_r_lfe2emprt_2019_w_nutscodes", 
    "lfst_r_lfe2emprt_1_Data.csv"
  )

regional_employ <- read.csv(
  regional_employ_data, 
  stringsAsFactors = FALSE
)
```

```{r clean regional employment rates data, echo = TRUE, message = FALSE, warning = FALSE}
rgn_employ_clean <- regional_employ %>%
  select(
    year = TIME, nuts2 = GEO,
    unit = UNIT, value = Value, 
    flag_notes = Flag.and.Footnotes
  ) %>% 
  mutate(
    flag_notes = case_when(
      flag_notes == "b" ~ "break in timeseries",
      flag_notes == "u" ~ "low reliability",
      flag_notes == "bu" ~ "break in timeseries and low reliability",
      !flag_notes %in% c("b", "u", "bu") ~ ""
    )
  )

## check dataflags
# unique(rgn_employ_clean$flag_notes)
# filter(rgn_employ_clean, flag_notes == "low reliability") # not Baltic country
# filter(rgn_employ_clean, flag_notes == "break in timeseries and low reliability") # not Baltic country
# filter(rgn_employ_clean, flag_notes == "break in timeseries") # not such a concern

## remove flags_notes
rgn_employ_clean <- rgn_employ_clean %>% 
  select(-flag_notes, -unit) %>% 
  mutate(value = ifelse(value == ":", NA, value)) %>% 
  mutate(value = as.numeric(value)) %>% 
  filter(substr(nuts2, 1, 2) %in% c("SE", "FI", "EE", "LV", "LT", "PL", "DE", "DK"))
# summary(rgn_employ_clean$value)
```


**Join employment info with NUTS2 shapefiles**

```{r join employ data with nuts2 shapefile, results = "show", fig.width = 9.5, fig.height = 4}

rgn_employ_nuts_spatial <- tidyr::pivot_wider(
  rgn_employ_clean, 
  names_from = year, names_prefix = "emp_", 
  values_from = value
)
rgn_employ_nuts_spatial <- nuts2_2016_shp %>% 
  sf::st_transform(crs = 4326) %>% 
  rename(nuts2 = NUTS_ID) %>% 
  left_join(rgn_employ_nuts_spatial, by = "nuts2")

## plotting initial values, raw NUTS2 regions employment rates
gridExtra::grid.arrange(
  basemap + 
    geom_sf(data = rgn_employ_nuts_spatial, aes(fill = emp_2019), size = 0.1) +
    geom_sf(data = buffer_25k_simple, fill = NA, color = "red", size = 0.2) +
    scale_fill_gradient2(mid = "thistle", midpoint = 65, limits = c(45, 85)) +
    labs(fill = "Emp. Rate %") +
    theme(legend.position = c(0.18, 0.8)) +
    ggtitle("Employment 2019"),
  basemap + 
    geom_sf(data = rgn_employ_nuts_spatial, aes(fill = emp_2009), size = 0.1, show.legend = FALSE) +
    geom_sf(data = buffer_25k_simple, fill = NA, color = "red", size = 0.2) +
    scale_fill_gradient2(mid = "thistle", midpoint = 65, limits = c(45, 85)) +
    ggtitle("Employment 2009"),
  nrow = 1
)
## loop through all years to check employment data alignment with reporting regions...
```

<br>

**Check NUTS2 regions alignment with reported Employment Data**

For Finland NUTS2 version 2016 works for employment data after and including 2005. Because of this we will keep and use only employment data from 2005 onwards.

Missing rates for Lithuania before 2013-- reported for LT00 which includes LT01 and LT02. Will assign to LT01 and LT02 the same employment rate reported for LT00 for years before 2013.

For Denmark, don't have pop data by nuts2016 level2 reporting regions before 2007. Changes are small though - population overall increases less than 1% from 2007 to 2005 (5447084 in 2007 and 5411405 in 2005). Will use ratios between country populations in different years to approximate regional population for 2005 and 2006.


```{r aligning NUTS regions with employment data across years and changes in regions reporting, echo = TRUE, message = FALSE, warning = FALSE}
## fix rgn_employ_clean to address these issues and then recreate rgn_employ_nuts_spatial 
rgn_employ_clean <- rbind(
  rgn_employ_clean %>% 
    filter(!(year %in% 2005:2012 & nuts2 %in% c("LT00", "LT01", "LT02"))) %>% 
     filter(!(year %in% 2005:2007 & substr(nuts2, 1, 2) == "DK")) %>%
    filter(year >= 2005), 
  ## Lithuania before 2013: give regions the overal employment rate
  as_tibble(rgn_employ_clean) %>% 
    filter(year %in% 2005:2012, nuts2 == "LT00") %>% 
    mutate(nuts2 = list(c("LT00", "LT01", "LT02"))) %>% 
    tidyr::unnest(cols = c(nuts2)) %>% 
    data.frame(),
  ## approximated regional employment rates for denmarks nuts2 regions, for 2005 and 2006
  rgn_employ_clean %>% 
    filter(year %in% 2005:2007 & substr(nuts2, 1, 2) == "DK") %>% 
    tidyr::pivot_wider(names_from = year, values_from = value) %>% 
    mutate(
      `2005` = `2007`*(75.9/77), 
      `2006` = `2007`*(77.4/77)
    ) %>%
    tidyr::pivot_longer(cols = 2:4, names_to = "year", values_to = "value")
)

nuts2_employ <- tidyr::pivot_wider(
  rgn_employ_clean, 
  names_from = year, names_prefix = "emp_", 
  values_from = value
)
```

<br>

#### 2.2.3 Clean Population Data: standardize units, rename fields/variables {-}

```{r alternative BHI coastal population estimation from HYDE population density rasters}
## to sum over pop densities within 25km-buffer-NUTS2 intersection polygons...
# library(raster)
# pop_den_zips <- file.path(dir_rawdata, "Hyde", "hyde3.1", "hyde31_final")
# unzip(file.path(pop_den_zips, "2000ad_pop.zip"))
# unzip(file.path(pop_den_zips, "2005ad_pop.zip"))
# 
# pop_den_rast <- raster(file.path(pop_den_zips, "2005ad_pop?"))
# nuts_bhibuff_sppolys <- as_Spatial(nuts_bhibuff_polys)
# pop_vals <- extract(pop_den_rast, nuts_bhibuff_sppolys)
# nuts_bhibuff_sppolys$sum_in_poly <- unlist(lapply(pop_vals , function(x) if (!is.null(x)) sum(x, na.rm=TRUE) else NA))
```


```{r BHI coastal population estimation from NUTS2 population data, echo = TRUE, message = FALSE, warning = FALSE}
## regional employment per NUTS2 assessemnt areas, raw data from eurostat
regional_pop_data <- file.path(
    dir_rawdata, "Population_Jan1", 
    "demo_r_gind3", 
    "demo_r_gind3_1_Data.csv"
  )
regional_pop <- read.csv(
  regional_pop_data, 
  stringsAsFactors = FALSE
)
```

```{r clean regional population data, echo = TRUE, message = FALSE, warning = FALSE}
rgn_pop_clean <- regional_pop %>%
  dplyr::select(
    year = TIME, nuts2 = GEO,
    value = Value,
    flag_notes = Flag.and.Footnotes
  ) %>%
  mutate(
    value = as.numeric(stringr::str_remove_all(value, ",")),
    flag_notes = case_when(
      flag_notes == "b" ~ "break in timeseries",
      flag_notes == "u" ~ "low reliability",
      flag_notes == "bu" ~ "break in timeseries and low reliability",
      !flag_notes %in% c("b", "u", "bu") ~ ""
    )
  )

## check dataflags
# unique(rgn_pop_clean$flag_notes)

## no rows "low reliability" or "break in timeseries and low reliability"
## breaks in timeseries a concern?
# filter(rgn_pop_clean, flag_notes == "break in timeseries")

## remove flags_notes
rgn_pop_clean <- rgn_pop_clean %>% 
  dplyr::select(-flag_notes) %>% 
  filter(substr(nuts2, 1, 2) %in% c("SE", "FI", "EE", "LV", "LT", "PL", "DE", "DK"))
# summary(rgn_pop_clean$value)
```


**Join population info with NUTS2 shapefiles**

```{r join pop data with nuts2 shapefile, results = "show", fig.width = 9.5, fig.height = 8}

rgn_pop_nuts_spatial <- tidyr::pivot_wider(
  rgn_pop_clean,
  names_from = year, names_prefix = "pop_", 
  values_from = value
)
rgn_pop_nuts_spatial <- nuts2_2016_shp %>% 
  sf::st_transform(crs = 4326) %>% 
  rename(nuts2 = NUTS_ID) %>% 
  left_join(rgn_pop_nuts_spatial, by = "nuts2")

## plotting initial values, raw NUTS2 regions employment rates
## loop through all years to check population data alignment with reporting regions...
basemap + 
  geom_sf(data = rgn_pop_nuts_spatial, aes(fill = pop_2019), size = 0.1) +
  geom_sf(data = buffer_25k_simple, fill = NA, color = "red", size = 0.2) +
  scale_fill_gradient2(
    high = "royalblue4", 
    mid = "thistle1", 
    midpoint = 3*10^4, 
    limits = c(2*10^4, 5*10^6)
  ) +
  labs(fill = "Population") +
  theme(legend.position = c(0.15, 0.85)) +
  ggtitle("Population 2019")
```

**Check NUTS2 regions alignment with reported Population Data**

For Denmark, don't have pop data by nuts2016 level2 reporting regions before 2007. Changes are small though - population overall increases less than 1% from 2007 to 2005 (5447084 in 2007 and 5411405 in 2005). Will use ratios between country populations in different years to approximate regional population for 2005 and 2006.

Population data for Lithuania, as for employment data, is reported at the country level before 2013. Will use LT0 values to derive population estimates in each LT01 and LT02 for year 2005-2012, based on the relative proprtions in LT01 and LT02 in 2013...

```{r aligning NUTS regions with population data across years and changes in regions reporting, echo = TRUE, message = FALSE, warning = FALSE}

rgn_pop_clean <- rbind(
  rgn_pop_clean %>% 
    filter(!(year %in% 2005:2007 & substr(nuts2, 1, 2) == "DK")) %>% 
    filter(!(year %in% 2005:2013 & substr(nuts2, 1, 2) == "LT")) %>% 
    filter(year >= 2005),
  ## approximated regional pops for denmarks nuts2 regions, for 2005 and 2006
  rgn_pop_clean %>% 
    filter(year %in% 2005:2007 & substr(nuts2, 1, 2) == "DK") %>% 
    tidyr::pivot_wider(names_from = year, values_from = value) %>% 
    mutate(
      `2005` = `2007`*(5411405/5447084), 
      `2006` = `2007`*(5427459/5447084)
    ) %>%
    tidyr::pivot_longer(cols = 2:4, names_to = "year", values_to = "value"),
  ## approximate regional pops for lithuania nuts2 regions, for 2005-2012
  rgn_pop_clean %>% 
    filter(year %in% 2005:2013 & substr(nuts2, 1, 2) == "LT") %>%
    tidyr::pivot_wider(names_from = year, values_from = value) %>% 
     mutate(
       `2005` = `2013`*(3355220/2971905),        
       `2006` = `2013`*(3289835/2971905),
       `2007` = `2013`*(3249983/2971905),
       `2008` = `2013`*(3212605/2971905),
       `2009` = `2013`*(3183856/2971905),
       `2010` = `2013`*(3141976/2971905),
       `2011` = `2013`*(3052588/2971905),
       `2012` = `2013`*(3003641/2971905)
     ) %>%
    tidyr::pivot_longer(cols = 2:10, names_to = "year", values_to = "value")
)

nuts2_pop <- tidyr::pivot_wider(
  rgn_pop_clean,
  names_from = year, names_prefix = "pop_", 
  values_from = value
)
```


<br>

#### 2.2.4 BHI Region Employment Rates and Coastal Populations {-}

A BHI region's coastal employment rate is derived from populations and employment rates reported for the NUTS2 regions that overlap with the BHI region's coastal zone -- the area between the waters edge and 25km inland. The calculation assumes uniform population density and employment rates across NUTS2 regions. The BHI region's coastal employment rate is estimated as follows:

$$BHIempl.rate = \frac{BHIcoast.empl.num}{BHIcoast.pop} = \frac{\sum_{buffNUTS2}(buffNUTS2.coast.pop \times NUTS2empl.rate) }{\sum_{buffNUTS2}{buffNUTS2.coast.pop}}$$

$$buffNUTS2.coast.pop = \frac{NUTS2pop}{NUTS2area}\times buffNUTS2.area$$

$$buffNUTS2 = \text{polygons within the given BHI region, created by intersecting the BHI region with NUTS2 polygons}$$

The below code will calculate in order: 

(1) average population density over NUTS2 regions
(2) population in each of intersected buffer-NUTS2 polygons, using polygon area and the average population density
(3) coastal employment estimate (number of employed) in each of the intersected buffer-NUTS2 polygons
(4) estimated total coastal population in BHI region, as sum of pops in intersected buffer-NUTS2 polygons per BHI rgn
(5) estimated total coastal employment (number of employed) per BHI region
(6) coastal employment rate in BHI regions



**Join NUTS2 Data (Employ and Pop) with NUTS2 Spatial Information**

```{r join cleaned population and employment datasets with spatial nuts2 regions, echo = TRUE, message = FALSE, warning = FALSE}
nuts2_2016_shp <- cbind(
  nuts2_2016_shp,
  ## calculate area of NUTS2 regions in 2016 version
  ## convert area to sq km, also convert class from units to numeric
  nuts2_area_km2 = as.numeric(sf::st_area(nuts2_2016_shp)/(1000^2))
) 

nuts2spatial_popdensity_employ <- nuts2_2016_shp %>% 
  sf::st_transform(crs = 4326) %>% 
  rename(nuts2 = NUTS_ID) %>% 
  left_join(nuts2_employ, by = "nuts2") %>% 
  left_join(nuts2_pop, by = "nuts2") %>% 
  ## get average population density from nuts population and areas
  ## population per km2 in 2016 nuts2
  mutate_at(vars(starts_with("pop")), function(x){x/.$nuts2_area_km2})
```


**NUTS Regions with Data, Intersected with BHI 25km Inland Buffer**

```{r intersect relevant NUTS2 polygons and BHI 25km inland buffer, results = "show", fig.width = 9.5, fig.height = 6}
## create polygons intersecting NUTS2 regions and 25km-BHI Buffer

## may need to fix nuts2 polygon geometry with st_buffer so intersection works...
## https://github.com/r-spatial/sf/issues/347
nuts2_bhibuff_polys <- sf::st_intersection(
  st_buffer(nuts2spatial_popdensity_employ, 0), 
  buffer_25km_inland
)
nuts2_bhibuff_polys <- cbind(
  nuts2_bhibuff_polys,
  ## calculate area of the intersected polygons
  area_km2 = as.numeric(sf::st_area(nuts2_bhibuff_polys)/(1000^2))
) 

## check the intersected polygons on a map
pal <- colorRampPalette(RColorBrewer::brewer.pal(8, "Set2"))(42)[sample(1:42)]
basemap +
  geom_sf(
    data = nuts2_2016_shp,
    fill = NA,
    show.legend = FALSE
  ) +
  geom_sf(
    data = mutate(nuts2_bhibuff_polys, BHI_ID = as.character(BHI_ID)),
    aes(fill = BHI_ID), alpha = 0.7,
    color = "red", size = 0.2,
    show.legend = FALSE
  ) +
  scale_fill_manual(values = pal)
```

<br>

```{r calculate coastal employment and coastal population, echo = TRUE, message = FALSE, warning = FALSE}

## combine population and employment and calculate info for intersected polygons
coastal_pop_employ_combine <- nuts2_bhibuff_polys %>% 
  st_drop_geometry() %>% 
  select(nuts2, nuts2_area_km2, starts_with("pop"), starts_with("emp"), bhi_id = BHI_ID, area_km2) %>% 
  
  ## calculate population of intersected polygons
  ## note: using mutate_at, the pop vars now represent populations of the intersected polygons!
  mutate_at(vars(starts_with("pop")), function(x){x*.$area_km2}) %>% 
  tidyr::pivot_longer(cols = matches("[a-z]{3}_[0-9]{4}")) %>% 
  mutate(year = as.numeric(substr(name, 5, 8)), name = substr(name, 1, 3)) %>% 
  tidyr::pivot_wider(names_from = name, values_from = value) %>% 
  
  ## calculate coastal employment in each of the intersected buffer-NUTS2 polygons
  mutate(num_employed = emp/100*pop)
  

## calculate coastal employment and coastal population for BHI regions!
coastal_employ_bhi_rgns <- coastal_pop_employ_combine %>% 
  ## sum coastal employment and coastal population by BHI ID
  ## will use these two vars to get coastal employment rate per BHI region
  group_by(bhi_id, year) %>% 
  summarise(
    coastal_pop = sum(pop), 
    coastal_employ_num = sum(num_employed)
  ) %>% 
  ungroup() %>% 
  mutate(coastal_employ_rate = coastal_employ_num/coastal_pop) %>% 
  
  ## some regions should NOT have population, employed, employment rate
  ## due to spatial intersection of less than perfectly aligned buffer and NUTS2 polygons
  ## BHI regions 19, 22, and 33 are Russia so have no eurostat data -- set to NA
  ## BHI region 30 is ocean with no significant coastline (a few tiny islands) -- set to NA
  mutate_at(
    vars(starts_with("coastal")), 
    function(x){ifelse(.$bhi_id == 19|.$bhi_id == 22|.$bhi_id == 33|.$bhi_id == 30, NA, x)}
  )

mapsf <- buffer_25k_simple %>% 
  rename(bhi_id = BHI_ID) %>% 
  left_join(filter(coastal_employ_bhi_rgns, year == 2019), by = "bhi_id") %>% 
  mutate(bhi_id = as.character(bhi_id))

basemap +
  geom_sf(
    data = mapsf,
    aes(fill = coastal_employ_rate),
    color = "snow", size = 0.3,
    show.legend = FALSE
  ) +
  scale_fill_gradient2(mid = "thistle", midpoint = 0.65, limits = c(0.45, 0.85))
```


#### 2.2.5 Save LIV Data Layers

```{r save the liv subgoal regional employment data layer, eval = FALSE, echo = TRUE, message = FALSE, warning = FALSE}
write_csv(
  select(
    coastal_employ_bhi_rgns, 
    region_id = bhi_id,
    year,
    bhi_employ_rate = coastal_employ_rate
  ), 
  file.path(dir_layers, "le_liv_regional_employ_bhi2019.csv")
)
```

```{r save the liv subgoal national employment data layer, echo = TRUE, message = FALSE, warning = FALSE}
natl_employ_rates <- left_join(
  rgn_employ_clean %>%
    filter(stringr::str_length(nuts2) == 2) %>% 
    rename(eez = nuts2, employ_pop = value),
  read_csv(file.path(dir_prep, "supplement",  "lookup_tabs", "rgns_complete.csv")) %>% 
    select(eez, region_id) %>% 
    mutate(eez = case_when(
      eez == "Sweden" ~ "SE",
      eez == "Finland" ~ "FI",
      eez == "Estonia" ~ "EE",
      eez == "Latvia" ~ "LV",
      eez == "Lithuania" ~ "LT",
      eez == "Poland" ~ "PL",
      eez == "Denmark" ~ "DK",
      eez == "Germany" ~ "DE",
      eez == "Denmark" ~ "DK",
      eez == "Russia" ~ "RU"
    )),
  by = "eez"
)

#write_csv(
#  select(natl_employ_rates, region_id, year, employ_pop),
#  file.path(dir_layers, "le_liv_national_employ_bhi2019.csv")
#)
```

```{r save coastal employment intermediate dataset, eval = FALSE, echo = TRUE, message = FALSE, warning = FALSE}
write_csv(
  select(
    coastal_employ_bhi_rgns, 
    region_id = bhi_id,
    year,
    coastal_pop
  ), 
  file.path(dir_prep, "data", "LIV", version_year, "intermediate", "le_coastal_pop_bhi2019.csv")
)
```

<br/>


### 2.3 Initial Data Exploration

#### 2.3.1 Timeseries Plots: check regional employment time series

```{r plot and check regional employment time series, results = "show", fig.width = 9.5}
ggplot(coastal_pop_employ_combine) +
  geom_point(aes(x = year, y = emp, colour = nuts2), size = 0.8) +
  geom_line(aes(x = year, y = emp, colour = nuts2)) +
  facet_wrap(~bhi_id, ncol = 6) +
  labs(x = NULL, y = "Employment rate (%)\n", color = "NUTS2 2016 Region") + 
  scale_color_manual(values = pal)
```
