---
title: "Preparation of general spatial files"
output: html_document
---

# Linking the HELCOM data to baltic regions

This section is based on the HELCOM_2_baltic.R from initial BHI assessment of 2015. Some edits were made to replace some spatial functions from `sp` with their counterpart in `sf` package.

```{r set up}

source(file.path(here::here(), "R/spatial_prep_fun.R"))

## load data and reporject where necessary
helcom_loc <- "/var/data/ohi/git-annex/Baltic/HELCOM_to_BHIregion/raw/benthos/Agrypnetes crassicornis (DD)"
helcom <- sf::st_read(file.path(helcom_loc, "Join_Benthic_invertebrates")) %>% select("FID_1", "CELLCODE")

bhi <- sf::st_read("/var/data/ohi/git-annex/clip-n-ship/bhi/spatial/rgn_offshore_gcs.shp") %>% 
  st_transform(proj4string(helcom))

## make sure the transfomration seemed to go well
plot(helcom)
plot(bhi, add = TRUE, col = "red")


```


## http://gis.stackexchange.com/questions/140504/extracting-intersection-areas-in-r
pi <- raster::intersect(helcom, bhi)
areas <- data.frame(area_m2=sapply(pi@polygons, FUN=function(x) {slot(x, 'area')}))
row.names(areas) <- sapply(pi@polygons, FUN=function(x) {slot(x, 'ID')})
# Combine attributes info and areas
attArea <- spCbind(pi, areas)
data <- aggregate(area_m2~rgn_id + CELLCODE, data=attArea, FUN=sum)
data$prop_area <- data$area_m2/1.000000e+08
hist(data$area_m2)
hist(data$prop_area)
data$type <- "bhi_sea"

write.csv(data, "baltic2015/prep/spatial/helcom_to_rgn_bhi_sea.csv", row.names=FALSE)



############################################
###### 50 km inland ######################

## NOTE: My idea was to get the 50 km inland data and merge it with the sea data to ensure that
## species landing slightly on land would be included.
## However, I let this run for over 24 hours and had no success.  I think there is something strange about the
## spatial file that is preventing this from working.

#### BHI map
bhi_inland <- readOGR(dsn = "/var/data/ohi/git-annex/Baltic/regions/BHI_regions_inland_buffer/BHI_inland_50km",
               layer = "BHI_inland_50km")


### Using the template from
### Getting a HELCOM map
helcom <- readOGR(dsn= '/var/data/ohi/git-annex/Baltic/HELCOM_to_BHIregion/raw/benthos/Agrypnetes crassicornis (DD)',
                  layer='Join_Benthic_invertebrates')
helcom <- helcom[, c("FID_1", "CELLCODE")]

#### transform the sp_map data to have the same
#### coordinate reference system as the region boundaries
bhi_inland <- spTransform(bhi_inland, CRS(proj4string(helcom)))

## make sure the transfomration seemed to go well (overlap of the 2 spatial files)
plot(helcom)
# this is when using a subset of the data extracted above
# text(helcom, labels=as.character(helcom@data$CELLCODE),
#      cex=0.5, font=2, offset=0.5, adj=c(0,2))
plot(bhi_inland, add=TRUE, col="red")

## this code came from here:
## http://gis.stackexchange.com/questions/140504/extracting-intersection-areas-in-r
pi <- raster::intersect(helcom, bhi_inland)
areas <- data.frame(area_m2=sapply(pi@polygons, FUN=function(x) {slot(x, 'area')}))
row.names(areas) <- sapply(pi@polygons, FUN=function(x) {slot(x, 'ID')})
# Combine attributes info and areas
attArea <- spCbind(pi, areas)
data <- aggregate(area_m2~rgn_id + CELLCODE, data=attArea, FUN=sum)
data$prop_area <- data$area_m2/1.000000e+08
hist(data$area_m2)
hist(data$prop_area)
data$type <- "bhi_inland50km"

write.csv(data, "baltic2015/prep/spatial/helcom_to_rgn_bhi_inland50km.csv", row.names=FALSE)

***


# Create EEZ and SUBBASIN maps 


This will be done using `maptools::unionSpatialPolygons` to 'dissolve' the boundaries between BHI regions. 
This function won't preserve the data frame because there's not the same number of rows.

```{r, message = FALSE}
## setup ----
library(tidyverse)
library(rgdal)
library(maptools)

dir_spatial <- path.expand('~/github/bhi/baltic2015/spatial')

## prepare rgns data.frame information for EEZ and SUBBASIN
rgns <- read.csv(file.path(dir_spatial, 'regions_lookup_complete_wide.csv'), stringsAsFactors = FALSE) %>%
  mutate(eez_ID      = eez_id      - 300,
         subbasin_ID = subbasin_id - 500)

rgns_eez <- rgns %>%
  group_by(eez_id, eez_name) %>%
  summarize(area_km2 = sum(area_km2_rgn)) %>%
  ungroup()

rgns_subbasin <- rgns %>%
  group_by(subbasin_id, subbasin_name) %>%
  summarize(area_km2 = sum(area_km2_rgn)) %>%
  ungroup()

```

## Map of 42 BHI regions
```{r}
## prepare BHI spatial file ----
bhi  <-  rgdal::readOGR(
  dsn = dir_spatial,
  layer = 'rgn_offshore_gcs', stringsAsFactors = FALSE)

plot(bhi)

## join bhi@data with rgns information
bhi@data <- bhi@data %>%
  left_join(rgns %>%
              dplyr::rename(rgn_id  = region_id,
                            rgn_name = region_name),
            by = c('rgn_id', 'rgn_name'))
```

## Map of 9 BHI EEZs
Here we create just the EEZ boundaries by dissolving the subbasin boundaries
```{r}

## EEZ: unite spatial polygons ----
bhi_eez <- maptools::unionSpatialPolygons(bhi, ID = bhi@data$eez_ID)
plot(bhi_eez)

## recreate as SpatialPolygonsDataFrame. Note: writeOGR doesn't like tbl_dfs, only data.frames
bhi_eez <- SpatialPolygonsDataFrame(bhi_eez,  as.data.frame(rgns_eez))

## standardize dataframe
bhi_eez@data <- bhi_eez@data %>%
  dplyr::rename(rgn_id   = eez_id, 
                rgn_name = eez_name)
```

## Map of 17 BHI SUBBASINS
Here we create just the SUBBASIN boundaries by dissolving the EEZ boundaries
```{r}
## SUBBASIN: unite spatial polygons ----
bhi_subbasin <- unionSpatialPolygons(bhi, ID = bhi@data$subbasin_ID)
plot(bhi_subbasin)

## recreate as SpatialPolygonsDataFrame. Note: writeOGR doesn't like tbl_dfs, only data.frames
bhi_subbasin <- SpatialPolygonsDataFrame(bhi_subbasin,  as.data.frame(rgns_subbasin))

## standardize dataframe
bhi_subbasin@data <- bhi_subbasin@data %>%
  dplyr::rename(rgn_id   = subbasin_id, 
                rgn_name = subbasin_name)
```

```{r, eval=FALSE}
## save EEZ map
writeOGR(obj    = bhi_eez,
         dsn    = dir_spatial,
         layer  = 'BHI_EEZ_regions',
         driver = 'ESRI Shapefile', overwrite=TRUE)


## save SUBBASIN map
writeOGR(obj    = bhi_subbasin,
         dsn    = dir_spatial,
         layer  = 'BHI_SUBBASIN_regions',
         driver = 'ESRI Shapefile', overwrite=TRUE)
```

### TECHNICAL NOTE: 

This error was because `writeOGR` expects an object with class `data.frame`, not the tidyverse's `tibble`, `tbl`, `tbl_df` classes. Therefore, to use `writeOGR` without error above, I used `as.data.frame`.
