---
title: "Biodiversity Goal Data Preparation"
output:
  html_document:
    toc: true
    toc_depth: 4
    code_folding: hide
---

<br>
<br>

```{r bd preamble prep, message = FALSE}
source(here::here("R", "setup.R"))
knitr::opts_chunk$set(message = FALSE, warning = FALSE, results = "hide")

loc <- here::here("prep", "BD")
dir_rawdata <- file.path(dir_B, "Goals", "BD")
dir_intermediate <- here::here("data", "BD", version_year, "intermediate")

bkgd_path <- here::here("supplement", "goal_summaries", "BD.Rmd")
data_path <- here::here("data", "ECO", version_year, "bd_data.rmd")
refs_path <- file.path(loc, "bd_references.Rmd")
# file.exists(c(loc, bkgd_path, data_path, refs_path))
```

## 1. Background

```{r bd background, child = bkgd_path, results = "asis", echo = FALSE}
```

<br/>

## 2. Data

This prep document is used to generate and explore the following data layers:

- `bd_hab_benthic_bhi2019.csv` 
- `bd_hab_pelagic_bhi2019.csv`
- `bd_spp_fish_bhi2019.csv` 
- `bd_spp_seals_bhi2019.csv`

These are saved to the `layers` folder. All these are derived from or informed by the raw datasets from the HELCOM 2018 integrated biodiversity status assessments.

<br>

```{r bd data, child = data_path, results = "asis", echo = FALSE}
```

<br/>

## 3. Prep: Wrangling & Derivations, Checks/Evaluation, Gapfilling {-}


### 3.1 Load datasets, Subset, replace zeros with NAs where zone Not Assessed {-}

```{r load habitats and fish and seals spatial data layers}
files <- grep(
  "helcom_.*\\.shp$", 
  list.files(dir_rawdata, recursive = TRUE, full.names = TRUE), 
  value = TRUE
)
lapply(
  files,
  function(x){
    assign(
      sprintf("%s_shp", str_extract(x, "benthic|Fish|Seals|pelagic")),
      sf::st_read(x) %>% 
        dplyr::select(helcom_id = HELCOM_ID, country, subbasin = level_2, area_km2 = Area_km2, BQR, status = STATUS) %>% 
        mutate(BQR = ifelse(status == "Not assessed", NA, BQR)) %>% 
        ## these are just the 2018 integrated assessment
        mutate(year = 2018) %>% 
        select(-status)
    )
  }
)
```


### 3.2 Match the BHI regions by Country and Subbasin {-}

```{r match bhi regions to biodiversity datasets}
rgns_complete <- read_csv(file.path(dir_prep, "supplement", "lookup_tabs", "rgns_complete.csv")) %>% 
  select(region_id, subbasin, country = eez, region_name, region_area_km2) %>% 
  mutate(subbasin = ifelse(subbasin == "Aland Sea", "Åland Sea", subbasin))

## to match BHI regions, need to split the 'opensea' areas between adjacent countries
benthic_shp <- bind_rows(
  benthic_shp %>% 
    filter(country != "Opensea") %>% 
    left_join(rgns_complete, by = c("subbasin", "country")) %>% 
    mutate(coastal = TRUE),
  benthic_shp %>% 
    filter(country == "Opensea") %>% 
    left_join(tidyr::nest(rgns_complete, data = c(region_id, region_name, country, region_area_km2)), by = "subbasin") %>% 
    select(-country) %>% 
    tidyr::unnest(cols = c(data)) %>% 
    mutate(coastal = FALSE)
)
benthic_shp <- benthic_shp %>% 
  group_by(region_id) %>% 
  mutate(coastal_area  = sum(area_km2*coastal)) %>% 
  ungroup() %>% 
  mutate(area_km2 = ifelse(coastal, area_km2, region_area_km2 - coastal_area)) %>% 
  select(-coastal, -coastal_area, -region_area_km2)
```


```{r save bd layers}

```


```{r trial bd status calculations to test calculations and visualize}
bd_all_data <- bind_rows(
  benthic_shp %>% 
    st_drop_geometry() %>% 
    mutate(indicator = "hab_benthic"),
  pelagic_shp %>% 
    st_drop_geometry() %>% 
    mutate(indicator = "hab_pelagic"),
  Fish_shp %>% 
    st_drop_geometry() %>% 
    mutate(indicator = "spp_fishes"),
  Seals_shp %>% 
    st_drop_geometry() %>% 
    mutate(indicator = "spp_seals")
)

bd_status <- bd_all_data %>% 
  group_by(region_id, region_name, indicator) %>% 
  summarize(averageBQR = weighted.mean(BQR, area_km2, na.rm = TRUE)) %>% 
  ungroup()
```


<br/>

## 4. Visualizing Data Layers {-}

### 4.1 Maps {-}

```{r simplify bd shapefiles and plot for visual inspection}
source(here::here("R", "spatial.R"))
bhirgn_simple <- rmapshaper::ms_simplify(input = BHI_rgns_shp) %>% sf::st_as_sf()

## simplify the polygons before creating maps
benthic_shp <- rmapshaper::ms_simplify(input = benthic_shp) %>% sf::st_as_sf()
pelagic_shp <- rmapshaper::ms_simplify(input = pelagic_shp) %>% sf::st_as_sf()
Fish_shp <- rmapshaper::ms_simplify(input = Fish_shp) %>% sf::st_as_sf()
Seals_shp <- rmapshaper::ms_simplify(input = Seals_shp) %>% sf::st_as_sf()

## basemap and color palettes
basemap <- ggplot2::ggplot(rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")) +
  geom_sf(size = 0.1, color = "burlywood", alpha = 0.4) +
  theme(panel.background = element_rect(fill = "#F8FBFC", color = "#E2EEF3")) +
  scale_x_continuous(limit = c(4, 32)) +
  scale_y_continuous(limit = c(53.5, 66)) 

cols <- c("indianred", "coral", "goldenrod1", "khaki", "lightblue", "steelblue")
```


#### 4.1.1 Benthic Habitats {-}

```{r map of benthic habitats, eval = FALSE, results = "show", fig.width = 9.5, fig.height = 8.5}
## maps of BQR for each layer: habitats (benthic and pelagic), fish, seals
# ggplot2::ggplot() +
#   geom_sf(
#     data = bhirgn_simple,
#     size = 0.1, color = "burlywood", fill = NA
#   ) +
#   geom_sf(data = , mapping = )
```


#### 4.1.2 Pelagic Habitats {-}

#### 4.1.3 Fish {-}

#### 4.1.4 Seals {-}


<br>

## 5. Considerations for `BHI3.0`



<br>

## 6. References

```{r bd references, child = refs_path, results = "asis", echo = FALSE}
```
