---
title: "Biodiversity Goal Data Preparation"
output:
  html_document:
    toc: true
    toc_depth: 4
    code_folding: hide
---

<br>
<br>

```{r bd preamble prep, message = FALSE}
source(here::here("R", "setup.R"))
source(here::here("R", "spatial.R"))
knitr::opts_chunk$set(message = FALSE, warning = FALSE, results = "hide")

loc <- here::here("prep", "BD")
dir_rawdata <- file.path(dir_B, "Goals", "BD")
dir_intermediate <- here::here("data", "BD", version_year, "intermediate")

bkgd_path <- here::here("supplement", "goal_summaries", "BD.Rmd")
data_path <- here::here("data", "ECO", version_year, "bd_data.rmd")
refs_path <- file.path(loc, "bd_references.Rmd")
# file.exists(c(loc, bkgd_path, data_path, refs_path))
```

## 1. Background

```{r bd background, child = bkgd_path, results = "asis", echo = FALSE}
```

<br/>

## 2. Data

This prep document is used to generate and explore the following data layers:

- `bd_hab_benthic_bhi2019.csv` 
- `bd_hab_pelagic_bhi2019.csv`
- `bd_spp_fish_bhi2019.csv` 
- `bd_spp_seals_bhi2019.csv`
- `bd_spp_winter_birds_bhi2019.csv` 
- `bd_spp_breeding_birds_bhi2019.csv`

These are saved to the `layers` folder. All these are derived from or informed by the raw datasets from the HELCOM 2018 integrated biodiversity status assessments.

<br>

```{r bd data, child = data_path, results = "asis", echo = FALSE}
```

<br/>

## 3. Prep: Wrangling & Derivations, Checks/Evaluation, Gapfilling {-}


### 3.1 Initial Data Wrangling/Harmonizing {-}

**Load all Biodiversity Datasets**

```{r load habitats and fish and seals spatial data layers, warning = FALSE, message = FALSE, results = "hide"}
files <- grep(
  "helcom_.*\\.shp$", 
  list.files(dir_rawdata, recursive = TRUE, full.names = TRUE), 
  value = TRUE
)
lapply(
  X = files,
  FUN = function(x){
    tmp <- sf::st_read(x)
    colnames(tmp) <- str_to_lower(names(tmp))
    nm <- sprintf("%s_shp", str_extract(x, "benthic|Fish|Seals|pelagic|wintering|briding")) %>% 
      str_replace("wintering", "wintering_birds") %>% 
      str_replace("briding", "breeding_birds") %>% 
      str_to_lower()
    assign(nm, tmp, envir = .GlobalEnv)
  }
)
```


**Subset, replace zeros with NAs where zone Not Assessed, add Year column**

```{r initial wrangling of bd datasets}
## HABITATS
## benthic and pelagic habitats have the same columns and spatial units
for(shp in c("benthic_shp", "pelagic_shp")){
  
  habshp <- get(shp)
  colnames(habshp) <- str_to_lower(names(habshp))
  
  habshp <- habshp %>% 
    dplyr::select(helcom_id, country, subbasin = level_2, area_km2, BQR = bqr, status) %>% 
    mutate(BQR = ifelse(str_detect(status, "Not assessed"), NA, BQR)) %>% 
    ## these are just the 2018 integrated assessment
    mutate(year = 2018) %>% 
    select(-status)
  
  assign(shp, habshp)
}


## FISH
## fish have data by coastal area (by country) and open sea sub-basins, similar to habitats spatial units
colnames(fish_shp) <- str_to_lower(names(fish_shp))
fish_shp$country <- as.character(fish_shp$country)
fish_shp <- fish_shp %>% 
  mutate(country = ifelse(country == "Open sea", "Opensea", country)) %>% 
  ## create subbasins column from name; need some corrections
  mutate(subbasin = ifelse(
    country == "Opensea",
    "Opensea",
    str_extract(name, "^.* (?=Finnish|Swedish|Estonian|Latvian|Lithuanian|Danish)"))
  ) %>% 
  mutate(subbasin = str_remove(subbasin, "(?=-).*$")) %>% 
  mutate(subbasin = str_remove(subbasin, " $")) %>%
  mutate(subbasin = case_when(
    subbasin == "Western Gotlland Basin" ~ "Western Gotland Basin",
    subbasin == "Mecklenburg Bight" ~ "Bay of Mecklenburg",
    subbasin == "Belts" ~ "Great Belt",
    TRUE ~ subbasin
  )) %>% 
  ## filter to keep assessed, select only cols needed for the bhi status calculations
  filter(!(is.na(subbasin) & is.na(ecosystemc))) %>% 
  select(helcom_id, country, subbasin, BQR = bqr) %>% 
  ## these data are just the 2018 integrated assessment
  mutate(year = 2018)

## calculate area of the polygons - coastal and open sea regions
fish_shp <- cbind(fish_shp, area_km2 = as.numeric(sf::st_area(fish_shp)/(1000^2))) 


## SEALS
## seal data is by subbasin rather than country and subbasin,
## will expand to have data by bhi regions to be on same scale as other data, 
## but original data is at subbasin level!
seals_shp <- seals_shp %>%
  dplyr::select(helcom_id, subbasin = level_2, area_km2, BQR = bqr) %>% 
  ## these are just the 2018 integrated assessment
  mutate(year = 2018) %>% 
  ## will join subbasins as for the seal data as for other layers' opensea data
  mutate(country = "Opensea")


## WATERBIRDS
for(shp in c("wintering_birds_shp", "breeding_birds_shp")){
  
  birdshp <- get(shp)
  colnames(birdshp) <- str_to_lower(names(birdshp))
  
  birdshp <- birdshp %>% 
    mutate_if(is.factor, as.character) %>% 
    dplyr::select(
      helcom_id, country, subbasin = level_2, area_km2, 
      wading_feeders = wading_fee, wading_status = wading_f_1,
      surface_feeders = surface_fe, surface_status = surface__1,
      pelagic_feeders = pelagic_fe, pelagic_status = pelagic__1,
      benthic_feeders = benthic_fe, benthic_status = benthic_fe,
      grazing_feeders = grazing_fe, grazing_status = grazing__1,
      name
    ) %>% 
    mutate(
      wading_feeders = ifelse(str_detect(wading_status, "Not Assessed"), NA, wading_feeders),
      surface_feeders = ifelse(str_detect(surface_status, "Not Assessed"), NA, surface_feeders),
      pelagic_feeders = ifelse(str_detect(pelagic_status, "Not Assessed"), NA, pelagic_feeders),
      benthic_feeders = ifelse(str_detect(benthic_status, "Not Assessed"), NA, benthic_feeders),
      grazing_feeders = ifelse(str_detect(grazing_status, "Not Assessed"), NA, grazing_feeders)
    ) %>% 
    ## these are just the 2018 integrated assessment
    mutate(year = 2018) %>% 
    select(-wading_status, -surface_status, -pelagic_status, -benthic_status, -grazing_status) %>% 
    mutate(
      wading_feeders = as.numeric(wading_feeders), 
      surface_feeders = as.numeric(surface_feeders),
      winter_or_breeding = str_extract(shp, "-.*(?<=_)")
    )
  
  assign(shp, birdshp)
}

```


### 3.2 Match BHI regions by Country and/or Subbasin {-}

```{r wrangle habitats datasets including matching to bhi regions}
regions_shape()

rgns_complete <- read_csv(file.path(dir_prep, "supplement", "lookup_tabs", "rgns_complete.csv")) %>% 
  select(region_id, subbasin, country = eez, region_name, region_area_km2) %>% 
  mutate(subbasin = ifelse(subbasin == "Aland Sea", "Åland Sea", subbasin))

for(shp in c("benthic_shp", "pelagic_shp")){# "fish_shp", "seals_shp")){
  
  coastal_w_rgns <- get(shp) %>% 
    filter(country != "Opensea") %>% 
    left_join(rgns_complete, by = c("subbasin", "country")) %>% 
    mutate(coastal = TRUE)
  
  ## to match BHI regions, need to split the 'opensea' areas between adjacent countries
  ## for fish will need to spatially intersect opensea areas with BHI regions to evaluate overlap
  if(shp == "fish_shp"){
    
    ## need to fix weird geometries and also invalid geometries of bhi regions, then intersect...
    ## https://strimas.com/smoothr/reference/drop_crumbs.html
    opensea_w_rgns <- sf::st_intersection(
      lwgeom::st_make_valid(st_transform(BHI_rgns_shp, crs = 3035)),
      smoothr::drop_crumbs(filter(get(shp), country == "Opensea"), 300*(1000^2), drop_empty = TRUE)
    )
    
    ## view map of intersection between fish open sea and BHI regions
    # rmapshaper::ms_simplify(input = opensea_w_rgns) %>%
    #   sf::st_as_sf() %>%
    #   mutate(BHI_ID = as.character(BHI_ID)) %>%
    #   ggplot() +
    #   geom_sf(aes(fill = BHI_ID)) +
    #   scale_fill_manual(values = colorRampPalette(RColorBrewer::brewer.pal(6, "Dark2"))(80)[sample(1:80, 42)])
    
    opensea_w_rgns <- cbind(
      select(opensea_w_rgns, -area_km2, -subbasin),
      ## recalculate area of opensea after intersection with BHI regions
      area_km2 = as.numeric(sf::st_area(opensea_w_rgns)/(1000^2))
    )
    opensea_w_rgns <- opensea_w_rgns %>% 
      mutate(region_name = sprintf("%s, %s", subbasin, rgn_nam)) %>% 
      select(
        helcom_id = HELCOM_ID, subbasin = Subbasin, 
        area_km2, BQR, year, 
        region_id = BHI_ID, region_name, country = rgn_nam, 
        region_area_km2 = Area_km2
      ) %>% 
      mutate(coastal = FALSE)
    
  } else {
    opensea_w_rgns <- get(shp) %>% 
      filter(country == "Opensea") %>% 
      left_join(tidyr::nest(rgns_complete, data = c(region_id, region_name, country, region_area_km2)), by = "subbasin") %>% 
      select(-country) %>% 
      tidyr::unnest(cols = c(data)) %>% 
      mutate(coastal = FALSE)
  }

  ## calculate areas for opensea fractions of BHI regions
  shp_w_rgns <- bind_rows(opensea_w_rgns, coastal_w_rgns) %>% 
    group_by(region_id) %>% 
    mutate(coastal_area = sum(area_km2*coastal)) %>% 
    ungroup() %>% 
    mutate(area_km2 = ifelse(coastal, area_km2, region_area_km2 - coastal_area)) %>% 
    select(-coastal, -coastal_area, -region_area_km2)
  
  assign(shp, shp_w_rgns)
}
```


### 3.3 Waterbirds Data {-}

```{r wrangle water birds data}
## need to merge and reshape as the waterbirds data is by bird groups: 
## wading, surface, pelagic, benthic, grazing feeders
## and also wintering vs breeding

birds_data <- bind_rows(wintering_birds_shp, breeding_birds_shp) %>% 
  tidyr::pivot_longer(
    cols = c(),
    names_to = "feeder_group",
    values_to = "BQR"
  ) %>% 
  mutate(feeder_group = sprintf("%s, %s", feeder_group, winter_or_breeding)) %>% 
  select(-winter_or_breeding)
```


### 3.4 Save Biodiversity Layers {-}

```{r save bd layers}
write_csv(
  benthic_shp %>% 
    st_drop_geometry() %>% 
    select(region_id, year, BQR, area_km2), 
  file.path(dir_layers, "bd_hab_benthic_bhi2019.csv")
)
write_csv(
  pelagic_shp %>% 
    st_drop_geometry() %>% 
    select(region_id, year, BQR, area_km2), 
  file.path(dir_layers, "bd_hab_pelagic_bhi2019.csv")
)
write_csv(
  Fish_shp %>% 
    st_drop_geometry() %>% 
    select(region_id, year, BQR, area_km2), 
  file.path(dir_layers, "bd_spp_fish_bhi2019.csv")
)
write_csv(
  Seals_shp %>% 
    st_drop_geometry() %>% 
    select(region_id, year, BQR, area_km2), 
  file.path(dir_layers, "bd_spp_seals_bhi2019.csv")
)
write_csv(
  Seals_shp %>% 
    st_drop_geometry() %>% 
    select(region_id, year, feeder_group, wintBQR, area_km2), 
  file.path(dir_layers, "bd_spp_birds_bhi2019.csv")
)
```


### 3.5 Biodiversity Trial Status & Trend Calculations {-}

```{r trial bd status calculations to test calculations and visualize}
bd_all_data <- bind_rows(
  benthic_shp %>% 
    st_drop_geometry() %>% 
    mutate(indicator = "hab_benthic"),
  pelagic_shp %>% 
    st_drop_geometry() %>% 
    mutate(indicator = "hab_pelagic"),
  Fish_shp %>% 
    st_drop_geometry() %>% 
    mutate(indicator = "spp_fishes"),
  Seals_shp %>% 
    st_drop_geometry() %>% 
    mutate(indicator = "spp_seals")
)

## is it better to calculate status (which involves capping at 1) first and then aggregate indicators, 
## or aggregate indicators first? 
## aggregating first would mean status over ref point in some areas couldn't go towards compensating for areas under ref point...
## also, will use geometric mean to represent how components are not interchangeable

## geometric mean function
## https://stackoverflow.com/questions/2602583/geometric-mean-is-there-a-built-in
gm_mean = function(x, na.rm = TRUE, zero.propagate = FALSE){
  if(any(x < 0, na.rm = TRUE)){
    return(NaN)
  }
  if(zero.propagate){
    if(any(x == 0, na.rm = TRUE)){
      return(0)
    }
    exp(mean(log(x), na.rm = na.rm))
  } else {
    exp(sum(log(x[x > 0]), na.rm = na.rm) / length(x))
  }
}

## at BQR of 0.6 is the reference point
## will rescale so BQR = 0.6 or above is a status of 1, while BQR = 0 is a status of zero
bd_status <- bd_all_data %>% 
  group_by(region_id, region_name, indicator) %>% 
  summarize(averageBQR = weighted.mean(BQR, area_km2, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(status = pmin(1, (5/3)*averageBQR)) %>% 
  select(-averageBQR)

bd_status <- bind_rows(
  bd_status,
  bd_status %>% 
    group_by(region_id, region_name) %>% 
    summarize(status = gm_mean(indicator_status)) %>% 
    ungroup() %>% 
    mutate(indicator = "combined")
)
```

<br/>

## 4. Visualizing Data Layers {-}

### 4.1 Maps {-}

```{r simplify bd shapefiles and plot for visual inspection}
bhirgn_simple <- rmapshaper::ms_simplify(input = BHI_rgns_shp) %>% sf::st_as_sf()

## simplify the polygons before creating maps
## first transforming to match rnaturalearth basemap crs...
benthic_shp <- rmapshaper::ms_simplify(input = sf::st_transform(benthic_shp, crs = 4326)) %>%
  sf::st_as_sf()
pelagic_shp <- rmapshaper::ms_simplify(input = sf::st_transform(pelagic_shp, crs = 4326)) %>%
  sf::st_as_sf()
Fish_shp <- rmapshaper::ms_simplify(input = sf::st_transform(Fish_shp, crs = 4326)) %>% 
  sf::st_as_sf()
Seals_shp <- rmapshaper::ms_simplify(input = sf::st_transform(Seals_shp, crs = 4326)) %>% 
  sf::st_as_sf()

## basemap and color palettes
basemap <- ggplot2::ggplot(rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")) +
  geom_sf(size = 0.1, color = "burlywood", alpha = 0.4) +
  theme(panel.background = element_rect(fill = "#F8FBFC", color = "#E2EEF3")) +
  scale_x_continuous(limit = c(4, 32)) +
  scale_y_continuous(limit = c(53.5, 66)) 

cols <- c("indianred", "coral", "goldenrod1", "khaki", "lightblue", "steelblue")
```


#### 4.1.1 Benthic Habitats {-}

```{r map of benthic and pelagic habitats, eval = FALSE, results = "show", fig.width = 9.5, fig.height = 8.5}


## maps of BQR for each layer: benthic and pelagic habitats
basemap +
  geom_sf(
    data = benthic_shp, 
    mapping = aes(fill = BQR),
    alpha = 0.7, color = "steelblue", size = 0.3
  ) +
  # geom_sf(
  #   data = bhirgn_simple,
  #   size = 0.1, color = "burlywood", fill = NA
  # ) +
  scale_fill_viridis_c()
```


```{r map of fish and seal assessments, eval = FALSE, results = "show", fig.width = 9.5, fig.height = 8.5}
## Fish Open Sea areas Map
basemap + 
  geom_sf(
    data = mapsf %>% 
      filter(country == "Opensea"), 
    mapping = aes(fill = BQR), 
    alpha = 0.7, size = 0.2
  ) +
  scale_fill_viridis_c()


```


#### 4.1.2 Pelagic Habitats {-}

#### 4.1.3 Fish {-}

#### 4.1.4 Seals {-}


<br>

## 5. Considerations for `BHI3.0`



<br>

## 6. References

```{r bd references, child = refs_path, results = "asis", echo = FALSE}
```
