
```{r munitions setup}
## root location of the raw data
source(here::here("R", "setup.R"))
source(here::here("R", "spatial.R"))

dir_A <- file.path(dirname(dir_prep), "bhi-data")
dir_B <- file.path(dirname(dir_prep), "bhi-data", "BHI 2.0")
dir_rawdata <- file.path(dir_B, "Pressure", "munitions")
dir_intermediate <- file.path(here::here("data", "pressures", "bottom_trawling", version_year, "intermediate"))

bhi_basin_lookup <- read_delim(file.path(dir_prep, "supplement", "lookup_tabs", "bhi_basin_country_lookup.csv"), delim = ";") %>% 
  rename(subbasin = Subbasin) 
```

### 2.3 Data exploration Needed
Need to plot evaluate information associated with each data layer available.  

## 3. Pressure model
Option: total area in in BHI region associated with one or more munitions impact.

### 3. 1 Current conditions
Percent of BHI area with one or more munitions impact.  

### 3.2 Rescaling between 0 to 1

min value = 0

max value = BHI region with greatest percent munitions impact  


## Layer prep

```{r TODO, echo=FALSE, message=FALSE}
## read in data...

## plot data...

```
