
```{r nutrient load pressure data preamble, echo = FALSE, include = FALSE, error = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = TRUE, results = "asis", fig.width = 9.5, fig.height = 6)
source(here::here("R", "setup.R"))
source(here::here("R", "spatial.R"))
```

### 2.1 Datasets with Sources {-}
<br/>

Data used are official HELCOM reported data. 


#### 2.1.1 Nitrogen & Phosphorus (N & P) Loads {-}
<!-- dataset save location BHI_share/BHI 2.0/Pressure/nutrient_load -->

Data tables on total annual input of nitrogren and phosphorous (tonnes) as well as water flow into the Baltic Sea can be found [here](https://helcom.fi/wp-content/uploads/2020/06/BSEFS-Waterborne-nitrogen-and-phosphorus-inputs-and-water-flow-to-the-Baltic-Sea.pdf). Data on annual deposition of total nitrogen can be found [here](https://emep.int/publ/helcom/2019/B_BSEFS_N_dep_v3.pdf).

Actual total (e.g. air and waterborne, direct and point source discharges) annual input of nitrogren and phosphorous (tonnes) are accessed as part of the HELCOM Core Indicators, in the Pollution Load Compilation (PLC); see online resource about the most recent compilation [PLC-6 here](https://helcom.fi/helcom-at-work/projects/plc-6/). 


#### 2.1.2 Maximum Allowable Inputs (MAI) {-}
<!-- dataset save location BHI_share/BHI 2.0/Pressure/nutrient_load -->

MAI values for each basin are found in [Summary report on the development of revised Maximum Allowable Inputs and updated Country Allocated Reduction Targets of the Baltic Sea Action Plan](http://www.helcom.fi/Documents/Ministerial2013/Associated%20documents/Supporting/Summary%20report%20on%20MAI-CART.pdf) in Table 3 'Maximum Allowable nutrient Inputs to main Baltic Sea sub-basins', on page 8.  

These are the same values as presented in Table 2.2 page 11 in the [Review of the Fifth Baltic Sea PLC for the 2013 HELCOM Ministerial Meeting](https://helcom.fi/media/publications/BSEP141.pdf)

<br/>


### 2.2 Centralization & Normalization

```{r nutrient load pressure load raw data, echo = TRUE, message = FALSE, warning = FALSE, results = "hide"}
## root location of the pdfs containing the data
dir_rawdata <- file.path(dir_B, "Pressure", "nutrient_load")

library(pdftools)
library(stringr)
```

#### 2.2.1 Extract Data from PDFs

```{r extract the tonnes nutrients input and water inflow values}

pdf_file_mai <- file.path(dir_B, "Pressure", "nutrient_load", "Summary-report-on-MAI-CART.pdf")
txt <- pdf_text(pdf_file_mai)

mai_txt <- str_extract(txt[9], regex("(?<=Maximum Allowable Inputs\n  ).*(?=4 HOW ARE THE UPDATED CART WAS DETERMINED)", dotall = TRUE))

mai_tab <- as.data.frame(str_split(mai_txt, "\n")[[1]][2:9])
colnames(mai_tab) <- "alldata"

newcolnames <- str_split(str_split(mai_txt, "\n")[[1]][1], "(?<=[a-z])  +")[[1]] %>% 
  str_replace_all(" ", "_") %>% 
  str_remove_all(",|-") %>% 
  str_to_lower()

mai_tab <- tidyr::separate(mai_tab, col = "alldata", into = newcolnames, sep = "(?<=[a-z0-9])  +") %>% 
  rename(baltic_sea_basin = baltic_sea_subbasin) %>% 
  mutate(
    baltic_sea_basin = str_extract(baltic_sea_basin, "[A-Z].*"),
    total_nitrogen_tonnes = str_remove(total_nitrogen_tonnes, ","),
    total_phosphorus_tonnes = str_remove(total_phosphorus_tonnes, ",")
  ) %>% 
   mutate(
    total_nitrogen_tonnes = as.numeric(total_nitrogen_tonnes),
    total_phosphorus_tonnes = as.numeric(total_phosphorus_tonnes)
  ) %>% 
  filter(baltic_sea_basin != "Baltic Sea") 
```


```{r extract the tonnes nitrogen and phosphorus input and water inflow values}
pdf_file_inflows <- file.path(dir_B, "Pressure", "nutrient_load", "Waterborne-N-and-P-inputs-and-water-flow.pdf")
txt <- pdf_text(pdf_file_inflows)


## extract total annual waterborne flow 
## sum of riverine flow and direct flow (flow for point sources discharging direct into the Baltic Sea)

flows_txt <- str_extract(txt[15], regex("(?<=Flow \\(m3/s-1\\)\n  ).*(?=\n)", dotall = TRUE))
flows_tab <- as.data.frame(str_split(flows_txt, "\n")[[1]][2:24])
colnames(flows_tab) <- "alldata"

newcolnames <- str_split(str_split(flows_txt, "\n")[[1]][1], "(?<=[a-zA-Z])  +")[[1]] %>% 
  str_remove_all(" ")

flows_tab <- tidyr::separate(flows_tab, col = "alldata", into = newcolnames, sep = "(?<=[a-z0-9])  +") %>% 
  mutate_all(as.numeric) %>% 
  rename(year = Sum)


## extract total waterborne (riverine + direct) nitrogen inputs

n_txt <- str_extract(txt[18], regex("(?<= TN \\(tonnes\\)\n  ).*(?=\n)", dotall = TRUE))
n_tab <- as.data.frame(str_split(n_txt, "\n")[[1]][2:24])
colnames(n_tab) <- "alldata"

n_tab <- tidyr::separate(n_tab, col = "alldata", into = newcolnames, sep = "(?<=[a-z0-9])  +") %>% 
  mutate_all(as.numeric) %>% 
  rename(year = Sum)


## extract total waterborne (riverine + direct) phosphorus inputs

p_txt <- str_extract(txt[21], regex("(?<= TP \\(tonnes\\)\n  ).*(?=\n)", dotall = TRUE))
p_tab <- as.data.frame(str_split(p_txt, "\n")[[1]][2:24])
colnames(p_tab) <- "alldata"

p_tab <- tidyr::separate(p_tab, col = "alldata", into = newcolnames, sep = "(?<=[a-z0-9])  +") %>% 
  mutate_all(as.numeric) %>% 
  rename(year = Sum)
```


```{r tonnes nitrogen atmospheric deposition}
# pdf_atmos_n <- file.path(dir_B, "Pressure", "nutrient_load", "Atmospheric-deposition-N.pdf")
pdf_atmos_n <- file.path(dir_B, "Pressure", "nutrient_load", "Atmospheric-deposition-N.pdf")
txt <- pdf_text(pdf_atmos_n)


## extract total annual deposition of total nitrogen to the subbasins 
## units are ktonnes(N)/year

atmos_n_txt <- str_extract(txt[12], regex("(?<=Baltic Sea \\(sum of all sub-basins\\)\\.\n  ).*(?=\n)", dotall = TRUE))
atmos_n_tab <- as.data.frame(str_split(atmos_n_txt, "\n")[[1]][4:26])
colnames(atmos_n_tab) <- "alldata"

newcolnames <- str_split(str_split(atmos_n_txt, "\n")[[1]][3], "(?<=[a-zA-Z])  +")[[1]] %>% 
  str_remove_all(" ")
newcolnames <- c("year", newcolnames, "BAS")

atmos_n_tab <- tidyr::separate(atmos_n_tab, col = "alldata", into = newcolnames, sep = "(?<=[a-z0-9])  +") %>% 
  mutate_all(as.numeric)
```

 
#### 2.2.2 Standardize Units

```{r normalize nutrient input by river flows}
## not sure how to normalize...

n_normalized <- left_join(
  n_tab %>% 
    tidyr::pivot_longer(
      cols = matches("^[A-Z]{2,3}$"),
      names_to = "basin_code", 
      values_to = "input_tonnes"
    ),
  flows_tab %>% 
    tidyr::pivot_longer(
      cols = matches("^[A-Z]{2,3}$"),
      names_to = "basin_code", 
      values_to = "flow_cubic_meters_per_second"
    ),
  by = c("year", "basin_code")
)
n_normalized <- n_normalized %>% 
  mutate(input_per_flow = input_tonnes/flow_cubic_meters_per_second) %>% 
  mutate(basin_code = as.character(basin_code))
  
p_normalized <- left_join(
  p_tab %>% 
    tidyr::pivot_longer(
      cols = matches("^[A-Z]{2,3}$"),
      names_to = "basin_code", 
      values_to = "input_tonnes"
    ),
  flows_tab %>% 
    tidyr::pivot_longer(
      cols = matches("^[A-Z]{2,3}$"),
      names_to = "basin_code", 
      values_to = "flow_cubic_meters_per_second"
    ),
  by = c("year", "basin_code")
)
p_normalized <- p_normalized %>% 
  mutate(input_per_flow = input_tonnes/flow_cubic_meters_per_second) %>% 
  mutate(basin_code = as.character(basin_code))
```


```{r match basin codes with subbasin names as found in mai table}

## codes from figure 1 page 5 of pdf containing the input and flow data                                               
basinsdf <- tibble(
  baltic_sea_basin = c(
    "Kattegat", "Danish Straits", "Baltic Proper", "Bothnian Sea", 
    "Bothnian Bay", "Gulf of Riga", "Gulf of Finland"
  ),
  basin_code = c(
    "KAT", "DS", "BAP", "BOS",
    "BOB", "GUR", "GUF"
  )
)
n_normalized <- left_join(
  filter(n_normalized, basin_code != "BAS"), 
  basinsdf, 
  by = "basin_code"
)
p_normalized <- left_join(
  filter(p_normalized, basin_code != "BAS"), 
  basinsdf, 
  by = "basin_code"
)


## wrangle atmospheric N deposition to match basin codes with subbasins
## codes from figure 4 page 8 of pdf containing the atmospheric nitrogen data   
basinsdf <- tibble(
  baltic_sea_subbasin = c(
    "Archipelago Sea", "Baltic Proper", "Bothnian Bay", "Bothnian Sea", 
    "Gulf of Finland", "Gulf of Riga", "Kattegat", "The Sound", "Western Baltic"
  ),
  basin_code = c(
    "ARC", "BAP", "BOB", "BOS",
    "GUF", "GUR", "KAT", "SOU", "WEB"
  ),
  baltic_sea_basin = c(
    "Bothnian Sea", "Baltic Proper", "Bothnian Bay", "Bothnian Sea", 
    "Gulf of Finland", "Gulf of Riga", "Kattegat", "Danish Straits", "Danish Straits"
  )
)

## but need to aggregate to match the MAI and waterborne inputs spatial units...
atmos_n_tab <- atmos_n_tab %>% 
  tidyr::pivot_longer(
    cols = matches("^[A-Z]{3}$"),
    names_to = "basin_code", 
    values_to = "input"
  ) %>% 
  mutate(input_tonnes = input*1000) %>% 
  filter(basin_code != "BAS") %>% 
  left_join(basinsdf, by = "basin_code") %>% 
  group_by(year, baltic_sea_basin) %>% 
  summarise(input_tonnes = sum(input_tonnes))
```


#### 2.2.3 Save to BHI Database

```{r save wrangled nutrient load raw datasets}
dir_intermediate <- file.path(dir_prep, "data", "pressures", "nutrient_load", version_year, "intermediate")
write_csv(
  n_normalized,
  file.path(dir_intermediate, "basin_n_load.csv")
)
write_csv(
  atmos_n_tab,
  file.path(dir_intermediate, "basin_n_atmos_load.csv")
)
write_csv(
  p_normalized,
  file.path(dir_intermediate, "basin_p_load.csv")
)
write_csv(
  mai_tab,
  file.path(dir_intermediate, "n_p_load_targets.csv")
)
```

<br/>