
```{r nutrient load pressure data preamble, echo = FALSE, include = FALSE, error = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = TRUE, results = "asis", fig.width = 9.5, fig.height = 6)
source(here::here("R", "setup.R"))
source(here::here("R", "spatial.R"))
```

### 2.1 Datasets with Sources {-}
<br/>

Data used are official HELCOM reported data. 


#### 2.1.1 Nitrogen & Phosphorus (N & P) Loads {-}
<!-- dataset save location BHI_share/BHI 2.0/Pressure/nutrient_load -->

Data tables on total annual input of nitrogren and phosphorous (tonnes) as well as water flow into the Baltic Sea can be found [here](https://helcom.fi/wp-content/uploads/2020/06/BSEFS-Waterborne-nitrogen-and-phosphorus-inputs-and-water-flow-to-the-Baltic-Sea.pdf)

Actual total (e.g. air and waterborne, direct and point source discharges) annual input of nitrogren and phosphorous (tonnes) are accessed as part of the HELCOM Core Indicators, in the Pollution Load Compilation (PLC); see online resource about the most recent compilation [PLC-6 here](https://helcom.fi/helcom-at-work/projects/plc-6/). 


#### 2.1.2 Maximum Allowable Inputs (MAI) {-}
<!-- dataset save location BHI_share/BHI 2.0/Pressure/nutrient_load -->

MAI values for each basin are found in [Summary report on the development of revised Maximum Allowable Inputs and updated Country Allocated Reduction Targets of the Baltic Sea Action Plan](http://www.helcom.fi/Documents/Ministerial2013/Associated%20documents/Supporting/Summary%20report%20on%20MAI-CART.pdf) in Table 3 'Maximum Allowable nutrient Inputs to main Baltic Sea sub-basins', on page 8.  


---

<br/>


### 2.2 Centralization & Normalization

```{r nutrient load pressure load raw data, echo = TRUE, message = FALSE, warning = FALSE, results = "hide"}
## root location of the pdfs containing the data
dir_rawdata <- file.path(dir_B, "Pressure", "nutrient_load")

library(pdftools)
library(stringr)
```

#### 2.2.1 Extract Data from PDFs

```{r extract the tonnes nutrients input and water inflow values}

pdf_file_mai <- file.path(dir_B, "Pressure", "nutrient_load", "Summary-report-on-MAI-CART.pdf")
txt <- pdf_text(pdf_file_mai)

mai_txt <- str_extract(txt[9], regex("(?<=Maximum Allowable Inputs\n  ).*(?=4 HOW ARE THE UPDATED CART WAS DETERMINED)", dotall = TRUE))

mai_tab <- as.data.frame(str_split(mai_txt, "\n")[[1]][2:9])
colnames(mai_tab) <- "alldata"

newcolnames <- str_split(str_split(mai_txt, "\n")[[1]][1], "(?<=[a-z])  +")[[1]] %>% 
  str_replace_all(" ", "_") %>% 
  str_remove_all(",|-") %>% 
  str_to_lower()

mai_tab <- tidyr::separate(mai_tab, col = "alldata", into = newcolnames, sep = "(?<=[a-z0-9])  +") %>% 
  mutate(
    baltic_sea_subbasin = str_extract(baltic_sea_subbasin, "[A-Z].*"),
    total_nitrogen_tonnes = str_remove(total_nitrogen_tonnes, ","),
    total_phosphorus_tonnes = str_remove(total_nitrogen_tonnes, ",")
  ) %>% 
   mutate(
    total_nitrogen_tonnes = as.numeric(total_nitrogen_tonnes),
    total_phosphorus_tonnes = as.numeric(total_nitrogen_tonnes)
  )
```


```{r extract the tonnes nitrogen and phosphorus input and water inflow values}
pdf_file_inflows <- file.path(dir_B, "Pressure", "nutrient_load", "Waterborne-N-and-P-inputs-and-water-flow.pdf")
txt <- pdf_text(pdf_file_inflows)


## extract total annual waterborne flow 
## sum of riverine flow and direct flow (flow for point sources discharging direct into the Baltic Sea)

flows_txt <- str_extract(txt[15], regex("(?<=Flow \\(m3/s-1\\)\n  ).*(?=\n)", dotall = TRUE))
flows_tab <- as.data.frame(str_split(flows_txt, "\n")[[1]][2:24])
colnames(flows_tab) <- "alldata"

newcolnames <- str_split(str_split(flows_txt, "\n")[[1]][1], "(?<=[a-zA-Z])  +")[[1]] %>% 
  str_remove_all(" ")

flows_tab <- tidyr::separate(flows_tab, col = "alldata", into = newcolnames, sep = "(?<=[a-z0-9])  +") %>% 
  mutate_all(as.numeric) %>% 
  rename(year = Sum)


## extract total waterborne (riverine + direct) nitrogen inputs

n_txt <- str_extract(txt[18], regex("(?<= TN \\(tonnes\\)\n  ).*(?=\n)", dotall = TRUE))
n_tab <- as.data.frame(str_split(n_txt, "\n")[[1]][2:24])
colnames(n_tab) <- "alldata"

n_tab <- tidyr::separate(n_tab, col = "alldata", into = newcolnames, sep = "(?<=[a-z0-9])  +") %>% 
  mutate_all(as.numeric) %>% 
  rename(year = Sum)


## extract total waterborne (riverine + direct) phosphorus inputs

p_txt <- str_extract(txt[21], regex("(?<= TP \\(tonnes\\)\n  ).*(?=\n)", dotall = TRUE))
p_tab <- as.data.frame(str_split(p_txt, "\n")[[1]][2:24])
colnames(p_tab) <- "alldata"

p_tab <- tidyr::separate(p_tab, col = "alldata", into = newcolnames, sep = "(?<=[a-z0-9])  +") %>% 
  mutate_all(as.numeric) %>% 
  rename(year = Sum)
```

 
#### 2.2.2 Standardize Units

```{r normalize nutrient input by river flows}
n_normalized <- left_join(
  n_tab %>% 
    tidyr::pivot_longer(
      cols = matches("^[A-Z]{2,3}$"),
      names_to = "basin_code", 
      values_to = "input_tonnes"
    ),
  flows_tab %>% 
    tidyr::pivot_longer(
      cols = matches("^[A-Z]{2,3}$"),
      names_to = "basin_code", 
      values_to = "flow_cubic_meters_per_second"
    ),
  by = c("year", "basin_code")
)
n_normalized <- n_normalized %>% 
  mutate(flow_normalized_input = input_tonnes/flow_cubic_meters_per_second) %>% 
  mutate(basin_code = as.character(basin_code))
  
p_normalized <- left_join(
  p_tab %>% 
    tidyr::pivot_longer(
      cols = matches("^[A-Z]{2,3}$"),
      names_to = "basin_code", 
      values_to = "input_tonnes"
    ),
  flows_tab %>% 
    tidyr::pivot_longer(
      cols = matches("^[A-Z]{2,3}$"),
      names_to = "basin_code", 
      values_to = "flow_cubic_meters_per_second"
    ),
  by = c("year", "basin_code")
)
p_normalized <- p_normalized %>% 
  mutate(flow_normalized_input = input_tonnes/flow_cubic_meters_per_second) %>% 
  mutate(basin_code = as.character(basin_code))
```


```{r match basin codes with subbasin names as found in mai table}

## codes from figure 1 page 5 of pdf containing the input and flow data                                               
basinsdf <- tibble(
  baltic_sea_subbasinbasin_code = c(
    "Kattegat", "Danish Straits", "Baltic Proper", "Bothnian Sea", 
    "Bothnian Bay", "Gulf of Riga", "Gulf of Finland", "Baltic Sea"
  ),
  basin_code = c(
    "KAT", "DS", "BAP", "BOS",
    "BOB", "GUR", "GUF", "BAS"
  )
)
n_normalized <- left_join(n_normalized, basinsdf, by = "basin_code")
p_normalized <- left_join(p_normalized, basinsdf, by = "basin_code")
```


#### 2.2.3 Save to BHI Database

```{r save wrangled nutrient load raw datasets}
dir_intermediate <- file.path(dir_prep, "data", "pressures", "nutrient_load", version_year, "intermediate")
write_csv(
  n_normalized,
  file.path(dir_intermediate, "basin_n_load.csv")
)
write_csv(
  p_normalized,
  file.path(dir_intermediate, "basin_p_load.csv")
)
write_csv(
  mai_tab,
  file.path(dir_intermediate, "N_P_load_targets.csv")
)
```

<br/>

### 2.3 Initial Data Exploration

