## 4. Prepare bottom trawl pressure layer


```{r setup, echo = FALSE, message = FALSE}

## Libraries
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(RMySQL)
library(stringr)
library(tools)
library(rprojroot) # install.packages('rprojroot')
library(stringr) # install.packages("stringr")
library(rgdal)  # install.packages('rgdal')
library(raster) # install.packages('raster')
library(rgeos)  # install.packages('rgeos')

source('~/github/bhi/baltic2015/prep/common.r')

## rprojroot
root <- rprojroot::is_rstudio_project

## make_path() function to 
make_path <- function(...) rprojroot::find_root_file(..., criterion = is_rstudio_project)

dir_layers = make_path('baltic2015/layers') # replaces  file.path(dir_baltic, 'layers')


# root$find_file("README.md")
# 
# root$find_file("ao_need_gl2014.csv")
# 
# root <- find_root_file("install_ohicore.r", 
# 
# withr::with_dir(
#   root_file("DESCRIPTION"))

dir_layers = make_path('baltic2015/layers') # replaces  file.path(dir_baltic, 'layers')
dir_prep = make_path('baltic2015/prep') # replaces  file.path(dir_baltic, 'layers')
dir_trawl = file.path(dir_prep, 'pressures/bottom_trawling')

## add a README.md to the prep directory with the rawgit.com url for viewing on GitHub
create_readme(dir_trawl, 'bottom_trawling_prep.rmd') 
```

### 4.1 Read in Data and Explore

```{r read in data, cache = TRUE}

## directory setup
dir_M <- c('Windows' = '//mazu.nceas.ucsb.edu/ohi',
           'Darwin'  = '/Volumes/ohi',    ### connect (cmd-K) to smb://mazu/ohi
           'Linux'   = '/home/shares/ohi')[[ Sys.info()[['sysname']] ]]

if (Sys.info()[['sysname']] != 'Linux' & !file.exists(dir_M)){
  warning(sprintf("The Mazu directory dir_M set in src/R/common.R does not exist. Do you need to mount Mazu: %s?", dir_M))
  # TODO: @jennifergriffiths reset dir_M to be where the SRC hold your shapefiles
}

dir_shp <- file.path(dir_M, 'git-annex/Baltic')

# list.files(file.path(dir_shp, 'Bottom_trawling_pressure'))
# contains data from 2009 to 2013

## read in bottom trawling pressure data from 2010

trawling_raw_2009 <- rgdal::readOGR(dsn = path.expand(file.path(dir_shp, 'Bottom_trawling_pressure')),
                      layer = 'HELCOM_effort_year_MBCG_VMS_2009')  

trawling_raw_2010 <- rgdal::readOGR(dsn = path.expand(file.path(dir_shp, 'Bottom_trawling_pressure')),
                      layer = 'HELCOM_effort_year_MBCG_VMS_2010')  

trawling_raw_2011 <- rgdal::readOGR(dsn = path.expand(file.path(dir_shp, 'Bottom_trawling_pressure')),
                      layer = 'HELCOM_effort_year_MBCG_VMS_2011') 

trawling_raw_2012 <- rgdal::readOGR(dsn = path.expand(file.path(dir_shp, 'Bottom_trawling_pressure')),
                      layer = 'HELCOM_effort_year_MBCG_VMS_2012') 

trawling_raw_2013 <- rgdal::readOGR(dsn = path.expand(file.path(dir_shp, 'Bottom_trawling_pressure')),
                      layer = 'HELCOM_effort_year_MBCG_VMS_2013') 


##  extract crs_trawling for scaling 

# trawling_raw_2010@proj4string
# +proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs 
crs_trawling = CRS("+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs")

## read in bhi shape files
bhi = rgdal::readOGR(dsn = path.expand(file.path(dir_shp, 'BHI_MCG_shapefile')),
                      layer = 'BHI_MCG_11052016')  
bhi_data = bhi@data

## change bhi coord. system to the same as trawling efforts
bhi = spTransform(bhi, crs_trawling)

```

### plot trawling efforts with BHI boundaries

```{r plot data, fig.width= 4, fig.height = 16}

par(mfrow = c(5, 1), mar = c(1,1,1,1))

plot(trawling_raw_2009, col = 'blue', border = "blue"); plot(bhi, border = "grey", main = "BHI regions and Trawling Efforts overlay", add = TRUE); legend('bottomright', c("BHI regions", "Trawling efforts  (2010)"), lty = c(1,1), lwd = c(2.5, 2.5, 2.5), col = c("grey", "blue"), text.font = 1, box.lty = 0 )

plot(trawling_raw_2010, col = 'blue', border = "blue"); plot(bhi, border = "grey", main = "BHI regions and Trawling Efforts overlay", add = TRUE); legend('bottomright', c("BHI regions", "Trawling efforts (2010)"), lty = c(1,1), lwd = c(2.5, 2.5, 2.5), col = c("grey", "blue"), text.font = 1, box.lty = 0 )

plot(trawling_raw_2011, col = 'blue', border = "blue"); plot(bhi, border = "grey", main = "BHI regions and Trawling Efforts overlay", add = TRUE); legend('bottomright', c("BHI regions", "Trawling efforts (2011)"), lty = c(1,1), lwd = c(2.5, 2.5, 2.5), col = c("grey", "blue"), text.font = 1, box.lty = 0 )

plot(trawling_raw_2012, col = 'blue', border = "blue"); plot(bhi, border = "grey", main = "BHI regions and Trawling Efforts overlay", add = TRUE); legend('bottomright', c("BHI regions", "Trawling efforts (2012)"), lty = c(1,1), lwd = c(2.5, 2.5, 2.5), col = c("grey", "blue"), text.font = 1, box.lty = 0 )

plot(trawling_raw_2013, col = 'blue', border = "blue"); plot(bhi, border = "grey", main = "BHI regions and Trawling Efforts overlay", add = TRUE); legend('bottomright', c("BHI regions", "Trawling efforts (2013)"), lty = c(1,1), lwd = c(2.5, 2.5, 2.5), col = c("grey", "blue"), text.font = 1, box.lty = 0 )

```

### 4.2 Intersect trawling effort with BHI shapefiles  

```{r read all pressure data layers and intersect with bhi shape file, echo = FALSE, eval = FALSE}
## read all pressure data layers and save as .csv in temp_data folder 

trawling_file_list = list.files(file.path(dir_shp, 'Bottom_trawling_pressure'), 
                                pattern = 'HELCOM_effort_year_MBCG_VMS_[0-9]+.dbf')

for (i in trawling_file_list) {
  
  i = file_path_sans_ext(i)
  
  # read in shape file 
  trawl = rgdal::readOGR(dsn = path.expand(file.path(dir_shp, 'Bottom_trawling_pressure')), 
                         layer = i) 
  
  # intersect trawling and bhi shape file
  trawling_bhi <- raster::intersect(trawl, bhi)
  
  # save as Polygon file
  rgdal::writeOGR(trawling_bhi,
                  dsn = path.expand(file.path(dir_trawl, "temp_data")),
                  driver = "ESRI Shapefile",
                  layer = paste0('trawling_bhi_intersect_', substr(i, 29, 32)),
                  overwrite_layer = T)
  
  # extract data and save as .csv
  write_csv(trawling_bhi@data, file.path(dir_trawl, "temp_data", paste0('trawling_bhi_intersect_', substr(i, 29, 32), '.csv')))

}


```



### 4.4 Extract total fishing hours by BHI region for each year  

```{r extract total hours, eval = FALSE, echo = FALSE}

trawling_bhi_intersect_file_list = list.files(file.path(dir_trawl, 'temp_data'), 'trawling_bhi_intersect_[0-9]+.csv')

i = 1  
  
for (i in trawling_bhi_intersect_file_list) {
  
dat = read_csv(file.path(dir_trawl, 'temp_data', i))

hours_per_area = dat %>%
  dplyr::select(fishing_ho, BHI_ID) %>%
  group_by(BHI_ID) %>%
  summarize(total_hrs = sum(fishing_ho)) %>%
  left_join(dplyr::select(bhi_data, BHI_ID, Area_km2), 
            by = 'BHI_ID') %>%
  mutate(hours_per_area = total_hrs/Area_km2) %>%
  dplyr::select(BHI_ID, hours_per_area)

write_csv(hours_per_area, file.path(dir_trawl, 'temp_data', sprintf('hours_per_area_%s.csv', substr(hours_per_area_file_list[i], 16, 19))))
  
}

```


### 4.5 Calculate and Plot Fising Hours per km2

Plot below shows the trawling efforts per area in each region in each year. Note that each year's data covers slightly different BHI regions, and contain 33-37 regions out of 42. 

For years when a region reported no trawling data, the data gap was filled with the average trawling efforts of the rest of the years for that region. 

```{r calculate and plot hours per area region}

## calculate hours per km2, per year, in each region

hours_per_area_file_list = list.files(file.path(dir_trawl, 'temp_data'), 'hours_per_area_[0-9]+.csv' )

hours_per_area_all = data.frame()

for (i in hours_per_area_file_list) {

 d = read_csv(file.path(dir_trawl, 'temp_data', i)) %>%
   mutate(year = substr(i, 16, 19))

 hours_per_area_all = rbind(hours_per_area_all, d) %>%
   mutate(year = as.factor(year))

}

# ## fill in the missing years for each region; TODO: also fill in the missing BHI_IDs
# hours_per_area_filled = hours_per_area_all %>%
#   tidyr::complete(year = full_seq(year, 1), nesting(BHI_ID), fill = list(hours_per_area = 0)) %>%
#   arrange(BHI_ID, year) %>%
#   mutate(year = as.factor(year))

plot_hours_area = ggplot(hours_per_area_all, aes(x = BHI_ID, y = hours_per_area, fill = year)) +
  geom_bar(position="dodge", stat = 'identity') +
  labs(title = 'Fishing efforts per km2 2009 - 2014', 
       x = 'BHI ID', 
       y = 'Hours per km2')

print(plot_hours_area)

# fill in the NA's in hours_per_area, select the most recent year 2014
hours_per_area_filled <-  hours_per_area_all %>% 
  mutate(year = as.integer(as.character(year))) %>% 
  tidyr::complete(year = full_seq(year, 1), nesting(BHI_ID)) %>% 
  group_by(BHI_ID) %>% 
  mutate(ave_hours = mean(hours_per_area, na.rm = TRUE), 
         hours_per_area = ifelse(is.na(hours_per_area), ave_hours, hours_per_area)) %>% 
  ungroup
  

# rescale hours_per_area

bottom_trawling_score <- hours_per_area_filled %>% 
  group_by(BHI_ID) %>% 
  mutate(max = max(hours_per_area), 
         min = min(hours_per_area),
         score = round((hours_per_area - min) / (max - min), 1)) %>% 
  filter(year == 2014) %>% 
  dplyr::select(BHI_ID, 
         score)

# fill in the missing BHI_IDs, as this data set only contains 37 out of 42 BHI regions
BHI_rgns <- read.csv(file.path(dir_prep, 'bhi_basin_country_lookup.csv'), sep = ";") %>% 
  dplyr::select(BHI_ID)

bottom_trawling_score <- full_join(BHI_rgns, bottom_trawling_score, 
                                   by = "BHI_ID") %>% 
  mutate(score = ifelse(is.na(score), 0, score), 
         score = ifelse(BHI_ID == '32', 0, score)) %>% 
  dplyr::select(rgn_id = BHI_ID, 
         pressure_score = score)

write_csv(bottom_trawling_score, file.path(dir_layers, 'hab_bottom_trawl_bhi2015.csv')) 

```
 
# plot bottom trawling scores per region 

```{r plot scores}

ggplot(bottom_trawling_score, aes(x = rgn_id, y = pressure_score)) +
  geom_bar(stat = 'identity')


```

