
```{r eut data preamble, include = FALSE, echo = TRUE, message = FALSE, warning = FALSE}
source(here::here("R", "setup.R"))
source(here::here("R", "data.R"))
source(here::here("R", "spatial.R"))

## root location of the raw data
dir_A <- file.path(dirname(dir_prep), "bhi-data")
dir_B <- file.path(dirname(dir_prep), "bhi-data", "BHI 2.0")
dir_intermediate <- file.path(here::here("data", "CS", version_year, "intermediate"))
dir_rawdata <- file.path(dir_B, "Goals", "CS")
```

### 2.1 Datasets with Sources {-}
<br/>

#### 2.1.1 Zostera marina Data {-}

**Zostera marina observations (LC)**

Observations of Zostera marina were collected from the Baltic Sea area for HELCOM Red List species list.
Zostera marina has been placed to the Red List category of Least Concern (LC) species.

Dataset for download contains spatial grid of the Baltic Sea. 
Distribution of the species can be found in corresponding name column. 
Values are coded:
0 = no observations
1 = present before year 1995 or in 1995
2 = present after year 1995
3 = present before and after year 1995

<!-- dataset save location BHI_share/BHI 2.0/Goals/CS/ -->
```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Biodiversity:", "Red List"), 
  c("Macrophytes:", "Least Concern (LC)"),
  c("Least Concern (LC)", "Zostera marina (LC)")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [HELCOM Map and Data Service](http://metadata.helcom.fi/geonetwork/srv/eng/catalog.search#/metadata/dd90a6aa-1d65-4158-a2e7-7ca3ee87d615) <br/> Downloaded 13 July 2020 by Andrea De Cervo") 
```

<br/>

**Zostera marina distribution**

Distribution of eelgrass based on data submission by HELCOM contracting parties. 
Mainly pointwise occurrences of eelgrass were submitted, originally gathered in national mapping and monitoring campaigns, or for scientific research. Polygon data from Puck Bay (Poland) was digitized based on Polish Marine Atlas and Orlowo cliff area was added based on expert knowledge. From Estonian waters, a predictive model was used (200m resolution), that was converted to presence/absence using minimized difference threshold (MDT) criteria. 
Spatial resolution: 1 km x 1 km grid (originally 5 x 5 km).

Raster value representing presence (1) or absence (0) of Zostera marina.

<!-- dataset save location BHI_share/BHI 2.0/Goals/CS/ -->
```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Biodiversity:", "Ecosystem components (BSII)"), 
  c("Benthic species:", "Zostera marina distribution")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [HELCOM Map and Data Service](http://metadata.helcom.fi/geonetwork/srv/eng/catalog.search#/metadata/ca327bb1-d3cb-46c2-8316-f5f62f889090) <br/> Downloaded 13 July 2020 by Andrea De Cervo") 
```

<br/>

**Seagrass Meadows**

The data set is showing seagrass meadows in the Baltic Sea area as points used for the BRISK project (Sub-regional risk of spill of oil and hazardous substances in the Baltic Sea; learn more [here](http://www.brisk.helcom.fi/)). This dataset has been produced by [COWI](http://www.cowi.dk) based on data from HELCOM.

<!-- dataset save location BHI_share/BHI 2.0/Goals/CS/ -->
```{r echo = FALSE, results = "asis"}
tab <- t(data.frame(
  c("Shipping:", "Risk assessment (BRISK project)"), 
  c("Environmental parameters:", "Seagrass meadows point (BRISK)")))

colnames(tab) <- c("Option", "Specification")
rownames(tab) <- NULL

knitr::kable(tab, caption = "Source: [HELCOM Map and Data Service](http://metadata.helcom.fi/geonetwork/srv/eng/catalog.search#/metadata/6a0c8a15-3ff0-48e3-a878-0aa388122867) <br/> Downloaded 13 July 2020 by Andrea De Cervo") 
```

<br/>


### 2.2 Centralization & Normalization {-}

#### 2.2.1 Zostera marina shapefile {-}

We will only explore the dataset chosen by BHI 1.0 (Alternative 4), unless new available better data.

```{r load cs raw data, echo = TRUE, message = FALSE, warning = FALSE}
# library(readr)
library(raster)
# load all raw data (containing the newest data)
macrophytes <- sf::st_read(file.path(
  dir_rawdata, "Join_macrophytes_20200713",
  "Join_macrophytes.shp"
))

## extent:
extent(macrophytes)
# xmin        : 4e+06 
# xmax        : 5900000 
# ymin        : 2800000 
# ymax        : 5500000 

bhi <- sf::st_read(file.path(
  dir_A, "Shapefiles", "BHI_shapefile",
  "BHI_shapefile.shp"
))

## extent:
extent(bhi)
# xmin       : 9.420778 
# xmax       : 30.34708 
# ymin       : 53.60164 
# ymax       : 65.90708 

seagrass_meadows <- sf::st_read(file.path(
  dir_rawdata, "Seagrass_meadows_point_(BRISK)_20200713",
  "Seagrass meadows point BRISK.shp"
))

## extent:
extent(seagrass_meadows)
# xmin       : 4213701 
# xmax       : 5331130 
# ymin       : 3424054 
# ymax       : 4317448 
```

<br/>

```{r polygon data explore, echo=TRUE, message=FALSE}
## inspect BHI regions ----
# bhi <- sf::st_read(file.path(dir_A, "Shapefiles", "BHI_shapefile", "BHI_shapefile.shp"))
bhi <- rgdal::readOGR(dsn = path.expand(file.path(dir_A, "Shapefiles", "BHI_shapefile")),
                      layer = 'BHI_shapefile') 
# bhi # coord. ref. : +proj=longlat
## explore/plot
# plot(bhi) #takes ~15 seconds
# bhi@data
head(bhi@data)

# macrophytes <- sf::st_read(file.path(dir_rawdata, "Join_macrophytes_20200713", "Join_macrophytes.shp"))
macrophytes <- rgdal::readOGR(dsn = path.expand(file.path(dir_rawdata, "Join_macrophytes_20200713")),
                      layer = 'Join_macrophytes')

# mpa # coord. ref. : +proj=laea

## plot
# plot(macrophytes) # quick
# head(macrophytes@data)

## transform bhi to macrophytes coordinate reference system (CRS) ----
## LAEA: http://spatialreference.org/ref/epsg/etrs89-etrs-laea/
## https://www.nceas.ucsb.edu/~frazier/RSpatialGuides/OverviewCoordinateReferenceSystems.pdf
bhi_transf <- spTransform(bhi, macrophytes@proj4string)
# bhi # coord. ref. : +proj=laea

raster::extent(bhi_transf)
# xmin        : 4283762 
# xmax        : 5442611 
# ymin        : 3397830 
# ymax        : 4818369 

## save transformed shapefile
#writeOGR(bhi_transf, dsn = path.expand(dir_intermediate), layer = 'bhi_transf', driver = 'ESRI Shapefile', overwrite = TRUE) 
```

<br/>

#### 2.2.2  Intersect the BHI shapefiles and zostera data

**Problem**: The area listed in the Zostera file only contained the location of each sampling point, but not the area of vegetation coverage. The area data seems to be the total area of the BHI region. Therefore, we can't calculate the total area of Zostera as planned.

```{r intersect and save csv, eval = FALSE, message = FALSE}
library(rgeos)
bhi_buff <- gBuffer(bhi_transf, byid=TRUE, width=0)
macrophytes_buff <- gBuffer(macrophytes, byid=TRUE, width=0)

cs_bhi_intersect <- raster::intersect(macrophytes_buff, bhi_buff)
# head(cs_bhi_intersect)
# plot(cs_bhi_intersect)

## save intersected shapefile
#writeOGR(cs_bhi_intersect, dsn = path.expand(dir_intermediate), layer = 'cs_bhi_intersect', driver = 'ESRI Shapefile', overwrite = TRUE) 
```
