---
title: "wgi_pressure_prep.rmd"
output:
  github_document:
    toc: true
    toc_depth: 3
params: 
    datasource: csv
---

# Prepare Worldwide Governance Indicators (WGI) for Social pressure data layer 

## 1. Background 
The [Worldwide Governance Indicators (WGI) project](http://info.worldbank.org/governance/wgi/index.aspx#home) reports aggregate and individual governance indicators for 215 economies over the period 1996–2014, for six dimensions of governance: 

Voice and Accountability  
Political Stability and Absence of Violence/Terrorism  
Government Effectiveness  
Regulatory Quality  
Rule of Law  
Control of Corruption  

## 2. Data  
OHI has already extracted these data by country.  For BHI, we obtained these data on 13 July 
2016 from [the file 'wgi_combined_scores_by_country.csv' on the OHI github site](https://github.com/OHI-Science/ohiprep/tree/master/globalprep/worldbank_wgi/intermediate).

### 2.1 Data attributes  
Colname names in 'wgi_combined_scores_by_country.csv'
    "country": reporting country
    "year": reporting year
    "score_wgi_scale": mean score of six indicators:
        Voice and Accountability
        Political Stability and Absence of Violence/Terrorism
        Government Effectiveness
        Regulatory Quality
        Rule of Law
        Control of Corruption
    "score_ohi_scale": score_wgi_scale rescaled between 0 and 1 (score_wgi_scale range was -
2.5 and 2.5)


## 3. Pressure layer 
The rescaled value of the WGI score is a resilience data layer (eg. higher WGI, greater resilience).  The inverse is the WHI pressure score.  

### 3.1 Current value
The value for each country in 2013  

### 3.2 Rescaling
The WGI scores have a range -2.5 to 2.5. They need to be rescaled to between 0 and 1. Use minimum and maximum value in the dataset.  We compare the pressure layer difference if all countries are included in the dataset when selecting min and max to rescale or if only Baltic region values are used obtain min and max values to rescale.    

## 4. Prepare Data layer

```{r setup}

## Libraries
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(RMySQL)
library(stringr)
library(tools)
library(rprojroot) # install.packages('rprojroot')

## source common libraries, directories, functions, etc
source('~/github/bhi/baltic2015/prep/common.r')

## rprojroot
root <- rprojroot::is_rstudio_project


## make_path() function to 
make_path <- function(...) rprojroot::find_root_file(..., criterion = is_rstudio_project)



dir_layers = make_path('baltic2015/layers') # replaces  file.path(dir_baltic, 'layers')


# root$find_file("README.md")
# 
# root$find_file("ao_need_gl2014.csv")
# 
# root <- find_root_file("install_ohicore.r", 
# 
# withr::with_dir(
#   root_file("DESCRIPTION"))



dir_wgi    = file.path(dir_prep,'pressures/wgi_social')


## add a README.md to the prep directory with the rawgit.com url for viewing on GitHub
create_readme(dir_wgi, 'wgi_pressure_prep.rmd') 
```

### 4.1 Read in and organize data

#### 4.1.1 Read in data
```{r read in data}
## read in data
wgi_global = read.csv(file.path(dir_wgi, 'data_database/wgi_combined_scores_by_country.csv'), sep=";", stringsAsFactors = FALSE)
str(wgi_global)

bhi_lookup = read.csv(file.path(dir_wgi, 'bhi_basin_country_lookup.csv'), sep=";", stringsAsFactors = FALSE) %>%
            select(BHI_ID, rgn_nam) %>%
            dplyr::rename(rgn_id=BHI_ID,
                          country=rgn_nam)


```

#### 4.1.2 Select Baltic Sea countries
```{r select baltic countries}

wgi_baltic = wgi_global %>%
            filter(grepl("Denmark|Estonia|Finland|Germany|Latvia|Lithuania|Poland|Sweden|Russia",country))

wgi_baltic %>% select(country) %>% distinct()
```


### 4.2 Plot WGI scores

####  4.2.1  Plot WGI scores (WGI scale)
```{r Plot raw Baltic WGI scores}
ggplot(wgi_baltic) + 
  geom_point(aes(year, score_wgi_scale, colour = country ))+
   geom_line(aes(year, score_wgi_scale, colour = country ))+
  ylab("WGI Score")+
  ylim(-2.5,2.5)+
  ggtitle("WGI Scores (WGI scale")

```


####  4.2.2  Plot WGI scores  - OHI global rescale
```{r Plot Baltic WGI ohi global rescale scores}
ggplot(wgi_baltic) + 
  geom_point(aes(year, score_ohi_scale, colour = country ))+
   geom_line(aes(year, score_ohi_scale, colour = country ))+
  ylab("Score")+
  ylim(0,1)+
  ggtitle("WGI Scores (OHI global rescale")

```


### 4.3 Rescale to the Baltic region

#### 4.3.1 Select Min and Max
```{r select min and max from the Baltic}
## min
baltic_min = wgi_baltic %>%
             select(score_wgi_scale)%>%
             min()%>%
             as.numeric()
baltic_min

## country and year of min
filter(wgi_baltic, score_wgi_scale == baltic_min)
#         country year score_wgi_scale score_ohi_scale
# Russian Federation 2000      -0.8633578       0.3273284


##max
baltic_max = wgi_baltic %>%
             select(score_wgi_scale)%>%
             max()%>%
             as.numeric()

## country and year of max
filter(wgi_baltic, score_wgi_scale == baltic_max)

# country year score_wgi_scale score_ohi_scale
#Finland 2004        1.985394       0.8970788

```

#### 4.3.2 Rescale Baltic
```{r rescale baltic}
wgi_baltic = wgi_baltic %>%
             mutate(min = baltic_min,
                    max=baltic_max,
             score_bhi_scale = (score_wgi_scale - baltic_min)/(baltic_max - baltic_min))%>%
             select(-min,-max)


```


#### 4.3.3 Plot rescaled to Baltic Region
```{r plot rescaled to baltic }
ggplot(wgi_baltic) + 
  geom_point(aes(year, score_bhi_scale, colour = country ))+
   geom_line(aes(year, score_bhi_scale, colour = country ))+
  ylab("Score")+
  ylim(0,1)+
  ggtitle("WGI Scores (BHI region rescale")


```


#### 4.3.4 Plot and Compare Rescaling options
Using the BHI region rescaling penalizes Russia. Does also spread the scores somewhat.  

```{r Plot and Compare Rescaling options}
ggplot(gather(wgi_baltic, score_type, score, -country,-year) %>% filter(score_type != "score_wgi_scale")) + 
  geom_point(aes(year, score, colour = country ))+
   geom_line(aes(year, score, colour = country ))+
  facet_wrap(~score_type)+
  ylab("Score")+
  ylim(0,1)+
  ggtitle("Rescale comparison WGI Scores")

```


### 4.4 Convert to Pressure score  

#### 4.4.1 Convert to pressure score
Inverse of the score  
```{r convert to pressure score}

wgi_baltic_pressure = wgi_baltic %>%
                      select(-score_wgi_scale)%>%
                      gather(., score_type, score, -country,-year)%>%
                      mutate(pressure_score = 1 - score) %>%
                      select(-score)

```

#### 4.4.2 Plot pressure score
```{r Plot and Compare Rescaling options with Pressure score}
ggplot(wgi_baltic_pressure) + 
  geom_point(aes(year, pressure_score, colour = country ))+
   geom_line(aes(year, pressure_score, colour = country ))+
  facet_wrap(~score_type)+
  ylab("Pressure Score")+
  ylim(0,1)+
  ggtitle("Rescale comparison WGI Pressure Scores")

```

### 4.5 Assign country scores to BHI regions

#### 4.5.1 Join BHI region lookup to country scores
```{r join BHI region lookup to country scores}
wgi_pressure_rgn = wgi_baltic_pressure %>%
                   mutate(country = ifelse(country == "Russian Federation", "Russia", country))%>%
                   full_join(., bhi_lookup, by = "country")

```


## 5. Prepare and save Objects
Select global rescaling so that Russia is not so heavily penalized.  

### 5.1 Prepare object
```{r prepare object}

ss_wgi = wgi_pressure_rgn %>%
         filter(score_type == "score_ohi_scale")%>%
         select(-score_type, -country)%>%
         group_by(rgn_id) %>%
         filter(year == max(year))%>%
         ungroup() %>%
         select(rgn_id, year, pressure_score) %>%
         arrange(rgn_id)

ss_wgi = ss_wgi %>%
        select(-year)
        

```

#### 5.2 Save object
```{r save object}

write.csv(ss_wgi, file.path(dir_layers, 'ss_wgi_bhi2015.csv'), row.names=FALSE)
```

